---
title: "MAS Review - surveyframe 0.3.2"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: false
execute:
  echo: true
  warning: true
  error: true
  eval: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```


**Reviewer:** Mohammed Ali Sharafuddin  
**Date:** 2026-06-13  
**Version under review:** 0.3.2 (implemented locally, not yet on CRAN)  
**Purpose:** Full A-Z review of every feature, UI screen, UX flow, and logical
path before R CMD check and CRAN submission. Sign off after completing every
checklist item.

**Revised 2026-06-15** to match the GUI and report co-review: the builder gained
in-browser Export survey and Generate collector buttons and three distinct
Analyse stages (Part F); SurveyStudio dropped the Build screen, previews the real
survey in a frame, and now contains the response dashboard (Part G); the static
survey gained a footer and a multi-page progress bar (Part H); the HTML report
renders through Quarto with distribution plots and formatted tables (Part Q).

Work through this document sequentially in a fresh RStudio session.

---

## Prerequisites

- [ ] RStudio is open with a clean R session (restart R before starting).
- [ ] A modern browser is available (Chrome or Firefox).
- [ ] `devtools`, `remotes`, `psych`, `rmarkdown`, `shiny`, and `googlesheets4`
  are installed (the last two can be absent — they are Suggests only).
- [ ] Internet access is available for the GitHub install step.

---

## Part A — Installation

### Step A1 — Install the development version from GitHub

This is the exact path a developer or collaborator would use to get the latest
code before it lands on CRAN.

```{r chunk-01, eval=FALSE}
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
remotes::install_github("MohammedAliSharafuddin/surveyframe", force = TRUE)
```

- [ ] Installation completes with no errors.
- [ ] No warning about missing Imports packages (jsonlite, rlang, openssl must
  all install or already be present).

### Step A2 — Confirm version

```{r chunk-02}
packageVersion("surveyframe")
# Expected: '0.3.2'
```

- [ ] Version reads `0.3.2`.

### Step A3 — Confirm available functions

```{r chunk-03}
ls("package:surveyframe")
```

- [ ] The function list is non-empty and includes `sf_instrument`, `sf_item`,
  `sf_choices`, `sf_scale`, `sf_branch`, `sf_check`, `sf_model`,
  `export_static_survey`, `launch_builder`, `launch_studio`, `launch_dashboard`,
  `run_analysis_plan`, `render_report`, `render_results`.

---

## Part B — Package loading and citation

### Step B1 — Load silently

```{r chunk-04}
library(surveyframe)
```

- [ ] No startup messages or warnings.

### Step B2 — Citation (Change 1 of 0.3.2)

```{r chunk-05}
citation("surveyframe")
```

- [ ] Title reads: `surveyframe: Survey Instrument Workflows` (no "for R").
- [ ] The note field shows `R package version 0.3.2`, not the old hard-coded `0.3.0`.
- [ ] Author, ORCID, and URL are correct.

### Step B3 — Help pages load

```{r chunk-06, eval=FALSE}
?sf_instrument
?sf_item
?export_static_survey
?run_analysis_plan
?render_report
```

- [ ] Each help page opens without error.
- [ ] Examples in each page are present and readable.

---

## Part C — S3 methods for component classes (Change 2 of 0.3.2)

### Step C1 — Methods are registered

```{r chunk-07}
methods(class = "sf_choices")
methods(class = "sf_item")
methods(class = "sf_scale")
methods(class = "sf_branch")
methods(class = "sf_check")
methods(class = "sf_model")
```

- [ ] Each call returns at least `format.<class>`, `print.<class>`, `summary.<class>`.

### Step C2 — sf_choices display

```{r chunk-08}
cs <- sf_choices("agree5", 1:5,
                 c("Strongly disagree", "Disagree", "Neutral",
                   "Agree", "Strongly agree"))
cs               # calls print.sf_choices
format(cs)
summary(cs)
```

- [ ] `print(cs)` shows a header line and a two-column value/label table.
- [ ] `format(cs)` returns a compact single character string.
- [ ] `summary(cs)` shows `allow_other` and `randomise` flags.

### Step C3 — sf_item display

```{r chunk-09}
it <- sf_item("q1", "How satisfied are you?",
              type = "likert", choice_set = "agree5", required = TRUE)
it
format(it)
summary(it)
```

- [ ] `print(it)` shows id, type, label, choice set.
- [ ] `summary(it)` shows required and reverse flags.

### Step C4 — sf_scale display

```{r chunk-10}
sc <- sf_scale("sat", "Satisfaction", items = c("q1", "q2", "q3"))
sc
format(sc)
summary(sc)
```

- [ ] `print(sc)` shows id, item count, items list, and scoring method.
- [ ] `format(sc)` says "3 item(s)".

### Step C5 — sf_branch display (real field names: item_id, depends_on)

```{r chunk-11}
br <- sf_branch("q2", depends_on = "q1", operator = "==",
                value = "yes", action = "show")
br
format(br)
summary(br)
```

- [ ] `print(br)` shows the rule in readable English: "show when q1 == yes".
- [ ] No error about missing or wrong field names.

### Step C6 — sf_check display

```{r chunk-12}
ck <- sf_check("attn1", item_id = "q5", type = "attention",
               pass_values = 3, fail_action = "flag")
ck
format(ck)
summary(ck)
```

- [ ] `print(ck)` shows id, type, item_id, and pass values.
- [ ] `summary(ck)` shows `fail_action`.

### Step C7 — sf_model display

```{r chunk-13}
mdl <- sf_model("m1", label = "Measurement model", type = "cfa")
mdl
format(mdl)
summary(mdl)
```

- [ ] `print(mdl)` shows id, type, 0 construct(s), label, and engine.
- [ ] `summary(mdl)` shows engine and structural path count.

---

## Part D — Instrument construction

### Step D1 — Build a minimal instrument in code

```{r chunk-14}
agree5 <- sf_choices("agree5", 1:5,
                     c("Strongly disagree", "Disagree", "Neutral",
                       "Agree", "Strongly agree"))

q1 <- sf_item("q1", "The service met my expectations.",
              type = "likert", choice_set = "agree5",
              scale_id = "sat", required = TRUE)
q2 <- sf_item("q2", "I would recommend this service.",
              type = "likert", choice_set = "agree5",
              scale_id = "sat", required = TRUE)
q3 <- sf_item("attn_q", "Please select 'Agree' for this item.",
              type = "single_choice", choice_set = "agree5", required = TRUE)

sat <- sf_scale("sat", "Satisfaction", items = c("q1", "q2"), method = "mean")
chk <- sf_check("attn1", item_id = "attn_q", type = "attention",
                pass_values = 4, fail_action = "flag")

instr <- sf_instrument(
  title       = "MAS Review Test Survey",
  version     = "1.0.0",
  description = "Minimal instrument for the 0.3.2 review.",
  authors     = "Mohammed Ali Sharafuddin",
  components  = list(agree5, q1, q2, q3, sat, chk)
)

print(instr)

# Header and branding (shown on the exported survey in Part H).
# The logo is pulled from the public GitHub repo for this test. Internet is
# needed for this one step. openssl is a hard dependency, so base64_encode is
# always available.
logo_url <- paste0("https://raw.githubusercontent.com/MohammedAliSharafuddin/",
                   "surveyframe/main/man/figures/readme-logo.png")
logo_png <- tempfile(fileext = ".png")
download.file(logo_url, logo_png, mode = "wb", quiet = TRUE)

instr$render$theme                  <- "#16B3B1"
instr$render$header$institution     <- "surveyframe demo"
instr$render$header$logo_base64     <- openssl::base64_encode(
  readBin(logo_png, "raw", file.info(logo_png)$size))
instr$render$header$logo_media_type <- "image/png"
instr$render$header$show_progress   <- TRUE
```

- [ ] `print(instr)` shows a tidy summary: title, item count, scale count,
  check count, and validation status.
- [ ] No error during construction.
- [ ] The logo downloads without error and
  `nchar(instr$render$header$logo_base64)` is a few thousand characters.
- [ ] The institution, logo, and `#16B3B1` theme appear on the exported survey
  header in Part H. (If offline, skip the download lines; the rest still runs.)

### Step D2 — Validation

`validate_sframe()` defaults to `strict = TRUE`, which returns the instrument
(and errors if it is invalid). Pass `strict = FALSE` to get a result list with
`valid` and `problems`.

```{r chunk-15}
v <- validate_sframe(instr, strict = FALSE)
v$valid
v$problems
```

- [ ] `v$valid` is `TRUE`.
- [ ] `v$problems` is empty (length 0) or contains only expected informational notes.

### Step D3 — Declare the methodological contract (model and analysis plan)

The instrument is a pre-declared contract, so the measurement model and the
analysis plan are part of design, not an afterthought. A measurement model is
attached in code with `add_model()`. The analysis plan is authored in the
builder's Analyse tab (Part F) and saved into the `.sframe`, so a plan declared
there travels with the instrument and is executed later in Part O.

```{r chunk-16}
m <- sf_model("sat_cfa", type = "cfa",
  constructs = list(list(id = "sat", items = c("q1", "q2"))))
instr <- add_model(instr, m)
length(instr$models)        # 1
length(instr$analysis_plan) # plans authored in the builder appear here
```

- [ ] `length(instr$models)` is 1 after `add_model()`.
- [ ] You understand that the analysis plan is declared in the builder Analyse
  tab (Part F) or loaded from a designed `.sframe`, and executed on data in
  Part O. There is no separate code constructor for plan blocks in 0.3.2.

---

## Part E — Serialisation (write and read round-trip)

### Step E1 — Write and read back

```{r chunk-17}
tmp <- tempfile(fileext = ".sframe")
write_sframe(instr, tmp)
instr2 <- read_sframe(tmp)

identical(instr$meta$title, instr2$meta$title)
identical(length(instr$items), length(instr2$items))
```

- [ ] Both `identical()` calls return `TRUE`.
- [ ] No hash mismatch error on `read_sframe()`.

### Step E2 — Hash integrity check

```{r chunk-18}
# Tamper with the file to confirm the check fires
raw <- readLines(tmp)
raw[2] <- gsub("a", "b", raw[2], fixed = TRUE)
tmp2 <- tempfile(fileext = ".sframe")
writeLines(raw, tmp2)
tryCatch(read_sframe(tmp2), error = function(e) message("Caught: ", conditionMessage(e)))
```

- [ ] An error is caught mentioning hash or integrity.

---

## Part F — SurveyBuilder HTML

### Step F1 — Launch and explore

```{r chunk-19, eval=FALSE}
launch_builder(open = TRUE)
```

A browser tab opens with the SurveyBuilder HTML. Work through the UI:

- [ ] The builder loads without a blank screen or JavaScript errors (open the
  browser console — no red errors).
- [ ] You can add a new question using the Add Item button.
- [ ] The item type dropdown includes all 13 types: Likert, Single choice,
  Multiple choice, Matrix, Numeric, Text, Long text, Date, Slider, Rating,
  Ranking, Section break, Text block.
- [ ] You can add a choice set and assign it to a Likert item.
- [ ] You can add a scale and assign items to it.
- [ ] The Save .sframe button downloads a `.sframe` file.
- [ ] The Open button accepts the saved file and restores the instrument.
- [ ] The Analyse tab shows three distinct stages with captions: Plan, Run
  preview (drops syntax-only methods), and Report outline (adds the models
  section). They do not show identical content.
- [ ] You can declare an analysis plan and drag plan cards to reorder them on
  the Plan tab. This is where the analysis plan is authored as part of design.
- [ ] The Export survey button downloads a deployable HTML survey.
- [ ] Survey settings > Google Sheets has a Generate collector (.gs) button and
  numbered deploy steps. The settings gear button is clearly visible.
- [ ] No broken images, garbled text, or mojibake anywhere in the UI.

### Step F3 — Exported output from the builder

- [ ] Open the survey HTML exported in Step F1 (Export survey). It opens as a
  working survey with the welcome screen, sections, and a "Built with
  surveyframe" footer, identical to `export_static_survey()` output.

### Step F2 — Demo mode

```{r chunk-20, eval=FALSE}
launch_builder_demo()
```

- [ ] The builder opens pre-loaded with the demo instrument.
- [ ] Items, choice sets, and scales are all visible without a manual load step.

---

## Part G — SurveyStudio Shiny app

SurveyStudio is the analysis hub. Survey design (questions, welcome, logo,
thank-you, analysis plan, model) lives in the builder; the studio opens a
designed `.sframe`, previews it, and analyses responses.

### Step G1 — Launch studio

```{r chunk-21, eval=FALSE}
launch_studio_demo()
```

- [ ] Shiny app launches in the browser with the demo instrument and 120
  responses already loaded.
- [ ] The left navigation shows eight screens, in order: Open Instrument,
  Preview Survey, Upload Responses, Quality Dashboard, Reliability, Analysis
  Plan, Dashboard, Export. There is NO "Build Survey" screen (design is done in
  the builder).
- [ ] The logo and brand colours display correctly. No R console error.

### Step G2 — Open and preview

- [ ] Open Instrument loads a `.sframe` (or the demo is already loaded). The hint
  says design happens in the SurveyBuilder.
- [ ] Preview Survey shows the EXACT deployable survey inside a frame, with the
  welcome screen, sections, progress, and "Built with surveyframe" footer. It
  matches the static export, not a set of native widgets.

### Step G3 — Sample data and analysis

- [ ] Upload Responses has a "Load sample survey and responses" button and a
  "Download sample CSV" link.
- [ ] Quality Dashboard, Reliability, and Analysis Plan all render content (not
  blank) once data is loaded.
- [ ] Analysis Plan shows three stages: Plan, Run preview, Report outline. The
  variables panel excludes section breaks and text blocks. The 36 demo plans
  appear in the saved-plans table.
- [ ] Run preview executes the data-driven plans and excludes the syntax-only
  methods with a note.

### Step G4 — Integrated dashboard and export

- [ ] The Dashboard screen shows Overview, Items, Scales, and Data inline (KPIs,
  item bar charts, scale histograms, a raw-data table). Charts render, not blank.
- [ ] The Export screen offers Download .sframe and Generate HTML report only,
  with a note that the deployable survey HTML and the Sheets collector are
  produced in the builder.

---

## Part H — Static HTML survey export and browser review

### Step H1 — Export a survey

```{r chunk-22, eval=FALSE}
out <- export_static_survey(
  instr,
  output_path = file.path(tempdir(), "mas_review_survey.html"),
  open        = TRUE,
  overwrite   = TRUE
)
cat("Written to:", out, "\n")
cat("Size:", round(file.size(out) / 1024, 1), "KB\n")
```

- [ ] File is written without error.
- [ ] A message confirms the path and size in KB.
- [ ] The browser opens automatically.

### Step H2 — Header and welcome screen

In the browser:

- [ ] The header band shows the GitHub logo set in Step D1 and the institution
  name "surveyframe demo". The logo is sized consistently (not stretched).
- [ ] The brand theme colour (`#16B3B1`) is applied to the accent and buttons.
- [ ] A welcome screen appears before any questions.
- [ ] The survey title "MAS Review Test Survey" is visible.
- [ ] A Start button is present.

### Step H3 — Question rendering

Click Start and review the question pages:

- [ ] The Likert item (`q1`) renders with options in a horizontal row (one
  column per option, label below each radio button).
- [ ] The second Likert item (`q2`) renders in the same horizontal layout.
- [ ] The attention check item (`attn_q`) — declared as `single_choice` but
  designated as a check — also renders in the horizontal Likert layout, not
  as a vertical stacked list. (This is Change 5 of 0.3.2.)
- [ ] The required marker (`*`) appears next to all required items.
- [ ] This minimal instrument is a single page, so NO progress bar shows (the
  bar appears only on multi-page surveys). Confirm with the multi-page input-types
  demo (`export_static_survey(sframe_input_types_demo_data()$instrument, ...)`):
  the bar reads "Page X of Y" and advances.
- [ ] A "Built with surveyframe" footer appears at the foot of the survey.

### Step H4 — Validation and submission

- [ ] Clicking Next without answering a required item shows an inline error
  below the item AND a page-level banner at the top of the card.
- [ ] Answering the items and clicking Submit downloads a CSV file.
- [ ] Open the downloaded CSV: it has a `respondent_id` column and one column
  per item.

### Step H5 — Mobile layout

Open the exported HTML file on a phone (share the file via QR code, AirDrop,
or email), or use Chrome DevTools > Toggle device toolbar to simulate a phone:

- [ ] All items are readable on a narrow screen without horizontal scrolling.
- [ ] The Likert row wraps gracefully on very small screens (360px wide).
- [ ] Buttons are large enough to tap.

---

## Part I — Shiny survey module

### Step I1 — Embed the module in a minimal app

```{r chunk-23, eval=FALSE}
library(shiny)

ui <- fluidPage(
  titlePanel("Module test"),
  survey_module_ui("test_survey")
)

server <- function(input, output, session) {
  resp <- survey_module_server("test_survey", instrument = instr)
  observeEvent(resp(), {
    showNotification(paste("Submitted:", length(resp()), "responses"))
  })
}

shinyApp(ui, server)
```

- [ ] The app launches and the survey renders inside the page.
- [ ] Completing and submitting the survey fires the `observeEvent` notification.
- [x] No console warnings. A `sprintf` "one argument not used by format" warning
  (the module CSS passed `theme` twice but the format has a single `%s`, the
  second reference uses the CSS variable `var(--cp)`) was found during this
  review and fixed in `R/survey_module.R`.

---

## Part J — Demo data and response pipeline

### Step J1 — Load demo data

```{r chunk-24}
demo  <- sframe_demo_data()
instr_d <- demo$instrument
resp    <- demo$responses

cat("Items:", length(instr_d$items), "\n")
cat("Scales:", length(instr_d$scales), "\n")
cat("Responses:", nrow(resp), "\n")
```

- [ ] Counts are non-zero and match the tourism demo (15 items, 5 scales, 120
  responses). `sframe_demo_data()` loads `tourism_services_demo.sframe`, which
  also carries 34 analysis plans and 2 models.

### Step J2 — read_responses with a bundled CSV

```{r chunk-25}
# Export a one-row CSV from the static survey, then read it back
resp_csv <- file.path(tempdir(), "test_response.csv")

# Simulate what the browser download produces
fake_row <- as.list(setNames(
  rep(3, length(instr$items)),
  sapply(instr$items, function(i) i$id)
))
fake_row$respondent_id <- "R001"
fake_row$started_at    <- as.character(Sys.time())

write.csv(as.data.frame(fake_row), resp_csv, row.names = FALSE)

# Declare the meta columns. read_responses() is strict by default and rejects
# columns that are neither survey items nor declared meta columns. The
# respondent id has its own argument; other meta columns go in meta_cols.
read_resp <- read_responses(
  resp_csv, instr,
  respondent_id = "respondent_id",
  meta_cols     = "started_at"
)
nrow(read_resp)
```

- [ ] Returns a one-row data frame.
- [ ] No error. (Without declaring `respondent_id` and `started_at`, strict mode
  errors with "undeclared column(s)" — pass them as shown, or use
  `strict = FALSE`.)

---

## Part K — Quality report

```{r chunk-26}
qr <- quality_report(resp, instr_d)

cat("Respondents:", qr$summary$n_respondents, "\n")
cat("Flagged:",     qr$summary$n_flagged, "\n")
print(qr)
```

- [ ] `$n_respondents` matches the number of rows in `resp`.
- [ ] `print(qr)` shows a readable summary — no raw list output.
- [ ] No error about attention check items.

---

## Part L — Scale scoring

```{r chunk-27}
scored <- score_scales(resp, instr_d)

# Derive the scored columns from the instrument's scales (score_scales adds one
# column per scale, named by the scale id). Do not hard-code names.
scale_cols <- vapply(instr_d$scales, function(s) s$id, character(1))

# Prof-10: should display as a labelled data frame with Mean and SD
scale_summary <- data.frame(
  Scale = scale_cols,
  Mean  = round(colMeans(scored[, scale_cols, drop = FALSE], na.rm = TRUE), 2),
  SD    = round(apply(scored[, scale_cols, drop = FALSE], 2, sd, na.rm = TRUE), 2),
  row.names = NULL
)
print(scale_summary)
```

- [ ] `scored` has all original columns plus one per scale.
- [ ] Scale means are between 1 and 5 (Likert range).
- [ ] The data frame prints with Scale, Mean, and SD columns.

---

## Part M — Psychometric reports

### Step M1 — Missing data report

```{r chunk-28}
mr <- missing_data_report(resp, instr_d)
print(mr)
```

- [ ] Prints a summary of missing values per item.
- [ ] `mcar` slot returns `available = FALSE` (Little's MCAR is not implemented
  in 0.3.2; this is documented behaviour, not a bug).

### Step M2 — Reliability report

```{r chunk-29}
if (requireNamespace("psych", quietly = TRUE)) {
  rr <- reliability_report(scored, instr_d, omega = FALSE)
  print(rr)
}
```

- [ ] One reliability row per scale.
- [ ] Alpha values are between 0 and 1.
- [ ] No psych internal warnings printed to the console.

### Step M3 — Item report

```{r chunk-30}
ir <- item_report(resp, instr_d)
print(ir)
```

- [ ] Returns item-level statistics (mean, SD, skewness for each item).

### Step M4 — EFA readiness report

```{r chunk-31}
if (requireNamespace("psych", quietly = TRUE)) {
  er <- efa_report(scored, instr_d)
  print(er)
}
```

- [ ] KMO overall value is printed.
- [ ] Bartlett's test chi-square and p-value are printed.
- [ ] Suggested number of factors is printed.
- [ ] No errors.

### Step M5 — Validity report

```{r chunk-32}
loadings_list <- list(
  DMRE = c(dm_1 = 0.78, dm_2 = 0.82, dm_3 = 0.75, dm_4 = 0.80, dm_5 = 0.77),
  TS   = c(ts_1 = 0.84, ts_2 = 0.80, ts_3 = 0.88)
)
vr <- validity_report(loadings_list)
print(vr$reliability)
```

- [ ] CR and AVE are computed for each construct.
- [ ] AVE is between 0 and 1.
- [ ] No error.

### Step M6 — Assumption report

```{r chunk-33}
if (requireNamespace("psych", quietly = TRUE)) {
  ar <- assumption_report(scored, instr_d)
  print(ar)
}
```

- [ ] Normality results are shown per scale variable.
- [ ] No error.

---

## Part N — CFA and SEM syntax generation

`cfa_lavaan_syntax()` is instrument-first and builds a CFA from the
instrument's scales. `sem_lavaan_syntax()` and `seminr_syntax()` are model-first,
so pass a model. The tourism demo defines a `cfa` model and a `cb_sem` model;
use the CB-SEM model for the structural syntax.

```{r chunk-34}
# CFA syntax from the instrument's scales (instrument-first)
syn_cfa <- cfa_lavaan_syntax(instr_d)
cat(syn_cfa)

# The CB-SEM model carries the structural paths
sem_model <- Filter(function(m) m$type == "cb_sem", instr_d$models)[[1]]

syn_sem <- sem_lavaan_syntax(sem_model, instr_d)
cat(syn_sem)

syn_pls <- seminr_syntax(sem_model)
cat(syn_pls)
```

- [ ] `cfa_lavaan_syntax()` returns a multi-line lavaan model string
  (e.g., `digital_marketing =~ dm_1 + dm_2 + ...`).
- [ ] `sem_lavaan_syntax()` includes the measurement model and the declared paths.
- [ ] `seminr_syntax()` returns the seminr measurement and structural model syntax.
- [ ] None of these calls require lavaan or seminr to be loaded.

---

## Part O — Analysis plan

### Step O1 — Declare and run the plan

```{r chunk-35}
results <- run_analysis_plan(scored, instr_d)
length(results)
```

- [ ] Returns a list with one element per research question.
- [ ] No error.

### Step O2 — Inspect a result

```{r chunk-36}
r1 <- results[[1]]
cat("APA:    ", r1$apa,    "\n")
cat("Prompt: ", r1$prompt, "\n")
cat("Test:   ", r1$test,   "\n")
```

- [ ] `$apa` is a formatted string (e.g., `r(60) = 0.43, p = .002`).
- [ ] `$prompt` is a complete sentence template.
- [ ] `$test` names the statistical method used.
- [ ] No `$table`, `$plot`, or `$reference` field is present at the runner level
  (these do not exist yet in 0.3.2).

---

## Part P — Codebook report

```{r chunk-37}
cb <- codebook_report(instr_d)
nrow(cb$items_table)
head(cb$items_table[, c("id", "type", "scale_id", "reverse")], 6)
```

- [ ] `nrow(cb$items_table)` equals the number of items.
- [ ] The first 6 rows show correct ids and types.
- [ ] No error. Note: `codebook_report()` has no `path` argument — it returns
  an object, not a file.

---

## Part Q — HTML reporting

### Step Q1 — render_results

```{r chunk-38, eval=FALSE}
results_path <- render_results(
  results,
  instr_d,
  output_file     = file.path(tempdir(), "mas_results.html"),
  citation_format = "apa"
)
cat("Path:", results_path, "\n")
cat("Size:", round(file.size(results_path) / 1024, 1), "KB\n")
```

- [ ] File is written without error.
- [ ] `file.size()` returns a positive number.
- [ ] Open the HTML: each research question has its own section with the APA
  statistic, effect size, and writing prompt.

### Step Q2 — render_report (full report)

```{r chunk-39, eval=FALSE}
report_path <- render_report(
  instr_d,
  data                 = scored,
  output_file          = file.path(tempdir(), "mas_full_report.html"),
  include_codebook     = TRUE,
  include_quality      = TRUE,
  include_missing      = TRUE,
  include_descriptives = TRUE,
  include_reliability  = TRUE,
  include_analysis     = TRUE,
  include_models       = FALSE
)
```

- [ ] File is written without error. `render_report()` renders through Quarto
  when it is installed (themed HTML with a left table of contents) and falls back
  to a built-in self-contained HTML when Quarto is absent. There is no PDF path.
- [ ] Open the HTML: sections appear in order (codebook, quality, missing data,
  descriptives, response distributions, reliability, analysis results).
- [ ] A "Response distributions" section shows one chart per item (bar chart for
  categorical, histogram for numeric) and one histogram per scale.
- [ ] No raw markdown or raw R output leaks in (all tables render as HTML tables,
  none show pipe characters).
- [ ] Tables are padded and readable; wide tables scroll within the column rather
  than overflowing; numeric values are rounded to two decimal places.

---

## Part R — Response dashboard

This standalone read-only dashboard still ships. Its core views (Overview, Items,
Scales, Data) are now also built into SurveyStudio (Part G, Step G4), so the
standalone launcher overlaps the studio. It is kept for the lightweight
view-only use case and is a candidate for soft-deprecation in a later release.

```{r chunk-40, eval=FALSE}
launch_dashboard_demo()
```

- [ ] Dashboard opens in the browser automatically.
- [ ] Five tabs are visible: Overview, Items, Scales, Quality, Raw data.
- [ ] Overview shows response count, date range, and instrument metadata.
- [ ] Items tab shows per-item charts. Charts appear when you click the tab
  (they do not stay blank).
- [ ] Scales tab shows scale score distributions. Charts appear when you click.
- [ ] Quality tab shows attention-check pass rates.
- [ ] Raw data tab shows a scrollable table with a CSV download button.
- [ ] No JavaScript errors in the browser console.

---

## Part S — lavaan in Suggests (Change 3 of 0.3.2)

```{r chunk-41}
# Confirm lavaan is listed in Suggests, not Imports
desc <- read.dcf(system.file("DESCRIPTION", package = "surveyframe"))
cat("Imports:", desc[, "Imports"], "\n\n")
cat("Suggests:", desc[, "Suggests"], "\n")
```

- [ ] `lavaan` appears in Suggests.
- [ ] `lavaan` does NOT appear in Imports.
- [ ] The three hard Imports are exactly: `jsonlite`, `rlang`, `openssl`.

---

## Part T — replicate.R review (surveyframe-dev, Change 4 of 0.3.2)

Open `~/Documents/GitHub/surveyframe-dev/jss-paper/replicate.R` in RStudio.

Read through the full script:

- [ ] No `str()` calls anywhere in the file.
- [ ] `export_static_survey()` call is present and inside `if (interactive()) { }`.
- [ ] `render_results()` call is present and inside `if (interactive()) { }`.
- [ ] Every other browser-opening or Shiny call (if any) is also guarded.

Run the script from the RStudio console (interactive mode, so the guarded
sections execute):

```{r chunk-42, eval=FALSE}
source("~/Documents/GitHub/surveyframe-dev/jss-paper/replicate.R", echo = TRUE)
```

- [ ] Script runs end to end without error.
- [ ] The `export_static_survey()` guard fires and prints "Static survey written to: ..."
  plus the file size.
- [ ] The `render_results()` guard fires and prints "Report written to: ..." plus
  the file size.
- [ ] `sessionInfo()` prints at the end.

---

## Part U — Vignette review (Change 6 of 0.3.2)

Knit the main vignette:

```{r chunk-43, eval=FALSE}
rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/surveyframe.Rmd",
  output_dir = tempdir(),
  quiet      = FALSE
)
```

- [ ] Knit completes with no errors. Warnings from psych internals are expected
  and acceptable.
- [ ] Open the output HTML in a browser.

Work through the vignette output from top to bottom:

#### Score and describe section

- [ ] A note below `score_scales()` explains that pre-scored columns are skipped.
- [ ] Scale means display as a labelled data frame with Scale, Mean, and SD
  columns — not as a raw named numeric vector.

#### Reliability section

- [ ] The interpretation note explains when to prefer omega over alpha:
  specifically when items within a scale differ in their factor loadings. The
  reason is stated, not just the recommendation.

#### EFA readiness section

- [ ] This section exists between Reliability and Analysis Plan.
- [ ] `efa_report()` is called inside a `psych` availability check.
- [ ] KMO threshold (0.60) and Bartlett's test are mentioned in plain language.

#### Construct validity section

- [ ] This section follows EFA readiness.
- [ ] `validity_report()` is called with a named loadings list.
- [ ] CR and AVE thresholds are stated (AVE above 0.50, CR above 0.70).
- [ ] The Fornell-Larcker criterion is explained in one sentence.

#### Assumption checks section

- [ ] This section appears between the Analysis Plan and Run the Analysis Plan.
- [ ] `assumption_report()` is called and its output is printed.
- [ ] A note explains how to use normality and homogeneity results to choose
  between parametric and non-parametric tests.

#### Results report section

- [ ] After `render_results()`, the code shows both the file path and the file
  size in KB via `file.size()`.

#### Other vignettes

Open each supporting vignette and skim for obvious errors:

```{r chunk-44, eval=FALSE}
rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/building-survey-instrument.Rmd",
  output_dir = tempdir(), quiet = TRUE)

rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/analysing-survey-responses.Rmd",
  output_dir = tempdir(), quiet = TRUE)

rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/scale-reliability-validity.Rmd",
  output_dir = tempdir(), quiet = TRUE)

rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/efa-cfa-sem-pls-syntax.Rmd",
  output_dir = tempdir(), quiet = TRUE)

rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/surveybuilder-gui-overview.Rmd",
  output_dir = tempdir(), quiet = TRUE)

rmarkdown::render(
  "~/Documents/GitHub/surveyframe/vignettes/deploying-and-collecting.Rmd",
  output_dir = tempdir(), quiet = TRUE)
```

- [ ] All six supporting vignettes knit without errors.
- [ ] `surveybuilder-gui-overview.Rmd` describes eight studio screens (no Build
  Survey screen) and the builder export buttons.
- [ ] `deploying-and-collecting.Rmd` (new) covers the Apps Script collector,
  GitHub Pages and Blogger hosting, and reading responses back.

---

## Part V — Full test suite

```{r chunk-45}
devtools::test()
```

- [ ] All 407 tests pass (FAIL 0, WARN 0, SKIP 0, PASS 407).
- [ ] No new unexpected warnings in the test output.

---

## Part W — Final logic and UX checks

Answer each question based on your review above:

- [ ] Every exported function that receives a non-sframe object returns a clear,
  actionable error message pointing the user to `sf_instrument()` or
  `read_sframe()`. No raw `inherits()` assertion failures.
- [ ] No output from any function goes to the console unintentionally (psych
  suppressions are in place, run_analysis_plan is silent on assignment).
- [ ] The static survey works on a phone-sized screen (Step H5).
- [ ] The dashboard charts actually render when their tab is clicked, not blank.
- [ ] The S3 methods for all 6 sub-classes produce readable, professional output
  that you would be comfortable showing in a JSS article.
- [ ] The CITATION is correct and will display correctly on the CRAN page after
  the 0.3.2 tarball is accepted.

---

## Sign-off

After all checkboxes above are ticked, record your sign-off here and inform the
Claude Code session to proceed to `R CMD check`.

### Partial sign-off (2026-06-16)

- **Parts A through H: SIGNED OFF.** Reviewed against the freshly installed
  0.3.2 from the public repo. Covers installation, citation, the six S3 methods,
  instrument construction (now with a header, logo, and brand theme in Step D1),
  serialisation and hash integrity, the SurveyBuilder (distinct Analyse stages,
  Export survey and Generate collector buttons, footer, academic tables), the
  SurveyStudio (eight screens, no Build screen, iframe Preview, integrated
  Dashboard with brand pill tabs and academic tables, Quality elaborated, brand
  teal, alignment fix), and the static survey export with its header logo.
- **Shiny survey module (Part I): IN REVIEW.** A `sprintf` "one argument not
  used" warning was found and fixed in `R/survey_module.R`.
- **Standalone dashboard (Part R):** logo serving fixed (`addResourcePath`);
  `launch_dashboard()` no longer silently loads demo data.

Remaining for full sign-off: the module submit flow (Part I, in progress), the
analysis/reporting/dashboard parts (J onward), and the vignette read-through
(Part U).

**Reviewed by:** Mohammed Ali Sharafuddin  
**Date signed off:** 2026-06-16 (Parts A through H)  
**All steps complete:** Parts A-H: Yes. Part I onward: in progress.  
**Notes for the developer (if any):**

Once signed off, paste the following prompt into Claude Code to proceed:

```text
Read CLAUDE.md and mas_review_032.md. The MAS co-review is complete and signed
off. Run R CMD build . to produce surveyframe_0.3.2.tar.gz, then run
R CMD check --as-cran on the tarball. The target is 0 errors, 0 warnings, at
most 1 NOTE. Update cran-comments.md with the verified results and give me the
text to paste on the CRAN submission form.
```
