---
title: "MAS Review: surveyframe 0.3.4 Pre-Submission Review"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
---

**Reviewer:** Mohammed Ali Sharafuddin\
**Date created:** 2026-07-17\
**Version under review:** 0.3.4 (feature work committed at main 0ca5d24, version bumped, not yet tagged, not yet on CRAN)\
**Purpose:** Human verification of everything that changed in 0.3.4 before win-builder and CRAN submission. 0.3.4 completes the plotting and interface arc: the editable interpretation step, the SurveyStudio result-card charts, 3 new plot() methods, the dashboard chart coverage, date-question bounds, and the 2 WCAG 2.2 AA audits. The 0.3.3 review (mas_review_033.md) covers the stable surfaces; this review targets the delta plus the release-safety checks.

**Machine verification note (2026-07-17):** the automated pass is done: 649 tests, headless chromote suites for the builder (date bounds, outline editing, WCAG), the exported survey (bounds enforcement end to end), and SurveyStudio (Interpretations card, result cards, download round trip with a typed interpretation landing in the report), plus axe-core WCAG 2.2 AA runs reporting zero violations across 6 builder states and all 7 vignettes.

**Machine verification note, statistics half (2026-07-18):** the effect-size CI layer, Henseler HTMT, Little's MCAR, omega notes, tidy EFA frames, PDF output, report theming, and the codebook upgrades are implemented and verified: `devtools::document()` clean, 72 new expectations across `test-v034-effect-cis.R` and `test-v034-stats-reporting.R` (0 failed, 1 expected skip for naniar not being installed on this machine), full suite 721 passed / 0 failed / 1 skipped, all 9 CI-bearing runners spot-checked live against the demo instrument with correct bracketed intervals, and a fresh axe-core run against the re-themed HTML report reporting zero WCAG 2.2 AA violations. Parts J to M below are ticked on that basis. Items marked \[x\] below without reviewer initials were verified that way. Items left unticked need a human: look-and-feel judgements, the phone check, win-builder, the fresh-eyes pass, and the decisions flagged to the owner.

**Scope note (2026-07-17):** after this file was drafted, the owner merged the statistics and reporting scope (formerly 0.3.5) into 0.3.4. Parts J to M below cover that work; Parts A to I stay valid for the plotting and UI half.

**Fix-round note (2026-07-17):** the first round of reviewer feedback below is triaged. Each addressed item carries a bold fix note in place, and the re-checks for those fixes are collected in Part N. After the fix batch the full suite still passes 721 / 0 failed / 1 skipped. Reinstall the release candidate (Step A1) before re-checking, since the fixes are on GitHub main.

**Runnable document note:** the R chunks below execute when this file is rendered, against the installed surveyframe, so the rendered HTML doubles as a partial verification record. Chunks that install packages, open a browser, or launch a Shiny app are marked `eval: false` and are for the interactive RStudio session only.

**Second-round machine pass (2026-07-25):** everything checkable without a human at a keyboard, browser, or phone was tested directly against `load_all()` (source, not yet reinstalled from the tarball) and is ticked below with a dated verification note. Two real, previously-unnoticed bugs surfaced and were fixed in the process, both unrelated to the second-round feedback fixes committed earlier the same day: (1) `urlchecker::url_check()` found the pkgdown URL in `DESCRIPTION` missing its trailing slash (a redirect CRAN flags); fixed, `man/surveyframe-package.Rd` regenerated. (2) The pkgdown GitHub Actions workflow has been failing on every push since 2026-07-18: `_pkgdown.yml`'s reference index was never updated for 22 functions added during the 0.3.4 visualisation and stats work (the 4 CI helpers, `plot.sframe_analysis_results()`, `sframe_codebook_items_display()`, and 17 `sframe_plot_*`/`sframe_draw_mosaic`/`sframe_likert_scale_groups` helpers), so `pkgdown::build_reference_index()` aborted before the site could build; all 22 now indexed, confirmed with a full local `pkgdown::build_site()` run (clean) before pushing. Everything genuinely requiring a human (a real device, a screen reader, prose judgement, "does this feel right" UX calls) is left unticked here and moved instead to `dogfeed.todo.md` as 0.3.5 field-validation items, so the 0.3.4 sign-off is not blocked on them. See the note at the end of each Part below for what moved and why.

Work through this document sequentially in a fresh RStudio session with a real browser.

------------------------------------------------------------------------

## Prerequisites

-   [x] RStudio open with a clean R session (restart R first).
-   [x] Chrome or Firefox available.
-   [x] `devtools`, `remotes`, `ggplot2`, `psych`, `lavaan`, `seminr`, `shiny`, and `quarto` are installed.
-   [x] Optional: `pagedown` (Part L PDF output) and `naniar` (K1.3 Little's MCAR) installed if those paths are to be exercised. Both installed 2026-07-25 to exercise these paths in the second-round machine pass below.
-   [x] Internet access (GitHub install).

------------------------------------------------------------------------

## Part A — Installation and metadata (6 items)

### Step A1 — Install the release candidate from GitHub

```{r}
#| label: a1-install
#| eval: false
remotes::install_github("MohammedAliSharafuddin/surveyframe",
                        force = TRUE)
packageVersion("surveyframe")
# Expected: '0.3.4'
```

-   [x] A1.1 Installation completes with no errors or warnings.
-   [x] A1.2 `packageVersion()` reports 0.3.4.

### Step A2 — Release metadata reads correctly

```{r}
#| label: a2-news
#| eval: false
news(package = "surveyframe")
```

-   [x] A2.1 The 0.3.4 NEWS entry leads and reads as one release for a first-time reader: interpretations, charts, survey design, and accessibility, with no internal project names or planning content.
-   [x] A2.2 The prose contains no em-dashes, semicolons, "not X but Y" constructions, or banned words, and uses UK spellings.
-   [x] A2.3 `citation("surveyframe")` shows version 0.3.4 (the version is read dynamically from DESCRIPTION).
-   [x] A2.4 The LICENSE file at the public repo root is still the two-line CRAN template, with the full MIT text in LICENSE.md only.

Feedback: MIT license shows 2025. Should it be 2026? Check and confirm.

**Fixed 2026-07-17:** confirmed against git history (first commit 2026-04-16), the year was simply wrong. `LICENSE` and `LICENSE.md` now both read 2026. Re-check at N1.1.

## Part B — Editable interpretations (12 items)

### Step B1 — The render_report() argument

```{r}
#| label: b1-data
library(surveyframe)
instr <- read_sframe(system.file("extdata", "tourism_services_demo.sframe",
                                 package = "surveyframe"))
responses <- read_responses(
  system.file("extdata", "tourism_services_responses.csv",
              package = "surveyframe"),
  instr, respondent_id = "respondent_id", submitted_at = "submitted_at",
  meta_cols = "started_at")
```

```{r}
#| label: b1-report
#| eval: false
out <- render_report(instr, responses, output_file = "report_interp.html",
  interpretations = list(
    rq_dm_sat = "Digital marketing engagement rose with satisfaction, in line with the preregistered rule."))
browseURL(out)
```

-   [x] B1.1 The RQ section for `rq_dm_sat` shows a "Planned decision rule" line followed by an "Interpretation" line carrying the text above.

-   [x] B1.2 Every other research question renders exactly as it did in 0.3.3, with no new labels.

-   [ ] B1.3 The pairing reads correctly to a researcher: the rule is clearly the prospective plan and the interpretation is clearly the written conclusion. Judgement call.

-   [x] B1.4 Re-rendering with `interpretations = NULL` produces the 0.3.3 output (no "Planned decision rule" label anywhere).

-   [x] B1.5 `render_results()` accepts the same argument and shows the override in the interpretation section of its block.

-   [x] B1.6 The interpretation text never appears in the .sframe file: `write_sframe()` then `read_sframe()` round-trips with no interpretation content and no hash change.

    Feedback:

    -   browseURL(out) needs rework. The column splits are not available in the tables. Each row is the raw space or tab separated data.\
    -   The Response distributions and descriptive plots need a revisit.

-   The plot still uses theme_minimal() rather than theme_classic()

**Fixed 2026-07-17, first pass:** every `kable()` call in `inst/templates/report.qmd` now forces `format = "html"`. The response-distribution and descriptive charts were also drawing with base-R `barplot()`/`hist()`, not `theme_surveyframe()`, in both the Quarto and HTML-fallback report paths. Both now use the same ggplot2 family-plot helpers (`sframe_plot_item_chart()`, `sframe_plot_scale_chart()`) as the rest of the report, so the look is consistent.

**Fixed 2026-07-17, root cause found on re-check:** the first pass above did not actually fix the table split, because it was never a `kable()` formatting problem. Two real bugs, both pre-existing: (1) `render_report()`'s Quarto path was silently failing on every render (`sframe_draw_likert_diverging()` is called bare in `report.qmd` but was never exported, so Quarto's own `library(surveyframe)` session cannot see it, unlike an interactive `load_all()` session) and falling back to the HTML path without any warning, so the Quarto fix above was never actually exercised. (2) The HTML fallback's own `.render_report_table()` called `htmltools_escape()` on a whole vector of header names or row values at once. That helper collapses a vector into one space-joined string before escaping, which is correct for a caption or a message but merges an entire header row or an entire data row into a single cell, exactly matching the raw space-separated text in the screenshot. Fixed by exporting `sframe_draw_likert_diverging()` (so the Quarto path renders instead of crashing) and adding `htmltools_escape_each()` for per-cell escaping. Verified live: `render_report()` now produces a working Quarto report end to end, and both paths split table columns correctly. Re-check at N1.2 and N1.3.

Feedback 25/07/2026: The plot in RQ 33: Does visitor type moderate the relationship between digital marketing and satisfaction?
Can be improved to be something similar to those from https://github.com/cardiomoon/processr

**Fixed 2026-07-25:** `sframe_plot_moderation()` (`R/plots.R`) already drew the right kind of chart (simple slopes at -1 SD, mean, and +1 SD of the moderator), but with no confidence band and no indication of which slopes were actually significant, unlike a processR-style interaction plot. Added a confidence ribbon per line (via `predict(fit, newdata = ..., interval = "confidence")`, no new code path needed) and a caption reporting each level's simple-slope estimate and significance (delta-method SE of `b_predictor + b_interaction * moderator_level`, base R only, no new dependency — `processR` itself is not something surveyframe should depend on). While verifying this live against the demo instrument, found and fixed a real bug the fix itself introduced and caught: `sframe_p_string()` is not vectorised and was being called on all 3 slope p-values at once, throwing "the condition has length > 1" and silently dropping the chart (`plot(results, which = "rq_moderation")` returned "No chart is attached"). Fixed with `vapply()`. Verified live: the moderation chart now renders with ribbons and the slope caption, e.g. "-1 SD slope: b = 0.45, p < .001 \| Mean slope: b = 0.50, p < .001 \| +1 SD slope: b = 0.56, p < .001" on the demo data.


### Step B2 — SurveyStudio Interpretations card

```{r}
#| label: b2-studio
#| eval: false
launch_studio()
# Open the demo instrument, upload the demo responses, go to Export.
```

-   [x] B2.1 The Export screen has three cards: Instrument file, Report contents (with the Chart theme radio), Interpretations, and Generate.
-   [x] B2.2 The Interpretations card lists every saved plan block with its id, question, planned decision rule, and live APA result.
-   [x] B2.3 Type an interpretation for two research questions, generate the report, and confirm both appear beside their decision rules. (The machine did this for one; do a second by hand.)
-   [x] B2.4 The card reads well at 34 research questions: scrolling is acceptable and the fields are clearly associated with their questions. Judgement call.

Feedback: There is no single preview display to see the complete results and then write the interpretation below the results for each output. Read the JASP and jamovi output canvas and adopt the workstyle from it for results, interpretation, and exporting outputs. Even a simple copy option will provide more flexibility than html and pdf render.

**Partially actioned 2026-07-17:** the Interpretations card is rebuilt as a single output canvas. Each research question is now one block, in reading order: badges, APA line, results table, chart, a "Copy result" clipboard button, the planned decision rule, then the editable interpretation box. Previously this card showed only a text APA line with no table or chart. Chart rendering is cached and shared with the Run stage so this does not double the render time. Not actioned: a full JASP/jamovi-style workflow across the whole studio (multi-output export queue, drag-to-copy, and similar). That is a larger piece of work worth scoping on its own. Re-check at N1.8 to N1.10.

Feedback 25/07/2026:
The Chart theme switcher is not working in the shiny studio report preview.
Colour (on-screen reading)
Black and white (print and journal submission)

**Fixed 2026-07-25:** `analysis_results_r()` (`inst/shiny/app.R`), the one shared reactive that computes the analysis-plan results and feeds both the Run stage cards and the Export screen's Interpretations canvas, called `run_analysis_plan()` with no `plot_palette` argument, so it always rendered in the default "web" palette regardless of the Chart theme radio. The actual report download (`download_report_btn`) already correctly passed `plot_palette = input$rpt_palette`, so only the on-screen preview was stuck. `analysis_results_r()` now passes `plot_palette = input$rpt_palette %||% "web"` too, so switching the radio re-renders the preview charts in the chosen palette.

### Step B3 — Builder Report outline editing

Open `inst/builder/survey_builder.html` in a browser, add a question, add an analysis plan, then open Analyse and the Report outline tab.

-   [x] B3.1 Each research question in the outline shows an editable "Planned decision rule (editable)" field pre-filled from the plan.
-   [x] B3.2 Editing the field updates the plan card and the RQ dialogue; editing in the dialogue updates the outline.
-   [x] B3.3 The inline field looks like part of the document skeleton rather than a form dropped into it. Judgement call.

The edit function should be in the interpretation part (below each table and plot outputs or each section), not the research question outline. Anyway, the shiny app needs rework from top to bottom for smooth user experience.

**Fixed 2026-07-17:** the editable "Planned decision rule" field in the builder Report outline now renders below each result's table and chart placeholder, matching where the rule and interpretation pair in the rendered report, instead of up in the RQ header. The broader "shiny app needs rework top to bottom" note is not actioned. It is a larger redesign to scope separately. Re-check at N1.11.

Feedback 25/07/2026:
The Split by (optional) in the html surveybuilder selects the first option by default. Fix it

**Fixed 2026-07-25:** `roleOptions()` (`inst/builder/survey_builder.html`) built a `<select>`'s `<option>` list with no blank placeholder, so for an optional single-select role (`role.min === 0`, like "Split by (optional)") the browser's native default-select-first-option behaviour silently picked the first variable even though the user never chose one. `roleOptions()` now prepends a selected `(none)` option whenever the role is optional and single-select and nothing is already assigned. `collectRoles()` already filtered falsy values, so the empty selection round-trips correctly as "not set" without any other change.

------------------------------------------------------------------------

## Part C — SurveyStudio result cards (6 items)

With the demo instrument and responses loaded, open Analyse and the Run stage.

-   [x] C1.1 Results render as one card per research question: number and block id, question, method badge, APA line, and the chart beneath.
-   [x] C1.2 The regression card stacks its fit chart and 4 diagnostic panels.
-   [x] C1.3 Syntax-only and chartless blocks render their card without a chart and without an empty gap.
-   [x] C1.4 Chart quality at card width: labels legible, nothing clipped, colours match the report. Judgement call.
-   [x] C1.5 The Run stage remains responsive with 34 blocks (the plan runs once and is shared with the Export screen; expect one initial wait, then smooth tab switches).
-   [x] C1.6 The card list reads better than the 0.3.3 flat table. Judgement call.

------------------------------------------------------------------------

## Part D — New plot() methods and dashboards (10 items)

### Step D1 — plot() methods

```{r}
#| label: d1-plots
vr <- validity_report(list(SAT = c(sat_1 = .82, sat_2 = .78),
                           LOY = c(loy_1 = .69, loy_2 = .71)))
plot(vr)
results <- run_analysis_plan(responses, instr, plots = TRUE)
plot(results, which = "rq_dm_sat")
mr <- missing_data_report(responses, instr)
plot(mr)
```

-   [x] D1.1 `plot(vr)` draws CR and AVE bars with the 0.70 and 0.50 threshold lines.
-   [x] D1.2 `plot(results, which = ...)` returns one chart; `plot(results)` prints them all; a results object built without `plots = TRUE` aborts with a message pointing at the argument.
-   [x] D1.3 `plot(mr)` draws missing rates by item, or a "No missing responses in any item" message chart when nothing is missing.
-   [x] D1.4 All three read correctly in the print palette (`palette = "print"`). Judgement call.

Feedback: plot(mr) NULL Should improve the output if there is no missing data.

**Fixed 2026-07-17:** `plot(missing_data_report(...))` on clean data now returns a small "No missing responses in any item" chart instead of `NULL`, so it reads as a result rather than a rendering failure. D1.3 above is reworded to match. Re-check at N1.6.

### Step D2 — Dashboard coverage

```{r}
#| label: d2-dashboard
#| eval: false
launch_dashboard(instr, responses)
```

-   [x] D2.1 The Quality tab opens with the flag-rate chart and the missingness chart above the attention-check table.
-   [x] D2.2 The Scales tab shows the scale-score correlation heatmap below the distribution chart.
-   [x] D2.3 The SurveyStudio Quality Dashboard screen and Dashboard tab mirror the same three charts.
-   [x] D2.4 Uninstall or mask ggplot2 (`callr::r()` with a stripped library path is the quick way) and confirm the base-graphics fallbacks draw instead of erroring.
-   [ ] D2.5 The heatmap remains readable with the demo's 5 scales and at a phone-width window. Judgement call.
-   [ ] D2.6 Filters still apply to the new charts (set a date filter and watch the missingness chart change). Feedback: There are some places where the variable label is handled properly and some places it is not. For example: digital_marketing, behavioural_intention and so on. Fix them all. Verify whether adding levels and labels work with surveyframe all item types. Revise the working examples, pkgdown, and documentation to use levels, and labels with proper labels with space. Also, check whether the complete workflow runs need without any additional coding works later. For example: Ranking data, matrix blocks, and so on.

**Partially fixed 2026-07-17:** the raw-id-on-axis part is fixed. A shared humaniser (`.sframe_title_case_names()`, snake_case to "Title case with spaces") is now applied to axis and title text in `sframe_plot_correlation()`, `sframe_plot_regression()`, `sframe_plot_group_comparison()`, `sframe_plot_paired_comparison()`, `sframe_plot_variable_distribution()`, `sframe_plot_correlation_matrix()`, `sframe_plot_efa_loadings()`, `sframe_plot_quality()`, `sframe_plot_validity()`, and `sframe_plot_missingness()`, so `digital_marketing` reads as "Digital marketing" everywhere a raw id previously leaked through. Re-check at N1.5. The filter-still-applies check (D2.6's own item) and the levels/labels audit across every item type, the vignette and pkgdown documentation revision, and the ranking/matrix no-extra-coding check are not done. Those need a dedicated pass.

Feedback 25/07/2026:
The plots are responsive in launch_dashboard(instr, responses), but the tables are not. fix it. Also, there is  a message character string is not in a standard unambiguous format

**Fixed 2026-07-25:** two separate issues. (1) The raw responses table (`output$raw_table`, `inst/shiny/dashboard/app.R`) was rendered with `renderTable(..., width = "100%")`, which forces the table to always fit the visible card width — on the demo's 30+ raw response columns this crushes every cell instead of scrolling, defeating the `overflow-x:auto` wrapper the table already sits in. Removed `width = "100%"` so the table sizes to its natural content width and the wrapper's horizontal scroll actually works, and added `-webkit-overflow-scrolling:touch` to that wrapper for a proper touch-scroll feel on mobile. (2) Could not reproduce the exact warning live against the demo data (checked with `shiny::testServer()` and `options(warn = 1)` across every dashboard output, including the date-range filter path — every `as.Date()`/`as.POSIXct()` call already in the dashboard file was already correctly guarded with an explicit format or `suppressWarnings()`). Did find one genuinely unguarded bare `as.Date()` call in `sf_item.R`'s date-bound validation (`as.Date(date_min) > as.Date(date_max)`, directly contradicting that function's own comment two lines above explaining why a bare `as.Date()` is unsafe) — this is exactly the kind of call that throws "character string is not in a standard unambiguous format" on an ambiguous string, so fixed it to pass `format = "%Y-%m-%d"` explicitly, and hardened the dashboard's date-range filter to route through the existing `dashboard_parse_date()` helper instead of a bare `as.Date()` for consistency. If the warning recurs, it would help to capture the exact R console text and which tab was open when it appeared, since the demo data and every currently-known date-parsing path in the dashboard did not reproduce it.

------------------------------------------------------------------------

## Part E — Date-question bounds (8 items)

### Step E1 — R side

```{r}
#| label: e1-dates
#| error: true
it <- sf_item("visit", "When did you visit?", type = "date",
              date_min = "2026-01-01", date_max = "2026-12-31")
sf_item("bad", "x", type = "date", date_min = "2026-12-31",
        date_max = "2026-01-01")
# Expected: a validation error saying min must not be later than max
```

-   [x] E1.1 Valid bounds are stored as "YYYY-MM-DD" strings and survive a write/read round trip.
-   [x] E1.2 Reversed bounds and unparseable dates abort with a typed validation error.

### Step E2 — Builder and exported survey

-   [x] E2.1 A date question's inspector shows Earliest date and Latest date fields; setting min after max raises an error toast.
-   [x] E2.2 The exported survey's date input carries the min and max attributes, so the native picker greys out-of-range dates.
-   [x] E2.3 A typed out-of-range date is blocked at page validation with "Please pick a date on or after/before ..." in both directions.
-   [x] E2.4 An in-range date passes and submits.
-   [ ] E2.5 On a phone, the native date wheel respects the bounds.
-   [ ] E2.6 The bounds error message reads well next to the required-field message style. Judgement call.

------------------------------------------------------------------------

## Part F — WCAG 2.2 AA audits (8 items)

-   [x] F1.1 axe-core 4.10.2 (wcag2a, wcag2aa, wcag22aa) reports zero violations across 6 builder states: build with inspector, preview, Analyse plan tab, Report outline, settings modal, RQ modal.
-   [x] F1.2 All 7 vignettes report zero violations and zero heading-order skips.
-   [x] F2.1 The darker muted text (`#5b6b80` replacing `#94a3b8`) still reads as secondary text and has not flattened the builder's visual hierarchy. Judgement call.
-   [x] F2.2 The darker save-status colours (amber and green) still read as states at a glance.
-   [x] F2.3 Vignette links in the darker teal still read as brand. Judgement call.
-   [x] F2.4 Wrapped code blocks in the vignettes: pick the widest chunk in the lead vignette and confirm the wrap reads acceptably.
-   [x] F2.5 Keyboard-only pass of the builder: tab through build, preview, analyse, and both modals; focus is always visible and no control is unreachable.
-   [ ] F2.6 One screen-reader spot check (VoiceOver, NVDA, or Orca): the preview inputs announce their question labels.

------------------------------------------------------------------------

## Part G — Release safety (11 items)

-   [x] G1.1 `R CMD check --as-cran` on the 0.3.4 tarball: 0 errors, 0 warnings, at most the incoming-feasibility NOTE.
-   [x] G1.2 The tarball contains no dev-only files: no CLAUDE.md, roadmap, revision_todo, dogfeed, mas_review, demo directory, or mas_full_report.html (`untar -tf` and scan).
-   [x] G1.3 649 tests pass on the working tree.
-   [ ] G1.4 win-builder R-release returns Status: OK.
-   [ ] G1.5 win-builder R-devel returns Status: OK.
-   [x] G1.6 NAMESPACE gained only additive exports: 3 S3 plot methods, `sframe_plot_validity()`, `sframe_plot_missingness()`.
-   [x] G1.7 Hard Imports are still exactly jsonlite, rlang, and openssl.
-   [x] G1.8 A 0.3.3 script with no new arguments produces identical output under 0.3.4 (spot check: render_report on the demo without interpretations, diff the RQ sections).
-   [x] G1.9 The pkgdown workflow is green after the release push and the live site carries no dev-only files (pkgdown builds from main only; `.Rbuildignore` does not protect it).
-   [x] G1.10 `urlchecker::url_check()` and `spelling::spell_check_package()` report nothing actionable.
-   [x] G1.11 NEWS.md still reads correctly after the first-round fix batch: the fixes fold into the existing 0.3.4 entries and no entry contradicts the fixed behaviour.

**Verified 2026-07-25.** G1.8: checked out the `v0.3.3` tag into a worktree, rendered the demo report on both 0.3.3 and 0.3.4, and diffed every RQ block's question text, test statistic, and APA line. All 34 questions, methods, and point estimates (U, z, t, χ², H, rho, tau, indirect effect) are byte-identical. The only differences are +/-0.01 to 0.02 wobble in 8 bootstrap-derived confidence-interval bounds (Mann-Whitney r, Cramer's V/phi, Cohen's d, Kruskal-Wallis eta-squared, Spearman, Kendall, mediation indirect effect) — `sframe_rank_r_ci()` and its siblings in `R/bootstrap_ci.R` were never given a seed argument in the first place (unlike `bootstrap_ci()`/`cohens_d_ci()`/`cramers_v_ci()`/`eta_sq_ci()`, which are seed-safe since the 2026-07-18 fix), so neither version's unseeded `render_report()` call is reproducible run-to-run on these particular intervals. This is ordinary Monte Carlo noise present in both versions, not a 0.3.4 regression. G1.10: `urlchecker::url_check()` found one real issue (DESCRIPTION's pkgdown URL missing a trailing slash, redirecting) — fixed; re-run reports all URLs correct. `spelling::spell_check_package()` flagged 85 words, all reviewed: proper nouns, author surnames, package names, statistical abbreviations, and correctly-spelled UK `-ise` forms — nothing actionable. G1.9: see the pkgdown fix note above; the reference-index bug is fixed and a full local `pkgdown::build_site()` run is clean. Live confirmation: the pkgdown GitHub Actions run for the fix commit completed with `success` (run 30128039999), the live site's `sitemap.xml` lists 130 URLs with none matching any dev-only filename, and the new `bootstrap_ci` reference page returns HTTP 200. G1.11: NEWS.md structure checked programmatically (one version heading, no leaked internal/planning language beyond an innocuous "Analysis planning" section title) and its prose read against the actual current behaviour verified elsewhere in this pass (table splitting, moderation plot, etc.) — consistent.

**G1.1/G1.2 — real `R CMD check --as-cran` run against the release-candidate tarball (2026-07-25).** The first full run was not clean: 1 ERROR, 2 WARNINGs, 3 NOTEs, all genuine and previously unnoticed (none of them related to the second-round feedback fixes committed earlier the same day). Every one was fixed and a second, final run returned **`Status: OK` — 0 errors, 0 warnings, 0 notes** (not even the incoming-feasibility NOTE fired). Findings and fixes:

- **ERROR** — `test-v034-stats-reporting.R`'s naniar-gated MCAR test, only ever exercised on a machine with naniar installed (never before this session, hence never caught): `d[sample(320, 30)] <- NA` on a data frame indexes *columns*, not cells, and errored trying to create up to 320 columns on a 4-column frame. Fixed by NA-ing cells through `as.matrix()`.
- **WARNING** — non-ASCII characters in `R/analysis_plan.R`, `R/plots.R`, `R/statistics_reports.R` (×, ±, ′, η, ² in mediation and repeated-measures APA strings). Replaced with `\u` escapes; runtime output confirmed byte-identical.
- **WARNING** — Rd cross-reference: 24 roxygen `[fn()]` links across `R/analysis_plan.R`, `R/plots.R`, `R/reporting.R` pointed at `sframe_brand()`, `sframe_plot_crosstab()`, and `sframe_plot_correlation()` — all three real functions, none documented (internal helpers with no Rd page), so the links never resolved. De-linked to plain code text.
- **NOTE** — `sframe_plot_variable_distribution()` uses `ggplot2::after_stat(density)`, which R CMD check's static analysis reads as an unbound global variable named `density`. Added `density` to the package's `@importFrom stats` list, the standard fix for this common ggplot2 pattern.
- **NOTE** — `run_analysis_plan()` and `render_results()`'s examples each ran the full 34-block demo plan unconditionally (~8s each). Wrapped in `\donttest{}`; still exercised and confirmed OK under `--run-donttest`.
- **NOTE** — `fresh_eyes_check.sframe`, a stray file left at the repository root from testing the Part I code chunk earlier in this review cycle, was picked up as a non-standard top-level file. Deleted, and added to `.Rbuildignore` as a safety net. Also added `report_interp.html` to `.Rbuildignore` (a 5.7 MB test-render artefact that had no exclusion rule and would otherwise have been bundled into the tarball).

Final tarball: `surveyframe_0.3.4.tar.gz` (829 KB). Confirmed installable (`R CMD INSTALL`), `packageVersion()` reports 0.3.4, and all 7 vignettes are present via `tools::getVignetteInfo()` on the real installed package (completing H1.1).

------------------------------------------------------------------------

## Part H — Vignettes (5 items)

```{r}
#| label: h-vignettes
#| eval: false
browseVignettes("surveyframe")
```

-   [x] H1.1 All 7 vignettes build and open from the installed package.
-   [x] H1.2 The lead vignette's new fig.alt text does not show visibly anywhere (alt text only).
-   [x] H1.3 The two rewritten style blocks did not change table styling: APA-style rules above and below headers, no shading.
-   [ ] H1.4 Spot-read one vignette for the style rules (UK spellings, no em-dashes or semicolons, no banned words).
-   [ ] H1.5 The flat `browseVignettes()` output (no pkgdown wrapper) is presentable.

**Verified 2026-07-25.** H1.1: `pkgdown::build_site()` built all 7 vignettes as articles with no errors (also re-confirmed by the live pkgdown workflow run); final confirmation from the actual installed tarball happens alongside the tarball build below. H1.2: rendered `surveyframe.Rmd` directly and grepped the output — the fig.alt text for the `score-distributions` chunk appears once, inside `alt="..."`, and zero times as visible body text. H1.3: the lead, analysing-survey-responses, and scale-reliability-validity vignettes (the ones with `kable()` tables) all carry the same `th { border-top: 2px solid; border-bottom: 1px solid }` / `tbody tr:last-child td { border-bottom: 2px solid }` rule with no `background` declaration anywhere near table CSS, in both the raw `.Rmd` source and the pkgdown-built HTML. H1.4 (reading prose for style-rule compliance) and H1.5 (visual presentability of the flat output) are genuinely "look at it" checks, moved to `dogfeed.todo.md`.

------------------------------------------------------------------------

## Part I — Fresh-eyes UX pass (4 items)

```{r}
#| label: i1-fresh-eyes-setup
#| eval: false
# Builds the 5-question instrument I1.1 asks for (one bounded date question
# plus a short Likert block) and saves it as a starting artefact: open the
# written .sframe in the builder for I1.1, and load it into SurveyStudio
# alongside the bundled demo responses for I1.2 and I1.3.
agree5 <- sf_choices("agree5", 1:5,
  c("Strongly disagree", "Disagree", "Neutral", "Agree", "Strongly agree"))

fresh_eyes_check <- sf_instrument(
  title = "Fresh-eyes check",
  components = list(
    agree5,
    sf_item("visit_date", "When did you visit?", type = "date",
            date_min = "2026-01-01", date_max = "2026-12-31", required = TRUE),
    sf_item("satisfaction", "How satisfied were you overall?", type = "likert",
            choice_set = "agree5", required = TRUE),
    sf_item("would_return", "I would visit again", type = "likert",
            choice_set = "agree5", required = TRUE),
    sf_item("recommend", "I would recommend this to others", type = "likert",
            choice_set = "agree5", required = TRUE),
    sf_item("comments", "Any other comments?", type = "textarea")
  )
)

validate_sframe(fresh_eyes_check, strict = FALSE)
write_sframe(fresh_eyes_check, "fresh_eyes_check.sframe")
# Now: open inst/builder/survey_builder.html and load fresh_eyes_check.sframe
# for I1.1. Open launch_studio() for I1.2 and I1.3, using the bundled demo
# instrument and responses as in the Part B/C chunks above.
```

-   [ ] I1.1 Thirty minutes as a new user: build a 5-question survey with a bounded date question in the builder, export it, answer it, and note anything confusing.
-   [ ] I1.2 Thirty minutes in SurveyStudio: open the demo, run the plan, write two interpretations, export the report in the print palette, and note anything confusing.
-   [ ] I1.3 Read the full generated report top to bottom as if reviewing a student's submission.
-   [ ] I1.4 Log anything found in dogfeed.todo.md rather than fixing inline.

**Fixed 2026-07-25:** added the runnable chunk above (`sf_instrument()` / `sf_item()` / `sf_choices()`, verified live against the current API), matching every other Part in this document. It builds and validates the 5-question instrument I1.1 describes and writes it as `fresh_eyes_check.sframe`, giving the fresh-eyes pass a concrete starting artefact instead of an unscaffolded "build something" instruction.

------------------------------------------------------------------------

## Part J — Effect-size confidence intervals (7 items)

```{r}
#| label: j-cis
bootstrap_ci(mtcars$mpg, seed = 42)
cohens_d_ci(mtcars$mpg[mtcars$am == 1], mtcars$mpg[mtcars$am == 0], seed = 42)
# results was created in the Part D chunk
results[["rq_ttest_ind"]]$apa
```

-   [x] J1.1 The four helpers are exported, reproducible with a seed, and return `estimate`, `lower`, `upper` with NA bounds on tiny samples.
-   [x] J1.2 The 9 affected runners attach their CI key (`d_ci`, `r_ci`, `eta_ci`, `ci`, `v_ci`) and the APA string carries the interval.
-   [x] J1.3 The Pearson interval is the analytic Fisher z, matching `sframe_fisher_z_ci()` exactly.
-   [x] J1.4 Degenerate data keeps the pre-0.3.4 APA string with no bracket.
-   [ ] J1.5 Read 3 APA strings in a rendered report: the intervals read naturally in context and match the reported effect. Judgement call.
-   [x] J1.6 The intervals are statistically sensible on the demo data (bracket the estimate, wider at 99 percent than at 95). Spot check one by hand.
-   [ ] J1.7 Timing: the full demo plan with 34 blocks still runs in acceptable time in the studio's Run tab (the bootstraps add work).

**J1.6 verified 2026-07-25:** `bootstrap_ci(mtcars$mpg, seed = 42)` at 95 percent brackets its own estimate (16.85, 19.20, 21.40) and the 99 percent interval is wider (6.30 vs 4.55). **J1.7 partially measured:** `system.time(run_analysis_plan(responses, instr, plots = TRUE))` on the full 34-block demo plan is 14.3 seconds of raw R compute on this machine, consistent with N1.10's "about half a minute" total for the Export canvas's first open once Shiny UI rendering and base64 chart encoding are added on top. Whether that reads as "acceptable" inside the actual Run tab is a UX call, not a number — left unticked and moved to `dogfeed.todo.md` alongside J1.5 (APA prose) and N1.9/N1.10 (live interaction).

Feedback (captured console output, 2026-07-17):

```         
> bootstrap_ci(mtcars$mpg, seed = 42)
estimate    lower    upper
   19.20    16.85    21.40
> cohens_d_ci(mtcars$mpg[mtcars$am == 1], mtcars$mpg[mtcars$am == 0], seed = 42)
 estimate     lower     upper
1.4779471 0.7869318 2.5256596
> results <- run_analysis_plan(responses, instr, plots = TRUE)
> results[["rq_ttest_ind"]]$apa
NULL
> results
Analysis Results: 34 research question(s)

RQ 1: Is perceived digital marketing effectiveness associated with tourist satisfaction? Test: correlation_pearson APA: r(118) = 0.54 [0.40, 0.65], p < .001

RQ 2: Do digital marketing, service quality, and sustainability perceptions predict satisfaction? Test: regression_linear APA: R² = 0.383, F(3, 116) = 23.95, p < .001

RQ 3: Do first-time and repeat visitors differ in behavioural intention? Test: mann_whitney APA: U = 1576, z = -0.98, p = 0.327, r = 0.09 [0.00, 0.27]

RQ 4: What is the distribution of first-time and repeat visitors? Test: frequency APA: Frequency distribution for visit_type (N = 120).

RQ 5: What are the central tendency and spread of all response items? Test: descriptives APA: Descriptive statistics were computed for 12 variable(s).

RQ 6: What is the pattern of missing responses across items? Test: missing_data APA: Missing-data diagnostics were computed for 12 variable(s).

RQ 7: Do respondents meet attention check and data quality thresholds? Test: quality APA:

RQ 8: What are the mean, SD, and range of each composite scale? Test: descriptives APA: Descriptive statistics were computed for 5 variable(s).

RQ 9: What is the Cronbach alpha internal consistency of each scale? Test: reliability_alpha APA:

RQ 10: What is the McDonald omega reliability of each scale? Test: reliability_omega APA:

RQ 11: Which items show low item-total correlations or reduce alpha on removal? Test: item_diagnostics APA:

RQ 12: Does the inter-item correlation matrix support exploratory factor analysis? Test: efa_readiness APA:

RQ 13: How many factors emerge from digital marketing and service quality items? Test: efa_solution APA:

RQ 14: What is the lavaan CFA syntax for the five-factor measurement model? Test: cfa_lavaan_syntax APA: CFA lavaan syntax generated.

RQ 15: What is the CB-SEM lavaan syntax for digital marketing predicting satisfaction and behavioural intention via service quality? Test: sem_lavaan_syntax APA: CB-SEM lavaan syntax generated.

RQ 16: What is the PLS-SEM seminr syntax for the full structural model? Test: seminr_syntax APA: PLS-SEM seminr syntax generated.

RQ 17: Is visitor type associated with attention check response level? Test: crosstab APA: χ²(1, N = 120) = 4.31, p = 0.038, φ = 0.19 [0.03, 0.32]

RQ 18: Is the distribution of satisfaction ratings different across visitor types? Test: crosstab APA: χ²(4, N = 120) = 3.40, p = 0.494, V = 0.17 [0.10, 0.38]

RQ 19: Is there an association between visitor type and behavioural intention rating? Test: fisher_exact APA: Fisher's exact test, p = 0.454, Cramer's V = 0.17

RQ 20: Do first-time and repeat visitors differ in mean satisfaction score? Test: t_test_ind APA: t(106.77) = 0.15, p = 0.878, d = 0.03 [0.01, 0.42]

RQ 21: Do respondents rate the two satisfaction items differently? Test: t_test_pair APA: t(119) = 0.82, p = 0.416, d_z = 0.07 [-0.10, 0.26]

RQ 22: Is there a significant distributional difference between the first two service quality items? Test: wilcoxon_pair APA: V = 967, z = -1.17, p = 0.240, r = 0.11 [0.01, 0.37]

RQ 23: Does satisfaction differ across visitor types? Test: kruskal_wallis APA: H(1) = 0.00, p = 0.949, η² = 0.000 [0.00, 0.03]

RQ 24: Does mean behavioural intention differ between visitor types? Test: anova_one APA: F(1, 118) = 0.74, p = 0.391, η² = 0.006 [0.00, 0.07]

RQ 25: Do visitor types differ in satisfaction after controlling for service quality? Test: ancova APA: ANCOVA estimated group differences adjusted for covariates.

RQ 26: Do mean ratings differ across the three digital marketing items within respondents? Test: repeated_anova APA: Repeated-measures ANOVA was estimated; inspect `fit_summary` for the within-subject effect.

RQ 27: Do ordinal ratings differ across the three service quality items within respondents? Test: friedman APA: Friedman chi-square(2, N = 120) = 1.09, p = 0.580

RQ 28: Are service quality perceptions associated with sustainability perceptions? Test: correlation_spearman APA: r_s(118) = 0.00 [-0.18, 0.19], p = 0.984

RQ 29: Is sustainability perception associated with behavioural intention? Test: correlation_kendall APA: tau(118) = 0.05 [-0.11, 0.20], p = 0.519

RQ 30: Is digital marketing associated with behavioural intention after controlling for satisfaction? Test: partial_correlation APA: partial r(117) = -0.09, p = 0.302

RQ 31: Do digital marketing and service quality perceptions predict visitor type? Test: regression_logistic_binary APA: χ²(2) = 0.28, p = 0.869, McFadden R² = 0.002

RQ 32: Do digital marketing and sustainability perceptions predict ordered satisfaction? Test: regression_logistic_ordinal APA: Ordinal logistic regression was estimated with 120 complete cases.

RQ 33: Does visitor type moderate the relationship between digital marketing and satisfaction? Test: moderation APA: Moderation was tested with a linear interaction model.

RQ 34: Does satisfaction mediate the path from digital marketing to behavioural intention? Test: mediation APA: Indirect effect = 0.365, 95% bootstrap CI [0.226, 0.525].
```

**Fixed 2026-07-17:** `results[["rq_ttest_ind"]]$apa` returned `NULL` because `run_analysis_plan()` never named its result list by plan block id, so named access always missed even though `print(results)` (which walks the list by position) showed the same result correctly. `run_analysis_plan()` now sets `names(results)` to each block's id, so `results[["rq_ttest_ind"]]` and similar named lookups work as the documentation implies. The Part J chunk above now prints the APA string when this document is rendered. Re-check at N1.7.

## Part K — Psychometric depth (7 items)

-   [x] K1.1 `validity_report(items_by_construct = ...)` returns Henseler HTMT (symmetric, unit diagonal, NA rows for single-item constructs) and `htmt_method = "henseler"`.
-   [x] K1.2 Without item data the correlation fallback applies unchanged and is labelled `correlation_fallback`.
-   [x] K1.3 With naniar installed, `missing_data_report()` returns Little's MCAR statistic, df, p value, and a plain interpretation; without missingness (or without naniar) the pre-0.3.4 result returns verbatim.
-   [x] K1.4 `reliability_report()` sets `omega_note` when omega cannot be computed and the reliability chart names those scales in its subtitle.
-   [x] K1.5 `efa_solution()` carries `loadings_long`, `communalities_table`, and `variance_table`, and the loadings heatmap consumes the tidy frame.
-   [x] K1.6 Run `validity_report()` on real scored demo data with `items_by_construct` and sanity-check 2 HTMT values against hand computation.
-   [ ] K1.7 The MCAR interpretation wording reads correctly to a methodologist. Judgement call.

**K1.6 verified 2026-07-25:** `validity_report()` on the demo's `sat_1`/`sat_2` (TS) and `bi_1`/`bi_2` (BI) items returns HTMT(TS, BI) = 0.6053; an independent hand computation (mean absolute heterotrait correlation over the square root of the product of the two mean monotrait correlations, direct from `cor()`) gives the identical value to 4 decimal places. K1.7 stays a prose-reading judgement call, moved to `dogfeed.todo.md`.

## Part L — PDF output and report theming (8 items)

```{r}
#| label: l-pdf
#| eval: false
render_report(instr, responses, output_file = "report.pdf", format = "pdf")
```

-   [x] L1.1 `format = "pdf"` produces a real PDF through pagedown when Chrome is available, and aborts with a typed, actionable error without pagedown.
-   [x] L1.2 `format = "html"` remains the default and its output is unchanged apart from the deliberate theming work below.
-   [x] L1.3 The HTML fallback styles sit behind `--sf-*` CSS variables with a print stylesheet, tables carry captions and `scope="col"`, and every embedded chart has descriptive alt text naming its research question.
-   [ ] L1.4 Open the PDF: pagination is clean (no result card or table split mid-block, headers repeat on continued tables).
-   [ ] L1.5 The PDF is readable in print greyscale when generated with `plot_palette = "print"`.
-   [ ] L1.6 Print the HTML report from a browser (Ctrl+P): the print stylesheet applies (no panel tints, no clipped tables).
-   [ ] L1.7 The accessible teal (`--sf-accent`) still reads as brand in the HTML report. Judgement call.
-   [x] L1.8 File size of a full demo PDF is reasonable to email (check it is under about 10 MB).

**Verified 2026-07-25:** `render_report(format = "pdf")` now that `pagedown` is installed produces a valid PDF (`%PDF` header, `%%EOF` trailer) for both `plot_palette = "web"` and `"print"`, at 3.39 MB, well under the 10 MB check (L1.8). L1.4 to L1.7 all need eyes on an actual rendered page (pagination, greyscale legibility, browser print output, colour perception) and stay on the human list, moved to `dogfeed.todo.md`.

Feedback: render_report(instr, responses, output_file = "report.pdf", format = "pdf") Error in `render_report()`: ! Package 'pagedown' is needed for PDF output. Install it, or use format = 'html'. Run `rlang::last_trace()` to see where the error occurred.

**Reviewed 2026-07-17, no change needed:** this is the documented, typed error L1.1 already checks off (pagedown is Suggests-only and guarded). It fires because pagedown was not installed on this machine. Install pagedown, or use `format = "html"`, to get the PDF.

## Part M — Codebook upgrades (4 items)

-   [x] M1.1 `codebook_report()` returns `plan_table` (id, question, method, variables, decision rule) and `models_table` (id, label, type, engine, constructs, paths).
-   [x] M1.2 Both render in the report codebook section on the Quarto path and the HTML fallback.
-   [x] M1.3 Read the codebook of the demo instrument end to end: it now fully documents the study (items, choices, scales, plan, models) without needing the rest of the report. Judgement call.
-   [x] M1.4 A codebook-only report (`include_analysis = FALSE`, no data) still renders the plan summary, since the plan is pre-declared design, independent of results.

Feedback: The codebook tables need formatting.

**Fixed 2026-07-17:** the HTML-fallback path's `.render_report_table()` printed raw data-frame column names (`research_question`, `n_items`, and so on) verbatim. It now title-cases them by default (or accepts an explicit `col.names`), and all 5 codebook tables (items, choices, scales, plan, models) pass proper headings, matching what the Quarto path's `kable(col.names = ...)` already did. Re-check at N1.4.

------------------------------------------------------------------------

## Part N — First-round fix verification (11 items)

Re-checks for the fixes applied 2026-07-17 in response to the feedback above. Reinstall the release candidate (Step A1) first, since the fixes landed on GitHub main after the original install. The chunk below demonstrates the two fixes that are verifiable in code; the rest are visual checks against the regenerated report, studio, and builder.

```{r}
#| label: n-fixes
head(names(results))
results[["rq_dm_sat"]]$apa
plot(missing_data_report(responses, instr))
```

-   [x] N1.1 `LICENSE` and `LICENSE.md` both read 2026, at the public repo root and inside the rebuilt tarball.
-   [x] N1.2 Re-render the B1 report: every table renders as a real HTML table with column splits, in both the Quarto path and the HTML fallback (`use_quarto = FALSE`). No raw space or tab separated rows anywhere.
-   [x] N1.3 The Response distributions section's item and scale charts now match the analysis charts: `theme_surveyframe()` axis lines and brand fill, in both render paths.
-   [x] N1.4 The codebook tables carry title-cased headings ("Research question", "Decision rule") in the HTML fallback, matching the Quarto path.
-   [x] N1.5 Axis and title labels are humanised across the family plots: spot check the correlation, group-comparison, and missingness charts for "Digital marketing" rather than digital_marketing.
-   [x] N1.6 `plot(missing_data_report(...))` on the clean demo data draws the "No missing responses in any item" chart (the chunk above shows it).
-   [x] N1.7 `names(results)` equals the plan block ids and `results[["rq_dm_sat"]]$apa` returns the APA string (the chunk above shows both).
-   [x] N1.8 SurveyStudio Export, Interpretations canvas: each research question block shows badges, APA line, results table, chart, Copy result button, planned decision rule, and the interpretation box, in that order.
-   [ ] N1.9 Copy result puts the question, block id, decision rule, APA line, and table on the clipboard. Paste into a text editor to confirm.
-   [ ] N1.10 Canvas timing: the first open of the Export screen computes once (about half a minute for the 34-block demo), and returning to the screen after visiting another tab is close to instant.
-   [x] N1.11 Builder Report outline: the editable decision-rule field sits below each result's placeholder table and figure, and still syncs with the plan card and RQ dialogue both ways.

**Verification detail for the N1.x ticks above (2026-07-25):** N1.1 confirmed by reading `LICENSE`/`LICENSE.md` directly. N1.2/N1.4 confirmed by rendering the demo report on both paths (`render_report()` default and with `options(surveyframe.use_quarto = FALSE)`) and checking the output HTML for real `<table>` markup, title-cased codebook headers, and the absence of the old space-joined-row artefact. N1.3 confirmed at source: `render_report()`'s Response distributions section calls `sframe_plot_item_chart()`/`sframe_plot_scale_chart()` in both paths, the same helpers the analysis charts use, with the base-R fallback only firing when ggplot2 is absent. N1.5 confirmed live: a rendered correlation-plot object's axis labels read "Digital marketing" / "Satisfaction", not the raw ids. N1.6/N1.7 confirmed by running the exact chunk above. N1.8 and N1.11 confirmed by reading `output$export_interpretations_ui` (`inst/shiny/app.R`) and `reportSkeletonHtml()`/`saveRQ()`/`updatePlanInterp()` (`inst/builder/survey_builder.html`) directly: both build their DOM/HTML deterministically via synchronous templating, so source order is DOM order, and the outline's write path (`updatePlanInterp()`) and the RQ modal's write path (`saveRQ()`) write the same two fields (`decision_rule`, `interpretation`) on the same shared plan object, confirming the two-way sync structurally. N1.9 and N1.10 need a live browser (`navigator.clipboard` is not something headless Chrome grants without extra CDP permission wiring, and "feels instant" is not a thing to assert from source) — moved to `dogfeed.todo.md` for the 0.3.5 field-validation pass.

------------------------------------------------------------------------

## Deferred to 0.3.5 (logged in dogfeed.todo.md)

Everything below survived the 2026-07-25 machine pass unticked because it
genuinely needs a human: a real phone, a screen reader, a browser's Ctrl+P
dialogue, or a judgement about whether prose or a colour "reads right."
Deferring them does not weaken the 0.3.4 sign-off — none change behaviour
that the machine pass could not already verify structurally, and none are
release-blocking defects. Logged in `dogfeed.todo.md` for the 0.3.5
field-validation pass rather than left silently unticked here.

- **B1.3** — interpretation/decision-rule pairing reads correctly to a
  researcher (prose judgement).
- **D2.5** — scale-correlation heatmap readable at phone width.
- **D2.6 (remainder)** — filter-still-applies live check, the full
  levels/labels audit across every item type, the vignette/pkgdown
  documentation revision, and the ranking/matrix no-extra-coding check.
  This is real scoped feature/audit work, not a quick verification — the
  axis/title humanising half of D2.6 is already fixed and confirmed
  (N1.5).
- **E2.5** — phone native date-wheel bounds.
- **E2.6** — bounds error message styling next to the required-field
  message (visual judgement).
- **F2.6** — screen-reader spot check (VoiceOver/NVDA/Orca).
- **H1.4** — vignette prose spot-read for house style.
- **H1.5** — flat `browseVignettes()` output presentability.
- **I1.1 to I1.4** — the full 30-minute fresh-eyes UX pass (builder,
  SurveyStudio, and full-report read), already scoped for a dogfeed
  session by I1.4's own instruction.
- **J1.5** — APA strings reading naturally in context (prose judgement).
  J1.7's raw compute time is measured (14.3s) — only the perceived-speed
  half stays deferred.
- **K1.7** — MCAR interpretation wording reading correctly to a
  methodologist (prose judgement).
- **L1.4 to L1.7** — PDF pagination, greyscale legibility, browser print
  stylesheet, and brand-colour perception (all need eyes on a rendered
  page).
- **N1.9 to N1.10** — clipboard paste confirmation (blocked by headless
  Chrome's clipboard permission model, not attempted) and perceived
  canvas timing.
- **G1.4 to G1.5** — win-builder R-release and R-devel. Not a dogfeed
  item: this is the next release-process step, run by the owner after
  this file's tarball is built (see below), not deferred to 0.3.5.

------------------------------------------------------------------------

## Sign-off

-   [x] All parts above are complete or their exceptions are recorded. Every item checkable without a human is verified above (2026-07-25); every remaining item is a real "needs a human/device" case, explicitly recorded and moved to `dogfeed.todo.md` for 0.3.5, except **G1.4/G1.5 (win-builder)**, which are the immediate next release-process step, not a dogfeed deferral.
-   [ ] Decision recorded: submit 0.3.4 to CRAN or hold for fixes.

**Release-candidate status as of 2026-07-25:** `surveyframe_0.3.4.tar.gz` built from `main` at commit `6556f91`. `R CMD check --as-cran`: **Status OK, 0 errors, 0 warnings, 0 notes.** Full test suite: all tests pass (0 failed), naniar and pagedown both installed and exercised. Remaining before submission: win-builder R-release and R-devel (G1.4/G1.5), then this sign-off.

**Reviewer signature:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ **Date:** \_\_\_\_\_\_\_\_\_\_\_\_