---
title: "MAS Review - surveyframe 0.3.3"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: false
execute:
  echo: true
  warning: true
  error: true
  eval: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

**Reviewer:** Mohammed Ali Sharafuddin\
**Date created:** 2026-07-10\
**Version under review:** 0.3.3 (tagged v0.3.3, pushed to both remotes, not yet on CRAN)\
**Purpose:** Human verification of everything that changed in 0.3.3 before win-builder and CRAN submission, plus the 0.3.2 review areas this release touched again (vignettes, the Shiny surfaces, and the response pipeline). Machine verification (519 automated tests, headless browser drives, R CMD check clean) is done; this checklist is the human pass the machine cannot do.

This is the runnable companion to `mas_review_033.md`. Every chunk is
self-contained: work top to bottom in a fresh RStudio session and run each
chunk with Ctrl+Shift+Enter. Chunks that launch an app, open a browser, or
need your Google account are marked in their text. Rendering the whole file
executes nothing (`eval: false` globally); execution is chunk-by-chunk on
purpose.

------------------------------------------------------------------------

## Prerequisites

- [ ] RStudio open with a clean R session (restart R first).
- [ ] Chrome or Firefox available.
- [ ] `devtools`, `remotes`, `ggplot2`, `psych`, `lavaan`, `seminr`, `googlesheets4`, `shiny`, and `quarto` are installed.
- [ ] Internet access (GitHub install, Google Sheets read-back).
- [ ] Optional but recommended: a physical phone on the same network for Part C.

```{r chunk-00}
for (p in c("devtools", "remotes", "ggplot2", "psych", "lavaan",
            "seminr", "googlesheets4", "shiny")) {
  cat(sprintf("%-14s %s\n", p, requireNamespace(p, quietly = TRUE)))
}
```

------------------------------------------------------------------------

## Part A — Installation and metadata (6 items)

### Step A1 — Install the tagged release from GitHub

```{r chunk-01, eval=FALSE}
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
remotes::install_github("MohammedAliSharafuddin/surveyframe@v0.3.3",
                        force = TRUE)
```

Restart R after the install, then confirm:

```{r chunk-02}
packageVersion("surveyframe")
# Expected: '0.3.3'
```

- [ ] A1.1 Installation completes with no errors or warnings.
- [ ] A1.2 `packageVersion()` reports 0.3.3.

### Step A2 — Release metadata reads correctly

```{r chunk-03}
news(package = "surveyframe")
citation("surveyframe")
```

- [ ] A2.1 The 0.3.3 NEWS entry leads and reads as ONE release covering the plotting layer, the field-deployment fixes, and the survey redesign. No separate 0.3.4 heading, no raw markdown table anywhere in the file, and the analysis content comes first.
- [ ] A2.2 The prose contains no em-dashes, semicolons, "not X but Y" constructions, or banned words, and uses UK spellings.
- [ ] A2.3 `citation("surveyframe")` shows version 0.3.3.
- [ ] A2.4 The LICENSE file at the public repo root is the two-line CRAN template (`YEAR:` / `COPYRIGHT HOLDER:`), with the full MIT text in LICENSE.md only.

------------------------------------------------------------------------

## Part B — Branching engine fixes (10 items)

### Step B1 — Two-rule cascade instrument

The instrument mirrors the eligibility cascade that drove the fixes. The
export opens in your browser; click through it there.

```{r chunk-04}
library(surveyframe)
yn  <- sf_choices("yn", c("yes","no"), c("Yes","No"))
q1  <- sf_item("q1", "Are you 18 or older?", type = "single_choice",
               choice_set = "yn", required = TRUE)
q2  <- sf_item("q2", "Did you order room service?", type = "single_choice",
               choice_set = "yn", required = TRUE)
q3  <- sf_item("q3", "How many times?", type = "single_choice",
               choice_set = "yn", required = FALSE)
b1  <- sf_branch("q2", depends_on = "q1", operator = "==", value = "yes",
                 action = "show")
b2  <- sf_branch("q3", depends_on = "q1", operator = "==", value = "yes",
                 action = "show")
b3  <- sf_branch("q3", depends_on = "q2", operator = "==", value = "yes",
                 action = "show")
instr_b <- sf_instrument("Branch check",
                         components = list(yn, q1, q2, q3, b1, b2, b3))
export_static_survey(instr_b,
                     output_path = file.path(tempdir(), "branch.html"),
                     open = TRUE, overwrite = TRUE)
```

- [ ] B1 On load, only q1 is visible.
- [ ] B2 Answer q1 = Yes: q2 appears; q3 stays hidden (its second rule is unmet). This is the multi-rule AND fix; 0.3.2 would have shown q3 here.
- [ ] B3 Answer q2 = Yes: q3 appears.
- [ ] B4 Change q1 to No: q2 AND q3 both disappear, even though q2 still holds a stale Yes. This is the cascade fix; 0.3.2 kept q3 visible.
- [ ] B5 Change q1 back to Yes: q2 and q3 return with their earlier answers still selected.
- [ ] B6 With q1 = No, submit: no required-field error blocks the hidden questions, and the thank-you screen appears.
- [ ] B7 The downloaded CSV contains columns only for q1/q2/q3 plus the metadata columns, and no section-break columns.
- [ ] B8 Nothing in the flow feels laggy or flickers when visibility changes.

### Step B2 — Complex walkthrough: numeric coding and a screen-out message

From the review feedback: numeric-coded choice values (so the CSV needs no
recoding in R) and a screen-out message shown when the gate fails, built
from a branched `text_block`.

```{r chunk-05}
library(surveyframe)
yn2  <- sf_choices("yn2", 1:2, c("Yes","No"))       # numeric values, text labels
freq <- sf_choices("freq", 1:4,
                   c("Never","Once","Two or three times","Weekly"))
g1 <- sf_item("g1", "Are you 18 or older?", type = "single_choice",
              choice_set = "yn2", required = TRUE)
so <- sf_item("screen_out",
              "Sorry. You should be above 18 to participate in this survey.",
              type = "text_block")
g2 <- sf_item("g2", "Did you order room service?", type = "single_choice",
              choice_set = "yn2", required = TRUE)
g3 <- sf_item("g3", "How often did you order?", type = "single_choice",
              choice_set = "freq", required = FALSE)
r_so <- sf_branch("screen_out", depends_on = "g1", operator = "==",
                  value = "2", action = "show")     # value 2 = "No"
r_g2 <- sf_branch("g2", depends_on = "g1", operator = "==", value = "1",
                  action = "show")
r_g3a <- sf_branch("g3", depends_on = "g1", operator = "==", value = "1",
                   action = "show")
r_g3b <- sf_branch("g3", depends_on = "g2", operator = "==", value = "1",
                   action = "show")
instr_b2 <- sf_instrument("Branch check 2",
  components = list(yn2, freq, g1, so, g2, g3, r_so, r_g2, r_g3a, r_g3b))
export_static_survey(instr_b2,
                     output_path = file.path(tempdir(), "branch2.html"),
                     open = TRUE, overwrite = TRUE)
```

- [ ] B9 Answering g1 = No shows the screen-out message block and hides g2/g3; answering Yes hides the message and reveals the chain.
- [ ] B10 The downloaded CSV stores numeric codes (1/2 for g1 and g2, 1 to 4 for g3), analysable in R without recoding.

------------------------------------------------------------------------

## Part C — Theme B survey design (18 items)

### Step C1 — Desktop pass

```{r chunk-06}
library(surveyframe)
demo <- sframe_input_types_demo_data()
export_static_survey(demo$instrument,
                     output_path = file.path(tempdir(), "themeb.html"),
                     open = TRUE, overwrite = TRUE)
```

- [ ] C1 Welcome screen: serif title, intro text, tinted consent box, and a full-width start button in the theme colour. The branded header (logo or study name, no decorative dot) shows above.
- [ ] C2 Question text is serif and clearly larger than body text; help text is italic; the required marker is a small theme-coloured asterisk.
- [ ] C3 Single-choice options are bordered cards with A/B/C key chips; the selected card shows a tinted background, a theme border, a filled chip, and a right-side tick.
- [ ] C4 Multiple-choice cards behave the same and allow several selections.
- [ ] C5 Likert items are numbered squares with small labels underneath; the selected square fills with the theme colour and its number turns white.
- [ ] C6 Short text/number/date inputs are underline style; the underline turns the theme colour on focus. The textarea keeps a boxed style. The numeric nights question refuses values below 0 and above 365 (typed values clamp).
- [ ] C7 Star rating: large stars, theme colour when active, subtle grow on hover.
- [ ] C8 Matrix (tinted header row, horizontal rules, hover highlight), slider (card container, large serif value, both end labels visible including 0), and ranking (option-card rows with theme rank badges) all carry the Theme B design.
- [ ] C9 Section breaks show the short theme-coloured rule above a serif heading with italic intro.
- [ ] C10 The sticky top bar shows the progress line, a label, and the percentage, with no dash or duplicated page counter. On a single-page survey the label counts answered questions and updates as you answer.

### Step C2 — Theme colour selector

```{r chunk-07}
demo2 <- sframe_input_types_demo_data()
demo2$instrument$render$theme <- "#7c3aed"   # any colour you like
export_static_survey(demo2$instrument,
                     output_path = file.path(tempdir(), "themeb2.html"),
                     open = TRUE, overwrite = TRUE)
```

- [ ] C11 Every accent re-tints from the one colour: progress bar, chips, selected cards, Likert fills, rank badges, stars, buttons, focus states. No element kept the old colour.
- [ ] C12 Repeat once from the SurveyBuilder's Settings colour picker (launch_builder, change the theme colour, Export survey) and confirm the exported file matches the picked colour.

### Step C3 — Phone pass (real device if possible)

Serve the exported file and open it on the phone, or use browser device
mode at 390 x 844:

```{r chunk-08, eval=FALSE}
# In a terminal, from the folder holding themeb.html:
#   python3 -m http.server 8000
# then open http://<your-ip>:8000/themeb.html on the phone.
tempdir()    # the folder holding the exports from the chunks above
```

- [ ] C13 No horizontal scrolling anywhere.
- [ ] C14 Option cards and Likert squares are comfortably tappable with a thumb; no mis-taps on adjacent options.
- [ ] C15 The on-screen keyboard does not permanently cover a focused text input.
- [ ] C16 A logo set in the header keeps a stable box: try one wide and one tall logo and confirm the study name does not jump horizontally.
- [ ] C17 Conversational mode (chunk below): one question per page, a "Question X of Y" eyebrow with no dash line, the study title in the sticky bar, and working Back/Next.
- [ ] C18 Overall judgement: the survey looks like the approved Theme B mock, and you would be happy sending this to a respondent today.

```{r chunk-09}
demo3 <- sframe_input_types_demo_data()
demo3$instrument$render$mode <- "conversational"
export_static_survey(demo3$instrument,
                     output_path = file.path(tempdir(), "themeb_conv.html"),
                     open = TRUE, overwrite = TRUE)
```

------------------------------------------------------------------------

## Part D — SurveyBuilder fixes (10 items)

This part is manual clicking; the chunk only launches the builder.

```{r chunk-10, eval=FALSE}
surveyframe::launch_builder()
```

- [ ] D1 Empty canvas shows accessible CTA buttons at 44px minimum: a filled "+ Add your first question" and outlined "Open an existing .sframe" / "Set up survey settings", all clearly readable against the dark sidebar.
- [ ] D2 Open a `.sframe` saved by R or by the builder: the toast reports "integrity verified (sha256 ...)". A file edited by hand in a text editor reports an integrity mismatch instead. (Known: the Preview tab still shows the old simplified design; Export survey and SurveyStudio carry Theme B. The Preview restyle is scheduled in the 0.3.4 builder rework.)
- [ ] D3 Select a Section item: the inspector labels its text field "Section header" (a Text block reads "Block text"); ordinary questions still read "Question text".
- [ ] D4 Add a Multiple choice question, edit its choices to something distinctive, then add another Multiple choice question: the new one gets a fresh Option 1/2/3 set.
- [ ] D5 Edit the choices of a choice question that shares its set with another question: the other question's choices are untouched (the set forks).
- [ ] D6 Analyse tab, Add Analysis Plan, Pearson correlation: pick the same variable for X and Y. The modal blocks it and Save stays disabled until fixed.
- [ ] D7 Suggestion box with one Likert item and one numeric item suggests Spearman correlation (never Mann-Whitney).
- [ ] D8 A single-choice item with more than two categories plus a numeric outcome suggests Kruskal-Wallis; a two-category grouping stays Mann-Whitney.
- [ ] D9 Known and accepted: the Preview tab lags Theme B until the 0.3.4 builder rework. Confirm it is logged and does not block submission.
- [ ] D10 Export survey and Generate collector still download files without errors after all the above.

------------------------------------------------------------------------

## Part E — Model syntax fixes (7 items)

### Step E1 — Label sanitisation

```{r chunk-11}
library(surveyframe)
con1 <- sf_construct("SAT", "Satisfaction", paste0("sat_", 1:3))
con2 <- sf_construct("LOY", "Loyalty", paste0("loy_", 1:3))
m <- sf_model("m1", "Labelled", type = "cb_sem",
              constructs = list(con1, con2),
              paths = list(sf_path("SAT", "LOY",
                                   "H1: SAT positively influences LOY")))
cat(sem_lavaan_syntax(m))
```

- [ ] E1 The structural line reads `LOY ~ H1*SAT`. The free-text label was reduced to the `H1` tag; no spaces or colons appear in the parameter.

```{r chunk-12}
# lavaanify() takes the generated syntax STRING, never the model object
pt <- lavaan::lavaanify(sem_lavaan_syntax(m))
head(pt[, c("lhs", "op", "rhs", "label")], 10)
```

- [ ] E2 The call parses with no warnings about identifiers with spaces, and the `label` column shows `H1` on the structural row.

### Step E2 — seminr syntax runs as generated

```{r chunk-13}
pls <- sf_model("p1", type = "pls_sem",
  constructs = list(
    sf_construct("SAT", "Satisfaction", paste0("sat_", 1:3),
                 mode = "composite"),
    sf_construct("LOY", "Loyalty", "loy_1", mode = "single_item")),
  paths = list(sf_path("SAT", "LOY")))
cat(seminr_syntax(pls, nboot = 50))
```

- [ ] E3 The generated code includes `library(seminr)` after the check_installed line.
- [ ] E4 The assessment tail uses `summary(pls_model)` accessors (`$reliability`, `$validity$htmt`, bootstrapped paths). No `reliability()`/`ave()`/`htmt()` calls.

```{r chunk-14}
# E5: run the generated block exactly as produced, against simulated data
# named `data` (the generated code's default data name)
set.seed(42)
data <- as.data.frame(matrix(sample(1:5, 120 * 4, replace = TRUE),
                             nrow = 120,
                             dimnames = list(NULL, c("sat_1", "sat_2",
                                                     "sat_3", "loy_1"))))
eval(parse(text = seminr_syntax(pls, nboot = 50)))
summary(pls_model)$reliability
```

- [ ] E5 The generated block runs end to end with no "could not find function".

```{r chunk-15}
# E6: an analysis plan declaring method = "pls_sem" (the model-type
# spelling) resolves to the seminr syntax runner
cs  <- sf_choices("ag5", 1:5,
                  c("Strongly disagree", "Disagree", "Neutral",
                    "Agree", "Strongly agree"))
its <- lapply(c("sat_1", "sat_2", "sat_3", "loy_1"), function(id)
  sf_item(id, paste("Item", id), type = "likert", choice_set = "ag5"))
instr_e <- sf_instrument("PLS alias check",
                         components = c(list(cs), its))
instr_e <- add_model(instr_e, pls)
instr_e$analysis_plan <- list(list(
  id = "RQ_pls", research_question = "Does satisfaction predict loyalty?",
  family = "measurement", method = "pls_sem",
  roles = list(model = "p1"), options = list(alpha = 0.05)))
res_e <- run_analysis_plan(cbind(respondent_id = paste0("R", 1:120),
                                 submitted_at = as.character(Sys.time()),
                                 data),
                           instr_e, scored = FALSE)
cat("error:", res_e[[1]]$error %||% "none", "\n")
cat(substr(res_e[[1]]$syntax, 1, 120), "...\n")
```

- [ ] E6 `method = "pls_sem"` returns the seminr syntax rather than "Test 'pls_sem' is unavailable".
- [ ] E7 The Model Builder panel in the builder exports the same syntax through its Export syntax button (manual, with the builder open from Part D).

------------------------------------------------------------------------

## Part F — Google Sheets collection (9 items)

### Step F1 — Static checks on the export and collector

```{r chunk-16}
library(surveyframe)
demo_f <- sframe_input_types_demo_data()
f_html <- file.path(tempdir(), "cors_check.html")
export_static_survey(demo_f$instrument, output_path = f_html, open = FALSE,
                     overwrite = TRUE,
                     endpoint_url = "https://script.google.com/macros/s/X/exec")
html <- paste(readLines(f_html, warn = FALSE), collapse = "\n")
cat("no-cors present:      ", grepl("mode:'no-cors'", html, fixed = TRUE), "\n")
cat("text/plain present:   ", grepl("'Content-Type':'text/plain'", html,
                                    fixed = TRUE), "\n")
cat("application/json gone:", !grepl("'Content-Type':'application/json'",
                                     html, fixed = TRUE), "\n")

gs <- export_google_sheet(demo_f$instrument,
                          sheet_url = "https://docs.google.com/spreadsheets/d/demo",
                          output_dir = tempdir())
cols_line <- grep("EXPECTED_COLUMNS", readLines(gs), value = TRUE)[1]
cat("sec_ columns gone:    ", !grepl("sec_", cols_line), "\n")
cat("ranking expanded:     ", grepl("__", cols_line), "\n")
```

- [ ] F1 All five printed checks are TRUE.
- [ ] F2 `EXPECTED_COLUMNS` contains no `sec_*` columns and expands ranking/matrix items to `item__` columns.

### Step F2 — Live round trip (needs your Google account)

- [ ] F3 Deploy the fresh collector to a scratch Google Sheet, host the exported survey over HTTP (not file://), submit one response, and confirm the row lands in the sheet. This proves the CORS fix in the field.

```{r chunk-17, eval=FALSE}
# Fill in your scratch sheet URL after deploying the collector
sheet_url <- "PASTE-SHEET-URL"
resp_f <- read_sheet_responses(sheet_url, demo_f$instrument)
head(resp_f)
resp_f2 <- read_sheet_responses(sheet_url, demo_f$instrument,
                                meta_cols = c("extra1"))
```

- [ ] F4 `read_sheet_responses()` returns the submitted row with correct values.
- [ ] F5 The `meta_cols` variant accepts an extra sheet column without the undeclared-column warning.
- [ ] F6 SurveyStudio, Upload screen, Google Sheet card: the same sheet imports, and the completion time card shows a median.
- [ ] F7 Live AIC-RSAM read-back (read-only, own account): `read_sheet_responses()` on the live sheet returns the real responses cleanly. Do NOT submit test data to the live endpoint.
- [ ] F8 Decision recorded: re-export and republish the live AIC-RSAM survey mid-collection (CORS fix + Theme B), or leave the hand-patched deployment until collection ends.
- [ ] F9 The deploy kit lives in AI_Room_service/surveyframe_Integration/deploy_kit and its README says so.

------------------------------------------------------------------------

## Part G — Visualisation foundation (12 items)

### Step G1 — Plots on the bundled demo

```{r chunk-18}
library(surveyframe)
library(ggplot2)
demo_g <- sframe_demo_data()
resp_g <- read_responses(demo_g$responses_path, demo_g$instrument,
                         respondent_id = "respondent_id",
                         submitted_at  = "submitted_at",
                         meta_cols     = "started_at")
res_g <- run_analysis_plan(resp_g, demo_g$instrument, plots = TRUE)
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_surveyframe()
```

```{r chunk-19}
# Show every plot the plan produced, one after another
for (r in res_g) {
  if (!is.null(r$plot)) {
    cat("\n---", r$block_id %||% "", "-", r$test %||% "", "---\n")
    print(r$plot)
  }
}
```

- [ ] G1 `theme_surveyframe()` renders branded: ink text, quiet grids, bottom legend.
- [ ] G2 A frequency block's plot is a horizontal teal bar chart with no bar for missing values.
- [ ] G3 A chi-square/cross-tab plot is a dodged bar chart in the fixed series order (teal, amber, blue, pink, violet) with white gaps and a bottom legend.
- [ ] G4 A correlation plot is a scatter with an ink regression line, a shaded band, and the APA string as subtitle.
- [ ] G5 A regression plot shows predictor vs outcome (one predictor) or observed vs fitted with a dashed identity line (several predictors).

```{r chunk-20}
# G6/G7/G8: nulls, defaults, and tables
sapply(res_g, function(r) c(test = r$test %||% "",
                            has_plot = !is.null(r$plot),
                            has_table = !is.null(r$table)))
res_off <- run_analysis_plan(resp_g, demo_g$instrument)   # plots = FALSE
cat("plots when off:", sum(!sapply(res_off, function(r) is.null(r$plot))),
    "(expect 0)\n")
for (r in res_off) {
  if (is.data.frame(r$table)) {
    cat("\n---", r$block_id %||% "", "---\n")
    print(knitr::kable(r$table))
  }
}
```

- [ ] G6 Block types outside the family have `$plot = NULL` and no error.
- [ ] G7 With the default `plots = FALSE` no block carries a plot.
- [ ] G8 Correlation tables are one statistic row; regression tables are coefficient tables; t-test tables are two-row group summaries. All print through `kable()`.

```{r chunk-21}
# G9: the report picks the tables up automatically
out_g <- file.path(tempdir(), "report_g.html")
render_report(demo_g$instrument, resp_g, output_file = out_g)
utils::browseURL(out_g)
```

```{r chunk-22}
# G10: empty strings count as missing, never as a category
df <- data.frame(respondent_id = paste0("R", 1:10),
                 submitted_at  = as.character(Sys.time()),
                 gender = c("m", "f", "", "m", "", "f", "m", "f", "m", ""),
                 member = c("yes", "", "no", "yes", "no",
                            "yes", "", "no", "yes", "no"),
                 stringsAsFactors = FALSE)
cs_g  <- sf_choices("mf", c("m", "f"), c("Male", "Female"))
yn_g  <- sf_choices("yn", c("yes", "no"), c("Yes", "No"))
i1_g  <- sf_item("gender", "Gender?", type = "single_choice", choice_set = "mf")
i2_g  <- sf_item("member", "Member?", type = "single_choice", choice_set = "yn")
instr_g <- sf_instrument("Empty strings",
                         components = list(cs_g, yn_g, i1_g, i2_g),
                         analysis_plan = list(
  list(id = "Q1", research_question = "Distribution of gender",
       family = "descriptive", method = "frequency",
       roles = list(variable = "gender")),
  list(id = "Q2", research_question = "Gender by membership",
       family = "categorical", method = "chi_square",
       roles = list(row = "gender", column = "member"))))
res_es <- run_analysis_plan(df, instr_g, scored = FALSE)
print(res_es[[1]]$table)     # frequency: NA counted, no "" row
print(res_es[[2]]$table)     # crosstab: no unnamed category
```

- [ ] G9 The rendered report shows the result tables in the analysis sections.
- [ ] G10 Neither table shows a blank or unnamed category for the empty strings.
- [ ] G11 Eyeball the plots at 150 percent zoom: fonts legible, no clipped labels.
- [ ] G12 Optional: without ggplot2 installed, `plots = TRUE` fails with the friendly check_installed prompt and everything else works.

------------------------------------------------------------------------

## Part H — Report checks (6 items)

```{r chunk-23}
# A small instrument with a reliability block and a CFA syntax block, so
# the report exercises both new analysis-section renderings
library(surveyframe)
cs_h <- sf_choices("ag5", 1:5,
                   c("Strongly disagree", "Disagree", "Neutral",
                     "Agree", "Strongly agree"))
items_h <- lapply(paste0("sat_", 1:4), function(id)
  sf_item(id, paste("Satisfaction item", id), type = "likert",
          choice_set = "ag5", scale_id = "sat", required = TRUE))
scale_h <- sf_scale("sat", "Satisfaction", paste0("sat_", 1:4),
                    method = "mean")
model_h <- sf_model("cfa1", "Satisfaction CFA", type = "cfa",
                    constructs = list(
                      sf_construct("sat", "Satisfaction",
                                   paste0("sat_", 1:4))))
instr_h <- sf_instrument("Report check",
  components = c(list(cs_h), items_h, list(scale_h)),
  models = list(model_h),
  analysis_plan = list(
    list(id = "RQ1", research_question = "Is the scale reliable?",
         family = "measurement", method = "reliability_alpha",
         roles = list(items = paste0("sat_", 1:4))),
    list(id = "RQ2", research_question = "Does the CFA syntax generate?",
         family = "measurement", method = "cfa_lavaan_syntax",
         roles = list(model = "cfa1"))))
set.seed(7)
dat_h <- as.data.frame(matrix(sample(1:5, 60 * 4, replace = TRUE), 60,
                              dimnames = list(NULL, paste0("sat_", 1:4))))
dat_h <- cbind(respondent_id = paste0("R", 1:60),
               submitted_at = as.character(Sys.time()), dat_h)
out_h <- file.path(tempdir(), "report_h.html")
render_report(instr_h, dat_h, output_file = out_h)
utils::browseURL(out_h)
```

- [ ] H1 The RQ2 section prints the full CFA syntax in a monospaced code block.
- [ ] H2 The RQ1 section renders a compact per-scale alpha table instead of an empty "Result:" line.
- [ ] H3 The reliability section's "Omega h" and "Omega total" headers do not wrap mid-word.
- [ ] H4 View the report at 1280 x 720: headings, tables, and plots readable from presentation distance.
- [ ] H5 The references show italic journal names (no literal asterisks).
- [ ] H6 The TOC navigates correctly to every section.

------------------------------------------------------------------------

## Part I — Release safety (8 items)

Everything here runs from RStudio. The first chunk swaps the web-installed
package for the local development checkout so the remaining chunks test
exactly what will be submitted.

```{r chunk-24}
# Path to the local development checkout; adjust if yours differs
pkg <- path.expand("~/Documents/GitHub/surveyframe-dev")
file.exists(file.path(pkg, "DESCRIPTION"))

# Drop the GitHub-installed copy and install the local tree in its place
try(detach("package:surveyframe", unload = TRUE), silent = TRUE)
try(remove.packages("surveyframe"), silent = TRUE)
devtools::install(pkg, upgrade = "never", quick = TRUE)
library(surveyframe)
packageVersion("surveyframe")   # expect 0.3.3, now from the local tree
```

```{r chunk-24b}
# I1: full suite from the local checkout (expect 519; the CRAN tarball
# runs fewer because the AIC-RSAM fixture tests are dev-only)
devtools::test(pkg)
```

```{r chunk-24c}
# I2: build the tarball and run the CRAN check on it (several minutes)
tarball <- devtools::build(pkg)
tarball
devtools::check_built(tarball, cran = TRUE)
```

```{r chunk-24d}
# I3: the tarball must contain no dev files (expect character(0))
files <- utils::untar(tarball, list = TRUE)
grep("mas_|dogfeed|roadmap|aic_rsam|CODE_OF|SECURITY",
     files, value = TRUE)
```

```{r chunk-24e}
# I5/I6: tag position and remote agreement
system2("git", c("-C", pkg, "describe", "--tags"), stdout = TRUE)
system2("git", c("-C", pkg, "ls-remote", "origin",
                 "refs/heads/main", "refs/tags/v0.3.3"), stdout = TRUE)
system2("git", c("-C", pkg, "ls-remote", "public",
                 "refs/heads/main", "refs/tags/v0.3.3"), stdout = TRUE)
```

- [ ] I1 Full suite passes (519 in the dev checkout).
- [ ] I2 `R CMD check --as-cran` on the fresh tarball: 0 errors, 0 warnings, at most the incoming-feasibility NOTE.
- [ ] I3 The tarball contains no dev files (the grep above returns nothing).
- [ ] I4 The vignette builds offline.
- [ ] I5 `git describe --tags` on main says v0.3.3 and the tag includes the review fixes.
- [ ] I6 Public and private remotes agree on main and the tag.
- [ ] I7 Win-builder R-release and R-devel both return Status OK.
- [ ] I8 `cran-comments.md` carries the local and win-builder results and notes ggplot2 in Suggests only.

------------------------------------------------------------------------

## Part J — Fresh-eyes UX pass (4 items)

Thirty minutes as a first-time user, no checklist in hand.

- [ ] J1 Take a survey end to end on your phone as a respondent. Note anything off in dogfeed.todo.md rather than fixing on the spot.
- [ ] J2 Build a five-question instrument in the builder from scratch, export, open the export. Anything confusing goes to dogfeed.todo.md.
- [ ] J3 Read the 0.3.3 NEWS entry once more as a stranger: is it clear what changed and why it matters?
- [ ] J4 Confirm dogfeed.todo.md's "After this session" and deferred items still reflect reality.

------------------------------------------------------------------------

## Part K — Vignettes (6 items)

Added 2026-07-10: the 0.3.2 review's Part U (vignettes) was not revisited
for 0.3.3, and this release changed the survey design, the report, and the
collection path the vignettes describe.

```{r chunk-25}
browseVignettes("surveyframe")   # restart R first if this shows nothing
```

```{r chunk-26, eval=FALSE}
# Knit the lead vignette from the dev checkout
rmarkdown::render("vignettes/surveyframe.Rmd", output_dir = tempdir())
```

- [ ] K1 All vignettes list and open from `browseVignettes()` after a fresh install.
- [ ] K2 The lead vignette knits cleanly with the installed 0.3.3.
- [ ] K3 No vignette prose or screenshot contradicts the Theme B design, and analysis output claims match what 0.3.3 prints (including `$table`).
- [ ] K4 The deployment vignette's Google Sheets instructions match the 0.3.3 collector: no CORS patching, no section-break columns.
- [ ] K5 Decide and record (dogfeed.todo.md): a `plots = TRUE` example in the lead vignette now, or in the 0.3.4 breadth patch?
- [ ] K6 No vignette prose violates the writing rules.

------------------------------------------------------------------------

## Part L — Shiny renderer and dashboard (4 items)

Added 2026-07-10: Theme B restyled the static HTML export only. The Shiny
survey module has its own renderer and the dashboard its own charts;
neither was touched in 0.3.3.

```{r chunk-27, eval=FALSE}
demo_l <- sframe_demo_data()
render_survey(demo_l$instrument)       # Shiny renderer; Ctrl+C / Stop to end
```

```{r chunk-28, eval=FALSE}
launch_dashboard()                     # bundled demo dashboard
```

- [ ] L1 The Shiny renderer still works for every item type (it predates Theme B; confirm nothing broke).
- [ ] L2 The divergence is recorded: the Shiny renderer does not carry Theme B; convergence is scheduled with the 0.3.4 builder rework.
- [ ] L3 `launch_dashboard()` panels all render with the bundled demo, including the ranking item's data.
- [ ] L4 SurveyStudio end to end with the bundled sample: preview, upload, quality, reliability, analysis plan, and export screens all function.

------------------------------------------------------------------------

## Part M — Response pipeline round trip (4 items)

Added 2026-07-10: the ranking export change altered the data shape; the
0.3.2 review's pipeline parts assumed the old single-column form.

```{r chunk-29}
library(surveyframe)
demo_m <- sframe_input_types_demo_data()
# M1: the bundled CSV still uses the legacy single ranking column
resp_m <- read_responses(demo_m$responses_path, demo_m$instrument,
                         respondent_id = "respondent_id",
                         submitted_at  = "submitted_at",
                         meta_cols     = "started_at")
str(resp_m[, grep("ranking", colnames(resp_m)), drop = FALSE])
```

```{r chunk-30}
# M2/M3: a fresh CSV downloaded from the Theme B export (Part C) with
# item__option ranking columns. Point csv_path at the downloaded file.
csv_path <- file.path(path.expand("~"), "Downloads",
                      list.files(file.path(path.expand("~"), "Downloads"),
                                 pattern = "_response_.*\\.csv$")[1])
csv_path
resp_new <- read_responses(csv_path, demo_m$instrument,
                           respondent_id = "respondent_id",
                           meta_cols     = "started_at")
colnames(resp_new)[grep("__", colnames(resp_new))]
qr_m <- quality_report(resp_new, demo_m$instrument,
                       respondent_id = "respondent_id")
sc_m <- score_scales(resp_new, demo_m$instrument)
rr_m <- reliability_report(resp_new, demo_m$instrument)
```

- [ ] M1 The bundled legacy responses read without errors: backwards compatibility holds.
- [ ] M2 The fresh CSV (with `item__option` ranking columns) reads back with no missing-column or undeclared-column warnings.
- [ ] M3 `quality_report()`, `score_scales()`, and `reliability_report()` all run on the fresh CSV.
- [ ] M4 The codebook and full report render from the fresh data; ranking columns appear sensibly.

```{r chunk-31}
out_m <- file.path(tempdir(), "report_m.html")
render_report(demo_m$instrument, resp_new, output_file = out_m)
utils::browseURL(out_m)
```

------------------------------------------------------------------------

## Sign-off

Counts: Parts A-M, 104 checklist items (88 original, B9-B10 added with the
Part B rework, K1-K6, L1-L4, and M1-M4 added 2026-07-10 from the 0.3.2
review's uncovered areas).

**Reviewed by:**\
**Date signed off:**\
**All steps complete:**\
**Notes for the developer (if any):**

Once signed off, paste the following prompt into Claude Code to proceed:

```
Read CLAUDE.md and mas_review_033.md. The MAS review of 0.3.3 is complete
and signed off. Rebuild the tarball if anything changed, confirm
R CMD check --as-cran is 0 errors, 0 warnings, at most 1 NOTE, update
cran-comments.md with the local and win-builder results, and give me the
text to paste on the CRAN submission form.
```
