---
title: "Getting started with surveyframe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with surveyframe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(surveyframe)
```

surveyframe treats a survey instrument as a first-class R object. One
declarative source drives validation, deployment, quality checking, scoring,
psychometric diagnostics, and report generation. This vignette walks through
the complete workflow using a five-item customer satisfaction instrument.

---

## 1. Define the instrument

Every instrument is built from typed component objects assembled by
`sf_instrument()`. Start with choice sets, then items, then scales. Branching
rules and attention checks are optional but recommended.

```{r define}
# A five-point agreement scale used across all Likert items
agree5 <- sf_choices(
  id     = "agree5",
  values = 1:5,
  labels = c("Strongly disagree", "Disagree", "Neutral",
             "Agree", "Strongly agree")
)

# A yes/no set used for a screening item
yn <- sf_choices(
  id     = "yn",
  values = c("yes", "no"),
  labels = c("Yes", "No")
)

# Five satisfaction items, all linked to the "sat" scale
sat_1 <- sf_item(
  id         = "sat_1",
  label      = "The service was delivered on time.",
  type       = "likert",
  choice_set = "agree5",
  scale_id   = "sat",
  required   = TRUE
)

sat_2 <- sf_item(
  id         = "sat_2",
  label      = "Staff were knowledgeable and helpful.",
  type       = "likert",
  choice_set = "agree5",
  scale_id   = "sat",
  required   = TRUE
)

sat_3 <- sf_item(
  id         = "sat_3",
  label      = "The overall experience did not meet my expectations.",
  type       = "likert",
  choice_set = "agree5",
  scale_id   = "sat",
  required   = TRUE,
  reverse    = TRUE   # this item is reverse-coded
)

sat_4 <- sf_item(
  id         = "sat_4",
  label      = "I would use this service again.",
  type       = "likert",
  choice_set = "agree5",
  scale_id   = "sat",
  required   = TRUE
)

sat_5 <- sf_item(
  id         = "sat_5",
  label      = "I would recommend this service to others.",
  type       = "likert",
  choice_set = "agree5",
  scale_id   = "sat",
  required   = TRUE
)

# Demographic items
age    <- sf_item("age",    "What is your age?",   type = "numeric")
gender <- sf_item("gender", "What is your gender?",
                  type = "single_choice", choice_set = "yn")

# An attention check item
attn_item <- sf_item(
  id         = "attn_1",
  label      = "To confirm you are reading carefully, please select Agree.",
  type       = "likert",
  choice_set = "agree5"
)

# The satisfaction scale definition
sat_scale <- sf_scale(
  id            = "sat",
  label         = "Customer Satisfaction",
  items         = c("sat_1", "sat_2", "sat_3", "sat_4", "sat_5"),
  method        = "mean",
  min_valid     = 4L
)

# Attach the attention check to the attn_1 item
attn_check <- sf_check(
  id          = "chk_attn_1",
  item_id     = "attn_1",
  type        = "attention",
  pass_values = 4,
  fail_action = "flag",
  label       = "Attention check"
)

# Assemble the instrument
instr <- sf_instrument(
  title       = "Customer Satisfaction Survey",
  version     = "1.0.0",
  description = "Measures overall satisfaction with a five-item Likert scale.",
  authors     = "Mohammed Ali Sharafuddin",
  components  = list(
    agree5, yn,
    sat_1, sat_2, sat_3, sat_4, sat_5,
    age, gender, attn_item,
    sat_scale,
    attn_check
  )
)

print(instr)
```

---

## 2. Validate and save

`validate_sframe()` checks the instrument for structural problems before
anything else happens. `write_sframe()` saves the validated instrument as a
SHA-256 hashed JSON file. The file is human-readable, version-control
friendly, and portable across machines and R sessions.

```{r validate-save, eval = FALSE}
# Validate (raises an error with a detailed message if problems exist)
instr <- validate_sframe(instr)

# Save to a .sframe file
write_sframe(instr, "satisfaction_survey.sframe")

# Load back later in any R session
instr <- read_sframe("satisfaction_survey.sframe")
```

The `strict = FALSE` option returns a list of problems without stopping,
which is useful during instrument development:

```{r validate-soft}
result <- validate_sframe(instr, strict = FALSE)
cat("Valid:", result$valid, "\n")
cat("Problems:", length(result$problems), "\n")
```

---

## 3. Deploy as a Shiny survey

`render_survey()` turns the instrument into a running Shiny application. All
item types, required-field logic, and single-condition branching rules are
rendered automatically. When `save_responses = "csv"`, submitted responses are
written to a CSV file with `started_at` and `submitted_at` metadata columns.

```{r render, eval = FALSE}
render_survey(
  instr,
  save_responses = "csv",
  output_path = "responses.csv"
)
```

To launch the full SurveyStudio interface, which provides a visual shell for
the complete workflow, use `launch_studio()`:

```{r studio, eval = FALSE}
launch_studio(instrument = instr)
```

All remaining workflow steps also work in regular R scripts. Shiny is only
needed for interactive survey deployment and the SurveyStudio interface.

---

## 4. Load responses

Once data are collected (as a CSV from the Shiny app or any external tool),
load them with `read_responses()`. Column names must match instrument item IDs.
Files written by `render_survey()` include `started_at` and `submitted_at`
metadata columns, so load those through `meta_cols` and `submitted_at`.

```{r load-responses}
# Simulate a response dataset for this vignette
set.seed(101)
n <- 120
responses_raw <- data.frame(
  respondent_id = paste0("P", seq_len(n)),
  submitted_at  = as.character(
    as.POSIXct("2024-06-01") + sample(0:86400, n, replace = TRUE)),
  sat_1  = sample(1:5, n, replace = TRUE, prob = c(.05,.10,.15,.40,.30)),
  sat_2  = sample(1:5, n, replace = TRUE, prob = c(.05,.10,.20,.40,.25)),
  sat_3  = sample(1:5, n, replace = TRUE, prob = c(.30,.35,.20,.10,.05)),
  sat_4  = sample(1:5, n, replace = TRUE, prob = c(.05,.10,.15,.35,.35)),
  sat_5  = sample(1:5, n, replace = TRUE, prob = c(.05,.10,.20,.35,.30)),
  age    = sample(18:70, n, replace = TRUE),
  gender = sample(c("yes","no"), n, replace = TRUE),
  attn_1 = c(rep(4, round(n * 0.90)), rep(2, round(n * 0.10))),
  stringsAsFactors = FALSE
)

responses <- read_responses(
  x             = responses_raw,
  instrument    = instr,
  respondent_id = "respondent_id",
  submitted_at  = "submitted_at"
)

cat("Loaded:", nrow(responses), "responses,", ncol(responses), "columns\n")
```

---

## 5. Check data quality

`quality_report()` evaluates the response data against the checks defined in
the instrument. It covers attention check performance, straight-lining within
scale blocks, item and respondent-level missingness, and duplicate IDs.

```{r quality}
qr <- quality_report(
  data          = responses,
  instrument    = instr,
  respondent_id = "respondent_id",
  submitted_at  = "submitted_at"
)

print(qr)
```

The quality report object is structured, so individual results are accessible
programmatically:

```{r quality-access}
# Pass rate for the attention check
cat("Attention check pass rate:",
    round(qr$attention$chk_attn_1$pass_rate * 100, 1), "%\n")

# Number of respondents flagged for straight-lining
cat("Straight-line flags (sat scale):",
    length(qr$straightline$sat$flagged_rows), "\n")
```

---

## 6. Score scales

`score_scales()` applies the scale definitions from the instrument: reverse
coding, composite scoring, and minimum valid item rules. It returns a tibble
with one scored column per scale appended.

```{r score}
scored <- score_scales(
  data       = responses,
  instrument = instr,
  keep_items = TRUE,
  keep_meta  = TRUE
)

cat("Scale column added: 'sat'\n")
cat("Mean satisfaction score:", round(mean(scored$sat, na.rm = TRUE), 3), "\n")
cat("SD:", round(sd(scored$sat, na.rm = TRUE), 3), "\n")
```

---

## 7. Psychometric diagnostics

### Reliability

`reliability_report()` computes Cronbach's alpha and McDonald's omega for
each scale. Both statistics are derived from the instrument's scale
definitions, so item membership and reverse coding are already known.

```{r reliability}
rr <- reliability_report(responses, instr, omega = FALSE)
print(rr)
```

### Item diagnostics

`item_report()` produces item-total correlations, floor and ceiling
proportions, and descriptive statistics for every item within each scale.

```{r items}
ir <- item_report(responses, instr)
print(ir[[1]]$diagnostics)
```

### EFA readiness

`efa_report()` prepares a dataset for exploratory factor analysis by
computing KMO adequacy and Bartlett's test. It does not estimate a factor
solution. That step belongs to the researcher, using `psych::fa()` or a
similar package.

```{r efa}
er <- efa_report(responses, instr)
print(er)
```

### CFA syntax

`cfa_syntax()` generates a `lavaan` model string directly from the scale
structure in the instrument. Paste it into `lavaan::cfa()` when you are
ready to fit the confirmatory model.

```{r cfa}
syntax <- cfa_syntax(instr)
cat(syntax)
```

To fit the model:

```{r cfa-fit, eval = FALSE}
library(lavaan)
fit <- cfa(syntax, data = scored, std.lv = TRUE)
summary(fit, fit.measures = TRUE)
```

---

## 8. Generate a codebook and report

`codebook_report()` produces a structured codebook from the instrument
definition. `render_report()` assembles everything into a single reproducible
HTML file.

```{r codebook}
cb <- codebook_report(instr)
print(cb)
```

```{r report, eval = FALSE}
# Requires the quarto package
render_report(
  instrument          = instr,
  data                = responses,
  output_file         = "satisfaction_survey_report.html",
  include_codebook    = TRUE,
  include_quality     = TRUE,
  include_reliability = TRUE
)
```

The report embeds the instrument hash so every output is traceable to the
exact instrument version that produced it.

---

## The complete pipeline in twelve lines

```{r pipeline, eval = FALSE}
library(surveyframe)

# Build once
instr     <- read_sframe("satisfaction_survey.sframe")

# Load and check
responses <- read_responses("responses.csv", instr, respondent_id = "id")
qr        <- quality_report(responses, instr, respondent_id = "id")

# Score and measure
scored    <- score_scales(responses, instr)
rr        <- reliability_report(responses, instr)
syntax    <- cfa_syntax(instr)

# Report
render_report(instr, data = responses, output_file = "report.html")
```

---

## Citing surveyframe

If you use surveyframe in a publication, please cite the package and include
the instrument hash from your `.sframe` file in your supplementary materials.
This makes your measurement instrument fully reproducible and auditable by
reviewers.

```{r citation, eval = FALSE}
citation("surveyframe")
```
