---
title: "SurveyBuilder GUI overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SurveyBuilder GUI overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(surveyframe)
```

`surveyframe` has three graphical entry points. They are intentionally kept
separate in v0.3.

* `launch_builder()` is a standalone questionnaire builder. It runs in the
  browser and saves `.sframe` files for later use in R.
* `launch_studio()` is the workflow hub. It can build, preview, load
  responses, run quality and reliability checks, plan analyses, and export
  outputs.
* `launch_dashboard()` is a read-only response explorer for instruments and
  response data that are already available.

## Input-types demo

The input-types demo covers the main controls supported by SurveyBuilder and
SurveyStudio.

```{r input-types}
demo <- sframe_input_types_demo_data()
instr <- demo$instrument
responses <- demo$responses

table(vapply(instr$items, function(x) x$type, character(1)))
dim(responses)
```

## Recommended GUI workflow

1. Use `launch_builder()` to create or revise a questionnaire.
2. Save the instrument as a `.sframe` file.
3. Use `read_sframe()` to load the saved instrument in R.
4. Use `read_responses()` to import collected data.
5. Use `launch_studio()` to inspect the instrument and responses together.
6. Use `launch_dashboard()` for read-only response exploration.

## Demo launchers

The demo launchers are not evaluated in this vignette because CRAN examples
and vignettes should not open browsers.

```{r gui-demo, eval = FALSE}
launch_builder_demo()
launch_studio_demo()
launch_dashboard_demo()
```

`launch_builder_demo()` writes a demo `.sframe` file to a temporary folder and
opens the builder. The user can then load the file in the browser builder.
`launch_studio_demo()` opens the full workflow with an instrument and response
data already loaded. `launch_dashboard_demo()` opens the response dashboard
with the same demo data.

## Standalone builder

```{r builder-path, eval = FALSE}
builder_file <- launch_builder(open = FALSE)
builder_file
```

Use the builder when the immediate task is questionnaire authoring. Because
the builder is client-side, it is not the right place for the response
dashboard.

## Studio workflow hub

```{r studio, eval = FALSE}
launch_studio(
  instrument = instr,
  responses = responses,
  screen = "analysis",
  launch.browser = FALSE
)
```

SurveyStudio reads preloaded objects passed by `launch_studio()`. When response
data are present, `screen = "auto"` opens the dashboard screen. The dashboard
button in Studio explains how to open the separate dashboard launcher from the
R console.

## Dashboard

```{r dashboard, eval = FALSE}
launch_dashboard(
  instrument = instr,
  responses = responses,
  launch.browser = FALSE
)
```

Use the dashboard after data collection for read-only exploration. In v0.4,
the dashboard can be refactored into a native SurveyStudio tab after the v0.3
CRAN release is stable.

## Moving files between tools

The `.sframe` file is the shared object between the GUI and R workflow:

1. Build or edit the questionnaire in SurveyBuilder.
2. Save the `.sframe` file.
3. Load it in R with `read_sframe()`.
4. Import responses with `read_responses()`.
5. Use SurveyStudio or the dashboard for response checking and reporting.
