---
title: "A worked study: digital marketing effectiveness in Thailand's tourism"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{A worked study: digital marketing effectiveness in Thailand's tourism}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(surveyframe)
```

## About this vignette

This vignette rebuilds a published study inside surveyframe, from the research
design through to the reported results. The study is Sharafuddin, Madhavan, and
Wangtueai (2024), "Assessing the Effectiveness of Digital Marketing in Enhancing
Tourist Experiences and Satisfaction: A Study of Thailand's Tourism Services",
*Administrative Sciences*, 14(11), 273. It used partial least squares structural
equation modelling on 195 international tourists in coastal Thailand.

surveyframe holds the questionnaire, the analysis plan, and the measurement and
structural models in one instrument, so the design and the analysis stay
together. Everything below stays within the package as released at v0.3.0. No
new features are introduced.

The response data for the published study is not public, so the steps that need
the raw data are shown as the code you would run, followed by the results the
paper reported. The steps that do not need the raw data, which include the
instrument, the models, the generated syntax, the sample-size guidance, the
codebook, and the validity summary computed from the published loadings, run
here as written.

## Research design and objectives

The study took a quantitative approach and used PLS-SEM with a two-stage
treatment of two higher-order constructs. The population was international
tourists above the age of 20 who had visited the Chonburi and Rayong provinces
of Thailand and consumed at least one blue tourism service, such as a ferry or
speed-boat transfer. Convenience sampling was used. The survey was distributed
to 400 tourists and returned 195 complete responses, a response rate near 50
per cent.

The study set three objectives. The first was to develop a conceptual framework
linking digital marketing effectiveness, destination service quality,
destination sustainability quality, tourist satisfaction, and behavioural
intention. The second was to measure digital marketing effectiveness and its
impact on satisfaction and behavioural intention. The third was to analyse the
mediating role of service quality and sustainability quality in the path from
digital marketing effectiveness to behavioural intention.

## Research questions and hypotheses

The study addressed five research questions and tested thirteen hypotheses.

```{r hypotheses}
data.frame(
  hypothesis = paste0("H", 1:13),
  path = c(
    "DME -> DSQ", "DME -> DSUQ", "DME -> TS", "DME -> BI",
    "DSQ -> TS", "DSQ -> BI", "DSUQ -> TS", "DSUQ -> BI", "TS -> BI",
    "DME -> DSQ -> TS", "DME -> DSUQ -> TS",
    "DME -> DSQ -> TS -> BI", "DME -> DSUQ -> TS -> BI"
  ),
  type = c(rep("direct", 9), rep("indirect", 4)),
  stringsAsFactors = FALSE
)
```

Here DME is digital marketing effectiveness, DSQ is destination service
quality, DSUQ is destination sustainability quality, TS is tourist
satisfaction, and BI is behavioural intention.

## Sample-size guidance

The published study set its minimum sample at 155 using the inverse square root
method of Kock and Hadaya (2018), with a smallest path coefficient between 0.11
and 0.20, 80 per cent power, and a 5 per cent significance level, then collected
195. surveyframe's `sample_size_plan()` gives design-dependent guidance for an
SEM study rather than that specific calculation.

```{r sample-size}
sample_size_plan(type = "sem", power = 0.80, alpha = 0.05)
```

For a single bivariate question, the function returns a concrete figure.

```{r sample-size-corr}
sample_size_plan(type = "correlation", r = 0.20, power = 0.80, alpha = 0.05)
```

## The questionnaire

The final questionnaire held 33 rated items across nine lower-order constructs.
Digital marketing effectiveness was measured through relevance and engagement
(DMRE, four items), accessibility and usefulness (DMAU, three items), ease of
use (DMEU, three items), and perceived value (DMPV, four items). Destination
service quality was measured through accommodation (DSQA, three items) and local
transport (DSQT, five items). Destination sustainability quality (DSUQ) used
five items, tourist satisfaction (TS) three, and behavioural intention (BI)
three. All rated items used a Likert agreement scale.

The instrument is built programmatically from the construct structure. Item IDs
follow the construct name, so the link from item to construct stays visible.

```{r build-items}
likert5 <- sf_choices(
  "likert5",
  values = 1:5,
  labels = c("Strongly disagree", "Disagree",
             "Neither agree nor disagree", "Agree", "Strongly agree")
)

construct_labels <- c(
  DMRE = "Digital marketing: relevance and engagement",
  DMAU = "Digital marketing: accessibility and usefulness",
  DMEU = "Digital marketing: ease of use",
  DMPV = "Digital marketing: perceived value",
  DSQA = "Destination service quality: accommodation",
  DSQT = "Destination service quality: local transport",
  DSUQ = "Destination sustainability quality",
  TS   = "Tourist satisfaction",
  BI   = "Behavioural intention"
)
item_counts <- c(DMRE = 4, DMAU = 3, DMEU = 3, DMPV = 4,
                 DSQA = 3, DSQT = 5, DSUQ = 5, TS = 3, BI = 3)

make_items <- function(construct) {
  ids <- paste0(tolower(construct), "_", seq_len(item_counts[[construct]]))
  Map(function(id, k) {
    sf_item(
      id, sprintf("%s (item %d)", construct_labels[[construct]], k),
      type = "likert", required = TRUE, choice_set = "likert5",
      scale_id = construct
    )
  }, ids, seq_along(ids))
}

rated_items <- unlist(lapply(names(item_counts), make_items), recursive = FALSE)

make_scale <- function(construct) {
  ids <- paste0(tolower(construct), "_", seq_len(item_counts[[construct]]))
  sf_scale(construct, construct_labels[[construct]], items = ids, method = "mean")
}
scales <- lapply(names(item_counts), make_scale)
```

The form also collected the respondent profile. The items below follow the live
form and keep its order. The form's nationality question lists individual
countries grouped by region. It is summarised here to a representative set for
brevity.

```{r build-demographics}
gender <- sf_choices("gender", c("male", "female", "transgender"),
                     c("Male", "Female", "Transgender"))
age_band <- sf_choices("age_band",
  c("18_25", "26_40", "41_50", "51_60", "60_plus"),
  c("18-25", "26-40", "41-50", "51-60", "60 and above"))
visitor <- sf_choices("visitor", c("first_time", "repeat"),
                      c("First-time visitor", "Repeat visitor"))
frequency <- sf_choices("frequency",
  c("lt_1_yr", "once_yr", "2_3_yr", "gt_3_yr"),
  c("Less than once a year", "Once a year", "2 to 3 times a year",
    "More than 3 times a year"))
nationality <- sf_choices("nationality",
  c("thai", "china", "japan", "korea", "india", "maldives", "bangladesh",
    "australia", "uk", "germany", "france", "russia", "usa", "brazil",
    "saudi_arabia", "uae", "south_africa", "other"),
  c("Thai", "China", "Japan", "Korea (Republic of)", "India", "Maldives",
    "Bangladesh", "Australia", "United Kingdom", "Germany", "France",
    "Russian Federation", "United States of America", "Brazil",
    "Saudi Arabia", "U.A.E.", "South Africa", "Other"))
education <- sf_choices("education",
  c("high_school", "diploma", "undergraduate", "post_graduate"),
  c("High school", "Diploma", "Undergraduate", "Post graduate"))
profession <- sf_choices("profession",
  c("student", "business", "salaried", "freelancer", "not_working"),
  c("Student", "Business", "Salaried and working", "Freelancer",
    "Not working"))
companion <- sf_choices("companion", c("friends", "family", "other"),
                        c("Friends", "Family members", "Other"))
group_size <- sf_choices("group_size", c("lt_3", "3_5", "gt_5"),
                         c("Less than 3", "3 to 5", "More than 5"))

demo_items <- list(
  sf_item("gender", "Gender", type = "single_choice", required = TRUE, choice_set = "gender"),
  sf_item("age_band", "Age", type = "single_choice", required = TRUE, choice_set = "age_band"),
  sf_item("visitor_type", "I am a", type = "single_choice", required = TRUE, choice_set = "visitor"),
  sf_item("frequency_of_visit", "Frequency of visit", type = "single_choice", required = TRUE, choice_set = "frequency"),
  sf_item("nationality", "Nationality", type = "single_choice", required = TRUE, choice_set = "nationality"),
  sf_item("education", "Education level", type = "single_choice", required = TRUE, choice_set = "education"),
  sf_item("profession", "Profession", type = "single_choice", required = TRUE, choice_set = "profession"),
  sf_item("travel_companion", "I visit with my", type = "single_choice", required = TRUE, choice_set = "companion"),
  sf_item("group_size", "My travel group size is", type = "single_choice", required = TRUE, choice_set = "group_size")
)
```

The instrument is assembled from the choice sets, the rated items, the
demographic items, and the scales.

```{r assemble}
choice_sets <- list(likert5, gender, age_band, visitor, frequency, nationality,
                    education, profession, companion, group_size)

study <- sf_instrument(
  title       = "Digital Marketing Effectiveness of Tourism Services in Thailand",
  version     = "1.0.0",
  description = "Reconstruction of the published PLS-SEM study for surveyframe.",
  authors     = "Sharafuddin, Madhavan, and Wangtueai",
  languages   = "en",
  components  = c(choice_sets, rated_items, demo_items, scales)
)

study
```

## Validate the instrument

```{r validate}
validation <- validate_sframe(study, strict = FALSE)
validation$valid
length(validation$problems)
```

## The analysis plan

The plan records each research step as a block that binds a technique to the
variables that fill its roles. The plan below covers the screening, the
measurement assessment, the descriptive comparisons that the demographic items
make possible, and the model syntax. The role names are specific to each
technique.

```{r plan}
study$analysis_plan <- list(
  list(id = "D1", research_question = "What is the sample profile?",
       family = "descriptive", method = "descriptives",
       roles = list(variables = names(item_counts))),
  list(id = "D2", research_question = "How complete are the responses?",
       family = "data_quality", method = "missing_data",
       roles = list(variables = names(item_counts))),
  list(id = "M1", research_question = "Is each construct internally consistent?",
       family = "measurement", method = "reliability_alpha",
       roles = list(items = paste0("dmpv_", 1:4))),
  list(id = "M2", research_question = "Do the digital marketing items form coherent factors?",
       family = "measurement", method = "efa_readiness",
       roles = list(items = c(paste0("dmre_", 1:4), paste0("dmau_", 1:3),
                              paste0("dmeu_", 1:3), paste0("dmpv_", 1:4)))),
  list(id = "RQ_assoc", research_question = "Is service quality associated with satisfaction?",
       family = "association", method = "correlation_pearson",
       roles = list(x = "DSQA", y = "TS"), options = list(alpha = 0.05)),
  list(id = "RQ_reg", research_question = "Do service and sustainability quality predict satisfaction?",
       family = "regression", method = "regression_linear",
       roles = list(predictors = c("DSQA", "DSQT", "DSUQ"), dependent = "TS"),
       options = list(alpha = 0.05)),
  list(id = "RQ_group", research_question = "Do first-time and repeat visitors differ in satisfaction?",
       family = "group_comparison", method = "mann_whitney",
       roles = list(group = "visitor_type", outcome = "TS"),
       options = list(alpha = 0.05)),
  list(id = "RQ_cross", research_question = "Is visitor type associated with gender?",
       family = "categorical", method = "chi_square",
       roles = list(row = "gender", column = "visitor_type"),
       options = list(alpha = 0.05))
)

length(study$analysis_plan)
```

## Descriptives and data screening

With the 195-row dataset in hand, the screening blocks run as below. The paper
reported a sample of 195 tourists aged 18 to 65, 56 per cent female and 44 per
cent male, with 48 per cent holding an undergraduate degree and 37 per cent a
postgraduate degree. Working professionals made up 66 per cent of the sample.

```{r screening, eval = FALSE}
responses <- read_responses(
  "thailand_tourism_responses.csv",
  study,
  respondent_id = "respondent_id",
  submitted_at  = "submitted_at"
)

descriptives_report(responses, variables = names(item_counts))
missing_data_report(responses, study)
quality_report(responses, study, respondent_id = "respondent_id")
```

## Reliability and item diagnostics

The reliability and item reports read the scale definitions from the
instrument. The published lower-order constructs all exceeded the 0.7 threshold
for composite reliability. At the higher-order level the study reported Cronbach
alpha and composite reliability of 0.828 and 0.883 for DME, 0.792 and 0.906 for
DSQ, 0.769 and 0.843 for DSUQ, 0.897 and 0.936 for TS, and 0.921 and 0.950 for
BI.

```{r reliability, eval = FALSE}
reliability_report(responses, study, omega = TRUE)
item_report(responses, study)
```

## Validity from the published loadings

This step does not need the raw data. `validity_report()` computes composite
reliability and average variance extracted from standardised loadings. Feeding
the loadings reported in the paper reproduces the published reliability and
convergent validity, which is a direct check of the measurement model.

```{r validity}
published_loadings <- list(
  DMRE = c(dmre_1 = .815, dmre_2 = .899, dmre_3 = .668, dmre_4 = .838),
  DMAU = c(dmau_1 = .843, dmau_2 = .915, dmau_3 = .920),
  DMEU = c(dmeu_1 = .897, dmeu_2 = .929, dmeu_3 = .932),
  DMPV = c(dmpv_1 = .818, dmpv_2 = .916, dmpv_3 = .900, dmpv_4 = .863),
  DSQA = c(dsqa_1 = .904, dsqa_2 = .920, dsqa_3 = .869),
  DSQT = c(dsqt_1 = .778, dsqt_2 = .883, dsqt_3 = .879, dsqt_4 = .811, dsqt_5 = .713),
  DSUQ = c(dsuq_1 = .780, dsuq_2 = .855, dsuq_3 = .845, dsuq_4 = .551, dsuq_5 = .529),
  TS   = c(ts_1 = .912, ts_2 = .950, ts_3 = .869),
  BI   = c(bi_1 = .937, bi_2 = .911, bi_3 = .940)
)

validity <- validity_report(published_loadings)
validity$reliability
```

The composite reliability and AVE returned here match the values in the paper.
The study also reported Fornell-Larcker results, where the square root of the
AVE for each construct exceeded its correlations with the others, and HTMT
ratios that were all below 0.90, the highest being 0.766 between DMEU and DMAU.
Both confirm discriminant validity.

## Exploratory factor analysis

EFA is an exploratory step that screens whether the items group as intended
before the confirmatory model. `efa_report()` reports the Kaiser-Meyer-Olkin
measure and Bartlett's test, and `efa_solution()` returns loadings,
communalities, and cross-loading flags. Both use the optional `psych` package.

```{r efa, eval = FALSE}
efa_report(responses, study)
efa_solution(responses, study,
             items = c(paste0("dmre_", 1:4), paste0("dmau_", 1:3),
                       paste0("dmeu_", 1:3), paste0("dmpv_", 1:4)),
             nfactors = 4)
```

## The measurement model: CFA syntax

The confirmatory measurement model has the nine lower-order constructs as
reflective factors. `cfa_lavaan_syntax()` writes the lavaan syntax without
needing lavaan installed.

```{r cfa-model}
lower_order <- lapply(names(item_counts), function(construct) {
  ids <- paste0(tolower(construct), "_", seq_len(item_counts[[construct]]))
  sf_construct(construct, construct_labels[[construct]], items = ids, mode = "reflective")
})

cfa_model <- sf_model(
  id    = "dme_measurement",
  label = "Lower-order measurement model",
  type  = "cfa",
  constructs = lower_order
)

study <- add_model(study, cfa_model)
cat(cfa_lavaan_syntax(study, model = cfa_model))
```

## The structural model: CB-SEM syntax

The structural model carries the nine direct paths and the four indirect
effects from the hypotheses. This rendering treats the five constructs as
first-order reflective factors over their full item sets, which is the
covariance-based reading of the framework. `sem_lavaan_syntax()` writes the
measurement, structural, and indirect-effect lines.

```{r cbsem-model}
dm_items  <- c(paste0("dmre_", 1:4), paste0("dmau_", 1:3),
               paste0("dmeu_", 1:3), paste0("dmpv_", 1:4))
dsq_items <- c(paste0("dsqa_", 1:3), paste0("dsqt_", 1:5))

sem_model <- sf_model(
  id    = "tourism_structural",
  label = "Digital marketing structural model",
  type  = "cb_sem",
  constructs = list(
    sf_construct("DME",  "Digital marketing effectiveness", dm_items),
    sf_construct("DSQ",  "Destination service quality",     dsq_items),
    sf_construct("DSUQ", "Destination sustainability quality", paste0("dsuq_", 1:5)),
    sf_construct("TS",   "Tourist satisfaction",            paste0("ts_", 1:3)),
    sf_construct("BI",   "Behavioural intention",           paste0("bi_", 1:3))
  ),
  paths = list(
    sf_path("DME", "DSQ",  label = "h1"),
    sf_path("DME", "DSUQ", label = "h2"),
    sf_path("DME", "TS",   label = "h3"),
    sf_path("DME", "BI",   label = "h4"),
    sf_path("DSQ", "TS",   label = "h5"),
    sf_path("DSQ", "BI",   label = "h6"),
    sf_path("DSUQ", "TS",  label = "h7"),
    sf_path("DSUQ", "BI",  label = "h8"),
    sf_path("TS", "BI",    label = "h9")
  ),
  indirect = list(
    sf_indirect("DME", "DSQ", "TS",  label = "h10"),
    sf_indirect("DME", "DSUQ", "TS", label = "h11"),
    sf_indirect("DME", c("DSQ", "TS"), "BI",  label = "h12"),
    sf_indirect("DME", c("DSUQ", "TS"), "BI", label = "h13")
  ),
  options = list(estimator = "MLR", missing = "fiml", standardised = TRUE)
)

validate_model(sem_model, study)
cat(sem_lavaan_syntax(sem_model, study))
```

## The two-stage PLS-SEM: seminr syntax

The published study used PLS-SEM with a two-stage treatment of the higher-order
constructs. The first stage assesses the lower-order measurement model, which
is the CFA model above and the reliability and validity reports. The second
stage models digital marketing effectiveness and service quality as composites
of their lower-order construct scores, then estimates the structural paths.
`seminr_syntax()` writes the seminr code.

```{r pls-model}
pls_model <- sf_model(
  id    = "tourism_pls",
  label = "Two-stage higher-order PLS model",
  type  = "pls_sem",
  constructs = list(
    sf_construct("DME",  "Digital marketing effectiveness",
                 c("DMRE", "DMAU", "DMEU", "DMPV"), mode = "composite"),
    sf_construct("DSQ",  "Destination service quality",
                 c("DSQA", "DSQT"), mode = "composite"),
    sf_construct("DSUQ", "Destination sustainability quality",
                 paste0("dsuq_", 1:5), mode = "composite"),
    sf_construct("TS",   "Tourist satisfaction", paste0("ts_", 1:3), mode = "composite"),
    sf_construct("BI",   "Behavioural intention", paste0("bi_", 1:3), mode = "composite")
  ),
  paths = list(
    sf_path("DME", "DSQ"), sf_path("DME", "DSUQ"),
    sf_path("DME", "TS"),  sf_path("DME", "BI"),
    sf_path("DSQ", "TS"),  sf_path("DSQ", "BI"),
    sf_path("DSUQ", "TS"), sf_path("DSUQ", "BI"),
    sf_path("TS", "BI")
  ),
  options = list(bootstrap = 1000)
)

cat(seminr_syntax(pls_model))
```

When this model is fitted in seminr, the paper reported that digital marketing
effectiveness influenced service quality (0.573) and sustainability quality
(0.394), that both mediated the path to satisfaction (indirect effects of 0.155
and 0.152), and that satisfaction drove behavioural intention (0.713). The
direct paths from digital marketing effectiveness to satisfaction and
behavioural intention were not significant, which is why the mediating
constructs matter.

## Run the plan and render the report

With the dataset loaded, the plan runs in one pass and the report is rendered.
`run_analysis_plan()` returns one result per block, and `render_results()` or
`render_report()` writes the report.

```{r run, eval = FALSE}
results <- run_analysis_plan(responses, study)
results

render_results(results, study, output_file = "thailand-results.html")

render_report(
  study, data = responses, output_file = "thailand-report.html",
  include_codebook = TRUE, include_quality = TRUE, include_missing = TRUE,
  include_descriptives = TRUE, include_reliability = TRUE,
  include_analysis = TRUE, include_models = TRUE
)
```

## The visual route

The same study can be assembled and run through the two browser tools, which
have separate jobs.

SurveyBuilder, opened with `launch_builder()`, is a client-side authoring tool.
Its Build mode edits the questionnaire, its Analyse mode writes the analysis
plan and builds the model, and its model panel exports lavaan or seminr syntax
and the model JSON into the `.sframe` file. The builder saves the instrument and
configures a Google Sheet endpoint, but it computes no statistics.

SurveyStudio, opened with `launch_studio()`, is the R-backed workflow. Its tabs
run from Build Survey, Open Instrument, and Preview Survey through Upload
Responses, Quality Dashboard, Reliability, and Analysis Plan, to Dashboard and
Export. Once responses are uploaded, the Analysis Plan screen runs the saved
plan and shows a table of each question with its method and APA result. The
Export screen renders the full report.

```{r gui, eval = FALSE}
launch_builder()     # author the questionnaire, plan, and model
launch_studio()      # upload responses, run the plan, export the report
launch_dashboard()   # a read-only response view
```

## Reproducibility and citation

Keep the `.sframe` file with the analysis files. It records the questionnaire,
the analysis plan, and the three models used here, so the design and the
analysis travel together.

```{r citation, eval = FALSE}
citation("surveyframe")
```
