---
title: "0.4.1 review suite: start here"
subtitle: "Read this first, then 01 and 02"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

## What this suite is for

0.4.1 is a bug-fix release. Its scope is 4 defects found in the field plus
the testing work that followed, so this suite is not a repeat of
`review_040`'s 20-file sweep of every statistical method. It has 2 jobs the
earlier suites did not do.

**File 01 is a coverage ledger.** It enumerates the package's export list at
render time and calls every function in it, so "all functions are tested" is
a property that can be checked rather than a claim. A function added
tomorrow shows up as an uncovered row on the next render.

**File 02 is the Quarto audit.** The rendered report is the artefact a study
is judged on. This file checks that the engine which produced it is the one
the researcher thinks produced it, that the same inputs give the same
report, and that the chain from instrument to hash to artefact holds.

```{r}
#| label: files
files <- data.frame(
  file = c("01_function_coverage.qmd", "02_quarto_reproducibility.qmd"),
  covers = c("all 123 exports, smoke-called, with the gaps named",
             "render_report()'s 2 engines, reproducibility, the audit chain"),
  read_it = c("to see what is and is not exercised",
              "before trusting any rendered report as evidence"),
  stringsAsFactors = FALSE
)
knitr::kable(files, caption = "The suite")
```

## Read file 02 first if you are short of time

Its finding is the one that affects every study run with this package: the
report is **not reproducible**. `run_analysis_plan()` gives different answers
on the same inputs, because 5 bootstrap confidence-interval helpers and the
EFA parallel analysis run unseeded. The published interval moves between
renders. File 02 section 3.1 has the detail and the owner decision it needs.

## What this suite inherits, and what is new

`review_040/HANDOVER.md` documents the file template, the comparison helpers,
the token-saving rules, and 16 reference-call traps. All of it applies here
unchanged and is not repeated. `_setup.R` in this directory is
`review_040/_setup.R` plus 2 new sections.

```{r}
#| label: whats-new
newbits <- data.frame(
  helper = c("sf_exports()", "sf_smoke()", "sf_coverage_ledger()",
             "sf_coverage_verdict()", "sf_report_engine()",
             "sf_render_both_engines()", "sf_render_twice()"),
  purpose = c("the export list, minus S3 methods",
              "call one export and record what happened",
              "every export against what the suite touched",
              "the one-line summary and the uncovered rows",
              "which engine produced a rendered report, read off the artefact",
              "the same report down each engine, side by side",
              "the same inputs twice, for a reproducibility check"),
  stringsAsFactors = FALSE
)
knitr::kable(newbits, caption = "New in this suite's _setup.R")
```

## The build under review

```{r}
#| label: build
b <- data.frame(
  check = c("package version", "0.4.1 feature present",
            "quarto available", "quarto version"),
  value = c(as.character(utils::packageVersion("surveyframe")),
            as.character(sf_is_041()),
            as.character(sf_quarto_available()),
            sf_quarto_version()),
  stringsAsFactors = FALSE
)
knitr::kable(b, caption = "Build")
```

`sf_is_041()` tests for a feature rather than a version string, because
DESCRIPTION reads `0.4.0.9000` on `dev` and `0.4.0` on `main`, so the number
cannot tell the development build from the released one.

## How to run it

```r
# from the repository root
quarto render review_041/01_function_coverage.qmd
quarto render review_041/02_quarto_reproducibility.qmd
```

Each file loads the working tree with `pkgload::load_all()` if an installed
0.4.1 is not on the library path, so no install step is needed.

Rendered `.html` output is gitignored, as in the other suites. Only the
`.qmd` sources, `_setup.R`, and `HANDOVER.md` are tracked.

## What to do with what you find

Log it in `dogfeed.todo.md` rather than leaving it in a review file, so it
survives into a release lane. Findings from this suite's first run are
already logged there.

## What is not here yet

This suite covers breadth and the report. It does not repeat `review_040`'s
per-method accuracy files, which remain valid for the statistics, nor
`review_050`'s text-analysis files. `HANDOVER.md` lists the files that would
complete it and what each would cover.
