---
title: "01. Function coverage"
subtitle: "Every exported function, called at least once, with the gaps named"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

"Comprehensive" has to be a property that can be checked, or it is a claim.
This file enumerates the package's own export list **at render time** and
calls every function in it, so a function added tomorrow appears as an
uncovered row on the next render rather than quietly going untested.

This is a smoke pass, not an accuracy pass. It asks whether each function
can be called and what it returns. Whether the answer is *right* is what
the per-family files are for, here and in `review_040`. Both matter, and
they fail differently: an accuracy file cannot find a defect in a function
nobody ever calls, and several of this package's defects lived in exactly
such paths.

## 1. The export list

```{r}
#| label: exports
ex <- sf_exports()
cat(sprintf("%d exported objects, excluding S3 methods, which are exercised\nthrough their generics rather than by name.\n", length(ex)))
```

```{r}
#| label: export-families
family <- function(f) {
  if (grepl("^sf_(item|choices|scale|branch|check|instrument|model|construct|path|indirect|covariance|conjoint)", f)) "construction"
  else if (grepl("^(sf_(meta|items|scales|choice_sets|branches|checks|models|plan|id|label|apa|flagged|is_valid|problems|object)|as_sframe)", f)) "accessors"
  else if (grepl("^(read_|write_|export_|link_git|amend|amendment)", f)) "io and provenance"
  else if (grepl("^(validate_|sframe_check)", f)) "validation"
  else if (grepl("^(run_analysis|score_|assumption_|posthoc_|efa_|cfa_|sem_|seminr_|item_report|reliability_|validity_|quality_|codebook_|missing)", f)) "analysis"
  else if (grepl("^(bootstrap_ci|cohens_d_ci|eta_sq_ci|cramers_v_ci|sample_size)", f)) "inference helpers"
  else if (grepl("(topsis|ahp|anp|dematel|vikor|moora|smart|waspas|promethee|electre|decision|sensitivity)", f, ignore.case = TRUE)) "decision analysis"
  else if (grepl("(term_|ngram|cooccur|co_occurrence|sentiment|topic|quote|clean_text)", f)) "text analysis"
  else if (grepl("^(sframe_plot|sframe_draw|theme_surveyframe|sframe_likert_scale_groups)", f)) "plotting"
  else if (grepl("^(render_|launch_|survey_module|studio)", f)) "reporting and apps"
  else "other"
}
fam <- vapply(ex, family, character(1))
knitr::kable(as.data.frame(table(family = fam), responseName = "functions"),
             caption = "Exports by family")
```

## 2. Fixtures the smoke pass calls against

One small instrument with a plan, one bundled demo, one decision instrument.
Between them they satisfy the argument requirements of nearly every export.

```{r}
#| label: fixtures
set.seed(41)

inst <- sf_instrument(
  title = "Coverage", version = "1.0.0",
  components = list(
    sf_choices("ag5", 1:5, c("Strongly disagree", "Disagree", "Neutral",
                             "Agree", "Strongly agree")),
    sf_choices("grp", c("a", "b"), c("Group A", "Group B")),
    sf_item("g",  "Group",  type = "single_choice", choice_set = "grp"),
    sf_item("q1", "Item 1", type = "likert", choice_set = "ag5", scale_id = "S"),
    sf_item("q2", "Item 2", type = "likert", choice_set = "ag5", scale_id = "S"),
    sf_item("q3", "Item 3", type = "likert", choice_set = "ag5", scale_id = "S"),
    sf_item("age", "Age", type = "numeric"),
    sf_item("open", "Anything else?", type = "textarea"),
    sf_scale("S", "Satisfaction", items = c("q1", "q2", "q3"))
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question = "What are the distributions?",
         family = "descriptive", method = "frequencies",
         roles = list(variables = "q1")),
    list(id = "RQ2", research_question = "Do the groups differ?",
         family = "group_comparison", method = "t_test_ind",
         roles = list(outcome = "age", group = "g"))
  )
)

n <- 80
dat <- data.frame(
  respondent_id = sprintf("r%02d", seq_len(n)),
  g   = sample(c("a", "b"), n, TRUE),
  q1  = sample(1:5, n, TRUE), q2 = sample(1:5, n, TRUE), q3 = sample(1:5, n, TRUE),
  age = round(rnorm(n, 40, 10)),
  open = sample(c("the service was good", "slow but polite",
                  "very helpful staff", "too expensive"), n, TRUE),
  stringsAsFactors = FALSE
)
resp <- read_responses(dat, inst, respondent_id = "respondent_id")
demo <- sframe_demo_data()

fx <- data.frame(
  fixture = c("inst", "resp", "demo$instrument", "demo$responses"),
  what = c("3-item scale, 1 group, 1 numeric, 1 open text, 2-block plan",
           paste(nrow(resp), "simulated responses"),
           "bundled tourism instrument", "bundled tourism responses"),
  stringsAsFactors = FALSE
)
knitr::kable(fx, caption = "Fixtures")
```

## 3. The smoke pass

Each call is recorded with `sf_smoke()`. `ok` means it returned a value.
`ERROR` means it raised where a value was expected, and is a finding.
`skipped` means calling it needs a browser, a network, or a human, and the
reason is recorded rather than the row being omitted.

```{r}
#| label: smoke-construction
sf_smoke("sf_instrument", inst)
sf_smoke("sf_item", sf_item("x", "X", type = "text"))
sf_smoke("sf_choices", sf_choices("c1", 1:2, c("a", "b")))
sf_smoke("sf_scale", sf_scale("S2", "S2", items = c("q1", "q2")))
sf_smoke("sf_branch", sf_branch("q2", "q1", "%in%", c(4, 5)))
sf_smoke("sf_check", sf_check("chk1", item_id = "q1", type = "attention",
                              pass_values = 3))
sf_smoke("sf_model", sf_model(id = "m", type = "cfa",
  constructs = list(sf_construct("S", items = c("q1", "q2", "q3")))))
sf_smoke("sf_construct", sf_construct("S", items = c("q1", "q2")))
sf_smoke("sf_path", sf_path("A", "B"))
sf_smoke("sf_indirect", sf_indirect("A", "B", "C"))
sf_smoke("sf_covariance", sf_covariance("A", "B"))
```

```{r}
#| label: smoke-accessors
sf_smoke("sf_meta", sf_meta(inst))
sf_smoke("sf_items", sf_items(inst))
sf_smoke("sf_scales", sf_scales(inst))
sf_smoke("sf_choice_sets", sf_choice_sets(inst))
sf_smoke("sf_branches", sf_branches(inst))
sf_smoke("sf_checks", sf_checks(inst))
sf_smoke("sf_models", sf_models(inst))
sf_smoke("sf_plan", sf_plan(inst))
sf_smoke("sf_plan<-", { i2 <- inst; sf_plan(i2) <- sf_plan(inst); sf_plan(i2) })
sf_smoke("sf_id", sf_id(sf_items(inst)[[1]]))
sf_smoke("sf_label", sf_label(sf_items(inst)[[1]]))
sf_smoke("as_sframe", as_sframe(validate_sframe(inst, strict = FALSE)))
sf_smoke("sf_is_valid", sf_is_valid(validate_sframe(inst, strict = FALSE)))
sf_smoke("sf_problems", sf_problems(validate_sframe(inst, strict = FALSE)))
sf_smoke("sf_object", sf_object(validate_sframe(inst, strict = FALSE)))
```

```{r}
#| label: smoke-io
p_sf <- tempfile(fileext = ".sframe")
sf_smoke("write_sframe", write_sframe(inst, p_sf, overwrite = TRUE))
sf_smoke("read_sframe", read_sframe(p_sf))
sf_smoke("validate_sframe", validate_sframe(inst, strict = FALSE))
sf_smoke("read_responses", read_responses(dat, inst, respondent_id = "respondent_id"))
sf_smoke("export_static_survey",
         export_static_survey(inst, tempfile(fileext = ".html"),
                              open = FALSE, overwrite = TRUE))
sf_smoke("export_google_sheet",
         export_google_sheet(inst, sheet_url = "https://example.com/x",
                             output_dir = tempdir()))
sf_smoke("read_sheet_responses", NULL, expect = "skip",
         note = "needs a live Google Sheet and network")
sf_smoke("amend_sframe", {
  rev <- inst; rev$meta$title <- "Coverage v2"
  # instrument_revision is a design-tier reason, and the package rightly
  # refuses one without a deviation report. That refusal is the behaviour
  # worth having, so the smoke call satisfies it rather than working round it.
  amend_sframe(inst, rev, reason_code = "instrument_revision",
               reason_text = "smoke test", author = "reviewer",
               deviation_report = "Title reworded during coverage smoke test.")
})
```

```{r}
#| label: smoke-analysis
sf_smoke("run_analysis_plan", run_analysis_plan(resp, inst))
sf_smoke("score_scales", score_scales(resp, inst))
sf_smoke("reliability_report", reliability_report(resp, inst))
sf_smoke("item_report", item_report(resp, inst))
sf_smoke("quality_report", quality_report(resp, inst))
sf_smoke("codebook_report", codebook_report(inst))
sf_smoke("missing_report", missing_report(resp, inst))
sf_smoke("assumption_report", assumption_report(resp, inst))
sf_smoke("descriptives_report", descriptives_report(resp, variables = c("age", "q1")))
sf_smoke("cfa_syntax", cfa_syntax(inst))
sf_smoke("sample_size_plan", sample_size_plan("mean", margin_error = 0.5, sd = 1))
```

```{r}
#| label: smoke-inference
sf_smoke("bootstrap_ci", bootstrap_ci(resp$age, seed = 1))
sf_smoke("cohens_d_ci", cohens_d_ci(resp$age[resp$g == "a"], resp$age[resp$g == "b"]))
sf_smoke("eta_sq_ci", eta_sq_ci(resp$age, resp$g))
sf_smoke("cramers_v_ci", cramers_v_ci(table(resp$g, resp$q1)))
```

```{r}
#| label: smoke-text
txt <- resp$open[!is.na(resp$open)]
sf_smoke("clean_text_responses", clean_text_responses(data.frame(open = txt), "open"))
sf_smoke("term_frequency", term_frequency(txt))
sf_smoke("ngram_frequency", ngram_frequency(txt, n = 2))
sf_smoke("extract_quotes", NULL, expect = "skip",
         note = "needs a fitted stm topic model")
```

```{r}
#| label: smoke-plots
# Plot functions are the least-covered family in the package: 5 of them were
# called by no test and no review file before this suite existed.
rel <- reliability_report(resp, inst)
sf_smoke("theme_surveyframe", theme_surveyframe())
sf_smoke("sframe_plot_descriptives",
         sframe_plot_descriptives(descriptives_report(resp, variables = "age"), resp))
sf_smoke("sframe_plot_missingness",
         sframe_plot_missingness(missing_data_report(resp, inst)))
sf_smoke("sframe_plot_quality", sframe_plot_quality(quality_report(resp, inst)))
sf_smoke("sframe_plot_reliability", sframe_plot_reliability(rel))
sf_smoke("sframe_plot_term_frequency", NULL, expect = "skip",
         note = "takes an analysis-plan result, not term_frequency() output, see 4.1")
sf_smoke("sframe_plot_ngram_frequency", NULL, expect = "skip",
         note = "takes an analysis-plan result, not ngram_frequency() output, see 4.1")
sf_smoke("sframe_likert_scale_groups", sframe_likert_scale_groups(inst))
```

```{r}
#| label: smoke-reporting
sf_smoke("render_report",
         render_report(inst, resp, output_path = tempfile(fileext = ".html")))
sf_smoke("launch_builder", NULL, expect = "skip", note = "opens a browser")
sf_smoke("launch_studio", NULL, expect = "skip", note = "starts a Shiny app")
sf_smoke("launch_dashboard", NULL, expect = "skip", note = "starts a Shiny app")
sf_smoke("launch_builder_demo", NULL, expect = "skip", note = "opens a browser")
sf_smoke("launch_dashboard_demo", NULL, expect = "skip", note = "starts a Shiny app")
sf_smoke("survey_module_ui", NULL, expect = "skip", note = "Shiny module, needs a session")
sf_smoke("survey_module_server", NULL, expect = "skip", note = "Shiny module, needs a session")
sf_smoke("sframe_demo_data", sframe_demo_data())
sf_smoke("sframe_input_types_demo_data", sframe_input_types_demo_data())
```

```{r}
#| label: smoke-decision
hotel <- read_sframe(sf_fixture("hotel_supplier_mcdm.sframe"))
sf_smoke("sframe_decision_options", sframe_decision_options(hotel))
sf_smoke("sf_conjoint_design",
         sf_conjoint_design("cj1",
                            attributes = list(brand = c("A", "B"),
                                              price = c("1", "2")),
                            method = "full"))
```

```{r}
#| label: smoke-models
mdl <- sf_model(id = "m", type = "cfa", engine = "lavaan",
  constructs = list(sf_construct("S", items = c("q1", "q2", "q3"))))
# Both engines require at least 1 structural path, which validate_model()
# says plainly rather than generating empty syntax. A 2-construct model is
# the smallest thing that satisfies it.
two_con <- list(sf_construct("S", items = c("q1", "q2")),
                sf_construct("T", items = c("q3", "age")))
sem_mdl <- sf_model(id = "sm", type = "cb_sem", engine = "lavaan",
  constructs = two_con, paths = list(sf_path("S", "T")))
pls_mdl <- sf_model(id = "pm", type = "pls_sem", engine = "seminr",
  constructs = list(sf_construct("S", items = c("q1", "q2"), mode = "composite"),
                    sf_construct("T", items = c("q3", "age"), mode = "composite")),
  paths = list(sf_path("S", "T")))
sf_smoke("validate_model", validate_model(mdl, instrument = inst, strict = FALSE))
sf_smoke("add_model", add_model(inst, mdl))
sf_smoke("model_json", model_json(mdl))
sf_smoke("model_report_template", model_report_template(mdl))
sf_smoke("cfa_lavaan_syntax", cfa_lavaan_syntax(inst, mdl))
sf_smoke("sem_lavaan_syntax", sem_lavaan_syntax(sem_mdl, inst))
sf_smoke("seminr_syntax", seminr_syntax(pls_mdl))
sf_smoke("efa_syntax", efa_syntax(items = c("q1", "q2", "q3"), nfactors = 1))
```

```{r}
#| label: smoke-reports-2
sf_smoke("missing_data_report", missing_data_report(resp, inst))
sf_smoke("outlier_report", outlier_report(resp, variables = "age"))
sf_smoke("posthoc_report",
         posthoc_report(resp, method = "anova", outcome = "age", group = "g"))
sf_smoke("amendment_log", amendment_log(inst))
sf_smoke("link_git_commit", link_git_commit(inst, repo_path = "."))
sf_smoke("term_context", term_context(txt, term = "good"))
sf_smoke("sf_apa", sf_apa(run_analysis_plan(resp, inst)))
sf_smoke("sf_flagged", sf_flagged(quality_report(resp, inst)))
sf_smoke("sframe_codebook_items_display",
         sframe_codebook_items_display(codebook_report(inst)))
```

```{r}
#| label: smoke-efa
# The demo instrument has enough items for a factor solution. The coverage
# fixture's 3-item scale does not.
efa_ok <- tryCatch({
  ef <- efa_solution(demo$responses, demo$instrument, nfactors = 2)
  TRUE
}, error = function(e) FALSE)
if (efa_ok) {
  sf_smoke("efa_solution", efa_solution(demo$responses, demo$instrument, nfactors = 2))
  sf_smoke("efa_report", efa_report(demo$responses, demo$instrument, nfactors = 2))
  er <- efa_report(demo$responses, demo$instrument, nfactors = 2)
  sf_smoke("sframe_plot_efa_loadings", sframe_plot_efa_loadings(ef))
  sf_smoke("sframe_plot_efa_scree", sframe_plot_efa_scree(er))
} else {
  for (f in c("efa_solution", "efa_report", "sframe_plot_efa_loadings",
              "sframe_plot_efa_scree")) {
    sf_smoke(f, NULL, expect = "skip", note = "no EFA solution on this fixture")
  }
}
```

```{r}
#| label: smoke-builder-state
sf_smoke("sframe_builder_empty_state", sframe_builder_empty_state())
sf_smoke("sframe_builder_state_from_instrument",
         sframe_builder_state_from_instrument(inst))
sf_smoke("sframe_builder_validate_draft", {
  st <- sframe_builder_state_from_instrument(inst)
  sframe_builder_validate_draft(st$meta, st$choices, st$items, st$scales,
                                st$branching, st$checks, st$analysis_plan,
                                st$models, st$render)
})
```

```{r}
#| label: smoke-decision-2
sf_smoke("sframe_assemble_pairwise", NULL, expect = "skip",
         note = "needs collected pairwise judgements, covered in review_040 file 13")
sf_smoke("sframe_aggregate_judgements", NULL, expect = "skip",
         note = "needs collected pairwise judgements, covered in review_040 file 13")
sf_smoke("sframe_collected_weights", NULL, expect = "skip",
         note = "needs a criteria_weight item with data, covered in review_040 file 13")
sf_smoke("sframe_rated_matrix", NULL, expect = "skip",
         note = "needs a rated performance matrix, covered in review_040 file 13")
sf_smoke("sensitivity_analysis", NULL, expect = "skip",
         note = "needs an MCDA result, covered in review_040 file 15")
```

```{r}
#| label: smoke-plots-2
# The remaining plot builders. Several take a result object rather than data,
# so each is fed the result its own family produces.
res <- run_analysis_plan(resp, inst)
cs  <- sf_choice_sets(inst)[["ag5"]]
sf_smoke("sframe_plot_correlation_matrix",
         sframe_plot_correlation_matrix(resp, vars = c("q1", "q2", "q3")))
sf_smoke("sframe_plot_variable_distribution",
         sframe_plot_variable_distribution(resp, "age"))
sf_smoke("sframe_plot_scale_chart",
         sframe_plot_scale_chart(score_scales(resp, inst)[["S"]], "Satisfaction"))
sf_smoke("sframe_plot_item_chart",
         sframe_plot_item_chart(sf_items(inst)[["q1"]], resp$q1, cs))
sf_smoke("sframe_plot_likert_scale", NULL, expect = "skip",
         note = "needs the choice-set object shape used inside render_report()")
sf_smoke("sframe_draw_likert_diverging",
         sframe_draw_likert_diverging(table(resp$q1)))
sf_smoke("sframe_plot_validity", NULL, expect = "skip",
         note = "needs a validity_report(), which needs a fitted factor solution")
sf_smoke("validity_report", NULL, expect = "skip",
         note = "needs loadings and construct scores from a fitted model")
for (f in c("sframe_plot_group_comparison", "sframe_plot_paired_comparison",
            "sframe_plot_regression_diagnostics", "sframe_plot_decision_ranking",
            "sframe_plot_dematel_influence", "sframe_plot_cooccurrence",
            "sframe_plot_cooccurrence_network", "sframe_plot_sentiment",
            "sframe_plot_topics", "sframe_plot_likert_matrix",
            "sframe_draw_mosaic")) {
  sf_smoke(f, NULL, expect = "skip",
           note = "takes a result object from its own family, covered where that family is")
}
```

```{r}
#| label: smoke-render
sf_smoke("render_survey", NULL, expect = "skip",
         note = "mode is \"shiny\" only, the static route is export_static_survey()")
sf_smoke("render_results",
         render_results(res, inst, output_path = tempfile(fileext = ".html")))
sf_smoke("launch_studio_demo", NULL, expect = "skip", note = "starts a Shiny app")
```

## 4. The ledger

Every export, and what this suite did with it. Rows reading `NOT EXERCISED`
are the honest gaps.

```{r}
#| label: ledger
ledger <- sf_coverage_ledger(
  also_called_in = list.files(".", pattern = "\\.qmd$", full.names = TRUE))
knitr::kable(ledger, caption = "Coverage ledger, all exports")
```

```{r}
#| label: ledger-verdict
uncovered <- sf_coverage_verdict(ledger)
```

### 4.1 What the ledger found on its first run, and how it was resolved

Writing this file was itself a test of the argument for it, and it found 2
things on functions **no test and no review file had ever called**.

**A regression in 0.4.1, from a fix in 0.4.1.**
`sframe_plot_quality()` returns `NULL` on the bundled demo.

```{r}
#| label: quality-plot-regression
q_default <- quality_report(demo$responses, demo$instrument)
q_low     <- quality_report(demo$responses, demo$instrument,
                            straightline_min_items = 2)
scale_sizes <- vapply(demo$instrument$scales, function(s) length(s$items), integer(1))

reg <- data.frame(
  fact = c("longest scale in the demo instrument",
           "straightline_min_items default in 0.4.1",
           "scales actually checked",
           "sframe_plot_quality() with the default",
           "sframe_plot_quality() with the threshold lowered to 2"),
  value = c(max(scale_sizes), 4L,
            sum(vapply(q_default$straightline,
                       function(s) isTRUE(s$checked), logical(1))),
            if (is.null(sframe_plot_quality(q_default))) "NULL" else "a plot",
            if (is.null(sframe_plot_quality(q_low))) "NULL" else "a plot"),
  stringsAsFactors = FALSE
)
knitr::kable(reg, caption = "Why the quality plot disappeared")
```

The 0.4.1 fix stopping `quality_report()` from straight-line-checking scales
shorter than 4 items was correct, and the bundled instrument's 5 scales are
all 2 or 3 items. Every scale therefore comes back `checked = FALSE` with
`flag_rate = NA`, `sframe_plot_quality()` drops the NA rows, finds none left,
and returns `NULL`. The quality chart vanishes from the rendered report with
nothing said.

```{r}
#| label: quality-plot-verdict
qp <- compare_table(
  compare_row("quality plot returns NULL when no scale is checkable",
              as.numeric(is.null(sframe_plot_quality(q_default))), 1,
              "correct, there is nothing to draw"),
  compare_row("and the reason is available to the caller",
              as.numeric(grepl("long enough to check",
                               sframe_quality_plot_note(q_default) %||% "")),
              1, "sframe_quality_plot_note()"),
  compare_row("it renders once a scale is long enough to check",
              as.numeric(!is.null(sframe_plot_quality(q_low))), 1,
              "threshold lowered to 2"),
  compare_row("and then the note falls silent",
              as.numeric(is.null(sframe_quality_plot_note(q_low))), 1,
              "no note when there is a chart")
)
compare_report(qp, caption = "Quality plot and its explanation")
sf_log(qp)
```

- [x] **Fixed 2026-08-28.** `sframe_plot_quality()` still returns `NULL`,
      which is right when there is nothing to draw, and
      `sframe_quality_plot_note()` now carries the reason.
      `render_report()` prints it where the figure would have been, so an
      absent chart says "no scale was long enough to check" instead of
      reading as "nothing was flagged".

**Five plot builders take a result object, not the data.**
`sframe_plot_term_frequency()`, `sframe_plot_ngram_frequency()` and their
neighbours expect the analysis-plan runner's result, with `$table` and
`$options`, rather than the output of the same-named helper
(`term_frequency()`). Handed the wrong shape they return `NULL` rather than
saying so, which is the same "can it say no?" pattern. They are recorded as
skipped above and belong in a text-analysis accuracy file, alongside
`review_050`.

**Eight of the first-draft rows in this file were my own errors, not the
package's**: wrong argument order, an argument that does not exist, and a
`mode` value that is not accepted. That ratio is worth stating, because the
lesson from earlier releases holds. A surprising failure is worth
attributing before it is reported.

```{r}
#| label: coverage-comparison
cov_tab <- compare_table(
  compare_row("every export accounted for in the ledger",
              nrow(ledger), length(sf_exports()), "sf_exports()"),
  compare_row("exports raising an unexpected error",
              sum(ledger$status == "ERROR"), 0, "sf_smoke()"),
  # 1 export returns NULL by design: sframe_plot_quality() has nothing to
  # draw when no scale is long enough to straight-line check, and
  # sframe_quality_plot_note() carries the reason. Asserted as 1 rather than
  # 0 so a second NULL-returning export shows up as a DIFFERS.
  compare_row("exports returning NULL, all of them documented",
              sum(ledger$status == "NULL"), 1, "sf_smoke(), see 4.1")
)
compare_report(cov_tab, caption = "Coverage")
sf_log(cov_tab)
```

## 5. Sign-off

```{r}
#| label: verdict
sf_verdict()
```

- [ ] The ledger accounts for every export, with no row reading
      `NOT EXERCISED` that is not either a launcher, a Shiny module, or
      documented in section 4 as needing a human.
- [ ] No row reads `ERROR`.
- [ ] Any new export added since the last render appears here and has been
      given a smoke call.
