---
title: "02. Quarto rendering, audit and reproducibility"
subtitle: "Which engine actually produced the report in front of you"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

The rendered report is the audit artefact of a surveyframe study. It is what
a supervisor, a reviewer, or an ethics panel is shown, and the claim it
carries is that the numbers in it came from this instrument and this data,
reproducibly. This file checks that claim rather than assuming it.

`render_report()` has **2 engines**. It prefers Quarto, rendering
`inst/templates/report.qmd` with the instrument, the data, and the
instrument hash passed as parameters. If Quarto is missing, or its render
fails for any reason, it falls back to a base-R HTML writer in
`.render_report_html()`.

The fallback is the thing to look at closely. It is the correct engineering
choice, since a user without Quarto should still get a report. Whether the
user can **tell which one they got** is a separate question, and it is the
one that matters for an audit.

## 1. Is Quarto present on this machine, and is the template shipped?

```{r}
#| label: environment
env <- data.frame(
  check = c("quarto on PATH", "quarto version",
            "report.qmd shipped in the installed package",
            "surveyframe.use_quarto option"),
  value = c(
    sf_quarto_bin(),
    sf_quarto_version(),
    system.file("templates", "report.qmd", package = "surveyframe"),
    as.character(getOption("surveyframe.use_quarto", TRUE))
  ),
  stringsAsFactors = FALSE
)
knitr::kable(env, caption = "Rendering environment")
```

If `quarto on PATH` is empty, every result below describes the fallback and
the Quarto claims in this file are untested rather than passed. That
distinction is the whole point of the section, so read the table before the
rest.

```{r}
#| label: gate
quarto_here <- sf_quarto_available()
if (!quarto_here) {
  cat("Quarto is NOT installed on this machine.\n",
      "Everything below reports the fallback engine only.\n",
      "Install Quarto and re-render before signing this file off.\n")
}
```

## 2. Both engines, on the same instrument and the same data

The bundled tourism demo is the realistic case: a full instrument with
responses, the shape a real study hands to `render_report()`.

```{r}
#| label: both-engines
demo <- sframe_demo_data()
engines <- sf_render_both_engines(demo$instrument, demo$responses)
knitr::kable(engines, caption = "The same report down each engine")
```

Two things to read here.

`engine_used` is detected by inspecting the artefact, not by trusting the
request. `sf_report_engine()` looks for Quarto's own generator metadata and
runtime identifiers in the HTML. A row where `requested` is `quarto` and
`engine_used` is `fallback (base R)` means the fallback fired **silently**,
which is exactly the case an auditor needs to be able to see.

`hash_stamped` asks whether the instrument's SHA-256 appears in the rendered
document. That is the link between the artefact and the instrument it claims
to describe. A report without it cannot be tied back to anything.

```{r}
#| label: engine-comparison
tab <- compare_table(
  compare_row("quarto engine used when quarto requested",
              as.numeric(engines$engine_used[1] == "quarto"), 1,
              "sf_report_engine() on the artefact"),
  compare_row("instrument hash stamped, quarto engine",
              as.numeric(engines$hash_stamped[1]), 1, "grep on the artefact"),
  compare_row("instrument hash stamped, fallback engine",
              as.numeric(engines$hash_stamped[2]), 1, "grep on the artefact"),
  compare_row("both engines produced a file",
              as.numeric(all(!is.na(engines$size_kb))), 1, "file.exists()")
)
compare_report(tab, caption = "Engine selection and hash stamping")
sf_log(tab)
```

### The finding this file exists for

**A caller cannot tell which engine rendered their report.** Both paths end
in `return(invisible(dest))` and hand back a file path. There is no return
value, attribute, message, or metadata field naming the engine. The only way
to find out, the one used above, is to read the HTML afterwards and look for
Quarto's fingerprint.

The 2 artefacts are not close substitutes. On the run recorded while writing
this file the Quarto report was **5.8 MB** and the fallback **1.9 MB**, from
identical inputs. They are different documents.

The practical failure: a researcher runs `render_report()` on a machine
where Quarto is installed, gets the Quarto artefact, and writes their method
section around it. The same script later runs on a colleague's machine, or a
CI runner, or after a PATH change, and silently produces the fallback. Both
are called "the surveyframe report". Nothing in either says which it is.

```{r}
#| label: engine-observability
# Measured, not asserted. Each row is checked against a real render.
probe <- local({
  old <- options(surveyframe.use_quarto = FALSE); on.exit(options(old), add = TRUE)
  f <- tempfile(fileext = ".html")
  msgs <- character(0)
  path <- withCallingHandlers(
    suppressWarnings(render_report(demo$instrument, demo$responses,
                                   output_path = f)),
    message = function(m) { msgs <<- c(msgs, conditionMessage(m)); invokeRestart("muffleMessage") })
  list(path = path, msgs = msgs,
       body = paste(readLines(f, warn = FALSE), collapse = "\n"))
})
observable <- data.frame(
  route = c("message emitted during the render",
            "attribute on the returned path",
            "a line inside the rendered report",
            "return value of render_report()"),
  names_the_engine = c(
    if (any(grepl("engine", probe$msgs, ignore.case = TRUE))) "yes" else "no",
    if (!is.null(attr(probe$path, "engine"))) paste0("yes, engine = \"",
        attr(probe$path, "engine"), "\"") else "no",
    if (grepl("Rendering engine", probe$body, fixed = TRUE)) "yes, beside the instrument hash" else "no",
    "the destination path, as before"),
  stringsAsFactors = FALSE
)
knitr::kable(observable, caption = "Ways a caller can learn which engine ran")
```

- [ ] **Owner decision.** Should `render_report()` say which engine it used?
      The suggestion on the table is a `message()` naming the engine plus an
      `engine` attribute on the returned path, so a script can assert on it.
      A stricter option is an argument such as `engine = c("auto", "quarto",
      "html")`, where `"quarto"` errors rather than falling back, for users
      whose reproducibility claim depends on it.
- [ ] Whichever is chosen, the rendered report should name its own engine
      somewhere visible, next to the instrument hash it already carries.

## 3. Reproducibility: the same inputs twice

An audit artefact has to be reproducible. The same instrument and the same
data, rendered twice, should give the same document. Timestamps and
temporary paths legitimately differ between runs, so they are normalised
before comparison and reported separately. Normalising is itself a risk, so
the raw comparison is shown alongside.

```{r}
#| label: reproducible-quarto
rep_q <- if (quarto_here) sf_render_twice(demo$instrument, demo$responses,
                                          use_quarto = TRUE) else NULL
rep_f <- sf_render_twice(demo$instrument, demo$responses, use_quarto = FALSE)

rep_tab <- data.frame(
  engine = c(if (quarto_here) rep_q$engine, rep_f$engine),
  identical_raw = c(if (quarto_here) rep_q$identical_raw, rep_f$identical_raw),
  identical_after_normalising = c(if (quarto_here) rep_q$identical_normalised,
                                  rep_f$identical_normalised),
  stringsAsFactors = FALSE
)
knitr::kable(rep_tab, caption = "Same inputs, rendered twice")
```

`identical_raw` being `FALSE` while `identical_after_normalising` is `TRUE`
is the healthy result: the documents differ only in their render timestamp
and temp paths. Both being `FALSE` is a finding, and means the report varies
between runs on something other than time.

```{r}
#| label: reproducibility-check
# Fixed on 2026-08-28, so both rows now assert that reproducibility holds.
# The quarto row is asserted only when the installed build is the one under
# review, because Quarto renders in a separate session that loads the
# installed package. See 3.2.
matches <- sf_installed_matches_source()
rt <- compare_table(
  compare_row("fallback report reproducible after normalising",
              as.numeric(isTRUE(rep_f$identical_normalised)), 1,
              "2 renders, same inputs"),
  if (quarto_here && matches) compare_row(
              "quarto report reproducible after normalising",
              as.numeric(isTRUE(rep_q$identical_normalised)), 1,
              "2 renders, same inputs") else NULL
)
compare_report(rt, caption = "Render-to-render reproducibility")
sf_log(rt)
if (quarto_here && !matches) {
  cat("The quarto row is held back. The installed surveyframe is a different\n",
      "build from the one under review, and a Quarto render loads the installed\n",
      "one. Install this build and re-render to test that engine. See 3.2.\n")
}
```

### 3.1 The report is reproducible, and this is how it was fixed

Both engines now produce the same document twice from identical inputs. They
did not before 2026-08-28, and the cause was upstream of rendering:
`run_analysis_plan()` was non-deterministic, differing in **32 of 1768
values** between 2 runs on the same instrument and the same responses.

Three independent causes, all of them unseeded randomness, all now fixed.

```{r}
#| label: plan-nondeterminism
r1 <- run_analysis_plan(demo$responses, demo$instrument)
r2 <- run_analysis_plan(demo$responses, demo$instrument)
s1 <- capture.output(str(r1, max.level = 7))
s2 <- capture.output(str(r2, max.level = 7))
differing <- sum(s1 != s2[seq_along(s1)], na.rm = TRUE)

nd <- data.frame(
  source = c("bootstrap effect-size CIs", "parallel analysis in EFA",
             "jittered points on group-comparison charts"),
  what_moved = c("r_ci, d_ci, eta_ci, v_ci, and the APA strings quoting them",
                 "pc.sim, fa.sim and the simulated eigenvalue matrix",
                 "the scatter drawn over each boxplot"),
  fix = c("run_analysis_plan() gains seed, defaulting to a fixed value",
          "the same seeding, plus mc.cores pinned to 1 so forked workers draw from the parent stream",
          "position_jitter(seed = ) on the shared boxplot helper"),
  stringsAsFactors = FALSE
)
knitr::kable(nd, caption = "Three sources of non-determinism, all fixed")
cat(sprintf("\nstr() lines differing between 2 identical runs: %d of %d\n",
            differing, length(s1)))
```

`bootstrap_ci()` takes `seed` and defaulted it to `NULL`, so it seeded only
when asked, and nothing in the analysis path asked. The 5 call sites are
`R/analysis_plan.R` lines 315, 422, 492, 546, and 653, covering
`cramers_v_ci()`, `cohens_d_ci()`, `eta_sq_ci()`, a direct `bootstrap_ci()`,
and `sframe_kw_eta_sq_ci()`. The EFA path inherited `psych::fa.parallel()`'s
own simulation, which splits across forked `mclapply()` workers whose streams
`set.seed()` never reached, and which split by core count so the answer
depended on the machine.

**The third cause appeared only after the first 2 were fixed.** With the
numbers settled, 5 lines of the rendered report still moved, all of them
embedded chart images: `geom_jitter()` drew its offsets from the global stream
at plot time, which is after `run_analysis_plan()` restores the caller's
state. A report used as an audit artefact has to settle in its pictures as
well as its numbers, so the shared boxplot helper now uses
`position_jitter(seed = )`.

The visible effect before the fix, taken from 2 renders of the bundled demo:

```
render 1:  U = 1576, z = -0.98, p = 0.327, r = 0.09 [0.01, 0.27]
render 2:  U = 1576, z = -0.98, p = 0.327, r = 0.09 [0.00, 0.27]
```

The test statistic and the p value are stable, because they are computed
analytically. The interval a researcher would quote in a paper is not.

**Why this matters more than its size suggests.** A report that changes
between renders cannot be the audit artefact of a study. Re-running the
analysis to check a number produces a different number, and neither the
researcher nor a reviewer can tell whether that reflects a change in the
data, a change in the package, or nothing at all. It also undercuts the
instrument hash: the hash pins the *instrument*, and the report then varies
anyway.

- [x] **Owner decision taken 2026-08-28: seed by default and record it.**
      `run_analysis_plan()` gains `seed`, defaulting to a fixed value, applied
      once at the top of the run so it reaches anything random the plan
      touches. `seed = NULL` restores the earlier behaviour. This does change
      confidence intervals the package has already printed, the same class as
      `review_040`'s finding 1, so NEWS states plainly that intervals from
      earlier releases are reproducible from this release onward and move
      once when a study is re-run.
- [x] `render_report()` records the seed in the artefact beside the
      instrument hash, on both engines, so a report states what reproducing
      it would take.

### 3.2 A Quarto render measures the installed package

Worth knowing before reading the quarto row above, and it produced a phantom
defect on the day this section was rewritten.

Quarto renders `inst/templates/report.qmd` in a **separate R session**, which
resolves `library(surveyframe)` against the installed package. A working tree
loaded with `pkgload::load_all()` reaches the fallback engine and never
reaches the Quarto one. On 2026-08-28 that showed as the fallback rendering
reproducibly while Quarto appeared to keep moving its intervals, and the cause
was the subprocess running CRAN 0.4.0, which has no seeding, rather than any
defect in the build under review.

```{r}
#| label: installed-build
ib <- data.frame(
  check = c("installed build carries the feature under review",
            "so the quarto row above is",
            "to test the quarto engine"),
  value = c(as.character(sf_installed_matches_source()),
            if (sf_installed_matches_source()) "measuring this build" else
              "held back, see the note under section 3",
            "R CMD INSTALL this working tree, then re-render"),
  stringsAsFactors = FALSE
)
knitr::kable(ib, caption = "Which build a Quarto render actually loads")
```

Verified on 2026-08-28 by installing the working tree into a temporary
library and re-rendering: **0 differing lines on both engines**.

## 4. The audit chain: does the report bind to the instrument?

The hash is the instrument's identity. A report that carries it can be tied
to the exact `.sframe` that produced it, and a `.sframe` that fails its own
integrity check can be caught. This checks that the chain actually holds
end to end, rather than that a hash-shaped string appears somewhere.

```{r}
#| label: audit-chain
inst <- demo$instrument
stored <- sframe_hash_value(inst)

f <- tempfile(fileext = ".html")
suppressMessages(render_report(inst, demo$responses, output_path = f))
body <- paste(readLines(f, warn = FALSE), collapse = "\n")

# The instrument round-tripped through a file must still hash the same, or
# the value printed in the report identifies nothing stable.
p <- tempfile(fileext = ".sframe")
write_sframe(inst, p, overwrite = TRUE)
reread <- read_sframe(p)

chain <- compare_table(
  compare_row("hash appears in the rendered report",
              as.numeric(grepl(substr(stored, 1, 16), body, fixed = TRUE)), 1,
              "grep on the artefact"),
  compare_row("instrument survives a write/read with the same hash",
              as.numeric(identical(sframe_hash_value(reread), stored)), 1,
              "write_sframe() then read_sframe()"),
  compare_row("stored hash equals recomputed hash",
              as.numeric(identical(
                jsonlite::fromJSON(p, simplifyVector = FALSE)$hash$value,
                sframe_hash_value(reread))), 1,
              "file's own embedded hash")
)
compare_report(chain, caption = "Instrument to report audit chain")
sf_log(chain)
```

```{r}
#| label: format-field
# 0.4.1 writes the format version into the file, so a consumer can tell which
# shape of .sframe it holds. An audit trail that cannot version-negotiate is
# one release away from being unreadable.
written <- jsonlite::fromJSON(p, simplifyVector = FALSE)
fmt <- compare_table(
  compare_row("file states its own format version",
              as.numeric(!is.null(written$sframe_format)), 1, "sframe_format key"),
  compare_row("format version is the one this build writes",
              as.numeric(identical(written$sframe_format,
                                   surveyframe:::SFRAME_FORMAT_VERSION)), 1,
              "SFRAME_FORMAT_VERSION")
)
compare_report(fmt, caption = "Format declaration")
sf_log(fmt)
```

## 5. What happens when Quarto fails rather than being absent

Absence is the easy case. The harder one is a Quarto that exists and then
errors, which is what a broken install, a missing Chromium for PDF output,
or a template edit produces. `render_report()` treats both the same way and
falls back.

```{r}
#| label: quarto-broken
# Simulate a broken Quarto by pointing PATH at a directory with none in it.
# This exercises the same branch a genuinely failing render takes.
broken <- local({
  old_path <- Sys.getenv("PATH")
  on.exit(Sys.setenv(PATH = old_path), add = TRUE)
  Sys.setenv(PATH = tempdir())
  f <- tempfile(fileext = ".html")
  msgs <- character(0)
  ok <- tryCatch({
    withCallingHandlers(
      suppressWarnings(render_report(demo$instrument, demo$responses,
                                     output_path = f)),
      message = function(m) {
        msgs <<- c(msgs, conditionMessage(m)); invokeRestart("muffleMessage")
      })
    TRUE
  }, error = function(e) conditionMessage(e))
  data.frame(
    scenario = "quarto not reachable on PATH",
    render_succeeded = isTRUE(ok),
    engine_used = if (isTRUE(ok)) sf_report_engine(f) else paste("FAILED:", ok),
    # Detected, never hardcoded. A literal here would keep reading "match"
    # after the package started announcing the engine, which is a check that
    # can no longer fail.
    told_the_user = any(grepl("engine", msgs, ignore.case = TRUE)),
    stringsAsFactors = FALSE
  )
})
knitr::kable(broken, caption = "Behaviour when Quarto cannot be reached")
```

```{r}
#| label: fallback-silence
fb <- compare_table(
  compare_row("a report is still produced when quarto is unreachable",
              as.numeric(broken$render_succeeded), 1, "graceful degradation"),
  compare_row("the user is told which engine ran",
              as.numeric(broken$told_the_user), 1,
              "message class sframe_report_engine"),
  compare_row("the artefact names its own engine",
              as.numeric(grepl("Rendering engine", probe$body, fixed = TRUE)),
              1, "read off the rendered file")
)
compare_report(fb, caption = "Fallback behaviour")
sf_log(fb)
```

- [x] Fixed 2026-08-28. The fallback announces itself 3 ways, so a caller who
      expected Quarto and received the built-in writer can see that from the
      console, from the returned path, or from the report itself.

## 6. Does the Quarto template itself still run?

The template is a shipped file with its own chunks. It can break
independently of the R code, and a broken template shows up only as a silent
fallback, which section 2 established nobody sees.

```{r}
#| label: template-direct
tpl <- system.file("templates", "report.qmd", package = "surveyframe")
tpl_check <- data.frame(
  check = c("template exists", "template is non-empty",
            "declares params", "reads the instrument param",
            "stamps the instrument hash"),
  value = c(
    file.exists(tpl),
    file.exists(tpl) && file.size(tpl) > 0,
    file.exists(tpl) && any(grepl("^params:", readLines(tpl, warn = FALSE))),
    file.exists(tpl) && any(grepl("instrument_path", readLines(tpl, warn = FALSE))),
    file.exists(tpl) && any(grepl("instrument_hash", readLines(tpl, warn = FALSE)))
  ),
  stringsAsFactors = FALSE
)
knitr::kable(tpl_check, caption = "The shipped Quarto template")
```

```{r}
#| label: template-verdict
tv <- compare_table(
  compare_row("template ships with the package",
              as.numeric(tpl_check$value[1]), 1, "system.file()"),
  compare_row("template is parameterised",
              as.numeric(tpl_check$value[3]), 1, "params: block"),
  compare_row("template receives the instrument hash",
              as.numeric(tpl_check$value[5]), 1, "instrument_hash param")
)
compare_report(tv, caption = "Template integrity")
sf_log(tv)
```

## 7. Sign-off

```{r}
#| label: verdict
sf_verdict()
```

All 3 findings this file raised on 2026-08-28 are fixed and the sections now
assert the fixed behaviour. What a reviewer confirms:

- [ ] Section 1 shows Quarto present, and its version recorded here.
- [ ] Section 2 shows `engine_used` as `quarto` when Quarto is requested, and
      all 3 observability routes reading `yes`.
- [ ] Section 3 shows the fallback engine reproducible after normalising.
- [ ] **Section 3.2 shows the installed build matching this one.** If it
      reads `FALSE`, the quarto reproducibility row was held back and that
      engine is untested here. Install this working tree and re-render before
      signing the section off.
- [ ] Section 4 shows the audit chain intact: hash in the report, hash stable
      across a write and read, format version declared, and the analysis seed
      printed beside the hash.
- [ ] Section 5 shows the fallback announcing itself, so a caller who expected
      Quarto can tell what they received.
- [ ] Section 6 shows the template shipped and parameterised.
