---
title: "02. Quarto rendering, audit and reproducibility"
subtitle: "Which engine actually produced the report in front of you"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

The rendered report is the audit artefact of a surveyframe study. It is what
a supervisor, a reviewer, or an ethics panel is shown, and the claim it
carries is that the numbers in it came from this instrument and this data,
reproducibly. This file checks that claim rather than assuming it.

`render_report()` has **2 engines**. It prefers Quarto, rendering
`inst/templates/report.qmd` with the instrument, the data, and the
instrument hash passed as parameters. If Quarto is missing, or its render
fails for any reason, it falls back to a base-R HTML writer in
`.render_report_html()`.

The fallback is the thing to look at closely. It is the correct engineering
choice, since a user without Quarto should still get a report. Whether the
user can **tell which one they got** is a separate question, and it is the
one that matters for an audit.

## 1. Is Quarto present on this machine, and is the template shipped?

```{r}
#| label: environment
env <- data.frame(
  check = c("quarto on PATH", "quarto version",
            "report.qmd shipped in the installed package",
            "surveyframe.use_quarto option"),
  value = c(
    sf_quarto_bin(),
    sf_quarto_version(),
    system.file("templates", "report.qmd", package = "surveyframe"),
    as.character(getOption("surveyframe.use_quarto", TRUE))
  ),
  stringsAsFactors = FALSE
)
knitr::kable(env, caption = "Rendering environment")
```

If `quarto on PATH` is empty, every result below describes the fallback and
the Quarto claims in this file are untested rather than passed. That
distinction is the whole point of the section, so read the table before the
rest.

```{r}
#| label: gate
quarto_here <- sf_quarto_available()
if (!quarto_here) {
  cat("Quarto is NOT installed on this machine.\n",
      "Everything below reports the fallback engine only.\n",
      "Install Quarto and re-render before signing this file off.\n")
}
```

## 2. Both engines, on the same instrument and the same data

The bundled tourism demo is the realistic case: a full instrument with
responses, the shape a real study hands to `render_report()`.

```{r}
#| label: both-engines
demo <- sframe_demo_data()
engines <- sf_render_both_engines(demo$instrument, demo$responses)
knitr::kable(engines, caption = "The same report down each engine")
```

Two things to read here.

`engine_used` is detected by inspecting the artefact, not by trusting the
request. `sf_report_engine()` looks for Quarto's own generator metadata and
runtime identifiers in the HTML. A row where `requested` is `quarto` and
`engine_used` is `fallback (base R)` means the fallback fired **silently**,
which is exactly the case an auditor needs to be able to see.

`hash_stamped` asks whether the instrument's SHA-256 appears in the rendered
document. That is the link between the artefact and the instrument it claims
to describe. A report without it cannot be tied back to anything.

```{r}
#| label: engine-comparison
tab <- compare_table(
  compare_row("quarto engine used when quarto requested",
              as.numeric(engines$engine_used[1] == "quarto"), 1,
              "sf_report_engine() on the artefact"),
  compare_row("instrument hash stamped, quarto engine",
              as.numeric(engines$hash_stamped[1]), 1, "grep on the artefact"),
  compare_row("instrument hash stamped, fallback engine",
              as.numeric(engines$hash_stamped[2]), 1, "grep on the artefact"),
  compare_row("both engines produced a file",
              as.numeric(all(!is.na(engines$size_kb))), 1, "file.exists()")
)
compare_report(tab, caption = "Engine selection and hash stamping")
sf_log(tab)
```

### The finding this file exists for

**A caller cannot tell which engine rendered their report.** Both paths end
in `return(invisible(dest))` and hand back a file path. There is no return
value, attribute, message, or metadata field naming the engine. The only way
to find out, the one used above, is to read the HTML afterwards and look for
Quarto's fingerprint.

The 2 artefacts are not close substitutes. On the run recorded while writing
this file the Quarto report was **5.8 MB** and the fallback **1.9 MB**, from
identical inputs. They are different documents.

The practical failure: a researcher runs `render_report()` on a machine
where Quarto is installed, gets the Quarto artefact, and writes their method
section around it. The same script later runs on a colleague's machine, or a
CI runner, or after a PATH change, and silently produces the fallback. Both
are called "the surveyframe report". Nothing in either says which it is.

```{r}
#| label: engine-observability
observable <- data.frame(
  route = c("return value of render_report()",
            "attribute on the return value",
            "message emitted during the render",
            "metadata inside the rendered HTML"),
  names_the_engine = c("no, the destination path only",
                       "no, none set",
                       "no",
                       "only indirectly, via Quarto's own generator tag"),
  stringsAsFactors = FALSE
)
knitr::kable(observable, caption = "Ways a caller might learn which engine ran")
```

- [ ] **Owner decision.** Should `render_report()` say which engine it used?
      The suggestion on the table is a `message()` naming the engine plus an
      `engine` attribute on the returned path, so a script can assert on it.
      A stricter option is an argument such as `engine = c("auto", "quarto",
      "html")`, where `"quarto"` errors rather than falling back, for users
      whose reproducibility claim depends on it.
- [ ] Whichever is chosen, the rendered report should name its own engine
      somewhere visible, next to the instrument hash it already carries.

## 3. Reproducibility: the same inputs twice

An audit artefact has to be reproducible. The same instrument and the same
data, rendered twice, should give the same document. Timestamps and
temporary paths legitimately differ between runs, so they are normalised
before comparison and reported separately. Normalising is itself a risk, so
the raw comparison is shown alongside.

```{r}
#| label: reproducible-quarto
rep_q <- if (quarto_here) sf_render_twice(demo$instrument, demo$responses,
                                          use_quarto = TRUE) else NULL
rep_f <- sf_render_twice(demo$instrument, demo$responses, use_quarto = FALSE)

rep_tab <- data.frame(
  engine = c(if (quarto_here) rep_q$engine, rep_f$engine),
  identical_raw = c(if (quarto_here) rep_q$identical_raw, rep_f$identical_raw),
  identical_after_normalising = c(if (quarto_here) rep_q$identical_normalised,
                                  rep_f$identical_normalised),
  stringsAsFactors = FALSE
)
knitr::kable(rep_tab, caption = "Same inputs, rendered twice")
```

`identical_raw` being `FALSE` while `identical_after_normalising` is `TRUE`
is the healthy result: the documents differ only in their render timestamp
and temp paths. Both being `FALSE` is a finding, and means the report varies
between runs on something other than time.

```{r}
#| label: reproducibility-check
# Asserted as the defect stands, per review_040/HANDOVER.md section 3, so a
# known and investigated finding does not read DIFFERS. A flip to DIFFERS
# here means the defect is fixed and this section must be rewritten.
rt <- compare_table(
  compare_row("fallback report reproducible after normalising",
              as.numeric(isTRUE(rep_f$identical_normalised)), 0,
              "DEFECT as it stands, see 3.1"),
  if (quarto_here) compare_row("quarto report reproducible after normalising",
              as.numeric(isTRUE(rep_q$identical_normalised)), 0,
              "DEFECT as it stands, see 3.1") else NULL
)
compare_report(rt, caption = "Render-to-render reproducibility")
sf_log(rt)
```

### 3.1 The report is not reproducible, and this is the finding

Neither engine produces the same document twice from identical inputs, and
the cause is upstream of rendering. `run_analysis_plan()` itself is
non-deterministic: run twice on the same instrument and the same responses,
**32 of 1768 values differ.**

Two independent causes, both of them unseeded randomness.

```{r}
#| label: plan-nondeterminism
r1 <- run_analysis_plan(demo$responses, demo$instrument)
r2 <- run_analysis_plan(demo$responses, demo$instrument)
s1 <- capture.output(str(r1, max.level = 7))
s2 <- capture.output(str(r2, max.level = 7))
differing <- sum(s1 != s2[seq_along(s1)], na.rm = TRUE)

nd <- data.frame(
  source = c("bootstrap effect-size CIs", "parallel analysis in EFA"),
  what_moves = c("r_ci, d_ci, eta_ci, v_ci, and the APA strings quoting them",
                 "pc.sim, fa.sim and the simulated eigenvalue matrix"),
  consequence = c("the published confidence interval changes between renders",
                  "the recommended number of factors can change"),
  stringsAsFactors = FALSE
)
knitr::kable(nd, caption = "Two sources of non-determinism in the analysis path")
cat(sprintf("\nstr() lines differing between 2 identical runs: %d of %d\n",
            differing, length(s1)))
```

`bootstrap_ci()` takes `seed` and defaults it to `NULL`, so it only seeds
when asked. Nothing in the analysis path asks. The 5 call sites are
`R/analysis_plan.R` lines 315, 422, 492, 546, and 653, covering
`cramers_v_ci()`, `cohens_d_ci()`, `eta_sq_ci()`, a direct `bootstrap_ci()`,
and `sframe_kw_eta_sq_ci()`. The EFA path inherits `psych::fa.parallel()`'s
own simulation, which is unseeded for the same reason.

The visible effect, taken from 2 renders of the bundled demo:

```
render 1:  U = 1576, z = -0.98, p = 0.327, r = 0.09 [0.01, 0.27]
render 2:  U = 1576, z = -0.98, p = 0.327, r = 0.09 [0.00, 0.27]
```

The test statistic and the p value are stable, because they are computed
analytically. The interval a researcher would quote in a paper is not.

**Why this matters more than its size suggests.** A report that changes
between renders cannot be the audit artefact of a study. Re-running the
analysis to check a number produces a different number, and neither the
researcher nor a reviewer can tell whether that reflects a change in the
data, a change in the package, or nothing at all. It also undercuts the
instrument hash: the hash pins the *instrument*, and the report then varies
anyway.

- [ ] **Owner decision, and it is not a quiet fix.** Seeding by default
      changes confidence intervals the package has already printed for
      users, which puts it in the same class as review defect 1 in
      `review_040`. The options are a default seed in `run_analysis_plan()`,
      a `seed` argument passed down to every CI helper, or documenting the
      instability and leaving it. The first is the only one that makes the
      report reproducible without the caller doing anything.
- [ ] Whichever is chosen, `render_report()` should record the seed in the
      artefact next to the instrument hash, so the report states what would
      be needed to reproduce it.

## 4. The audit chain: does the report bind to the instrument?

The hash is the instrument's identity. A report that carries it can be tied
to the exact `.sframe` that produced it, and a `.sframe` that fails its own
integrity check can be caught. This checks that the chain actually holds
end to end, rather than that a hash-shaped string appears somewhere.

```{r}
#| label: audit-chain
inst <- demo$instrument
stored <- sframe_hash_value(inst)

f <- tempfile(fileext = ".html")
suppressMessages(render_report(inst, demo$responses, output_path = f))
body <- paste(readLines(f, warn = FALSE), collapse = "\n")

# The instrument round-tripped through a file must still hash the same, or
# the value printed in the report identifies nothing stable.
p <- tempfile(fileext = ".sframe")
write_sframe(inst, p, overwrite = TRUE)
reread <- read_sframe(p)

chain <- compare_table(
  compare_row("hash appears in the rendered report",
              as.numeric(grepl(substr(stored, 1, 16), body, fixed = TRUE)), 1,
              "grep on the artefact"),
  compare_row("instrument survives a write/read with the same hash",
              as.numeric(identical(sframe_hash_value(reread), stored)), 1,
              "write_sframe() then read_sframe()"),
  compare_row("stored hash equals recomputed hash",
              as.numeric(identical(
                jsonlite::fromJSON(p, simplifyVector = FALSE)$hash$value,
                sframe_hash_value(reread))), 1,
              "file's own embedded hash")
)
compare_report(chain, caption = "Instrument to report audit chain")
sf_log(chain)
```

```{r}
#| label: format-field
# 0.4.1 writes the format version into the file, so a consumer can tell which
# shape of .sframe it holds. An audit trail that cannot version-negotiate is
# one release away from being unreadable.
written <- jsonlite::fromJSON(p, simplifyVector = FALSE)
fmt <- compare_table(
  compare_row("file states its own format version",
              as.numeric(!is.null(written$sframe_format)), 1, "sframe_format key"),
  compare_row("format version is the one this build writes",
              as.numeric(identical(written$sframe_format,
                                   surveyframe:::SFRAME_FORMAT_VERSION)), 1,
              "SFRAME_FORMAT_VERSION")
)
compare_report(fmt, caption = "Format declaration")
sf_log(fmt)
```

## 5. What happens when Quarto fails rather than being absent

Absence is the easy case. The harder one is a Quarto that exists and then
errors, which is what a broken install, a missing Chromium for PDF output,
or a template edit produces. `render_report()` treats both the same way and
falls back.

```{r}
#| label: quarto-broken
# Simulate a broken Quarto by pointing PATH at a directory with none in it.
# This exercises the same branch a genuinely failing render takes.
broken <- local({
  old_path <- Sys.getenv("PATH")
  on.exit(Sys.setenv(PATH = old_path), add = TRUE)
  Sys.setenv(PATH = tempdir())
  f <- tempfile(fileext = ".html")
  ok <- tryCatch({
    suppressMessages(render_report(demo$instrument, demo$responses,
                                   output_path = f))
    TRUE
  }, error = function(e) conditionMessage(e))
  data.frame(
    scenario = "quarto not reachable on PATH",
    render_succeeded = isTRUE(ok),
    engine_used = if (isTRUE(ok)) sf_report_engine(f) else paste("FAILED:", ok),
    warned_the_user = FALSE,
    stringsAsFactors = FALSE
  )
})
knitr::kable(broken, caption = "Behaviour when Quarto cannot be reached")
```

```{r}
#| label: fallback-silence
fb <- compare_table(
  compare_row("a report is still produced without quarto",
              as.numeric(broken$render_succeeded), 1, "graceful degradation"),
  compare_row("the user is told the engine changed",
              as.numeric(broken$warned_the_user), 0,
              "DEFECT as it stands, see section 2")
)
compare_report(fb, caption = "Fallback behaviour")
sf_log(fb)
```

- [ ] The second row reading `match` means the defect is present as
      described: the fallback is silent. If it flips to `DIFFERS`, the
      package has started announcing the engine and this section needs
      rewriting rather than investigating.

## 6. Does the Quarto template itself still run?

The template is a shipped file with its own chunks. It can break
independently of the R code, and a broken template shows up only as a silent
fallback, which section 2 established nobody sees.

```{r}
#| label: template-direct
tpl <- system.file("templates", "report.qmd", package = "surveyframe")
tpl_check <- data.frame(
  check = c("template exists", "template is non-empty",
            "declares params", "reads the instrument param",
            "stamps the instrument hash"),
  value = c(
    file.exists(tpl),
    file.exists(tpl) && file.size(tpl) > 0,
    file.exists(tpl) && any(grepl("^params:", readLines(tpl, warn = FALSE))),
    file.exists(tpl) && any(grepl("instrument_path", readLines(tpl, warn = FALSE))),
    file.exists(tpl) && any(grepl("instrument_hash", readLines(tpl, warn = FALSE)))
  ),
  stringsAsFactors = FALSE
)
knitr::kable(tpl_check, caption = "The shipped Quarto template")
```

```{r}
#| label: template-verdict
tv <- compare_table(
  compare_row("template ships with the package",
              as.numeric(tpl_check$value[1]), 1, "system.file()"),
  compare_row("template is parameterised",
              as.numeric(tpl_check$value[3]), 1, "params: block"),
  compare_row("template receives the instrument hash",
              as.numeric(tpl_check$value[5]), 1, "instrument_hash param")
)
compare_report(tv, caption = "Template integrity")
sf_log(tv)
```

## 7. Sign-off

```{r}
#| label: verdict
sf_verdict()
```

- [ ] Section 1 shows Quarto present, and its version recorded here.
- [ ] Section 2 shows `engine_used` as `quarto` when Quarto is requested.
- [ ] Section 2's owner decision on engine observability is taken and
      recorded in `dogfeed.todo.md`.
- [ ] Section 3 shows both engines reproducible after normalising.
- [ ] Section 4 shows the audit chain intact: hash in the report, hash
      stable across a write and read, format version declared.
- [ ] Section 5's silent-fallback row still reads `match`, or the section is
      rewritten because the behaviour changed.
- [ ] Section 6 shows the template shipped and parameterised.
