---
title: "{{TITLE}}"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(surveyframe)
```

## The study

This notebook runs your instrument end to end: load the instrument and its
responses, run the pre-declared analysis plan, and render a report. It reads
`{{SFRAME_FILE}}` and `{{CSV_FILE}}` from the same folder as this notebook,
so keep the 3 files together.

```{r load}
instrument <- read_sframe("{{SFRAME_FILE}}")
# strict = FALSE: the CSV may carry extra columns (respondent_id,
# submitted_at, and similar) that aren't declared in the instrument itself.
responses  <- read_responses("{{CSV_FILE}}", instrument, strict = FALSE)
instrument
```

## The questions

```{r items}
codebook <- as.data.frame(codebook_report(instrument))
knitr::kable(codebook[, c("id", "label", "type")])
```

## The plan, declared before the data existed

This is what makes the analysis a contract rather than a search. The blocks
below were written into the instrument at design time.

```{r plan}
knitr::kable(do.call(rbind, lapply(sf_plan(instrument), function(b) {
  data.frame(id = b$id, question = b$research_question, method = b$method)
})))
```

## The results

```{r run}
results <- run_analysis_plan(responses, instrument)
status <- attr(results, "status")
for (b in results) {
  cat("\n###", b$block_id, b$research_question, "\n\n")
  if (!is.null(b$error)) {
    cat("**This analysis failed.** ", b$error, "\n\n", sep = "")
    next
  }
  if (!is.null(b$apa)) cat(b$apa, "\n\n")
  if (is.data.frame(b$table)) print(knitr::kable(b$table))
}
if (!is.null(status) && status$failed > 0) {
  cat("\n", status$failed, " of ", status$blocks, " analyses failed: ",
      paste(status$failed_blocks, collapse = ", "), ".\n", sep = "")
}
```

## A report you can hand to somebody

```{r report}
# The report is rendered only when at least one analysis succeeded.
if (!is.null(attr(results, "status")) && attr(results, "status")$succeeded > 0) {
  render_report(instrument, responses, output_file = "report.html")
} else {
  cat("Every analysis failed, so no report was rendered. See the errors above.\n")
}
```

## Checking these numbers against other software

```{r export}
#| eval: false
sframe_export_labelled(responses, instrument, "responses.sav")
```

The `.sav` arrives with the question wording and the response options already
attached, so your variables read as questions rather than as codes. Take it
into `psych`, SPSS, JASP, jamovi or Stata and compare.
