---
title: "0.4.0 review suite: start here"
subtitle: "Read this file first, then work through the numbered files in order"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: false
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

## What this suite is for

surveyframe 0.4.0 is a major release. It adds the whole small-sample track,
all 10 MCDM methods, 2 new item types, conjoint designs, an RStudio add-in,
and it makes 2 breaking changes to how responses are collected and hashed.

The goal of this suite is narrow and absolute: **confirm that every route
from questionnaire to result works, and that every number surveyframe
reports is the number an independent tool reports**, before the tarball goes
to CRAN.

It replaces the single long `mas_review_040.qmd`, which was one document of
roughly 120 chunks. That shape suited neither a reviewer working in sittings
nor a first-time user, so it is deferred rather than deleted. This suite is
written for someone meeting surveyframe for the first time, and each file
stands alone.

## How each file is built

Every file follows the same shape, so once you have run one you know how to
run all of them.

1.  **Questionnaire.** Build a small instrument from scratch. Nothing is
    loaded from a fixture unless the point of the file is the fixture.
2.  **Analysis plan.** Declare the research question before any data exists.
    This is the package's central claim, so every file exercises it.
3.  **HTML route.** Export the static survey and confirm every item reaches
    the file. The path is printed so you can open and click through it.
4.  **Shiny route.** Build every widget and drive the real collector, then
    read the row back. Both routes run in every file, because 0.4.0 fixed a
    defect where the Shiny collector wrote a shape the reader could not read.
5.  **Dummy data and statistical accuracy.** Simulate responses, run the
    plan, and compare each number against base R and against a reference
    package. Every comparison prints a table and a plot.

## Reading a comparison

Every accuracy chunk ends with a table and a plot from the same helper.

```{r}
#| label: demo-comparison
set.seed(1)
x <- rnorm(40, 100, 15)

tab <- compare_table(
  compare_row("mean", mean(x), 100.98, "hand-computed"),
  compare_row("sd",   sd(x),   sd(x),  "base R sd()"),
  compare_row("n",    length(x), 40L,  "by construction")
)
compare_report(tab, caption = "How to read a comparison table")
```

Scan the `verdict` column. Three words can appear.

-   `match`. The difference is at or below the stated tolerance.
-   `DIFFERS`. The difference is above tolerance. **Stop and investigate.**
    Do not tick the box.
-   `CHECK`. One side is `NA`. Usually the reference call failed, but treat
    it as unresolved until you know which.

The plot shows the size of each difference on a log scale, with the
tolerance drawn as a dashed line. It exists so a difference of 1e-16 and a
difference of 0.2 do not look alike in a table of rounded numbers. The first
row above is deliberately wrong, so you can see what `DIFFERS` looks like
before you meet a real one.

::: callout-important
## Tolerance is not a dial to turn

The default tolerance is `1e-6`. If a chunk loosens it, the chunk says why in
a comment. A tolerance loosened without a stated reason is how a real
difference hides, so if you see one, treat the missing reason as the finding.
:::

## A trap worth knowing before you start

Reference calls must match surveyframe's own options, or you will be
comparing 2 different estimators and reading the gap as a bug.

The clearest case is in file `12`. surveyframe fixes `exact = FALSE` on
`wilcox.test()`. Left on its own default, base R switches to the exact method
at small n, and the 2 disagree in the 5th decimal. Nothing is wrong. The
comparison was.

## Order to work through

**Written and passing.** Each renders with 0 errors and every comparison
reads `match`.

| File | Covers | Comparisons |
|------------------|--------------------------------------|--------|
| `01_first_workflow` | The whole package in one sitting | 17 |
| `04_categorical` | Chi-square, Fisher, McNemar, Cochran Q | 23 |
| `05_two_group` | t, Mann-Whitney, paired t, Wilcoxon | 28 |
| `06_multi_group` | ANOVA, Kruskal, two-way, ANCOVA, repeated, Friedman | 30 |
| `07_correlation` | Pearson, Spearman, Kendall, partial | 16 |
| `08_regression` | Linear, logistic, ordinal, multinomial, Firth | 32 |
| `09_psychometrics` | Alpha, omega, item diagnostics, EFA, validity | 23 |
| `12_small_sample` | The whole small-sample track | 19 |
| `13_mcdm` | All 10 MCDM methods | 15 |

That is 203 checked numbers across 43 of the 50 analysis methods.

**Still to write.** The pattern is settled, so these are mechanical rather
than exploratory.

| File | Covers | Needs a keyboard? |
|------------------|--------------------------------------|------------------|
| `02` | Every item type, one demo each | no |
| `03` | Descriptives, frequencies, quality, missingness | no |
| `10` | Model syntax: CFA, CB-SEM, PLS-SEM | no |
| `11` | Mediation and moderation | no |
| `15` | Sensitivity analysis and conjoint designs | no |
| `16` | Collection routes and branching | no |
| `17` | Reporting and the codebook | no |
| `18` | The 3 GUIs | **yes** |
| `19` | The RStudio add-in | **yes, release blocker** |
| `20` | Release safety and owner decisions | partly |

## What this suite has already found

Three defects, all logged in `dogfeed.todo.md`, none of which any test suite
had caught.

1. **`item_report()` ignores `reverse = TRUE`** while `score_scales()` and
   `reliability_report()` honour it. A reverse-keyed item reports a strongly
   negative item-rest correlation on a scale whose alpha reads 0.84, and it
   depresses every other item in the scale as well. Needs an owner decision
   for 0.4.0. See file `09`, section 5.
2. **`roles = list(variables = ...)`** works for `correlation_kendall`,
   fails with `Test failed.` for Pearson and Spearman, and leaks
   `undefined columns selected` for partial correlation, while 9 other
   methods accept it. See file `07`.
3. **`assumption_report()` prints "Assumption checks were computed."** when
   no check ran. Laned to 0.4.1. See file `05`, section 8.

## Environment this ran on

```{r}
#| label: env
data.frame(
  item = c("R", "platform", "surveyframe loaded from",
           "0.4.0 feature present"),
  value = c(R.version.string, R.version$platform,
            dirname(dirname(getNamespaceInfo("surveyframe", "path"))),
            as.character(sf_is_040())),
  stringsAsFactors = FALSE
)
```

```{r}
#| label: refs
pk <- c("jmv", "psych", "DescTools", "MASS", "nnet", "logistf", "naniar",
        "lavaan", "seminr", "RMCDA", "car")
data.frame(
  package = pk,
  version = vapply(pk, function(p)
    if (requireNamespace(p, quietly = TRUE))
      as.character(utils::packageVersion(p)) else "MISSING", character(1)),
  row.names = NULL
)
```

-   [ ] Every reference package above reports a version rather than
    `MISSING`. A `MISSING` package turns its comparisons into `CHECK`, which
    is not a pass.
-   [ ] `0.4.0 feature present` reads `TRUE`. If it reads `FALSE` you are
    reviewing the CRAN 0.3.4 build by mistake, and nothing below means
    anything.

## Where to record what you find

Log findings in `dogfeed.todo.md`, one entry each, status `open`. Do not fix
anything while reviewing. That is the dogfeed protocol, and it exists because
a fix made mid-review loses the observation that prompted it.
