---
title: "0.4.0 review suite: start here"
subtitle: "Read this file first, then work through the numbered files in order"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: false
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

## What this suite is for

surveyframe 0.4.0 is a major release. It adds the whole small-sample track,
all 10 MCDM methods, 2 new item types, conjoint designs, an RStudio add-in,
and it makes 2 breaking changes to how responses are collected and hashed.

The goal of this suite is narrow and absolute: **confirm that every route
from questionnaire to result works, and that every number surveyframe
reports is the number an independent tool reports**, before the tarball goes
to CRAN.

It replaces the single long `mas_review_040.qmd`, which was one document of
roughly 120 chunks. That shape suited neither a reviewer working in sittings
nor a first-time user, so it is deferred rather than deleted. This suite is
written for someone meeting surveyframe for the first time, and each file
stands alone.

## How each file is built

Every file follows the same shape, so once you have run one you know how to
run all of them.

1.  **Questionnaire.** Build a small instrument from scratch. Nothing is
    loaded from a fixture unless the point of the file is the fixture.
2.  **Analysis plan.** Declare the research question before any data exists.
    This is the package's central claim, so every file exercises it.
3.  **HTML route.** Export the static survey and confirm every item reaches
    the file. The path is printed so you can open and click through it.
4.  **Shiny route.** Build every widget and drive the real collector, then
    read the row back. Both routes run in every file, because 0.4.0 fixed a
    defect where the Shiny collector wrote a shape the reader could not read.
5.  **Dummy data and statistical accuracy.** Simulate responses, run the
    plan, and compare each number against base R and against a reference
    package. Every comparison prints a table and a plot.

## Reading a comparison

Every accuracy chunk ends with a table and a plot from the same helper.

```{r}
#| label: demo-comparison
set.seed(1)
x <- rnorm(40, 100, 15)

tab <- compare_table(
  compare_row("mean", mean(x), 100.98, "hand-computed"),
  compare_row("sd",   sd(x),   sd(x),  "base R sd()"),
  compare_row("n",    length(x), 40L,  "by construction")
)
compare_report(tab, caption = "How to read a comparison table")
```

Scan the `verdict` column. Three words can appear.

-   `match`. The difference is at or below the stated tolerance.
-   `DIFFERS`. The difference is above tolerance. **Stop and investigate.**
    Do not tick the box.
-   `CHECK`. One side is `NA`. Usually the reference call failed, but treat
    it as unresolved until you know which.

The plot shows the size of each difference on a log scale, with the
tolerance drawn as a dashed line. It exists so a difference of 1e-16 and a
difference of 0.2 do not look alike in a table of rounded numbers. The first
row above is deliberately wrong, so you can see what `DIFFERS` looks like
before you meet a real one.

::: callout-important
## Tolerance is not a dial to turn

The default tolerance is `1e-6`. If a chunk loosens it, the chunk says why in
a comment. A tolerance loosened without a stated reason is how a real
difference hides, so if you see one, treat the missing reason as the finding.
:::

## A trap worth knowing before you start

Reference calls must match surveyframe's own options, or you will be
comparing 2 different estimators and reading the gap as a bug.

The clearest case is in file `12`. surveyframe fixes `exact = FALSE` on
`wilcox.test()`. Left on its own default, base R switches to the exact method
at small n, and the 2 disagree in the 5th decimal. Nothing is wrong. The
comparison was.

## Order to work through

The suite is complete. Every file renders with 0 errors and every comparison
reads `match`.

**Run in any order, but `01` first if you are new to the package.**

| File | Covers | Comparisons |
|------------------|--------------------------------------|--------|
| `01_first_workflow` | The whole package in one sitting | 17 |
| `02_field_types` | Every item type, one demo each | 64 |
| `03_descriptives` | Frequencies, descriptives, missingness, quality, outliers | 60 |
| `04_categorical` | Chi-square, Fisher, McNemar, Cochran Q | 23 |
| `05_two_group` | t, Mann-Whitney, paired t, Wilcoxon | 28 |
| `06_multi_group` | ANOVA, Kruskal, two-way, ANCOVA, repeated, Friedman | 30 |
| `07_correlation` | Pearson, Spearman, Kendall, partial | 16 |
| `08_regression` | Linear, logistic, ordinal, multinomial, Firth | 32 |
| `09_psychometrics` | Alpha, omega, item diagnostics, EFA, validity | 23 |
| `10_model_syntax` | CFA, CB-SEM, PLS-SEM, and the model-type guard | 50 |
| `11_mediation_moderation` | Indirect effects and interactions | 38 |
| `12_small_sample` | The whole small-sample track | 19 |
| `13_mcdm` | All 10 MCDM methods | 17 |
| `15_sensitivity_conjoint` | Weight sensitivity and conjoint designs | 57 |
| `16_collection_routes` | Static, Shiny, Sheets, branching, the hash | 46 |
| `17_reporting` | Reports, codebook, PDF, interpretations | 46 |

That is 566 checked numbers.

**These 3 need you at a keyboard.** They carry a handful of automated rows
and a great many checkboxes, and the checkboxes are the point.

| File | Covers | What it needs |
|------------------|--------------------------------------|------------------|
| `18_guis` | SurveyBuilder, SurveyStudio, the dashboard | a browser |
| `19_rstudio_addin` | The add-in, task **H2** | RStudio, **release blocker** |
| `20_release_safety` | Pre-flight, `R CMD check`, owner decisions | a terminal |

::: callout-note
## There is no file `14`

The numbering leaves room for a second decision-analysis file that was not
needed once `13` and `15` between them covered the ground.
:::

## What this suite has found

Eight defects, all logged in `dogfeed.todo.md`, none of which any test suite
had caught. None is fixed here: that is the dogfeed protocol, and file `20`
section 7 is where the decisions get recorded.

| # | Finding | Lane | File |
|---|--------------------------------------------------|-------|------|
| 1 | `item_report()` ignores `reverse = TRUE` | 0.4.0 | `09` |
| 2 | Correlation roles: `variables` works for Kendall only | 0.4.0 | `07` |
| 3 | `assumption_report()` reports checks that never ran | 0.4.1 | `05` |
| 4 | Display-only items get a Shiny column, and it counts as missing | 0.4.0 | `02`, `16` |
| 5 | `sem_lavaan_syntax()` writes an indirect effect lavaan cannot parse | 0.4.0 | `10` |
| 6 | Conjoint `"balanced"` is rewarded for dropping a level | 0.4.0 | `15` |
| 7 | `render_results(citation_format = )` is validated then ignored | 0.4.0 | `17` |
| 8 | `codebook_report()` omits the item help text | 0.4.1 | `17` |

The 3 worth reading first:

-   **5** is a hard error. A `cb_sem` model with an indirect effect and
    unlabelled paths generates syntax lavaan refuses to parse, which is the
    single case a structural model is most often declared for.
-   **6** cannot be repaired after collection. A fielded conjoint design with
    a level nobody ever sees is partly inestimable, and `"balanced"` is worse
    than `"random"` at avoiding it.
-   **1** changes numbers the package has already printed for users, which is
    why it needs a decision rather than a fix.

Findings 1 to 3 came from the first sitting. Findings 4 to 8 came from the
second, and 4 of those 5 have the same shape as B11 and B13 before them:
**software returning something plausible instead of saying it has no
answer**. Where a result looks clean, ask whether it is clean or merely
quiet.

## Two ways this suite has been wrong about itself

Both are worth knowing, because both produced a confident `match` that meant
nothing.

1. **A field that does not exist compares equal to itself.** File `13`'s
   hash gate read `hotel$integrity$hash` twice. There is no `$integrity`
   element on an `sframe`, so it compared `NULL` to `NULL`, which is
   `identical()` and therefore a pass, whatever the package did. The hash
   lives in the `.sframe` file's JSON and `read_sframe()` verifies it on the
   way in. Both `13` and `16` now read the stored value and carry a mutation
   check.
2. **A mismatched reference looks exactly like a defect.** surveyframe
   reports the b1 and b2 shape estimators, which is `psych`'s default and
   `e1071`'s type 3. Comparing against `psych::skew(type = 1)`, the obvious
   choice, gives a gap of exactly `((n-1)/n)^1.5` that reads as a bug and is
   not one.

The rule both point at: **every new claim needs a mutation check.** Revert
the thing being tested and confirm the test fails. A check that cannot fail
is not a check.

## Environment this ran on

```{r}
#| label: env
data.frame(
  item = c("R", "platform", "surveyframe loaded from",
           "0.4.0 feature present"),
  value = c(R.version.string, R.version$platform,
            local({
              # Under devtools::load_all() the namespace path IS the package
              # root; for an installed package it is the library
              # subdirectory. Walk up until a DESCRIPTION appears, because
              # dirname(dirname()) covers only the second case and lands 2
              # levels too high on a loaded source tree.
              p <- getNamespaceInfo("surveyframe", "path")
              for (i in 1:3) {
                if (file.exists(file.path(p, "DESCRIPTION"))) break
                p <- dirname(p)
              }
              p
            }),
            as.character(sf_is_040())),
  stringsAsFactors = FALSE
)
```

```{r}
#| label: refs
pk <- c("jmv", "psych", "DescTools", "MASS", "nnet", "logistf", "naniar",
        "lavaan", "seminr", "RMCDA", "car", "e1071")
data.frame(
  package = pk,
  version = vapply(pk, function(p)
    if (requireNamespace(p, quietly = TRUE))
      as.character(utils::packageVersion(p)) else "MISSING", character(1)),
  row.names = NULL
)
```

-   [ ] Every reference package above reports a version rather than
    `MISSING`. A `MISSING` package turns its comparisons into `CHECK`, which
    is not a pass.
-   [ ] `0.4.0 feature present` reads `TRUE`. If it reads `FALSE` you are
    reviewing the CRAN 0.3.4 build by mistake, and nothing below means
    anything.

## Where to record what you find

Log findings in `dogfeed.todo.md`, one entry each, status `open`. Do not fix
anything while reviewing. That is the dogfeed protocol, and it exists because
a fix made mid-review loses the observation that prompted it.
