---
title: "05. Two-group comparisons"
subtitle: "t-test, Mann-Whitney, paired t-test, Wilcoxon signed-rank"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Four methods, each a complete workflow. The field combination is the same
throughout: **one numeric outcome and one 2-level categorical grouping
item** for the independent tests, and **2 numeric items measured on the same
people** for the paired tests.

Every number is checked against base R and against `jmv`.

## 1. Questionnaire and plan

```{r}
#| label: build
trial <- sf_instrument(
  title = "Training evaluation", version = "1.0.0",
  components = list(
    sf_choices("arm", c("control", "training"),
               c("Control group", "Training group")),
    sf_item("arm",  "Which group were you assigned to?",
            type = "single_choice", choice_set = "arm", required = TRUE),
    sf_item("pre",  "Confidence score before the course (0 to 100)",
            type = "slider", slider_min = 0, slider_max = 100),
    sf_item("post", "Confidence score after the course (0 to 100)",
            type = "slider", slider_min = 0, slider_max = 100)
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question =
           "Do the 2 groups differ on post-course confidence?",
         family = "group_comparison", method = "t_test_ind",
         roles = list(outcome = "post", group = "arm")),
    list(id = "RQ2", research_question =
           "Same question without assuming normality.",
         family = "group_comparison", method = "mann_whitney",
         roles = list(outcome = "post", group = "arm")),
    list(id = "RQ3", research_question =
           "Did confidence change from before to after?",
         family = "group_comparison", method = "t_test_pair",
         roles = list(before = "pre", after = "post")),
    list(id = "RQ4", research_question =
           "Same paired question without assuming normality.",
         family = "group_comparison", method = "wilcoxon_pair",
         roles = list(before = "pre", after = "post"))
  )
)
trial
```

::: callout-tip
## The field-to-method map this file demonstrates

| You have | You want | Method |
|-----------------------|-------------------|---------------------------|
| numeric or slider outcome, 2-level single choice | means differ | `t_test_ind` |
| the same, outcome skewed or ordinal | distributions differ | `mann_whitney` |
| 2 numeric items on the same people | change over time | `t_test_pair` |
| the same, not normal | change over time | `wilcoxon_pair` |

Swapping `method` is the only edit. The instrument does not change.
:::

## 2. Both collection routes

```{r}
#| label: routes
set.seed(4)
sf_both_routes(trial, function(i) list(
  arm  = sample(c("control", "training"), 1),
  pre  = sample(20:80, 1), post = sample(30:95, 1)
), n = 4)
```

-   [ ] Both routes report every item, and the Shiny row reads back.

## 3. Dummy data

```{r}
#| label: data
set.seed(101)
n   <- 140
arm <- rep(c("control", "training"), each = n / 2)
pre <- round(rnorm(n, 55, 12))
# The training arm gains more, so all 4 tests have something real to find.
post <- round(pre + ifelse(arm == "training", 12, 3) + rnorm(n, 0, 8))

responses <- data.frame(arm = arm,
                        pre  = pmin(100, pmax(0, pre)),
                        post = pmin(100, pmax(0, post)),
                        stringsAsFactors = FALSE)
dat <- read_responses(responses, trial)
res <- run_analysis_plan(dat, trial)
res
```

## 4. RQ1: independent-samples t-test

```{r}
#| label: acc-t
r  <- res[["RQ1"]]
br <- t.test(post ~ arm, data = dat, var.equal = FALSE)

tab <- compare_table(
  compare_row("t",  r$t,  br$statistic, "base R t.test()"),
  compare_row("df", r$df, br$parameter, "base R t.test()"),
  compare_row("p",  r$p,  br$p.value,   "base R t.test()"),
  compare_row("mean, control",  r$mean1, mean(dat$post[dat$arm == "control"]),
              "base R mean()"),
  compare_row("mean, training", r$mean2, mean(dat$post[dat$arm == "training"]),
              "base R mean()")
)
compare_report(tab, caption = "RQ1 vs base R")
sf_log(tab)
```

::: callout-note
## Sign is a convention, not a disagreement

surveyframe and `t.test()` agree on magnitude, df, and p, and they order the
groups the same way here. If you ever see a t that matches except for its
sign, check the group order before recording a finding. It cost this project
a round of investigation.
:::

```{r}
#| label: acc-t-jmv
j <- jmv::ttestIS(data = data.frame(post = dat$post, arm = factor(dat$arm)),
                  vars = "post", group = "arm",
                  welchs = TRUE, students = FALSE,
                  effectSize = TRUE, desc = TRUE)
jt <- as.data.frame(j$ttest)
jt
```

```{r}
#| label: acc-t-jmv-cmp
# jmv suffixes its columns [welc], and it orders the 2 groups the other way
# round, so its t and d carry the opposite sign. Magnitudes are compared
# here and the sign is checked separately just below, rather than hidden.
tab <- compare_table(
  compare_row("t, magnitude", abs(r$t), abs(jt[["stat[welc]"]][1]),
              "jmv::ttestIS()", tol = 1e-5),
  compare_row("df", r$df, jt[["df[welc]"]][1], "jmv::ttestIS()", tol = 1e-5),
  compare_row("p",  r$p,  jt[["p[welc]"]][1],  "jmv::ttestIS()", tol = 1e-5),
  compare_row("Cohen d, magnitude", abs(r$d), abs(jt[["es[welc]"]][1]),
              "jmv::ttestIS()", tol = 1e-4)
)
compare_report(tab, caption = "RQ1 against a third opinion")
sf_log(tab)

# The signs must be opposite, because the 2 tools order the groups
# differently. Same sign would mean one of them has the direction wrong.
data.frame(
  quantity = c("t", "Cohen d"),
  surveyframe = c(r$t, r$d),
  jmv = c(jt[["stat[welc]"]][1], jt[["es[welc]"]][1]),
  opposite_signs = c(sign(r$t) != sign(jt[["stat[welc]"]][1]),
                     sign(r$d) != sign(jt[["es[welc]"]][1]))
)
```

```{r}
#| label: acc-t-plot
boxplot(post ~ arm, data = dat, col = c("#deebf7", "#9ecae1"),
        border = "#3182bd", ylab = "Post-course confidence",
        main = sprintf("RQ1: t(%.1f) = %.2f", r$df, r$t))
points(c(1, 2), c(r$mean1, r$mean2), pch = 18, cex = 2, col = "#b2182b")
legend("topleft", pch = 18, col = "#b2182b", bty = "n",
       legend = "means as surveyframe reports them")
```

-   [ ] Every verdict reads `match`, including Cohen's d against `jmv`.
-   [ ] The red diamonds sit where the boxes say they should. If a reported
    mean lands outside its own box, the grouping is reversed somewhere.

## 5. RQ2: Mann-Whitney

```{r}
#| label: acc-mw
r  <- res[["RQ2"]]
a  <- dat$post[dat$arm == "control"]
b  <- dat$post[dat$arm == "training"]
# surveyframe fixes exact = FALSE, so the reference must too. See file 12.
w  <- wilcox.test(a, b, conf.int = TRUE, exact = FALSE)

tab <- compare_table(
  compare_row("U",  r$U, unname(w$statistic), "base R wilcox.test()"),
  compare_row("p",  r$p, w$p.value,           "base R wilcox.test()"),
  compare_row("Hodges-Lehmann shift", r$hl_shift, unname(w$estimate),
              "base R wilcox.test()"),
  compare_row("HL lower", r$hl_conf_int[1], w$conf.int[1],
              "base R wilcox.test()"),
  compare_row("HL upper", r$hl_conf_int[2], w$conf.int[2],
              "base R wilcox.test()")
)
compare_report(tab, caption = "RQ2 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-mw-plot
plot(density(a), col = "#3182bd", lwd = 2, xlim = range(c(a, b)),
     main = "RQ2: the 2 distributions being compared",
     xlab = "Post-course confidence")
lines(density(b), col = "#b2182b", lwd = 2)
abline(v = c(median(a), median(b)), col = c("#3182bd", "#b2182b"), lty = 2)
legend("topright", lwd = 2, col = c("#3182bd", "#b2182b"), bty = "n",
       legend = c("control", "training"))
```

-   [ ] U, p, and all 3 Hodges-Lehmann numbers match.
-   [ ] The 2 curves are shifted in the direction the APA sentence claims.

## 6. RQ3: paired t-test

```{r}
#| label: acc-tp
r  <- res[["RQ3"]]
bp <- t.test(dat$pre, dat$post, paired = TRUE)

tab <- compare_table(
  compare_row("t",  r$t,  bp$statistic, "base R t.test(paired)"),
  compare_row("df", r$df, bp$parameter, "base R t.test(paired)"),
  compare_row("p",  r$p,  bp$p.value,   "base R t.test(paired)"),
  compare_row("mean difference", r$mean_diff, unname(bp$estimate),
              "base R t.test(paired)"),
  compare_row("Cohen dz", r$d_z,
              unname(bp$statistic) / sqrt(nrow(dat)), "hand-computed")
)
compare_report(tab, caption = "RQ3 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-tp-plot
d <- dat$post - dat$pre
hist(d, breaks = 20, col = "#c7e9c0", border = "white",
     xlab = "post minus pre", main = "RQ3: the differences being tested")
abline(v = 0, lty = 2)
abline(v = mean(d), col = "#b2182b", lwd = 2)
```

-   [ ] All 5 verdicts read `match`.
-   [ ] The red line is the mean difference the table reports, and 0 is
    where the null sits.

## 7. RQ4: Wilcoxon signed-rank

```{r}
#| label: acc-wp
r  <- res[["RQ4"]]
ws <- wilcox.test(dat$pre, dat$post, paired = TRUE, conf.int = TRUE,
                  exact = FALSE)

tab <- compare_table(
  compare_row("V", r$V, unname(ws$statistic), "base R wilcox.test(paired)"),
  compare_row("p", r$p, ws$p.value,           "base R wilcox.test(paired)"),
  compare_row("pseudomedian", r$pseudomedian, unname(ws$estimate),
              "base R wilcox.test(paired)"),
  compare_row("pseudomedian lower", r$pseudomedian_conf_int[1],
              ws$conf.int[1], "base R wilcox.test(paired)"),
  compare_row("pseudomedian upper", r$pseudomedian_conf_int[2],
              ws$conf.int[2], "base R wilcox.test(paired)")
)
compare_report(tab, caption = "RQ4 vs base R")
sf_log(tab)
```

::: callout-note
## New in 0.4.0

The pseudomedian and its interval are part of the small-sample track and did
not exist before this release. No earlier review file covers them. File `12`
takes the track as a whole.
:::

-   [ ] All 5 verdicts read `match`.

## 8. Assumption checking

```{r}
#| label: assumptions
# Every check is driven by `variables`. Passing only `outcome` and `group`
# returns empty tables and still prints "Assumption checks were computed.",
# which is logged in dogfeed.todo.md. Pass `variables`.
ar <- assumption_report(dat, variables = "post", group = "arm")
ar
```

```{r}
#| label: assumptions-acc
sw <- shapiro.test(dat$post)
# car::leveneTest() centres on the median by default, which is the
# Brown-Forsythe test. surveyframe reports both, and its row labelled
# "Levene" is the mean-centred classic. Matching the wrong row here would
# look like a defect and would only be a mismatched reference.
lv_med  <- car::leveneTest(post ~ factor(arm), data = dat)
lv_mean <- car::leveneTest(post ~ factor(arm), data = dat, center = mean)
hg <- ar$homogeneity

tab <- compare_table(
  compare_row("Shapiro-Wilk W", ar$normality$shapiro_w, sw$statistic,
              "base R shapiro.test()"),
  compare_row("Shapiro-Wilk p", ar$normality$shapiro_p, sw$p.value,
              "base R shapiro.test()"),
  compare_row("Levene p (mean-centred)",
              hg$p[hg$test == "Levene"], lv_mean[["Pr(>F)"]][1],
              "car::leveneTest(center = mean)", tol = 1e-6),
  compare_row("Brown-Forsythe p (median-centred)",
              hg$p[hg$test == "Brown-Forsythe"], lv_med[["Pr(>F)"]][1],
              "car::leveneTest() default", tol = 1e-6)
)
compare_report(tab, caption = "Assumption tests vs their reference")
sf_log(tab)
```

-   [ ] All 4 assumption numbers match their reference.
-   [ ] **Worth knowing.** surveyframe reports Levene and Brown-Forsythe as
    2 rows. `car::leveneTest()` centres on the median by default, so its
    default output is the Brown-Forsythe row, not the Levene row. Comparing
    it to the Levene row would show a false `DIFFERS`.
-   [ ] The advisory, if one fired, points at the right alternative method.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] The 4 APA sentences are usable in a write-up without editing.
