---
title: "07. Correlation"
subtitle: "Pearson, Spearman, Kendall, partial correlation"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Four methods over the same field combination: **2 numeric, slider, rating, or
Likert items**, plus one more for the partial correlation to control for.

::: callout-warning
## Read this before writing your own plan

The correlation methods take the roles `x` and `y`. They do **not** accept
`variables`, although 9 other methods do and `correlation_kendall` does. A
block written with `variables` returns `Error: Test failed.` for Pearson and
Spearman, and `undefined columns selected` for partial correlation. This is
logged in `dogfeed.todo.md` and needs an owner decision before release.
:::

## 1. Questionnaire and plan

```{r}
#| label: build
wb <- sf_instrument(
  title = "Wellbeing and study habits", version = "1.0.0",
  components = list(
    sf_item("sleep",  "Average hours of sleep per night", type = "numeric"),
    sf_item("stress", "Stress level this week (0 to 100)",
            type = "slider", slider_min = 0, slider_max = 100),
    sf_item("grade",  "Most recent assessment mark (0 to 100)",
            type = "numeric"),
    sf_item("hours",  "Hours studied last week", type = "numeric")
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question = "Does sleep relate to stress?",
         family = "association", method = "correlation_pearson",
         roles = list(x = "sleep", y = "stress")),
    list(id = "RQ2", research_question = "The same on ranks.",
         family = "association", method = "correlation_spearman",
         roles = list(x = "sleep", y = "stress")),
    list(id = "RQ3", research_question = "The same with Kendall's tau.",
         family = "association", method = "correlation_kendall",
         roles = list(x = "sleep", y = "stress")),
    list(id = "RQ4", research_question =
           "Do study hours relate to grade once stress is controlled?",
         family = "association", method = "partial_correlation",
         roles = list(x = "hours", y = "grade", controls = "stress"))
  )
)
wb
```

## 2. Both collection routes

```{r}
#| label: routes
set.seed(33)
sf_both_routes(wb, function(i) list(
  sleep = round(runif(1, 4, 9), 1), stress = sample(0:100, 1),
  grade = sample(30:95, 1), hours = sample(0:40, 1)
), n = 4)
```

## 3. Dummy data

```{r}
#| label: data
set.seed(555)
n      <- 200
sleep  <- round(rnorm(n, 6.8, 1.1), 1)
# Stress falls as sleep rises, so RQ1 to RQ3 have a real negative relation.
stress <- round(pmin(100, pmax(0, 95 - 7 * sleep + rnorm(n, 0, 12))))
hours  <- round(pmax(0, rnorm(n, 14, 6)))
# Grade rises with hours and falls with stress, so the partial correlation
# is genuinely smaller than the raw one.
grade  <- round(pmin(100, pmax(0, 45 + 1.1 * hours - 0.20 * stress +
                                 rnorm(n, 0, 8))))

dat <- read_responses(
  data.frame(sleep = sleep, stress = stress, grade = grade, hours = hours),
  wb)
res <- run_analysis_plan(dat, wb)
res
```

## 4. RQ1: Pearson

```{r}
#| label: acc-pearson
r  <- res[["RQ1"]]
ct <- cor.test(dat$sleep, dat$stress, method = "pearson")

tab <- compare_table(
  compare_row("r",        r$r,  unname(ct$estimate),  "base R cor.test()"),
  compare_row("df",       r$df, unname(ct$parameter), "base R cor.test()"),
  compare_row("p",        r$p,  ct$p.value,           "base R cor.test()"),
  compare_row("CI lower", r$ci[["lower"]], ct$conf.int[1], "base R cor.test()"),
  compare_row("CI upper", r$ci[["upper"]], ct$conf.int[2], "base R cor.test()")
)
compare_report(tab, caption = "RQ1 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-pearson-jmv
j  <- jmv::corrMatrix(data = dat[, c("sleep", "stress")],
                      vars = c("sleep", "stress"), pearson = TRUE,
                      spearman = TRUE, kendall = TRUE)
jm <- as.data.frame(j$matrix)
jm
```

```{r}
#| label: acc-pearson-jmv-cmp
# jmv returns the matrix as characters with an em-rule on the diagonal, so
# the off-diagonal cell has to be pulled out and coerced. Row 2 is stress,
# column "sleep[r]" is its correlation with sleep.
jmv_cell <- function(col) as.numeric(jm[[col]][2])

tab <- compare_table(
  compare_row("Pearson r", r$r, jmv_cell("sleep[r]"), "jmv::corrMatrix()",
              tol = 1e-6),
  compare_row("Pearson p", r$p, jmv_cell("sleep[rp]"), "jmv::corrMatrix()",
              tol = 1e-8)
)
compare_report(tab, caption = "RQ1 against a third opinion")
sf_log(tab)
```

```{r}
#| label: acc-pearson-plot
plot(dat$sleep, dat$stress, pch = 19, col = "#41618980",
     xlab = "hours of sleep", ylab = "stress",
     main = sprintf("RQ1: r = %.3f [%.2f, %.2f]",
                    r$r, r$ci[["lower"]], r$ci[["upper"]]))
abline(lm(stress ~ sleep, data = dat), col = "#b2182b", lwd = 2)
```

-   [ ] All 5 base R verdicts and both `jmv` verdicts read `match`.
-   [ ] The line slopes the way r says. A positive r with a falling line
    means the variables were passed the wrong way round somewhere.

## 5. RQ2 and RQ3: Spearman and Kendall

```{r}
#| label: acc-rank
rs <- res[["RQ2"]]
rk <- res[["RQ3"]]
cs <- suppressWarnings(cor.test(dat$sleep, dat$stress, method = "spearman"))
ck <- suppressWarnings(cor.test(dat$sleep, dat$stress, method = "kendall"))

tab <- compare_table(
  compare_row("Spearman rho", rs$r, unname(cs$estimate), "base R cor.test()"),
  compare_row("Spearman p",   rs$p, cs$p.value,          "base R cor.test()"),
  compare_row("Kendall tau",  rk$r, unname(ck$estimate), "base R cor.test()"),
  compare_row("Kendall p",    rk$p, ck$p.value,          "base R cor.test()")
)
compare_report(tab, caption = "RQ2 and RQ3 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-rank-jmv
tab <- compare_table(
  compare_row("Spearman rho", rs$r, jmv_cell("sleep[rho]"),
              "jmv::corrMatrix()", tol = 1e-6),
  compare_row("Kendall tau",  rk$r, jmv_cell("sleep[tau]"),
              "jmv::corrMatrix()", tol = 1e-6)
)
compare_report(tab, caption = "Rank correlations against a third opinion")
sf_log(tab)
```

::: callout-note
## Why the rank intervals are not compared here

Pearson's interval is analytic, from the Fisher z transform, so it can be
checked against `cor.test()` exactly. The rank correlations bootstrap their
interval instead, so it is not reproducible to `1e-6` and there is no base R
number to compare it to. Check that the interval brackets the estimate and
is a sensible width, rather than looking for a `match`.
:::

```{r}
#| label: acc-rank-ci
data.frame(
  method   = c("Spearman", "Kendall"),
  estimate = c(rs$r, rk$r),
  lower    = c(rs$ci[["lower"]], rk$ci[["lower"]]),
  upper    = c(rs$ci[["upper"]], rk$ci[["upper"]]),
  brackets_the_estimate = c(
    rs$r >= rs$ci[["lower"]] && rs$r <= rs$ci[["upper"]],
    rk$r >= rk$ci[["lower"]] && rk$r <= rk$ci[["upper"]])
)
```

-   [ ] All 4 base R verdicts and both `jmv` verdicts read `match`.
-   [ ] Both bootstrap intervals bracket their own estimate.

## 6. RQ4: partial correlation

```{r}
#| label: acc-partial
r <- res[["RQ4"]]

# Partial correlation by hand: correlate the residuals of both variables
# after regressing each on the control.
rx <- residuals(lm(hours ~ stress, data = dat))
ry <- residuals(lm(grade ~ stress, data = dat))
pc <- cor.test(rx, ry)

raw <- cor(dat$hours, dat$grade)

tab <- compare_table(
  compare_row("partial r", r$r, unname(pc$estimate),
              "hand-computed from residuals"),
  compare_row("p",         r$p, pc$p.value,
              "hand-computed from residuals", tol = 1e-8),
  compare_row("n",         r$n, nrow(dat), "nrow()")
)
compare_report(tab, caption = "RQ4 vs a hand-computed partial correlation")
sf_log(tab)
```

```{r}
#| label: acc-partial-plot
barplot(c("raw r" = raw, "partial r" = r$r),
        col = c("#bdbdbd", "#3182bd"), border = NA, ylim = c(0, 1),
        ylab = "correlation of hours with grade",
        main = "RQ4: what controlling for stress removes")
```

-   [ ] All 3 verdicts read `match`.
-   [ ] The partial correlation is smaller than the raw one, which is the
    point of running it. If they are identical, the control is not being
    applied.
-   [ ] `r$controls` names `stress`, so the write-up can say what was
    controlled for.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
