---
title: "09. Psychometrics and factor analysis"
subtitle: "Alpha, omega, item diagnostics, EFA readiness, EFA solution, validity"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Everything here takes the same field combination: **a battery of Likert or
rating items grouped into one or more scales**. Section 5 proves a fix for a
bug that is live on CRAN right now.

## 1. Questionnaire and plan

Two constructs, 5 items each, so discriminant validity has something to
separate.

```{r}
#| label: build
lik <- function(id, label, scale, rev = FALSE)
  sf_item(id, label, type = "likert", choice_set = "a5",
          scale_id = scale, reverse = rev)

psy <- sf_instrument(
  title = "Learner motivation and anxiety", version = "1.0.0",
  components = c(
    list(sf_choices("a5", 1:5,
                    c("Strongly disagree", "Disagree", "Neutral",
                      "Agree", "Strongly agree"))),
    list(
      lik("mot1", "I enjoy the challenge of this subject.", "motivation"),
      lik("mot2", "I want to understand this topic deeply.", "motivation"),
      lik("mot3", "I put effort into this module.", "motivation"),
      lik("mot4", "I look forward to the classes.", "motivation"),
      lik("mot5", "I find this module a chore.", "motivation", rev = TRUE),
      lik("anx1", "I feel tense before assessments.", "anxiety"),
      lik("anx2", "I worry about failing this module.", "anxiety"),
      lik("anx3", "I feel calm during tests.", "anxiety", rev = TRUE),
      lik("anx4", "My mind goes blank under pressure.", "anxiety"),
      lik("anx5", "I dread the exam period.", "anxiety")
    ),
    list(
      sf_scale("motivation", "Motivation",
               items = paste0("mot", 1:5)),
      sf_scale("anxiety", "Test anxiety", items = paste0("anx", 1:5))
    )
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question = "Are the 2 scales reliable?",
         family = "psychometrics", method = "reliability_alpha"),
    list(id = "RQ2", research_question = "And by omega?",
         family = "psychometrics", method = "reliability_omega"),
    list(id = "RQ3", research_question = "Does any item weaken its scale?",
         family = "psychometrics", method = "item_diagnostics"),
    list(id = "RQ4", research_question = "Is this data suitable for EFA?",
         family = "psychometrics", method = "efa_readiness"),
    list(id = "RQ5", research_question = "Do 2 factors emerge as designed?",
         family = "psychometrics", method = "efa_solution",
         roles = list(items = c(paste0("mot", 1:5), paste0("anx", 1:5))),
         options = list(nfactors = 2))
  )
)
psy
```

## 2. Both collection routes

```{r}
#| label: routes
set.seed(12)
ids <- c(paste0("mot", 1:5), paste0("anx", 1:5))
sf_both_routes(psy, function(i)
  stats::setNames(as.list(sample(1:5, 10, TRUE)), ids), n = 4)
```

## 3. Dummy data with 2 real factors

```{r}
#| label: data
set.seed(2468)
n   <- 300
mot <- rnorm(n)
anx <- -0.3 * mot + rnorm(n, 0, 0.95)   # mildly negatively related

mk <- function(latent, load, reverse = FALSE) {
  v <- 3 + load * latent + rnorm(n, 0, 0.75)
  v <- pmin(5, pmax(1, round(v)))
  if (reverse) 6 - v else v
}

responses <- data.frame(
  mot1 = mk(mot, 0.9), mot2 = mk(mot, 0.85), mot3 = mk(mot, 0.8),
  mot4 = mk(mot, 0.75), mot5 = mk(mot, 0.8, reverse = TRUE),
  anx1 = mk(anx, 0.85), anx2 = mk(anx, 0.9), anx3 = mk(anx, 0.7, reverse = TRUE),
  anx4 = mk(anx, 0.8), anx5 = mk(anx, 0.75)
)
dat <- read_responses(responses, psy)
res <- run_analysis_plan(dat, psy)
res
```

## 4. RQ1 and RQ2: alpha and omega against `psych`

```{r}
#| label: acc-alpha
r1 <- res[["RQ1"]]
r2 <- res[["RQ2"]]

# The instrument declares mot5 and anx3 reversed, so the reference must
# reverse them too. check.keys = FALSE stops psych choosing its own keying
# and quietly making the 2 incomparable.
rev6 <- function(x) 6 - x
mot_d <- data.frame(dat[, paste0("mot", 1:4)], mot5 = rev6(dat$mot5))
anx_d <- data.frame(dat[, c("anx1", "anx2")], anx3 = rev6(dat$anx3),
                    dat[, c("anx4", "anx5")])

pa_m <- psych::alpha(mot_d, check.keys = FALSE, warnings = FALSE)
pa_a <- psych::alpha(anx_d, check.keys = FALSE, warnings = FALSE)

tab <- compare_table(
  compare_row("alpha, motivation", r1$motivation$alpha,
              pa_m$total$raw_alpha, "psych::alpha()"),
  compare_row("alpha, anxiety", r1$anxiety$alpha,
              pa_a$total$raw_alpha, "psych::alpha()"),
  compare_row("standardised alpha, motivation", r1$motivation$alpha_std,
              pa_m$total$std.alpha, "psych::alpha()")
)
compare_report(tab, caption = "RQ1 alpha vs psych")
sf_log(tab)
```

```{r}
#| label: acc-omega
om_m <- psych::omega(mot_d, nfactors = 1, plot = FALSE, warnings = FALSE)

tab <- compare_table(
  compare_row("omega total, motivation", r2$motivation$omega_t,
              om_m$omega.tot, "psych::omega()", tol = 1e-3)
)
compare_report(tab, caption = "RQ2 omega vs psych")
sf_log(tab)
```

```{r}
#| label: acc-alpha-plot
plot(reliability_report(dat, psy))
```

-   [ ] All 3 alpha verdicts read `match`.
-   [ ] Omega matches to `1e-3`. The tolerance is loosened here because
    `psych::omega()` fits a factor model and its optimiser is not bit-stable
    across versions. That is the stated reason, so the loosening is
    accountable rather than convenient.
-   [ ] Both scales clear 0.7, which they should, because the data was built
    with a single strong factor each.

## 5. RQ3: item diagnostics, and the bug this fixes

::: callout-important
## Live on CRAN right now

In 0.3.4 `item_report()` computed the item-rest correlation with
`rowMeans()` where the standard needs the **sum** of the other items
(`R/psychometrics.R:179`). On a scale with alpha 0.88 it returned
**negative** item-rest correlations, which would tell a researcher to delete
good items.

`psych::alpha()$item.stats$r.drop` is the independent oracle. If these
verdicts read `match`, the fix is real.
:::

```{r}
#| label: acc-itemrest
r3 <- res[["RQ3"]]
dg <- r3$motivation$diagnostics

# psych, run twice: once on the raw columns and once with mot5 reversed as
# the instrument declares. Only one of these should be the right reference.
ps_raw <- psych::alpha(dat[, paste0("mot", 1:5)],
                       check.keys = FALSE, warnings = FALSE)$item.stats
ps_rev <- pa_m$item.stats

data.frame(
  item                      = dg$item_id,
  surveyframe               = round(dg$item_rest_r, 4),
  psych_raw                 = round(ps_raw$r.drop, 4),
  psych_reverse_as_declared = round(ps_rev$r.drop, 4)
)
```

::: callout-warning
## Defect found while writing this file, 2026-08-02

surveyframe matches the **raw** column exactly and the **reversed** column
not at all. `item_report()` never applies `reverse = TRUE`, although
`score_scales()` and `reliability_report()` both do. So this scale reports
alpha 0.84 and, at the same time, a strongly negative item-rest correlation
for `mot5`.

Two things follow. The reverse-keyed item looks like a deletion candidate
and is not. And every other item in the scale is depressed as well, because
the unreversed item sits inside the rest-sum pulling against them, so an
item near a 0.3 retention threshold can be dropped on a keying error.

This is the same symptom A2 fixed in this release, reached by a different
route. A2 corrected the formula. The keying was never in scope. Logged in
`dogfeed.todo.md`, and it needs an owner decision before release.
:::

```{r}
#| label: acc-itemrest-cmp
# Against the raw reference, which is what surveyframe currently computes.
# This confirms the A2 formula fix is correct and intact: the sum of the
# other items, not rowMeans().
tab <- do.call(compare_table, lapply(seq_len(nrow(dg)), function(i) {
  compare_row(paste("item-rest r,", dg$item_id[i]),
              dg$item_rest_r[i], ps_raw$r.drop[i],
              "psych r.drop, raw columns")
}))
compare_report(tab, caption = "RQ3: the A2 formula fix, checked on raw columns")
sf_log(tab)
```

```{r}
#| label: acc-itemrest-plot
op <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
barplot(dg$item_rest_r, names.arg = dg$item_id, col = "#4292c6", border = NA,
        ylim = c(-1, 1), ylab = "item-rest correlation",
        main = "As surveyframe reports it")
abline(h = c(0, 0.3), lty = c(1, 2), col = c("#404040", "#b2182b"))
barplot(ps_rev$r.drop, names.arg = dg$item_id, col = "#1b7837", border = NA,
        ylim = c(-1, 1), ylab = "item-rest correlation",
        main = "With the declared reversal applied")
abline(h = c(0, 0.3), lty = c(1, 2), col = c("#404040", "#b2182b"))
par(op)
```

-   [ ] Every verdict reads `match` against the raw reference. That is the
    A2 formula fix confirmed: the sum of the other items rather than
    `rowMeans()`.
-   [ ] **The left plot has one bar below zero and the right plot has none.**
    That gap is the keying defect, not the A2 bug. Confirm you can see it
    before recording a decision.
-   [ ] **Owner decision.** Fix the keying in 0.4.0, which changes numbers
    `item_report()` currently reports and so belongs in NEWS as breaking, or
    lane it to 0.4.1 and document the limitation.
-   [ ] The floor and ceiling percentages in `dg` are plausible for a
    5-point scale.

## 6. RQ4: EFA readiness against `psych`

```{r}
#| label: acc-efa-ready
r4  <- res[["RQ4"]]
all_items <- c(paste0("mot", 1:5), paste0("anx", 1:5))
km  <- psych::KMO(dat[, all_items])
bt  <- psych::cortest.bartlett(cor(dat[, all_items]), n = nrow(dat))

tab <- compare_table(
  compare_row("KMO overall MSA", r4$kmo$MSA, km$MSA, "psych::KMO()"),
  compare_row("Bartlett chi-square", r4$bartlett$chisq, bt$chisq,
              "psych::cortest.bartlett()"),
  compare_row("Bartlett df", r4$bartlett$df, bt$df,
              "psych::cortest.bartlett()"),
  compare_row("Bartlett p", r4$bartlett$p.value, bt$p.value,
              "psych::cortest.bartlett()", tol = 1e-10)
)
compare_report(tab, caption = "RQ4 EFA readiness vs psych")
sf_log(tab)
```

```{r}
#| label: acc-efa-ready-plot
barplot(r4$kmo$MSAi, las = 2, col = "#66c2a5", border = NA, ylim = c(0, 1),
        ylab = "MSA", main = "RQ4: per-item sampling adequacy")
abline(h = 0.5, lty = 2, col = "#b2182b")
```

-   [ ] All 4 verdicts read `match`.
-   [ ] Every per-item MSA clears the 0.5 line.
-   [ ] `r4$suggested_nfactors` says 2, which is how the data was built. If
    it says something else, say so, because the parallel analysis is the
    part a user will trust without checking.

## 7. RQ5: the EFA solution

```{r}
#| label: acc-efa
r5 <- res[["RQ5"]]
# loadings carries an item_id column alongside the numeric factor columns.
r5$loadings
```

```{r}
#| label: acc-efa-cmp
pf <- psych::fa(dat[, all_items], nfactors = 2, rotate = r5$rotation,
                fm = "minres")
# Factor order and sign are arbitrary in any EFA, so the comparable quantity
# is the communality of each item, which is invariant to both.
tab <- do.call(compare_table, lapply(all_items, function(id) {
  compare_row(paste("communality,", id),
              r5$communalities[[id]],
              unname(pf$communality[[id]]),
              "psych::fa()", tol = 1e-3)
}))
compare_report(tab, caption = "RQ5 communalities vs psych::fa()")
sf_log(tab)
```

```{r}
#| label: acc-efa-plot
# The first column is item_id, so the factor columns are everything after it.
fac <- setdiff(names(r5$loadings), "item_id")
L   <- as.matrix(r5$loadings[, fac])
rownames(L) <- r5$loadings$item_id

plot(L[, 1], L[, 2], type = "n", xlim = c(-1, 1), ylim = c(-1, 1),
     xlab = fac[1], ylab = fac[2],
     main = "RQ5: do the 2 designed constructs separate?")
abline(h = 0, v = 0, col = "#cccccc")
text(L[, 1], L[, 2], labels = rownames(L),
     col = ifelse(grepl("^mot", rownames(L)), "#1b7837", "#b2182b"))
```

-   [ ] Every communality matches `psych::fa()` to `1e-3`.
-   [ ] **The plot is the real check.** The green `mot` labels cluster on one
    axis and the red `anx` labels on the other. If they mix, either the data
    or the rotation is not doing what the section claims.
-   [ ] `r5$variance_table` and `r5$item_flags` both print and are readable.

## 8. Validity

```{r}
#| label: validity
# validity_report() wants a long frame with `construct` and `loading`.
# Each item is assigned to the factor it loads most strongly on, and only
# that loading is used, which is what average variance extracted and
# composite reliability are defined on.
long <- r5$loadings_long
best <- do.call(rbind, lapply(split(long, long$item_id), function(g)
  g[which.max(abs(g$loading)), ]))
best$construct <- best$factor

vr <- validity_report(
  loadings = best[, c("item_id", "construct", "loading")],
  construct_scores = score_scales(dat, psy)[, c("motivation", "anxiety")],
  items_by_construct = list(motivation = paste0("mot", 1:5),
                            anxiety = paste0("anx", 1:5))
)
vr
```

```{r}
#| label: validity-assignment
# Which extracted factor each declared construct landed on. If the mot
# items and the anx items do not separate cleanly here, the numbers above
# are describing something other than the 2 designed constructs.
table(declared = ifelse(grepl("^mot", best$item_id), "motivation", "anxiety"),
      loaded_on = best$construct)
```

```{r}
#| label: validity-plot
plot(vr)
```

-   [ ] Average variance extracted and composite reliability print for both
    constructs.
-   [ ] The heterotrait-monotrait ratio is below the usual 0.85 threshold,
    which it should be, because the 2 constructs were built to be distinct.
-   [ ] **Owner judgement.** Would a first-time user know what to do with a
    failing HTMT from what is printed?

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] Section 5 in particular is clean, since that fix is the reason 0.4.0
    changes numbers earlier versions reported.
