---
title: "10. Measurement and structural model syntax"
subtitle: "cfa_lavaan_syntax, sem_lavaan_syntax, seminr_syntax, and the model-type guard"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

surveyframe does not fit structural equation models. It generates the syntax
for the package that does, from a model declared inside the instrument before
any data existed. That is the whole claim of this layer, and it is only worth
anything if the generated syntax estimates **the model that was declared**
rather than something close to it.

So the accuracy check in this file is not "does a number match a number". It
is: fit the generated syntax, fit a hand-written syntax for the same declared
model, and confirm the 2 fits are the same model to the last decimal. If they
are not, the generator has quietly changed the analysis.

## 1. Questionnaire

```{r}
#| label: build-items
lik <- function(id, scale) sf_item(
  id, sprintf("Item %s", id), type = "likert",
  choice_set = "agree5", scale_id = scale)

loyalty <- sf_instrument(
  title = "Service quality and loyalty", version = "1.0.0",
  components = c(
    list(sf_choices("agree5", as.character(1:5),
                    c("Strongly disagree", "Disagree", "Neutral", "Agree",
                      "Strongly agree"))),
    lapply(c("q1", "q2", "q3"), lik, scale = "qual"),
    lapply(c("v1", "v2", "v3"), lik, scale = "val"),
    lapply(c("l1", "l2", "l3"), lik, scale = "loy"),
    list(
      sf_scale("qual", "Perceived quality", items = c("q1", "q2", "q3")),
      sf_scale("val",  "Perceived value",   items = c("v1", "v2", "v3")),
      sf_scale("loy",  "Loyalty",           items = c("l1", "l2", "l3"))
    )
  )
)
loyalty
```

## 2. Three declared models

The same 9 items support 3 different models, and the point of declaring them
is that the package will refuse to generate the wrong syntax for any of them.

```{r}
#| label: build-models
constructs_reflective <- function() list(
  sf_construct("qual", label = "Perceived quality",
               items = c("q1", "q2", "q3"), mode = "reflective"),
  sf_construct("val",  label = "Perceived value",
               items = c("v1", "v2", "v3"), mode = "reflective"),
  sf_construct("loy",  label = "Loyalty",
               items = c("l1", "l2", "l3"), mode = "reflective")
)

m_cfa <- sf_model(
  "m_cfa", "Measurement model", type = "cfa",
  constructs = constructs_reflective()
)

m_cb <- sf_model(
  "m_cb", "Structural model, covariance-based", type = "cb_sem",
  constructs = constructs_reflective(),
  # Every path carries a label. That is not decoration, and section 7
  # explains why: without labels this model generates syntax lavaan cannot
  # parse. The unlabelled version is built there too, so you can see it.
  paths = list(
    sf_path("qual", "val", label = "a"),
    sf_path("val",  "loy", label = "b"),
    sf_path("qual", "loy", label = "c")
  ),
  indirect = list(sf_indirect("qual", through = "val", to = "loy"))
)

# The same model with the label argument left off, which is what the
# 2-argument call in every example produces.
m_cb_unlabelled <- sf_model(
  "m_cb_plain", "Structural model, paths unlabelled", type = "cb_sem",
  constructs = constructs_reflective(),
  paths = list(
    sf_path("qual", "val"),
    sf_path("val",  "loy"),
    sf_path("qual", "loy")
  ),
  indirect = list(sf_indirect("qual", through = "val", to = "loy"))
)

m_pls <- sf_model(
  "m_pls", "Structural model, partial least squares", type = "pls_sem",
  constructs = list(
    sf_construct("qual", items = c("q1", "q2", "q3"), mode = "composite"),
    sf_construct("val",  items = c("v1", "v2", "v3"), mode = "composite"),
    sf_construct("loy",  items = c("l1", "l2", "l3"), mode = "reflective")
  ),
  paths = list(sf_path("qual", "val"), sf_path("val", "loy"))
)

loyalty <- add_model(loyalty, m_cfa)
loyalty <- add_model(loyalty, m_cb)
loyalty <- add_model(loyalty, m_pls)

data.frame(
  id     = vapply(loyalty$models, function(m) m$id, character(1)),
  type   = vapply(loyalty$models, function(m) m$type, character(1)),
  engine = vapply(loyalty$models, function(m) m$engine %||% "", character(1)),
  row.names = NULL
)
```

-   [ ] Each model picked up the right engine from its type without being
    told: `lavaan` for CFA and CB-SEM, `seminr` for PLS-SEM.

```{r}
#| label: validate-models
data.frame(
  model = c("m_cfa", "m_cb", "m_pls"),
  validates = vapply(list(m_cfa, m_cb, m_pls), function(m)
    isTRUE(tryCatch({ validate_model(m, loyalty); TRUE },
                    error = function(e) FALSE)), logical(1)),
  row.names = NULL
)
```

## 3. Analysis plan

```{r}
#| label: build-plan
loyalty$analysis_plan <- list(
  list(id = "RQ1", research_question =
         "Do the 9 items measure the 3 constructs they claim to?",
       family = "model", method = "cfa_lavaan_syntax",
       roles = list(model = "m_cfa")),
  list(id = "RQ2", research_question =
         "Does quality raise loyalty, and does value carry part of it?",
       family = "model", method = "sem_lavaan_syntax",
       roles = list(model = "m_cb")),
  list(id = "RQ3", research_question =
         "The same structural question, estimated by partial least squares.",
       family = "model", method = "seminr_syntax",
       roles = list(model = "m_pls"))
)
v <- tryCatch(validate_sframe(loyalty), error = function(e) e)
if (inherits(v, "error")) conditionMessage(v) else
  list(problems = v$problems %||% "none")
```

::: callout-tip
## The field-to-method map this file demonstrates

| You have | You want | Method |
|-------------------------|---------------------|-----------------------|
| a declared `cfa` model | measurement syntax | `cfa_lavaan_syntax` |
| a declared `cb_sem` model | structural syntax for lavaan | `sem_lavaan_syntax` |
| a declared `pls_sem` model | syntax for seminr | `seminr_syntax` |
| declared scales, no model | quick measurement syntax | `cfa_syntax()` |
| no model at all | an exploratory starting point | `efa_syntax()` |
:::

## 4. Both collection routes

```{r}
#| label: routes
set.seed(10)
sf_both_routes(loyalty, function(i) {
  a <- function() as.character(sample(1:5, 1))
  list(q1 = a(), q2 = a(), q3 = a(), v1 = a(), v2 = a(), v3 = a(),
       l1 = a(), l2 = a(), l3 = a())
}, n = 4)
```

-   [ ] Both routes report 9 items and the Shiny row reads back.

## 5. Dummy data with a real factor structure

```{r}
#| label: data
set.seed(1010)
n <- 400
# A real path model is simulated, so the fits below have something to
# recover: quality drives value, value drives loyalty, and quality also
# reaches loyalty directly.
f_qual <- rnorm(n)
f_val  <- 0.60 * f_qual + rnorm(n, 0, sqrt(1 - 0.60^2))
f_loy  <- 0.45 * f_val + 0.30 * f_qual + rnorm(n, 0, 0.7)

ind <- function(f, loading) {
  z <- loading * f + rnorm(n, 0, sqrt(1 - loading^2))
  pmin(5, pmax(1, round(3 + z)))
}
responses <- data.frame(
  q1 = ind(f_qual, .80), q2 = ind(f_qual, .75), q3 = ind(f_qual, .70),
  v1 = ind(f_val,  .82), v2 = ind(f_val,  .78), v3 = ind(f_val,  .72),
  l1 = ind(f_loy,  .80), l2 = ind(f_loy,  .76), l3 = ind(f_loy,  .70)
)
dat <- read_responses(responses, loyalty)
res <- run_analysis_plan(dat, loyalty)
names(res)
```

```{r}
#| label: data-numeric
# lavaan and seminr both want numeric columns. read_responses() keeps the
# stored codes as character, which is right for a response file and wrong
# for a fitting function, so the coercion is explicit here rather than left
# to whichever package coerces first.
items <- c("q1", "q2", "q3", "v1", "v2", "v3", "l1", "l2", "l3")
num <- as.data.frame(lapply(dat[, items], as.numeric))
str(num, give.attr = FALSE)
```

## 6. RQ1: the measurement model

```{r}
#| label: cfa-syntax
syn_cfa <- res[["RQ1"]]$syntax
cat(syn_cfa)
```

```{r}
#| label: cfa-fit
# The test that matters: fit what surveyframe wrote, fit what a lavaan user
# would have written by hand for the same declared model, and compare.
hand_cfa <- '
qual =~ q1 + q2 + q3
val  =~ v1 + v2 + v3
loy  =~ l1 + l2 + l3
'
fit_gen  <- lavaan::cfa(syn_cfa,  data = num, std.lv = TRUE)
fit_hand <- lavaan::cfa(hand_cfa, data = num, std.lv = TRUE)

fm <- function(f, what) unname(lavaan::fitMeasures(f, what))
tab <- compare_table(
  compare_row("chi-square", fm(fit_gen, "chisq"), fm(fit_hand, "chisq"),
              "hand-written lavaan syntax"),
  compare_row("degrees of freedom", fm(fit_gen, "df"), fm(fit_hand, "df"),
              "hand-written lavaan syntax"),
  compare_row("CFI",   fm(fit_gen, "cfi"),   fm(fit_hand, "cfi"),
              "hand-written lavaan syntax"),
  compare_row("TLI",   fm(fit_gen, "tli"),   fm(fit_hand, "tli"),
              "hand-written lavaan syntax"),
  compare_row("RMSEA", fm(fit_gen, "rmsea"), fm(fit_hand, "rmsea"),
              "hand-written lavaan syntax"),
  compare_row("SRMR",  fm(fit_gen, "srmr"),  fm(fit_hand, "srmr"),
              "hand-written lavaan syntax"),
  compare_row("number of free parameters", fm(fit_gen, "npar"),
              fm(fit_hand, "npar"), "hand-written lavaan syntax")
)
compare_report(tab, caption = "Generated CFA syntax vs hand-written")
sf_log(tab)
```

```{r}
#| label: cfa-loadings
pe_gen  <- lavaan::parameterEstimates(fit_gen,  standardized = TRUE)
pe_hand <- lavaan::parameterEstimates(fit_hand, standardized = TRUE)
load_of <- function(pe, lhs, rhs)
  pe$std.all[pe$op == "=~" & pe$lhs == lhs & pe$rhs == rhs]

tab <- compare_table(
  compare_row("loading, qual on q1", load_of(pe_gen, "qual", "q1"),
              load_of(pe_hand, "qual", "q1"), "hand-written lavaan syntax"),
  compare_row("loading, val on v2",  load_of(pe_gen, "val", "v2"),
              load_of(pe_hand, "val", "v2"),  "hand-written lavaan syntax"),
  compare_row("loading, loy on l3",  load_of(pe_gen, "loy", "l3"),
              load_of(pe_hand, "loy", "l3"),  "hand-written lavaan syntax"),
  compare_row("every loading agrees",
              max(abs(pe_gen$std.all[pe_gen$op == "=~"] -
                        pe_hand$std.all[pe_hand$op == "=~"])), 0,
              "all 9 loadings, largest gap"),
  compare_row("latent covariances estimated, not fixed",
              sum(pe_gen$op == "~~" & pe_gen$lhs %in%
                    c("qual", "val", "loy") &
                    pe_gen$rhs %in% c("qual", "val", "loy") &
                    pe_gen$lhs != pe_gen$rhs), 3,
              "3 pairs of 3 constructs")
)
compare_report(tab, caption = "Loadings from the generated syntax")
sf_log(tab)
```

```{r}
#| label: cfa-plot
ld <- pe_gen[pe_gen$op == "=~", ]
op <- par(mar = c(4, 7, 3, 1))
barplot(rev(ld$std.all), horiz = TRUE, las = 1, border = NA,
        names.arg = rev(paste(ld$lhs, "->", ld$rhs)),
        col = ifelse(rev(ld$std.all) >= 0.7, "#1b7837", "#f4a582"),
        xlim = c(0, 1), xlab = "Standardised loading",
        main = "RQ1: measurement model from the generated syntax")
abline(v = 0.7, lty = 2)
par(op)
```

-   [ ] Every generated-vs-hand-written verdict reads `match`. Anything else
    means the generator wrote a different model from the declared one.
-   [ ] The declared constructs recovered loadings near the values the
    simulation built in, roughly 0.70 to 0.82.
-   [ ] The 3 latent covariances are estimated. surveyframe defaults
    `latent_covariances = TRUE`, which is what a measurement model normally
    wants. Set it to `FALSE` only deliberately.

## 7. RQ2: the structural model, covariance-based

```{r}
#| label: sem-syntax
syn_sem <- res[["RQ2"]]$syntax
cat(syn_sem)
```

::: callout-important
## Fifth finding, logged in `dogfeed.todo.md`

The syntax above works because **every path was given a label**. Leave the
labels off, which is what the plain 2-argument `sf_path()` call does, and the
generated syntax does not run at all. The next chunk shows it.
:::

```{r}
#| label: sem-unlabelled
syn_plain <- sem_lavaan_syntax(m_cb_unlabelled, loyalty)
cat(syn_plain)
```

```{r}
#| label: sem-unlabelled-fit
# lavaan is asked to fit it. This chunk is expected to error, and the error
# is the finding rather than a mistake in the review.
plain_msg <- tryCatch({
  lavaan::sem(syn_plain, data = num, std.lv = TRUE); NA_character_
}, error = function(e) conditionMessage(e))
plain_msg
```

```{r}
#| label: sem-unlabelled-cmp
# Isolating the trigger: the same model without the indirect effect fits,
# so it is the indirect line and nothing else.
m_no_ind <- sf_model(
  "m_cb_noind", "No indirect effect", type = "cb_sem",
  constructs = constructs_reflective(),
  paths = list(sf_path("qual", "val"), sf_path("val", "loy"),
               sf_path("qual", "loy"))
)
fits <- function(m) as.numeric(is.na(tryCatch({
  lavaan::sem(sem_lavaan_syntax(m, loyalty), data = num, std.lv = TRUE)
  NA_character_
}, error = function(e) conditionMessage(e))))

tab <- compare_table(
  compare_row("paths unlabelled, no indirect effect: fits",
              fits(m_no_ind), 1, "the control case"),
  compare_row("paths unlabelled, 1 indirect effect: fits",
              fits(m_cb_unlabelled), 0,
              "DEFECT, expected to fail until an owner decision"),
  compare_row("paths labelled, 1 indirect effect: fits",
              fits(m_cb), 1, "the workaround"),
  compare_row("the syntax invents a label nothing carries",
              as.numeric(grepl("qual__val", syn_plain) &&
                           !grepl("qual__val\\s*\\*\\s*qual", syn_plain)), 1,
              "gsub() on the edge name, R/model_layer.R:836")
)
compare_report(tab, caption = "Isolating the unlabelled-path defect")
sf_log(tab)
```

-   [ ] The second row reads `match` with a surveyframe value of **0**. That
    is the defect being confirmed, not a pass. If it reads `1`, the defect
    has been fixed and this section should be rewritten.
-   [ ] Labelling every path is the workaround, and it is what `m_cb` does.
-   [ ] **Worth knowing.** Labels also unlock the total-effect line. Without
    them the generator omits `total_qual_loy` silently, so a user who never
    labels a path loses the total effect and never learns it was on offer.

### The labelled model, fitted

```{r}
#| label: sem-fit
hand_sem <- '
qual =~ q1 + q2 + q3
val  =~ v1 + v2 + v3
loy  =~ l1 + l2 + l3
val ~ a*qual
loy ~ b*val + c*qual
indirect_qual_val_loy := a*b
total_qual_loy := c + indirect_qual_val_loy
'
fit_sem  <- lavaan::sem(syn_sem,  data = num, std.lv = TRUE)
fit_shan <- lavaan::sem(hand_sem, data = num, std.lv = TRUE)

pe_s <- lavaan::parameterEstimates(fit_sem,  standardized = TRUE)
pe_h <- lavaan::parameterEstimates(fit_shan, standardized = TRUE)
path_of <- function(pe, from, to)
  pe$std.all[pe$op == "~" & pe$lhs == to & pe$rhs == from]

tab <- compare_table(
  compare_row("chi-square", fm(fit_sem, "chisq"), fm(fit_shan, "chisq"),
              "hand-written lavaan syntax"),
  compare_row("degrees of freedom", fm(fit_sem, "df"), fm(fit_shan, "df"),
              "hand-written lavaan syntax"),
  compare_row("CFI",   fm(fit_sem, "cfi"),   fm(fit_shan, "cfi"),
              "hand-written lavaan syntax"),
  compare_row("RMSEA", fm(fit_sem, "rmsea"), fm(fit_shan, "rmsea"),
              "hand-written lavaan syntax"),
  compare_row("path qual to val", path_of(pe_s, "qual", "val"),
              path_of(pe_h, "qual", "val"), "hand-written lavaan syntax"),
  compare_row("path val to loy",  path_of(pe_s, "val", "loy"),
              path_of(pe_h, "val", "loy"),  "hand-written lavaan syntax"),
  compare_row("path qual to loy", path_of(pe_s, "qual", "loy"),
              path_of(pe_h, "qual", "loy"), "hand-written lavaan syntax")
)
compare_report(tab, caption = "Generated CB-SEM syntax vs hand-written")
sf_log(tab)
```

```{r}
#| label: sem-indirect
# The declared indirect effect is the reason for the := line in the
# generated syntax. If the generator dropped it, this row would be empty and
# the verdict would read CHECK rather than DIFFERS, which is why CHECK is
# never a pass.
def_of <- function(pe, label) pe$est[pe$op == ":=" & pe$label == label]
ind_gen  <- def_of(pe_s, "indirect_qual_val_loy")
ind_hand <- def_of(pe_h, "indirect_qual_val_loy")
a <- pe_h$est[pe_h$op == "~" & pe_h$lhs == "val" & pe_h$rhs == "qual"]
b <- pe_h$est[pe_h$op == "~" & pe_h$lhs == "loy" & pe_h$rhs == "val"]
cc <- pe_h$est[pe_h$op == "~" & pe_h$lhs == "loy" & pe_h$rhs == "qual"]

tab <- compare_table(
  compare_row("defined parameters generated", sum(pe_s$op == ":="), 2,
              "1 indirect effect plus its total"),
  compare_row("indirect effect value", ind_gen, ind_hand,
              "hand-written lavaan syntax"),
  compare_row("indirect effect is a times b", ind_gen, a * b,
              "hand-multiplied path estimates"),
  compare_row("total effect value", def_of(pe_s, "total_qual_loy"),
              def_of(pe_h, "total_qual_loy"), "hand-written lavaan syntax"),
  compare_row("total effect is c plus a times b",
              def_of(pe_s, "total_qual_loy"), cc + a * b,
              "hand-computed from the path estimates"),
  compare_row("standard errors are estimated too",
              sum(is.finite(pe_s$se[pe_s$op == ":="])), 2,
              "delta method, from lavaan")
)
compare_report(tab, caption = "The declared indirect effect")
sf_log(tab)
pe_s[pe_s$op == ":=", c("label", "est", "se", "pvalue")]
```

```{r}
#| label: sem-plot
paths <- pe_s[pe_s$op == "~", ]
op <- par(mar = c(4, 10, 3, 1))
barplot(rev(paths$std.all), horiz = TRUE, las = 1, border = NA,
        names.arg = rev(paste(paths$rhs, "->", paths$lhs)),
        col = "#4393c3", xlab = "Standardised path coefficient",
        main = "RQ2: structural paths")
abline(v = 0, lty = 2)
par(op)
```

-   [ ] Every fit index, path, and defined parameter matches the
    hand-written model.
-   [ ] The recovered paths are in the neighbourhood of the simulated 0.60,
    0.45, and 0.30. Attenuation from the 5-point rounding is expected.

## 8. RQ3: the same structure, partial least squares

```{r}
#| label: pls-syntax
syn_pls <- res[["RQ3"]]$syntax
cat(syn_pls)
```

::: callout-important
## The `multi_items()` shorthand is worth reading before you trust it

The generated measurement model writes `multi_items("q", 1:3)`, which seminr
expands to `q1`, `q2`, `q3`. That is correct here only because the items were
named with a common stem and consecutive numbers. The next chunk checks the
expansion against the declared item ids rather than assuming it.
:::

```{r}
#| label: pls-expansion
declared <- list(qual = c("q1", "q2", "q3"), val = c("v1", "v2", "v3"),
                 loy = c("l1", "l2", "l3"))
expanded <- list(qual = seminr::multi_items("q", 1:3),
                 val  = seminr::multi_items("v", 1:3),
                 loy  = seminr::multi_items("l", 1:3))

tab <- compare_table(
  compare_row("qual items expand to the declared ids",
              as.numeric(identical(expanded$qual, declared$qual)), 1,
              "seminr::multi_items()"),
  compare_row("val items expand to the declared ids",
              as.numeric(identical(expanded$val, declared$val)), 1,
              "seminr::multi_items()"),
  compare_row("loy items expand to the declared ids",
              as.numeric(identical(expanded$loy, declared$loy)), 1,
              "seminr::multi_items()"),
  compare_row("no declared item is left out",
              length(unlist(expanded)), length(unlist(declared)),
              "9 items were declared")
)
compare_report(tab, caption = "Does the shorthand name the right columns?")
sf_log(tab)
```

```{r}
#| label: pls-fit
# The generated script is executed as written, in its own environment, with
# `data` bound to the response frame. Executing it rather than reading it is
# the point: a script that looks right and does not run is not usable output.
env <- new.env(parent = globalenv())
env$data <- num
gen_script <- seminr_syntax(m_pls, data_name = "data", nboot = 100,
                            seed = 123)
ok <- tryCatch({ eval(parse(text = gen_script), envir = env); TRUE },
               error = function(e) { message(conditionMessage(e)); FALSE })
ok
```

```{r}
#| label: pls-cmp
# The same model built by hand with seminr, for comparison.
hand_mm <- seminr::constructs(
  seminr::composite("qual", seminr::multi_items("q", 1:3)),
  seminr::composite("val",  seminr::multi_items("v", 1:3)),
  seminr::composite("loy",  seminr::multi_items("l", 1:3))
)
hand_sm <- seminr::relationships(
  seminr::paths(from = "qual", to = c("val")),
  seminr::paths(from = "val",  to = c("loy"))
)
fit_hand_pls <- seminr::estimate_pls(data = num,
                                     measurement_model = hand_mm,
                                     structural_model = hand_sm)
gen_paths  <- env$pls_model$path_coef
hand_paths <- fit_hand_pls$path_coef

tab <- compare_table(
  compare_row("path qual to val", gen_paths["qual", "val"],
              hand_paths["qual", "val"], "hand-built seminr model"),
  compare_row("path val to loy",  gen_paths["val", "loy"],
              hand_paths["val", "loy"],  "hand-built seminr model"),
  compare_row("every path agrees", max(abs(gen_paths - hand_paths)), 0,
              "the whole path matrix, largest gap"),
  compare_row("R2 for val", summary(env$pls_model)$paths["R^2", "val"],
              summary(fit_hand_pls)$paths["R^2", "val"],
              "hand-built seminr model"),
  compare_row("R2 for loy", summary(env$pls_model)$paths["R^2", "loy"],
              summary(fit_hand_pls)$paths["R^2", "loy"],
              "hand-built seminr model"),
  compare_row("bootstrap ran as the script asked",
              as.numeric(inherits(env$boot_model, "boot_seminr_model")), 1,
              "the generated script's own bootstrap_model() call")
)
compare_report(tab, caption = "Generated seminr script vs hand-built model")
sf_log(tab)
```

```{r}
#| label: pls-summary
s <- summary(env$pls_model)
s$reliability
s$paths
```

```{r}
#| label: pls-plot
pm <- gen_paths[c("qual", "val"), c("val", "loy")]
vals <- c(`qual -> val` = pm["qual", "val"], `val -> loy` = pm["val", "loy"])
op <- par(mar = c(4, 9, 3, 1))
barplot(rev(vals), horiz = TRUE, las = 1, border = NA, col = "#4393c3",
        xlab = "PLS path coefficient",
        main = "RQ3: the same structure by partial least squares")
par(op)
```

-   [ ] The generated script **ran**, and its results equal the hand-built
    model exactly.
-   [ ] Composite reliability and AVE are printed, so the measurement
    assessment a PLS write-up needs is in the generated output rather than
    left to the reader.

## 9. The model-type guard

This is defect 5 from the independent validation round, fixed in 0.4.0.
Before the fix, a CB-SEM model handed to `seminr_syntax()` produced PLS-SEM
syntax without complaint, so a declared covariance-based model could be
estimated by a completely different estimator and nothing anywhere would say
so.

```{r}
#| label: guards
fires <- function(expr) {
  msg <- tryCatch({ force(expr); NA_character_ },
                  error = function(e) conditionMessage(e))
  msg
}
guard_msgs <- c(
  `seminr_syntax() on a cb_sem model` =
    fires(seminr_syntax(m_cb)),
  `sem_lavaan_syntax() on a pls_sem model` =
    fires(sem_lavaan_syntax(m_pls, loyalty)),
  `cfa_lavaan_syntax() on a pls_sem model` =
    fires(cfa_lavaan_syntax(instrument = loyalty, model = m_pls)),
  `seminr_syntax() on a cfa model` =
    fires(seminr_syntax(m_cfa))
)
data.frame(call = names(guard_msgs), message = unname(guard_msgs),
           row.names = NULL)
```

```{r}
#| label: guards-cmp
# A guard that does not fire returns NA here, so counting the non-NA
# messages counts the guards that did their job.
tab <- compare_table(
  compare_row("guards that fired", sum(!is.na(guard_msgs)), 4,
              "4 deliberate mismatches"),
  compare_row("messages name the type that was found",
              sum(grepl("is of type", guard_msgs)), 4,
              "every message says what it got"),
  compare_row("messages say why it matters",
              sum(grepl("estimate a different model", guard_msgs)), 4,
              "every message says the consequence"),
  # The mutation check: the same calls with the RIGHT model must not error.
  compare_row("correct pairings still work",
              sum(is.na(c(fires(seminr_syntax(m_pls)),
                          fires(sem_lavaan_syntax(m_cb, loyalty)),
                          fires(cfa_lavaan_syntax(instrument = loyalty,
                                                  model = m_cfa))))), 3,
              "3 correct pairings, none should error")
)
compare_report(tab, caption = "The model-type guard, both directions")
sf_log(tab)
```

-   [ ] All 4 mismatches error, and all 3 correct pairings do not. A guard
    that blocks everything is as broken as one that blocks nothing, which is
    why the last row is there.
-   [ ] Each message names the type it found and says what would have gone
    wrong. Read one and check you would understand it without this file open.

## 10. The 2 shortcuts that need no model

```{r}
#| label: shortcuts
# cfa_syntax() works straight off the declared scales, for anyone who has
# not built a model object yet.
cat(cfa_syntax(loyalty))
```

```{r}
#| label: shortcuts-efa
cat(efa_syntax(items = c("q1", "q2", "q3", "v1", "v2", "v3",
                         "l1", "l2", "l3"),
               nfactors = 3, extraction = "ml", rotation = "oblimin"))
```

```{r}
#| label: shortcuts-cmp
syn_scales <- cfa_syntax(loyalty)
fit_scales <- lavaan::cfa(syn_scales, data = num, std.lv = TRUE)

tab <- compare_table(
  compare_row("scale-based CFA has the same df as the model-based one",
              fm(fit_scales, "df"), fm(fit_gen, "df"),
              "the same 3 constructs either way"),
  compare_row("and the same chi-square", fm(fit_scales, "chisq"),
              fm(fit_gen, "chisq"), "the same 3 constructs either way"),
  compare_row("efa_syntax() names all 9 items",
              lengths(regmatches(efa_syntax(items = c("q1","q2","q3","v1",
                                                      "v2","v3","l1","l2",
                                                      "l3")),
                                 gregexpr('"[qvl][123]"',
                                          efa_syntax(items = c("q1","q2","q3",
                                                               "v1","v2","v3",
                                                               "l1","l2",
                                                               "l3"))))),
              9, "9 items were passed")
)
compare_report(tab, caption = "The shortcuts against the model route")
sf_log(tab)
```

-   [ ] Declaring 3 scales and declaring 3 reflective constructs produce the
    same measurement model. If they do not, one of the 2 routes is dropping
    an item.

## 11. The model as data

```{r}
#| label: model-json
cat(substr(model_json(m_cb), 1, 900))
```

```{r}
#| label: model-json-cmp
j <- jsonlite::fromJSON(model_json(m_cb), simplifyVector = FALSE)

tab <- compare_table(
  compare_row("constructs serialised",
              length(j$measurement$constructs), 3, "3 were declared"),
  compare_row("paths serialised",
              length(j$structural$paths), 3, "3 were declared"),
  compare_row("indirect effects serialised",
              length(j$structural$indirect), 1, "1 was declared"),
  compare_row("items serialised across all constructs",
              sum(vapply(j$measurement$constructs,
                         function(cn) length(cn$items), integer(1))), 9,
              "9 items were declared")
)
compare_report(tab, caption = "model_json() round trip")
sf_log(tab)
```

-   [ ] The JSON carries the whole declaration, including the indirect
    effect. This is what the builder and SurveyStudio read, so anything
    missing here is missing from both GUIs.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] Every generated script was **executed**, not only inspected, and
    reproduced the hand-written model exactly.
-   [ ] The model-type guard blocks all 4 wrong pairings and none of the 3
    right ones.
