---
title: "11. Mediation and moderation"
subtitle: "Indirect effects with a bootstrap interval, and interactions with simple slopes"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Two questions that look similar and are not. Mediation asks **how** an effect
travels: does workload reach burnout by way of sleep? Moderation asks **when**
an effect holds: is the workload effect stronger for people with less
support?

Both are declared in the instrument before data exists, which matters more
here than almost anywhere else in the package. Mediation and moderation are
where a post-hoc search does the most damage, because with 3 numeric items
there are 6 mediation models and 6 moderation models available and any of
them can be reported as though it were the plan.

## 1. Questionnaire and plan

```{r}
#| label: build
work <- sf_instrument(
  title = "Workload, sleep, and burnout", version = "1.0.0",
  components = list(
    sf_item("hours",   "Hours worked in a typical week", type = "numeric"),
    sf_item("sleep",   "Hours of sleep on a typical night", type = "numeric"),
    sf_item("support", "How supported do you feel at work? (0 to 100)",
            type = "slider", slider_min = 0, slider_max = 100),
    sf_item("burnout", "How burnt out do you feel? (0 to 100)",
            type = "slider", slider_min = 0, slider_max = 100)
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question =
           "Does long-hours working reach burnout by way of lost sleep?",
         family = "model", method = "mediation",
         roles = list(predictor = "hours", mediator = "sleep",
                      outcome = "burnout"),
         options = list(bootstrap = 2000)),
    list(id = "RQ2", research_question =
           "Is the effect of hours on burnout weaker for well-supported staff?",
         family = "model", method = "moderation",
         roles = list(predictor = "hours", moderator = "support",
                      outcome = "burnout"))
  )
)
work
```

```{r}
#| label: build-check
v <- tryCatch(validate_sframe(work), error = function(e) e)
if (inherits(v, "error")) conditionMessage(v) else
  list(problems = v$problems %||% "none")
```

::: callout-tip
## The field-to-method map this file demonstrates

| You have | You want | Method |
|-------------------------|--------------------|-----------------------|
| 3 numeric items, one on the path between the others | how the effect travels | `mediation` |
| 3 numeric items, one changing the strength of another | when the effect holds | `moderation` |

Both take 3 roles and no more. Mediation wants `predictor`, `mediator`,
`outcome`. Moderation wants `predictor`, `moderator`, `outcome`. Swapping
`mediator` for `moderator` in the roles list changes the question entirely
and nothing will warn you, which is why the plan is declared up front.
:::

## 2. Both collection routes

```{r}
#| label: routes
set.seed(11)
sf_both_routes(work, function(i) list(
  hours = sample(30:70, 1), sleep = round(runif(1, 4, 9), 1),
  support = sample(0:100, 1), burnout = sample(0:100, 1)
), n = 4)
```

-   [ ] Both routes report 4 items and the Shiny row reads back.

## 3. Dummy data with a real indirect effect

```{r}
#| label: data
set.seed(1111)
n <- 300

hours   <- round(rnorm(n, 45, 8))
# a path: more hours, less sleep.
sleep   <- 8.4 - 0.045 * hours + rnorm(n, 0, 0.8)
support <- round(pmin(100, pmax(0, rnorm(n, 55, 18))))
# b path: less sleep, more burnout. c' path: a direct effect on top.
# And a real interaction: support blunts the hours effect.
burnout <- 20 - 5.0 * sleep + 0.55 * hours -
  0.012 * (hours - mean(hours)) * (support - mean(support)) +
  rnorm(n, 0, 9)
burnout <- round(pmin(100, pmax(0, burnout)))

responses <- data.frame(hours = hours, sleep = round(sleep, 1),
                        support = support, burnout = burnout)
dat <- read_responses(responses, work)

# The bootstrap inside sframe_run_mediation() draws from the ambient RNG, so
# the seed is set immediately before the plan runs and again before the
# reference bootstrap below. Without that the 2 intervals differ by sampling
# noise and the comparison reads DIFFERS for no reason.
set.seed(2024)
res <- run_analysis_plan(dat, work)
res
```

```{r}
#| label: data-numeric
num <- as.data.frame(lapply(dat[, c("hours", "sleep", "support", "burnout")],
                            as.numeric))
summary(num)
```

## 4. RQ1: mediation

```{r}
#| label: med
r <- res[["RQ1"]]
data.frame(
  quantity = c("a path, hours to sleep", "b path, sleep to burnout",
               "direct effect c prime", "indirect effect a times b",
               "total effect c", "complete cases", "bootstrap draws"),
  value = c(r$a_path, r$b_path, r$direct, r$indirect, r$total, r$n,
            r$bootstrap),
  row.names = NULL
)
r$apa
```

```{r}
#| label: acc-med
# The 3 regressions surveyframe fits internally, fitted here by hand. This
# is the whole of the point estimate, so if any of these 4 rows disagrees
# the mediation decomposition is wrong.
fit_a <- lm(sleep ~ hours, data = num)
fit_b <- lm(burnout ~ hours + sleep, data = num)
fit_c <- lm(burnout ~ hours, data = num)

a  <- coef(fit_a)[["hours"]]
b  <- coef(fit_b)[["sleep"]]
cp <- coef(fit_b)[["hours"]]
ct <- coef(fit_c)[["hours"]]

tab <- compare_table(
  compare_row("a path", r$a_path, a, "base R lm(sleep ~ hours)"),
  compare_row("b path", r$b_path, b,
              "base R lm(burnout ~ hours + sleep)"),
  compare_row("direct effect", r$direct, cp,
              "base R lm(burnout ~ hours + sleep)"),
  compare_row("total effect", r$total, ct, "base R lm(burnout ~ hours)"),
  compare_row("indirect effect", r$indirect, a * b, "hand-multiplied"),
  compare_row("complete cases used", r$n, sum(complete.cases(num[, c(
    "burnout", "hours", "sleep")])), "base R complete.cases()")
)
compare_report(tab, caption = "RQ1 point estimates vs base R")
sf_log(tab)
```

```{r}
#| label: acc-med-identity
# The decomposition identity. In ordinary least squares with no missing
# data, total = direct + indirect exactly, not approximately. A gap here
# would mean the 3 models were fitted on different rows, which is the
# classic way a mediation table stops adding up.
tab <- compare_table(
  compare_row("total equals direct plus indirect",
              r$total, r$direct + r$indirect, "the OLS identity"),
  compare_row("proportion mediated", r$indirect / r$total,
              (ct - cp) / ct, "hand-computed 2 ways"),
  compare_row("the 3 models used the same rows",
              nobs(fit_a), nobs(fit_b), "base R nobs()"),
  compare_row("and the same rows as the total model",
              nobs(fit_b), nobs(fit_c), "base R nobs()")
)
compare_report(tab, caption = "The decomposition adds up")
sf_log(tab)
```

::: callout-note
## Why the identity is checked rather than assumed

`total = direct + indirect` holds in OLS only when all 3 models are fitted on
the same rows. surveyframe drops incomplete cases once, before fitting, which
is what makes it hold here. A package that dropped rows per model would give
a mediation table where the parts do not sum to the whole, and it would look
like rounding.
:::

```{r}
#| label: acc-med-boot
# The bootstrap interval, reproduced by a hand-written percentile bootstrap
# using the same algorithm and the same seed. Matching to the last decimal
# confirms the resampling is a plain non-parametric percentile bootstrap on
# complete cases and nothing more elaborate.
set.seed(2024)
df <- num[complete.cases(num[, c("burnout", "hours", "sleep")]),
          c("burnout", "hours", "sleep")]
nb <- r$bootstrap
reps <- numeric(nb)
for (i in seq_len(nb)) {
  idx <- sample.int(nrow(df), nrow(df), replace = TRUE)
  bd  <- df[idx, , drop = FALSE]
  reps[i] <- coef(lm(sleep ~ hours, data = bd))[["hours"]] *
    coef(lm(burnout ~ hours + sleep, data = bd))[["sleep"]]
}
ref_ci <- quantile(reps, c(.025, .975))

tab <- compare_table(
  compare_row("bootstrap CI lower", r$indirect_ci[[1]], ref_ci[[1]],
              "hand-written percentile bootstrap, same seed"),
  compare_row("bootstrap CI upper", r$indirect_ci[[2]], ref_ci[[2]],
              "hand-written percentile bootstrap, same seed"),
  compare_row("draws taken", nb, 2000,
              "options = list(bootstrap = 2000) was declared")
)
compare_report(tab, caption = "The bootstrap interval, reproduced")
sf_log(tab)
```

-   [ ] The declared `bootstrap = 2000` reached the runner. If it reads
    1000, the option was ignored and the interval is not what the plan asked
    for.

```{r}
#| label: acc-med-sobel
# An independent significance check that does not use the bootstrap at all.
# Sobel's delta-method standard error is more conservative than a percentile
# bootstrap and rests on a normality assumption the bootstrap avoids, so the
# 2 should agree on the verdict without agreeing on the number.
se_a <- summary(fit_a)$coefficients["hours", "Std. Error"]
se_b <- summary(fit_b)$coefficients["sleep", "Std. Error"]
sobel_se <- sqrt(b^2 * se_a^2 + a^2 * se_b^2)
sobel_z  <- (a * b) / sobel_se
sobel_p  <- 2 * pnorm(-abs(sobel_z))

tab <- compare_table(
  compare_row("bootstrap interval excludes zero",
              as.numeric(r$indirect_ci[[1]] * r$indirect_ci[[2]] > 0), 1,
              "both bounds on the same side of 0"),
  compare_row("Sobel test agrees the indirect effect is non-zero",
              as.numeric(sobel_p < 0.05), 1, "Sobel delta method"),
  # Two negative paths multiply to a positive indirect effect: more hours
  # costs sleep, and less sleep raises burnout, so the route through sleep
  # adds to burnout. Reading either path's sign as the indirect effect's
  # sign is a common slip and this row is here to catch it.
  compare_row("the indirect effect is positive, as 2 negative paths imply",
              as.numeric(r$indirect > 0), 1,
              "a below 0 times b below 0"),
  compare_row("and both paths are individually negative",
              as.numeric(r$a_path < 0 && r$b_path < 0), 1,
              "more hours costs sleep, less sleep raises burnout"),
  compare_row("bootstrap interval contains the point estimate",
              as.numeric(r$indirect > r$indirect_ci[[1]] &&
                           r$indirect < r$indirect_ci[[2]]), 1,
              "by construction of a percentile interval")
)
compare_report(tab, caption = "Two independent readings of the same effect")
sf_log(tab)
data.frame(sobel_z = sobel_z, sobel_p = sobel_p,
           boot_lower = r$indirect_ci[[1]], boot_upper = r$indirect_ci[[2]],
           row.names = NULL)
```

```{r}
#| label: acc-med-plot
op <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
hist(reps, breaks = 40, col = "#deebf7", border = "white",
     xlab = "Indirect effect, a times b",
     main = sprintf("%d bootstrap draws", nb))
abline(v = r$indirect, col = "#b2182b", lwd = 2)
abline(v = r$indirect_ci, col = "#2166ac", lty = 2)
abline(v = 0, col = "#404040", lty = 3)

barplot(c(direct = r$direct, indirect = r$indirect, total = r$total),
        col = c("#9ecae1", "#fdae61", "#4393c3"), border = NA,
        ylab = "Effect on burnout per extra hour",
        main = "The decomposition")
abline(h = 0)
par(op)
```

-   [ ] Every mediation number matches, including both bootstrap bounds.
-   [ ] The dotted zero line sits outside the dashed interval. If it sits
    inside, the APA sentence must not claim mediation.
-   [ ] The direct and indirect bars stack to the total bar.

## 5. RQ2: moderation

```{r}
#| label: mod
r <- res[["RQ2"]]
r$coefficients
r$conditional_effects
r$apa
```

```{r}
#| label: acc-mod
# surveyframe fits burnout ~ hours * support on the raw, uncentred
# variables. The reference must do the same. Centring first is the more
# common advice and gives different lower-order coefficients, so a centred
# reference would show 2 false DIFFERS and one true match on the
# interaction, which is a confusing way to discover the convention.
fit_m <- lm(burnout ~ hours * support, data = num)
cm <- summary(fit_m)$coefficients
sc <- r$coefficients

tab <- compare_table(
  compare_row("intercept", sc["(Intercept)", "Estimate"],
              cm["(Intercept)", "Estimate"], "base R lm(y ~ x * z)"),
  compare_row("hours coefficient", sc["hours", "Estimate"],
              cm["hours", "Estimate"], "base R lm(y ~ x * z)"),
  compare_row("support coefficient", sc["support", "Estimate"],
              cm["support", "Estimate"], "base R lm(y ~ x * z)"),
  compare_row("interaction coefficient", sc["hours:support", "Estimate"],
              cm["hours:support", "Estimate"], "base R lm(y ~ x * z)"),
  compare_row("interaction standard error",
              sc["hours:support", "Std. Error"],
              cm["hours:support", "Std. Error"], "base R lm(y ~ x * z)"),
  compare_row("interaction p value",
              sc["hours:support", "Pr(>|t|)"],
              cm["hours:support", "Pr(>|t|)"], "base R lm(y ~ x * z)"),
  compare_row("complete cases used", r$n, nobs(fit_m), "base R nobs()")
)
compare_report(tab, caption = "RQ2 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-mod-slopes
# The conditional effects are simple slopes at the moderator's mean and one
# standard deviation either side. Each is b_x + b_xz * z, evaluated at that
# z. They are hand-computed here from the base R fit, so the check is on
# surveyframe's arithmetic and on its choice of the 3 evaluation points.
mz  <- mean(num$support)
sdz <- sd(num$support)
bx  <- cm["hours", "Estimate"]
bxz <- cm["hours:support", "Estimate"]
ce  <- r$conditional_effects
slope_at <- function(lvl) ce$simple_slope[ce$moderator_level == lvl]
value_at <- function(lvl) ce$moderator_value[ce$moderator_level == lvl]

tab <- compare_table(
  compare_row("low moderator value", value_at("low"), mz - sdz,
              "mean minus 1 sd"),
  compare_row("mean moderator value", value_at("mean"), mz, "base R mean()"),
  compare_row("high moderator value", value_at("high"), mz + sdz,
              "mean plus 1 sd"),
  compare_row("simple slope at low support", slope_at("low"),
              bx + bxz * (mz - sdz), "hand-computed from lm()"),
  compare_row("simple slope at mean support", slope_at("mean"),
              bx + bxz * mz, "hand-computed from lm()"),
  compare_row("simple slope at high support", slope_at("high"),
              bx + bxz * (mz + sdz), "hand-computed from lm()"),
  compare_row("the slope at the mean equals the centred main effect",
              slope_at("mean"),
              coef(lm(burnout ~ scale(hours, scale = FALSE) *
                        scale(support, scale = FALSE),
                      data = num))[[2]],
              "base R lm() on mean-centred predictors")
)
compare_report(tab, caption = "Simple slopes vs hand-computed")
sf_log(tab)
```

::: callout-note
## The last row is the one worth understanding

Centring the predictors changes the lower-order coefficients but not the
interaction. The slope reported at the moderator's mean is exactly the
`hours` coefficient a centred model would print. That is why the 2 agree,
and it is the quickest way to satisfy yourself that surveyframe's uncentred
fit and the centred advice you have probably read are the same model.
:::

```{r}
#| label: acc-mod-recovery
# Did the interaction the simulation built in come back out? The data were
# generated with a coefficient of -0.012 on the centred product. Centred and
# uncentred models share the interaction term, so this is a fair comparison
# and a loose tolerance is right for a recovery check on n = 300.
tab <- compare_table(
  compare_row("interaction recovered from the simulation",
              cm["hours:support", "Estimate"], -0.012,
              "the value the data were built with", tol = 0.01),
  compare_row("and it is negative, as designed",
              as.numeric(cm["hours:support", "Estimate"] < 0), 1,
              "support blunts the hours effect"),
  compare_row("the slope is steeper at low support than at high",
              as.numeric(slope_at("low") > slope_at("high")), 1,
              "what a blunting moderator means")
)
compare_report(tab, caption = "Recovering the simulated interaction")
sf_log(tab)
```

-   [ ] The tolerance on the first row is loosened to 0.01, and the reason
    is stated: it is a sampling-recovery check on 300 rows, not an
    arithmetic check. Every other row in this file runs at 1e-6.

```{r}
#| label: acc-mod-plot
xs <- seq(min(num$hours), max(num$hours), length.out = 50)
cols <- c(low = "#b2182b", mean = "#404040", high = "#2166ac")
plot(num$hours, num$burnout, pch = 16, col = "#bdbdbd66",
     xlab = "Hours worked per week", ylab = "Burnout",
     main = "RQ2: the hours effect at 3 levels of support")
for (lvl in c("low", "mean", "high")) {
  z  <- value_at(lvl)
  b0 <- cm["(Intercept)", "Estimate"] + cm["support", "Estimate"] * z
  lines(xs, b0 + slope_at(lvl) * xs, col = cols[[lvl]], lwd = 2)
}
legend("topleft", lwd = 2, col = cols, bty = "n",
       legend = sprintf("%s support (%.0f)", names(cols),
                        vapply(names(cols), value_at, numeric(1))))
```

-   [ ] The 3 lines fan out rather than sitting parallel. Parallel lines
    mean no interaction, whatever the coefficient table says.
-   [ ] The red line, low support, is the steepest.

## 6. What happens when the roles are wrong

A role list that names a column that is not there, or a role the method does
not use, is the most common way one of these blocks fails. Both should say so
rather than returning something.

```{r}
#| label: roles
try_block <- function(roles, method) {
  b <- list(id = "X", research_question = "Q", family = "model",
            method = method, roles = roles)
  out <- sframe_run_one_block(b, dat, work)
  if (!is.null(out$error)) out$error else "returned a result"
}

data.frame(
  scenario = c("mediation, mediator column does not exist",
               "mediation, mediator role missing entirely",
               "moderation, moderator column does not exist",
               "moderation, predictor and outcome swapped"),
  outcome = c(
    try_block(list(predictor = "hours", mediator = "nosuch",
                   outcome = "burnout"), "mediation"),
    try_block(list(predictor = "hours", outcome = "burnout"), "mediation"),
    try_block(list(predictor = "hours", moderator = "nosuch",
                   outcome = "burnout"), "moderation"),
    try_block(list(predictor = "burnout", moderator = "support",
                   outcome = "hours"), "moderation")
  ),
  row.names = NULL
)
```

```{r}
#| label: roles-cmp
msgs <- c(
  try_block(list(predictor = "hours", mediator = "nosuch",
                 outcome = "burnout"), "mediation"),
  try_block(list(predictor = "hours", outcome = "burnout"), "mediation"),
  try_block(list(predictor = "hours", moderator = "nosuch",
                 outcome = "burnout"), "moderation")
)

tab <- compare_table(
  compare_row("all 3 bad role lists produce an error message",
              sum(msgs != "returned a result"), 3,
              "none should return a result"),
  compare_row("the messages name the missing column",
              sum(grepl("nosuch", msgs)), 2,
              "the 2 cases where a name was given"),
  # The mutation check: the correct role list must still work.
  compare_row("the correct role list still returns a result",
              as.numeric(try_block(list(predictor = "hours",
                                        mediator = "sleep",
                                        outcome = "burnout"),
                                   "mediation") == "returned a result"), 1,
              "the working case")
)
compare_report(tab, caption = "Bad role lists fail loudly")
sf_log(tab)
```

-   [ ] Read the 4 messages above. The last scenario, predictor and outcome
    swapped, is the one that **cannot** be caught: both columns exist, so it
    returns a perfectly good answer to a question nobody asked. That is what
    the pre-declared plan is for, and it is worth noticing that no amount of
    validation replaces it.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] The mediation decomposition adds up, the bootstrap interval
    reproduces exactly, and the simple slopes match a hand-computed model.
-   [ ] Both APA sentences are usable in a write-up without editing.
