---
title: "12. The small-sample track"
subtitle: "Hodges-Lehmann, pseudomedian, exact odds ratio, Firth, the advisory"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

::: callout-important
## No previous review covers any of this

`mas_review_032`, `033`, and `034` contain no mention of Hodges-Lehmann, the
pseudomedian interval, the exact odds-ratio interval, Firth regression, or
the small-sample advisory. For a release whose headline is small-sample
inference, this file closes the largest gap in the review history.
:::

The track exists for one situation: **n below 30**, where asymptotic p-values
and normal-theory intervals stop being trustworthy. Every method below
answers the same question its large-sample counterpart does, but with an
estimator that behaves at small n.

## 1. Questionnaire and plan

A small pilot study. Eighteen people, 2 arms.

```{r}
#| label: build
pilot <- sf_instrument(
  title = "Reading intervention pilot", version = "1.0.0",
  components = list(
    sf_choices("arm", c("usual", "new"), c("Usual practice", "New method")),
    sf_choices("yn", c("0", "1"), c("No", "Yes")),
    sf_item("arm",   "Which method were you taught with?",
            type = "single_choice", choice_set = "arm", required = TRUE),
    sf_item("pre",   "Reading score before", type = "numeric"),
    sf_item("post",  "Reading score after",  type = "numeric"),
    sf_item("passed", "Did you pass the end-of-term check?",
            type = "single_choice", choice_set = "yn")
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question =
           "Do the 2 arms differ on the post score?",
         family = "group_comparison", method = "mann_whitney",
         roles = list(outcome = "post", group = "arm")),
    list(id = "RQ2", research_question =
           "Did scores change from before to after?",
         family = "group_comparison", method = "wilcoxon_pair",
         roles = list(before = "pre", after = "post")),
    list(id = "RQ3", research_question =
           "Does the pass rate differ by arm?",
         family = "association", method = "fisher_exact",
         roles = list(row = "arm", column = "passed"))
  )
)
pilot
```

## 2. Both collection routes

```{r}
#| label: routes
set.seed(6)
sf_both_routes(pilot, function(i) list(
  arm = sample(c("usual", "new"), 1), pre = sample(30:70, 1),
  post = sample(35:80, 1), passed = sample(c("0", "1"), 1)
), n = 3)
```

## 3. Dummy data, deliberately small

```{r}
#| label: data
set.seed(5)
n   <- 18
arm <- rep(c("usual", "new"), each = n / 2)
pre  <- round(rnorm(n, 50, 8))
post <- round(pre + ifelse(arm == "new", 9, 2) + rnorm(n, 0, 5))
# A rare failure, so the exact odds-ratio interval has something to do.
passed <- c(rbinom(9, 1, 0.55), rbinom(9, 1, 0.90))

dat <- read_responses(data.frame(
  arm = arm, pre = pre, post = post,
  passed = as.character(passed), stringsAsFactors = FALSE), pilot)

data.frame(n = nrow(dat), per_arm = n / 2, passes = sum(passed))
res <- run_analysis_plan(dat, pilot)
res
```

-   [ ] n is 18, which is below the threshold the advisory watches for.

## 4. The advisory

```{r}
#| label: advisory
ar <- assumption_report(dat, variables = "post", group = "arm")
ar$advisory
```

```{r}
#| label: advisory-large
# The same call on a large sample, to confirm the advisory is conditional
# rather than always on. An advisory that never switches off is noise.
set.seed(5)
big <- data.frame(arm = rep(c("usual", "new"), each = 100),
                  post = c(rnorm(100, 52, 8), rnorm(100, 60, 8)))
ab <- assumption_report(big, variables = "post", group = "arm")

data.frame(
  sample = c("n = 18", "n = 200"),
  advisory_fired = c(!is.null(ar$advisory), !is.null(ab$advisory))
)
```

-   [ ] The advisory fires at n = 18 and stays silent at n = 200.
-   [ ] **Owner judgement.** Read the advisory text. Does it change what a
    researcher would actually do, or does it only say the sample is small?
    It names non-parametric alternatives and exact intervals, so the test is
    whether a reader knows which one to reach for next.

## 5. RQ1: Hodges-Lehmann shift and its interval

::: callout-warning
## Match the reference call, or you will report a bug that is not there

surveyframe fixes `exact = FALSE` on `wilcox.test()`
(`R/analysis_plan.R:359`). Left on its own default, base R switches to the
**exact** method at this n and no ties, and the 2 disagree in the 5th
decimal. Writing this suite produced exactly that false alarm. Both are
computed below so the difference is visible rather than hidden.
:::

```{r}
#| label: acc-hl
r <- res[["RQ1"]]
a <- dat$post[dat$arm == "usual"]
b <- dat$post[dat$arm == "new"]

w <- wilcox.test(a, b, conf.int = TRUE, exact = FALSE)

tab <- compare_table(
  compare_row("U", r$U, unname(w$statistic), "wilcox.test(exact = FALSE)"),
  compare_row("p", r$p, w$p.value,           "wilcox.test(exact = FALSE)"),
  compare_row("Hodges-Lehmann shift", r$hl_shift, unname(w$estimate),
              "wilcox.test(exact = FALSE)"),
  compare_row("HL lower", r$hl_conf_int[1], w$conf.int[1],
              "wilcox.test(exact = FALSE)"),
  compare_row("HL upper", r$hl_conf_int[2], w$conf.int[2],
              "wilcox.test(exact = FALSE)")
)
compare_report(tab, caption = "RQ1 like for like")
sf_log(tab)
```

```{r}
#| label: acc-hl-exact
# What base R gives a researcher on its own default at this n.
we <- wilcox.test(a, b, conf.int = TRUE)
data.frame(
  quantity = c("HL shift", "HL lower", "HL upper"),
  surveyframe_normal_approx = signif(c(r$hl_shift, r$hl_conf_int), 8),
  base_R_exact_default      = signif(c(unname(we$estimate), we$conf.int), 8),
  difference = signif(abs(c(r$hl_shift, r$hl_conf_int) -
                            c(unname(we$estimate), we$conf.int)), 3)
)
```

```{r}
#| label: acc-hl-plot
stripchart(list(usual = a, new = b), method = "jitter", pch = 19,
           col = c("#3182bd", "#b2182b"), vertical = TRUE,
           ylab = "post score",
           main = sprintf("RQ1: HL shift = %.2f [%.2f, %.2f]",
                          r$hl_shift, r$hl_conf_int[1], r$hl_conf_int[2]))
segments(1, median(a), 2, median(b), lty = 2, col = "#404040")
points(c(1, 2), c(median(a), median(b)), pch = 18, cex = 2)
```

-   [ ] All 5 like-for-like verdicts read `match`.
-   [ ] **Owner decision.** Compare the 2 tables. surveyframe reports the
    normal-approximation interval on a sample of 18, where base R's own
    default would give the exact one. Nothing is numerically wrong. The
    question is whether the release built for small samples should switch to
    the exact method precisely when n is small, or expose it as an option the
    way McNemar's correction is exposed in file `04`.
-   [ ] With only 9 per arm, every point is visible in the plot. Confirm the
    reported shift is plausible against what you can see.

## 6. RQ2: pseudomedian and its interval

The paired counterpart. The pseudomedian is the median of the Walsh averages
rather than the median of the differences, so it is not the same number as
`median(post - pre)` and should not be expected to be.

```{r}
#| label: acc-pm
r  <- res[["RQ2"]]
ws <- wilcox.test(dat$pre, dat$post, paired = TRUE, conf.int = TRUE,
                  exact = FALSE)

tab <- compare_table(
  compare_row("V", r$V, unname(ws$statistic), "wilcox.test(paired)"),
  compare_row("p", r$p, ws$p.value,           "wilcox.test(paired)"),
  compare_row("pseudomedian", r$pseudomedian, unname(ws$estimate),
              "wilcox.test(paired)"),
  compare_row("lower", r$pseudomedian_conf_int[1], ws$conf.int[1],
              "wilcox.test(paired)"),
  compare_row("upper", r$pseudomedian_conf_int[2], ws$conf.int[2],
              "wilcox.test(paired)")
)
compare_report(tab, caption = "RQ2 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-pm-walsh
# The pseudomedian from its own definition, as a second independent check.
d     <- dat$pre - dat$post
walsh <- outer(d, d, "+") / 2
tab <- compare_table(
  compare_row("pseudomedian from Walsh averages", r$pseudomedian,
              median(walsh[upper.tri(walsh, diag = TRUE)]),
              "median of Walsh averages", tol = 1e-3)
)
compare_report(tab, caption = "RQ2 against the definition")
sf_log(tab)

data.frame(pseudomedian = r$pseudomedian,
           median_of_differences = median(d),
           note = "these are different estimators and need not agree")
```

-   [ ] All 5 base R verdicts read `match`.
-   [ ] The Walsh-average check matches to `1e-3`. It uses a slightly
    different tie handling, so exact agreement is not expected here, and the
    tolerance says so rather than being quietly widened.

## 7. RQ3: exact odds-ratio interval

```{r}
#| label: acc-or
r  <- res[["RQ3"]]
tb <- table(dat$arm, dat$passed)
ft <- fisher.test(tb)
tb
```

```{r}
#| label: acc-or-cmp
tab <- compare_table(
  compare_row("p", r$p, ft$p.value, "base R fisher.test()"),
  compare_row("odds ratio", r$odds_ratio, unname(ft$estimate),
              "base R fisher.test()"),
  compare_row("OR lower", r$odds_ratio_conf_int[1], ft$conf.int[1],
              "base R fisher.test()"),
  compare_row("OR upper", r$odds_ratio_conf_int[2], ft$conf.int[2],
              "base R fisher.test()")
)
compare_report(tab, caption = "RQ3 vs base R")
sf_log(tab)
```

```{r}
#| label: acc-or-vs-sample
# The conditional MLE odds ratio Fisher reports is not the sample odds ratio.
# Showing both stops a reviewer recording a false finding.
sample_or <- (tb[1, 1] * tb[2, 2]) / (tb[1, 2] * tb[2, 1])
data.frame(
  quantity = c("conditional MLE odds ratio (reported)",
               "sample odds ratio (a*d)/(b*c)"),
  value = c(r$odds_ratio, sample_or)
)
```

-   [ ] All 4 verdicts read `match`.
-   [ ] The interval is wide, as it should be on 18 people. A narrow interval
    here would be the finding.
-   [ ] The 2 odds ratios above differ, and that is expected rather than a
    defect.

## 8. Firth logistic regression at small n

Firth's penalised likelihood is the small-sample answer to logistic
regression. File `08` checks it against `logistf` on a rare outcome at n =
260. Here is the case it was really built for: **complete separation**, where
ordinary logistic regression has no finite estimate at all.

```{r}
#| label: separation
sep <- data.frame(x = c(1, 2, 3, 4, 5, 6, 7, 8),
                  y = c(0, 0, 0, 0, 1, 1, 1, 1))

g  <- suppressWarnings(glm(y ~ x, data = sep, family = binomial()))
fl <- logistf::logistf(y ~ x, data = sep)

data.frame(
  method = c("ordinary glm()", "Firth logistf()"),
  slope  = c(coef(g)[["x"]], coef(fl)[["x"]]),
  finite = c(is.finite(coef(g)[["x"]]) && abs(coef(g)[["x"]]) < 100,
             is.finite(coef(fl)[["x"]]))
)
```

```{r}
#| label: separation-sf
sep_inst <- sf_instrument(
  title = "Separation demo", version = "1.0.0",
  components = list(
    sf_choices("yn", c("0", "1"), c("No", "Yes")),
    sf_item("x", "Predictor", type = "numeric"),
    sf_item("y", "Outcome", type = "single_choice", choice_set = "yn")),
  analysis_plan = list(list(
    id = "RQ1", research_question = "Does x predict y under separation?",
    family = "prediction", method = "firth_logistic",
    roles = list(outcome = "y", predictors = "x"))))

sep_dat <- read_responses(
  data.frame(x = sep$x, y = as.character(sep$y), stringsAsFactors = FALSE),
  sep_inst)
rs <- run_analysis_plan(sep_dat, sep_inst)[["RQ1"]]

tab <- compare_table(
  compare_row("slope", rs$coefficients["x", "Estimate"], coef(fl)[["x"]],
              "logistf::logistf()", tol = 1e-5),
  compare_row("CI lower", rs$coefficients["x", "ci_low"], fl$ci.lower[["x"]],
              "logistf::logistf()", tol = 1e-4),
  compare_row("CI upper", rs$coefficients["x", "ci_high"], fl$ci.upper[["x"]],
              "logistf::logistf()", tol = 1e-4),
  compare_row("p", rs$coefficients["x", "p"], fl$prob[["x"]],
              "logistf::logistf()", tol = 1e-5)
)
compare_report(tab, caption = "Firth under complete separation")
sf_log(tab)
```

-   [ ] Ordinary `glm()` produces an absurd slope and a warning. Firth
    produces a finite one.
-   [ ] All 4 verdicts read `match`.
-   [ ] **This is the strongest argument for the track.** A user with 8
    observations and perfect separation gets a usable answer instead of a
    number that only looks like one.

## 9. Sample-size planning

```{r}
#| label: ssp
plans <- list(
  t_test     = sample_size_plan("t_test", sd = 10, margin_error = 5),
  proportion = sample_size_plan("proportion", margin_error = 0.05, p = 0.5),
  correlation = sample_size_plan("correlation", r = 0.3)
)
plans
```

-   [ ] Each plan returns a number and says what assumptions produced it.
-   [ ] **Owner judgement.** Would a first-time user know which `type` to
    pick from the argument names alone?

## 10. The whole track on one page

```{r}
#| label: summary
data.frame(
  feature = c("small-sample advisory", "Hodges-Lehmann shift and CI",
              "pseudomedian and CI", "exact odds-ratio CI",
              "Firth logistic regression", "sample-size planning"),
  where = c("assumption_report()", "mann_whitney", "wilcoxon_pair",
            "fisher_exact", "firth_logistic", "sample_size_plan()"),
  new_in_040 = c(TRUE, TRUE, TRUE, TRUE, TRUE, FALSE),
  reviewed_before = FALSE
)
```

-   [ ] Every row above ran somewhere in this file.
-   [ ] `vignettes/small-sample.Rmd` tells the same story as this file, and
    does not contradict it.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] The 2 owner decisions in sections 4 and 5 are recorded in
    `dogfeed.todo.md` before this file is closed.
