---
title: "13. Decision analysis: all 10 MCDM methods"
subtitle: "TOPSIS, AHP, ANP, DEMATEL, VIKOR, MOORA, SMART, WASPAS, PROMETHEE, ELECTRE"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

The headline feature of 0.4.0. Ten multi-criteria decision-making methods,
2 new item types to collect their input, and a bundled worked example.

`RMCDA` is the independent oracle throughout. It already earned its place
during development: it caught WASPAS using SMART's normalisation.

## 1. The bundled instrument

```{r}
#| label: fixture
hotel <- read_sframe(sf_fixture("hotel_supplier_mcdm.sframe"))

data.frame(
  item = vapply(hotel$items, function(i) i$id, character(1)),
  type = vapply(hotel$items, function(i) i$type, character(1))
)
```

```{r}
#| label: fixture-plan
data.frame(
  block  = vapply(hotel$analysis_plan, function(b) b$id, character(1)),
  method = vapply(hotel$analysis_plan, function(b) b$method, character(1)),
  roles  = vapply(hotel$analysis_plan, function(b)
    paste(names(b$roles), vapply(b$roles, paste, character(1), collapse = ","),
          sep = " = ", collapse = "; "), character(1))
)
```

::: callout-important
## The hash gate

Every bundled instrument must still hash to what it stores. If this fails,
every stored `.sframe` file moved with it, and the hash stops being the
instrument's identity.

There is **no `$integrity` element** on an `sframe` object. The hash lives
inside the `.sframe` file's JSON, and `read_sframe()` verifies it on the way
in. An earlier draft of this chunk compared `hotel$integrity$hash` against
itself, which is `identical(NULL, NULL)`, reads `match` whatever the package
does, and proves nothing. File `16` section 6 covers the round trip in full.
:::

```{r}
#| label: hash
stored_hash <- function(path)
  jsonlite::fromJSON(paste(readLines(path, warn = FALSE), collapse = "\n"),
                     simplifyVector = FALSE)$hash$value

src <- sf_fixture("hotel_supplier_mcdm.sframe")
rewritten <- tempfile(fileext = ".sframe")
write_sframe(hotel, rewritten, overwrite = TRUE)

tampered <- sub("Hotel supplier selection", "Hotel supplier choice",
                paste(readLines(src, warn = FALSE), collapse = "\n"))
p_bad <- tempfile(fileext = ".sframe")
writeLines(tampered, p_bad)

tab <- compare_table(
  compare_row("the bundled file verifies on read",
              as.numeric(!inherits(tryCatch(read_sframe(src),
                                            error = function(e) e), "error")),
              1, "read_sframe()", tol = 0),
  compare_row("rewriting it reproduces the hash it already carried",
              as.numeric(stored_hash(src) == stored_hash(rewritten)), 1,
              "the shipped file's own stored value", tol = 0),
  # The mutation check. Without it the 2 rows above only show a constant
  # equals itself.
  compare_row("an edited copy is rejected",
              as.numeric(!inherits(tryCatch(read_sframe(p_bad),
                                            error = function(e) e), "error")),
              0, "the integrity check fires", tol = 0)
)
compare_report(tab, caption = "Hash gate")
sf_log(tab)
```

## 2. Both collection routes for the 2 new item types

```{r}
#| label: routes-html
sf_html_route(hotel)
```

```{r}
#| label: routes-widgets
sf_shiny_widgets(hotel)
```

```{r}
#| label: routes-inputs
# The Shiny input names for the 2 decision types, so nobody has to guess.
sf_answer_template(hotel)
```

-   [ ] Every item reaches the exported HTML.
-   [ ] Every widget builds, including `pairwise_comparison` and
    `criteria_weight`. Neither existed before 0.4.0.
-   [ ] **Open the exported survey** with `browseURL()` on the path from
    section 2, and confirm the constant-sum check on the
    `criteria_weight` question refuses to submit when the points do not
    total the required amount.

## 3. The bundled responses

```{r}
#| label: data
raw <- read.csv(sf_fixture("hotel_supplier_mcdm_responses.csv"),
                check.names = FALSE)
dat <- read_responses(raw, hotel, meta_cols = "respondent_id")
data.frame(respondents = nrow(dat), columns = ncol(dat))
```

::: callout-note
## `check.names = FALSE` matters here

`read.csv()` rewrites headers by default, so `crit_pairs__a__vs__b` becomes
something else and the instrument no longer recognises its own columns.
0.4.0 added a detection hint that explains this rather than just rejecting
the file. Drop `check.names = FALSE` and rerun if you want to see it.
:::

## 4. AHP, and the consistency ratio

```{r}
#| label: ahp
res  <- run_analysis_plan(dat, hotel)
ahp  <- res[["RQ1"]]
ahp$weights
```

```{r}
#| label: ahp-rmcda-exact
# Cross-check on a perfectly consistent matrix built from a known weight
# vector, which is what the package's own test suite does. On a consistent
# matrix every AHP weighting method must return the generating weights
# exactly, so this is an unambiguous check with a known right answer.
w0 <- c(0.40, 0.30, 0.20, 0.10)
mc <- outer(w0, w0, "/")
dimnames(mc) <- list(letters[1:4], letters[1:4])

sf_w <- sframe_principal_eigen(mc)$weights
rm_w <- RMCDA:::find.weight(mc)[[2]]

tab <- do.call(compare_table, lapply(seq_along(w0), function(i) {
  compare_row(paste("AHP weight on a consistent matrix,", letters[i]),
              sf_w[[i]], w0[i], "the generating weights")
}))
compare_report(tab, caption = "AHP recovers known weights exactly")
sf_log(tab)

data.frame(criterion = letters[1:4], surveyframe = as.numeric(sf_w),
           RMCDA = as.numeric(rm_w), generating = w0)
```

```{r}
#| label: ahp-rmcda-real
# On the real aggregated matrix the 2 tools differ slightly, because they
# approximate the principal eigenvector differently. This is a
# methodological difference and not a defect, so it is reported rather than
# compared at 1e-6, which would produce a false DIFFERS.
pw   <- sframe_assemble_pairwise(dat, hotel, "crit_pairs")
agg  <- sframe_aggregate_judgements(pw$matrices)
rm_r <- RMCDA:::find.weight(agg$matrix)[[2]]

data.frame(
  criterion   = names(ahp$weights),
  surveyframe = round(as.numeric(ahp$weights), 5),
  RMCDA       = round(as.numeric(rm_r), 5),
  difference  = signif(abs(as.numeric(ahp$weights) - as.numeric(rm_r)), 3),
  same_rank   = rank(-as.numeric(ahp$weights)) == rank(-as.numeric(rm_r))
)
```

```{r}
#| label: ahp-cr
data.frame(
  quantity = c("consistency ratio", "lambda max", "warning raised"),
  value = c(round(ahp$cr, 5), round(ahp$lambda_max, 5),
            as.character(!is.null(ahp$consistency_warning)))
)
```

```{r}
#| label: ahp-plot
barplot(sort(unlist(ahp$weights), decreasing = TRUE), col = "#4292c6",
        border = NA, ylab = "criterion weight",
        main = sprintf("AHP weights, CR = %.3f", ahp$cr))
```

-   [ ] On a perfectly consistent matrix surveyframe recovers the generating
    weights exactly, and `RMCDA` returns the same numbers.
-   [ ] On the real aggregated matrix the 2 tools differ in the 3rd decimal
    and **rank the criteria identically**. They approximate the principal
    eigenvector differently, so this is expected. A rank disagreement would
    not be.
-   [ ] The consistency ratio is below 0.1, and if it is not, the result says
    so rather than reporting weights silently.
-   [ ] The weights sum to 1.

## 5. The scale-mismatch guard

::: callout-important
## Bug 6, found in a live builder session

AHP and ANP read reciprocal relative importance on the Saaty 1 to 9 ratio
scale. A DEMATEL influence item collects directed 0 to 4 influence with a
zero diagonal and no reciprocity. They are not interchangeable, and pairing
an influence item into an AHP block returns plausible weights from
meaningless input.

Guarded on 4 surfaces. Two of them are checked here.
:::

```{r}
#| label: guard-validate
# Wire the DEMATEL influence item into the AHP block, which must not work.
bad <- hotel
bad$analysis_plan <- list(list(
  id = "BAD", research_question = "AHP fed an influence matrix.",
  family = "decision", method = "ahp",
  roles = list(pairwise = "crit_influence")))

# validate_sframe() aborts on a failing instrument rather than returning a
# problem list, so the guard is caught rather than read off a field.
v <- tryCatch(validate_sframe(bad), error = function(e) e)

data.frame(
  check = c("validation refused the instrument", "message names the scales"),
  value = c(as.character(inherits(v, "error")),
            as.character(inherits(v, "error") &&
              grepl("saaty|influence|scale", conditionMessage(v),
                    ignore.case = TRUE)))
)
if (inherits(v, "error")) cat(conditionMessage(v), "\n")
```

```{r}
#| label: guard-runner
# And the runner itself must error rather than compute.
out <- tryCatch(run_analysis_plan(dat, bad)[["BAD"]],
                error = function(e) list(error = conditionMessage(e)))
out$error
```

-   [ ] `validate_sframe()` reports the mismatch at design time.
-   [ ] The runner errors rather than returning weights.
-   [ ] Both messages name the 2 scales, so a user can tell what to change.

## 6. TOPSIS

```{r}
#| label: topsis
top <- res[["RQ3"]]
top$table
```

```{r}
#| label: topsis-rmcda
# `matrix_source` is a label ("collected"), not the matrix, so rebuild the
# rated performance matrix the runner used.
rm0 <- sframe_rated_matrix(dat, hotel, c("rate_service", "rate_location",
                                         "rate_price", "rate_delivery"))
mat <- as.matrix(rm0$matrix)
w   <- as.numeric(top$weights)

rm_s <- RMCDA::apply.TOPSIS(mat, w)

tab <- do.call(compare_table, lapply(seq_along(top$scores), function(i) {
  compare_row(paste("TOPSIS score,", top$alternatives[i]),
              top$scores[i], as.numeric(rm_s)[i], "RMCDA::apply.TOPSIS()",
              tol = 1e-6)
}))
compare_report(tab, caption = "TOPSIS scores vs RMCDA")
sf_log(tab)
```

```{r}
#| label: topsis-plot
o <- order(top$scores, decreasing = TRUE)
barplot(top$scores[o], names.arg = top$alternatives[o], las = 2,
        col = "#41ab5d", border = NA, ylab = "closeness coefficient",
        main = "TOPSIS ranking")
```

-   [ ] Every TOPSIS score matches `RMCDA`.
-   [ ] The rank column agrees with the plot.

## 7. DEMATEL

```{r}
#| label: dematel
dm <- res[["RQ4"]]
data.frame(criterion = dm$criteria, D = dm$D, R = dm$R,
           prominence = dm$prominence, relation = dm$relation,
           role = dm$role)
```

```{r}
#| label: dematel-plot
plot(dm$prominence, dm$relation, type = "n",
     xlab = "prominence (D + R)", ylab = "relation (D - R)",
     main = "DEMATEL: causes above the line, effects below")
abline(h = 0, lty = 2, col = "#b2182b")
text(dm$prominence, dm$relation, labels = dm$criteria)
```

-   [ ] `D`, `R`, prominence, and relation are all reported.
-   [ ] The `role` column calls a criterion a cause when its relation is
    positive and an effect when it is negative, matching the plot.

## 8. The other 7 methods on the same decision

Same performance matrix and the same weights, so the methods can be compared
like for like.

```{r}
#| label: others
methods <- c("vikor", "moora", "smart", "waspas", "promethee", "electre",
             "anp")
build <- function(m) {
  i <- hotel
  i$analysis_plan <- list(list(
    id = "RQ", research_question = paste("Which supplier, by", m, "?"),
    family = "decision", method = m,
    roles = if (m == "anp") list(pairwise = "crit_pairs") else
      list(performance_items = c("rate_service", "rate_location",
                                 "rate_price", "rate_delivery"),
           weights_item = "crit_points")))
  i
}
out <- lapply(methods, function(m)
  tryCatch(run_analysis_plan(dat, build(m))[["RQ"]],
           error = function(e) list(error = conditionMessage(e))))
names(out) <- methods

data.frame(
  method = methods,
  ran = vapply(out, function(r) is.null(r$error), logical(1)),
  message = vapply(out, function(r) r$error %||% "", character(1))
)
```

-   [ ] All 7 ran. Every one of the 10 methods in the release is now
    accounted for: AHP, TOPSIS, DEMATEL above, and these 7 here.

```{r}
#| label: others-tables
for (m in methods) {
  cat("\n###", m, "\n")
  print(utils::head(out[[m]]$table, 5))
}
```

```{r}
#| label: others-rmcda
# VIKOR and WASPAS against RMCDA. WASPAS is checked because it is where
# RMCDA already caught a real bug: it was using SMART's normalisation.
wa <- out[["waspas"]]
w  <- as.numeric(wa$weights)
ty <- wa$criteria_types

# RMCDA::apply.WASPAS() cannot take this matrix. Its normalisation step
# divides an alternatives-by-criteria block by a 1-row summary, which only
# recycles correctly when exactly 1 criterion is beneficial. All 4 are
# beneficial here, so it errors. The package's own test suite uses RMCDA's
# single-beneficial worked example for that reason.
#
# The oracle here is therefore WASPAS from its definition: normalise, take
# the weighted sum and the weighted product, and blend them at lambda.
norm <- mat
for (k in seq_len(ncol(mat))) {
  norm[, k] <- if (ty[k] == "benefit") {
    mat[, k] / max(mat[, k])
  } else {
    min(mat[, k]) / mat[, k]
  }
}
wsm   <- as.numeric(norm %*% w)
wpm   <- apply(norm, 1, function(r) prod(r^w))
rm_wa <- 0.5 * wsm + 0.5 * wpm

tab <- do.call(compare_table, lapply(seq_along(wa$scores),
  function(i) compare_row(
    paste("WASPAS score,", wa$alternatives[i]),
    wa$scores[i], as.numeric(rm_wa)[i],
    "WASPAS from its definition", tol = 1e-6)))
compare_report(tab, caption = "WASPAS vs its own definition")
sf_log(tab)
```

::: callout-note
## Why this one is not checked against RMCDA here

WASPAS is where RMCDA earned its place: it caught surveyframe using SMART's
normalisation. That cross-check lives in
`tests/testthat/test-decision-ranking.R`, on RMCDA's own worked example. It
cannot be repeated on the hotel matrix because
`RMCDA::apply.WASPAS()` only normalises correctly when exactly 1 criterion
is beneficial, and all 4 are beneficial here. Checking against the
definition instead is the stronger test anyway, since it shares no code with
either package.
:::

```{r}
#| label: others-agreement
# Do the methods agree on a winner? Disagreement is normal and is exactly
# what sensitivity analysis in file 15 exists to quantify.
ranked <- Filter(Negate(is.null), lapply(names(out), function(m) {
  r <- out[[m]]
  if (is.null(r$alternatives) || is.null(r$ranks)) return(NULL)
  data.frame(method = m,
             winner = r$alternatives[which.min(r$ranks)])
}))
do.call(rbind, ranked)
```

-   [ ] Every WASPAS score matches `RMCDA`.
-   [ ] The winners table prints. Methods disagreeing is informative rather
    than wrong, and file `15` quantifies how stable each answer is.

## 9. ELECTRE reporting a non-result

::: callout-important
## B11, found during this release

ELECTRE reported an empty outranking relation as "all 9 alternatives jointly
best", and `sensitivity_analysis()` then called that result stable. The
strongest robustness signal the function can produce, from the weakest input
it can take.
:::

```{r}
#| label: electre
el <- out[["electre"]]
el$table
data.frame(
  check = c("degenerate flag present", "notes present"),
  value = c(as.character(!is.null(el$degenerate)),
            as.character(!is.null(el$notes)))
)
```

-   [ ] If the outranking relation is empty, the result says so rather than
    declaring every alternative jointly best.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] All 10 methods ran. Three are checked number for number against an
    independent oracle: AHP against a matrix with a known right answer,
    TOPSIS against `RMCDA`, and WASPAS against its own definition.
-   [ ] `vignettes/mcdm-analysis.Rmd` tells the same story as this file.
