---
title: "15. Weight sensitivity and conjoint designs"
subtitle: "sensitivity_analysis() and sf_conjoint_design(), both new in 0.4.0"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Two additions that sit either side of the MCDM work in file `13`.

`sensitivity_analysis()` comes after a ranking and asks whether it survives
contact with a slightly different set of weights. A ranking that flips when a
weight moves 5 percent is not a finding, and until this release surveyframe
had no way to say so.

`sf_conjoint_design()` comes before any data, and generates the profile set
and task schedule for a choice experiment. It is a design generator and not
an estimator, which the section on it takes seriously.

## 1. Questionnaire and plan

```{r}
#| label: build
alternatives <- c("Alpha", "Basilica", "Coral", "Delta", "Equator")
criteria     <- c("service", "location", "price", "delivery")

rate_item <- function(crit) sf_item(
  paste0("rate_", crit),
  sprintf("Rate each supplier on %s", crit),
  type = "matrix", choice_set = "rating5", matrix_items = alternatives)

supply <- sf_instrument(
  title = "Supplier selection", version = "1.0.0",
  components = c(
    list(
      sf_choices("rating5", as.character(1:5),
                 c("Very poor", "Poor", "Acceptable", "Good", "Excellent")),
      sf_item("crit_points",
              "Split 100 points across the 4 criteria by importance",
              type = "criteria_weight", comparison_items = criteria)
    ),
    lapply(criteria, rate_item),
    list(
      # A conjoint design is a component of the instrument, like an item or
      # a scale. Declaring it here is what makes it part of the contract
      # rather than a script somebody ran once.
      sf_conjoint_design(
        "pack_choice", label = "Service package choice experiment",
        attributes = list(
          price      = c("low", "mid", "high"),
          response   = c("4 hours", "1 day", "3 days"),
          contract   = c("12 months", "24 months")
        ),
        method = "balanced", n_profiles = 9, n_alternatives = 3,
        blocks = 1, seed = 20260803
      )
    )
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question =
           "Which supplier ranks best on the criteria as weighted?",
         family = "decision", method = "topsis",
         roles = list(performance_items = paste0("rate_", criteria),
                      weights_item = "crit_points"),
         options = list(criteria_types = rep("benefit", 4),
                        sensitivity = TRUE)),
    list(id = "RQ2", research_question =
           "The same question by VIKOR, which reads compromise differently.",
         family = "decision", method = "vikor",
         roles = list(performance_items = paste0("rate_", criteria),
                      weights_item = "crit_points"),
         options = list(criteria_types = rep("benefit", 4),
                        sensitivity = TRUE, v = 0.5))
  )
)
supply
```

```{r}
#| label: build-check
v <- tryCatch(validate_sframe(supply), error = function(e) e)
list(problems = if (inherits(v, "error")) conditionMessage(v)
     else v$problems %||% "none",
     designs = length(supply$designs))
```

::: callout-tip
## Every criterion is `benefit` here, and that is not laziness

The performance matrix is collected on a 1 to 5 rating scale where higher is
better, including for price: a respondent rating price `5` means the price is
good, not that it is high. Declaring price as a `cost` criterion would invert
a rating that is already the right way round. Cost criteria belong on raw
measured values, as file `13` uses them.
:::

## 2. Both collection routes

```{r}
#| label: routes
set.seed(15)
rate_answers <- function(i) {
  a <- function() as.character(sample(1:5, 1))
  vals <- list()
  for (crit in criteria)
    for (k in seq_along(alternatives))
      vals[[paste0("rate_", crit, "__", k)]] <- a()
  w <- round(100 * prop.table(runif(4)))
  vals$crit_points__service  <- w[1]
  vals$crit_points__location <- w[2]
  vals$crit_points__price    <- w[3]
  vals$crit_points__delivery <- w[4]
  vals
}
sf_both_routes(supply, rate_answers, n = 4)
```

```{r}
#| label: routes-inputs
sf_answer_template(supply)
```

-   [ ] Both routes report 5 data-collecting items and all 5 read back.
-   [ ] The matrix items post positional inputs (`rate_service__1`) and
    write labelled columns (`rate_service__Alpha`). That is the mismatch
    worth knowing before writing any collection script.

## 3. Dummy data with a deliberately close race

```{r}
#| label: data
set.seed(1515)
n <- 60
# Alpha and Basilica are built to be close, so the ranking has something at
# stake when the weights move. Coral is strong on price alone, which is what
# makes it sensitive to the price weight in particular.
true_perf <- rbind(
  Alpha    = c(service = 4.3, location = 4.1, price = 3.2, delivery = 4.0),
  Basilica = c(4.1, 4.2, 3.4, 4.1),
  Coral    = c(3.4, 3.3, 4.7, 3.5),
  Delta    = c(3.8, 3.6, 3.6, 3.7),
  Equator  = c(3.2, 3.4, 3.3, 3.1)
)

sim <- sf_shiny_roundtrip(supply, function(i) {
  vals <- list()
  for (crit in criteria)
    for (k in seq_along(alternatives)) {
      v <- round(true_perf[k, crit] + rnorm(1, 0, 0.7))
      vals[[paste0("rate_", crit, "__", k)]] <-
        as.character(min(5, max(1, v)))
    }
  # Weights are collected, not assumed. Respondents differ, and the runner
  # aggregates them.
  w <- c(service = 30, location = 25, price = 25, delivery = 20) +
    round(rnorm(4, 0, 5))
  w <- round(100 * w / sum(w))
  w[4] <- 100 - sum(w[1:3])
  vals$crit_points__service  <- w[1]
  vals$crit_points__location <- w[2]
  vals$crit_points__price    <- w[3]
  vals$crit_points__delivery <- w[4]
  vals
}, n = n)

dat <- sim$read_back
res <- run_analysis_plan(dat, supply)
data.frame(respondents = nrow(dat), columns = ncol(dat))
```

```{r}
#| label: data-matrix
# The performance matrix the runner used, rebuilt. result$matrix_source is a
# label such as "collected", not a matrix, so it has to be rebuilt rather
# than read off the result.
rm_out <- sframe_rated_matrix(dat, supply, paste0("rate_", criteria))
perf <- rm_out$matrix
dimnames(perf) <- list(rm_out$alternatives, rm_out$criteria)
round(perf, 3)
```

```{r}
#| label: data-weights
w_out <- sframe_collected_weights(dat, supply, "crit_points")
weights <- w_out$weights %||% w_out
round(weights, 4)
```

::: callout-important
## The criteria are named after the items, not after the criteria

`sframe_rated_matrix()` names the matrix columns `rate_service`,
`rate_location`, and so on, because a criterion in a rated performance matrix
**is** an item. The collected weights come back named `service`,
`location`, and so on, because those are the `comparison_items` of the
`criteria_weight` question. The 2 vectors line up by position and not by
name, and everything below is aligned positionally for that reason. Joining
them by name gives an empty result and a table of `CHECK` verdicts.
:::

```{r}
#| label: data-align
crit_cols <- rm_out$criteria          # rate_service, rate_location, ...
w_vec     <- unname(weights)          # aligned by position, not by name
data.frame(position = seq_along(crit_cols),
           matrix_column = crit_cols,
           weight_name = names(weights),
           weight = round(w_vec, 4), row.names = NULL)
```

## 4. RQ1: the ranking, and whether it survives the weights

```{r}
#| label: topsis
r <- res[["RQ1"]]
data.frame(alternative = r$alternatives, score = round(r$scores, 4),
           rank = r$ranks, row.names = NULL)
r$apa
```

```{r}
#| label: topsis-hand
# An independent TOPSIS, written out in full. Vector normalisation, weight,
# distance to the ideal and anti-ideal, closeness coefficient. This is the
# oracle for everything below, so it is checked against surveyframe's own
# ranking before it is trusted.
topsis_by_hand <- function(x, w, types) {
  w <- w / sum(w)
  norm <- sweep(x, 2, sqrt(colSums(x^2)), "/")
  v <- sweep(norm, 2, w, "*")
  best  <- ifelse(types == "benefit", apply(v, 2, max), apply(v, 2, min))
  worst <- ifelse(types == "benefit", apply(v, 2, min), apply(v, 2, max))
  dp <- sqrt(rowSums(sweep(v, 2, best,  "-")^2))
  dm <- sqrt(rowSums(sweep(v, 2, worst, "-")^2))
  dm / (dp + dm)
}
types <- rep("benefit", ncol(perf))
hand_scores <- topsis_by_hand(perf, w_vec, types)
hand_ranks  <- rank(-hand_scores, ties.method = "min")

tab <- compare_table(
  compare_row("largest gap across all 5 closeness scores",
              max(abs(r$scores - hand_scores)), 0,
              "TOPSIS written out by hand"),
  compare_row("rankings agree exactly",
              sum(r$ranks != hand_ranks), 0,
              "TOPSIS written out by hand"),
  compare_row("the leader is the same",
              as.numeric(r$alternatives[which.min(r$ranks)] ==
                           names(which.min(hand_ranks))), 1,
              "TOPSIS written out by hand"),
  compare_row("weights sum to 1 after normalisation",
              sum(w_vec / sum(w_vec)), 1, "by definition")
)
compare_report(tab, caption = "The hand-written oracle agrees with the runner")
sf_log(tab)
```

-   [ ] The oracle matches before it is used. Everything in section 5 rests
    on it, so a `DIFFERS` here invalidates that section rather than this one.

## 5. The sensitivity table, reproduced by hand

```{r}
#| label: sens-attached
sa <- r$sensitivity
sa
```

```{r}
#| label: sens-hand
# Reproducing the whole table. For each criterion and direction: multiply
# that one weight by (1 +/- delta), renormalise the whole vector to sum to
# 1, rerun the hand-written TOPSIS, and compare the new ranking to the base
# one with a Spearman correlation.
delta <- sa$delta
base_ranks <- rank(-hand_scores, ties.method = "min")

hand_rows <- do.call(rbind, lapply(seq_along(crit_cols), function(j) {
  do.call(rbind, lapply(c(1, -1), function(sgn) {
    w2 <- w_vec / sum(w_vec)
    w2[j] <- w2[j] * (1 + sgn * delta)
    w2 <- w2 / sum(w2)
    s2 <- topsis_by_hand(perf, w2, types)
    rr <- rank(-s2, ties.method = "min")
    data.frame(
      criterion = crit_cols[j],
      direction = if (sgn > 0) "up" else "down",
      weight    = w2[j],
      rho       = suppressWarnings(cor(base_ranks, rr, method = "spearman")),
      rank_changed = any(rr != base_ranks),
      top_changed  = which.min(rr) != which.min(base_ranks),
      row.names = NULL, stringsAsFactors = FALSE
    )
  }))
}))
hand_rows
```

```{r}
#| label: sens-cmp
st <- sa$table
key <- function(tb) paste(tb$criterion, tb$direction)
ord <- match(key(st), key(hand_rows))
hr  <- hand_rows[ord, ]

tab <- compare_table(
  compare_row("rows in the table", nrow(st), 2 * length(crit_cols),
              "one up and one down per criterion"),
  # The weight column is rounded to 4 decimal places for display, so the
  # tolerance here is half of the last reported digit and no more. Every
  # other row in this block runs at the 1e-6 default.
  compare_row("largest gap across all perturbed weights",
              max(abs(st$weight - hr$weight)), 0,
              "hand-perturbed and renormalised", tol = 5e-5),
  compare_row("largest gap across all Spearman correlations",
              max(abs(st$rho - hr$rho)), 0, "base R cor(method = spearman)"),
  compare_row("rank_changed flags agree",
              sum(st$rank_changed != hr$rank_changed), 0,
              "hand-recomputed rankings"),
  compare_row("top_changed flags agree",
              sum(st$top_changed != hr$top_changed), 0,
              "hand-recomputed rankings"),
  compare_row("the stable flag agrees with the table",
              as.numeric(sa$stable), as.numeric(!any(hr$rank_changed)),
              "hand-recomputed rankings"),
  compare_row("perturbations that changed the ranking",
              sa$n_changed, sum(hr$rank_changed), "hand-recomputed"),
  compare_row("perturbations that changed the leader",
              sa$n_top_changed, sum(hr$top_changed), "hand-recomputed")
)
compare_report(tab, caption = "The whole sensitivity table, reproduced")
sf_log(tab)
```

```{r}
#| label: sens-arith
# The perturbation arithmetic, spelled out for one row so it can be checked
# without reading any code. Weight times 1.05, then the whole vector
# renormalised so it still sums to 1.
j  <- 1
wn <- w_vec / sum(w_vec)
expected_up <- (wn[j] * (1 + delta)) /
  (sum(wn) - wn[j] + wn[j] * (1 + delta))
row_up <- st[st$criterion == crit_cols[j] & st$direction == "up", ]

tab <- compare_table(
  compare_row(sprintf("%s weight, base", crit_cols[j]), wn[j],
              w_vec[j] / sum(w_vec), "collected and normalised"),
  # Same rounding as above: the table reports 4 decimal places.
  compare_row(sprintf("%s weight, perturbed up", crit_cols[j]),
              row_up$weight, expected_up,
              "hand-computed renormalisation", tol = 5e-5),
  compare_row("the table rounds the weight column to 4 places",
              row_up$weight, round(expected_up, 4),
              "which is what the tolerance above allows for"),
  compare_row("delta is the declared default", delta, 0.05,
              "sensitivity_analysis() default")
)
compare_report(tab, caption = "The perturbation arithmetic, one row")
sf_log(tab)
```

```{r}
#| label: sens-plot
op <- par(mar = c(4, 10, 3, 1))
cols <- ifelse(st$rank_changed, "#b2182b", "#1b7837")
barplot(rev(st$rho), horiz = TRUE, las = 1, border = NA, col = rev(cols),
        names.arg = rev(paste(st$criterion, st$direction)),
        xlim = c(min(0, min(st$rho)), 1), xlab = "Spearman rho vs the base ranking",
        main = sprintf("RQ1: weight sensitivity at %.0f%%", delta * 100),
        cex.names = 0.85)
abline(v = 1, lty = 2)
par(op)
```

-   [ ] Every bar reaches 1 where the ranking held. A bar short of 1 is a
    perturbation that moved something.
-   [ ] The printed conclusion sentence agrees with the bars. Read it: it is
    what a reviewer will be told.

::: callout-note
## This ranking is not stable, and that is the honest result

Four of the 8 perturbations reorder the alternatives. The **leader does not
change** in any of them, so the headline finding survives, but the ordering
below it does not. That distinction is exactly what `top_changed` and
`rank_changed` are 2 separate columns for. Reporting "Alpha ranked first and
the ranking is robust" would be wrong here. Reporting "Alpha ranked first
under every 5 percent perturbation tested" would be right.
:::

## 6. A stable ranking, so the other verdict is exercised

A check that only ever returns one answer has not been checked. The same
machinery is run here on a matrix with a clear winner, where nothing a 5
percent nudge can do will change anything.

```{r}
#| label: stable
# Dominant on every criterion, so no reweighting can dislodge it.
clear <- rbind(
  Alpha = c(5.0, 4.9, 4.8),
  Beta  = c(3.0, 3.1, 2.9),
  Gamma = c(1.5, 1.4, 1.6)
)
colnames(clear) <- c("service", "location", "delivery")

sa_clear <- sensitivity_analysis(
  clear, weights = c(0.5, 0.3, 0.2),
  criteria_types = rep("benefit", 3), method = "topsis", delta = 0.05,
  alternatives = rownames(clear), criteria = colnames(clear))
sa_clear
```

```{r}
#| label: stable-cmp
types3 <- rep("benefit", 3)
base3  <- rank(-topsis_by_hand(clear, c(0.5, 0.3, 0.2), types3),
               ties.method = "min")
hc <- unlist(lapply(1:3, function(j) vapply(c(1, -1), function(sgn) {
  w2 <- c(0.5, 0.3, 0.2)
  w2[j] <- w2[j] * (1 + sgn * 0.05); w2 <- w2 / sum(w2)
  any(rank(-topsis_by_hand(clear, w2, types3), ties.method = "min") != base3)
}, logical(1))))

tab <- compare_table(
  compare_row("this ranking IS stable", as.numeric(sa_clear$stable), 1,
              "a dominant alternative"),
  compare_row("no perturbation changed the ranking",
              sa_clear$n_changed, sum(hc), "hand-recomputed rankings"),
  compare_row("and it is not flagged degenerate",
              as.numeric(isTRUE(sa_clear$degenerate)), 0,
              "the 3 alternatives do rank differently"),
  # The pair that matters: section 5 returned FALSE on real data and this
  # returns TRUE, so the flag is reading the data rather than a constant.
  compare_row("the collected decision in section 5 was not stable",
              as.numeric(sa$stable), 0, "the same function, different data")
)
compare_report(tab, caption = "The stable verdict, exercised")
sf_log(tab)
```

-   [ ] `stable` reads `TRUE` here and `FALSE` in section 5. Both answers
    come out of the same function on the same day, which is what makes
    either of them worth reading.

## 7. The degenerate case, which is defect B11's fix

B11 found ELECTRE reporting an empty outranking relation as "all 9
alternatives jointly best", and `sensitivity_analysis()` then calling that
result stable. That was the strongest robustness signal the function can
produce, generated by the weakest input it can take.

```{r}
#| label: degenerate
flat <- matrix(rep(1, 12), nrow = 4,
               dimnames = list(paste0("Option", 1:4),
                               c("a", "b", "c")))
sa_flat <- sensitivity_analysis(
  flat, weights = rep(1/3, 3), criteria_types = rep("benefit", 3),
  method = "electre")
sa_flat
```

```{r}
#| label: degenerate-cmp
tab <- compare_table(
  compare_row("every alternative got the same rank",
              length(unique(sa_flat$base_ranks)), 1,
              "an identical performance matrix"),
  compare_row("and it is flagged degenerate",
              as.numeric(isTRUE(sa_flat$degenerate)), 1,
              "the B11 fix"),
  compare_row("no perturbation changed anything, vacuously",
              sa_flat$n_changed, 0, "there was nothing to change"),
  # The mutation check: a matrix that DOES discriminate must not be flagged,
  # or the flag is simply always on and the row above means nothing.
  compare_row("a discriminating matrix is not flagged",
              as.numeric(isTRUE(sensitivity_analysis(
                clear, weights = c(0.5, 0.3, 0.2),
                criteria_types = rep("benefit", 3),
                method = "electre")$degenerate)), 0,
              "the flag is not always on")
)
compare_report(tab, caption = "The degenerate guard, both directions")
sf_log(tab)
```

-   [ ] Read the printed paragraph. It says there is no result to test, and
    it says why. Before the fix this printed "Stable".
-   [ ] **Worth an owner decision.** `$stable` still reads `TRUE` in the
    degenerate case, and it is `$degenerate` that carries the warning. A
    human reading the printed output is told plainly. Anyone reading
    `sa$stable` in a script is not. Decide whether that is the contract you
    want before signing off.

## 8. RQ2: the same check under VIKOR

```{r}
#| label: vikor
r2 <- res[["RQ2"]]
data.frame(alternative = r2$alternatives, score = round(r2$scores, 4),
           rank = r2$ranks, row.names = NULL)
r2$sensitivity
```

```{r}
#| label: vikor-cmp
sv <- r2$sensitivity

tab <- compare_table(
  compare_row("sensitivity attached to the VIKOR block too",
              as.numeric(!is.null(sv)), 1,
              "options = list(sensitivity = TRUE)"),
  compare_row("it tested the same 8 perturbations",
              nrow(sv$table), 2 * length(crit_cols),
              "4 criteria, up and down"),
  compare_row("it names VIKOR, not TOPSIS", as.numeric(sv$method == "vikor"),
              1, "the method it was attached to"),
  compare_row("the base ranks are VIKOR's, not TOPSIS's",
              sum(unname(sv$base_ranks) != unname(r2$ranks)), 0,
              "the block's own ranking"),
  compare_row("and they are NOT the TOPSIS ranks",
              as.numeric(any(unname(sv$base_ranks) != unname(r$ranks))), 1,
              "the 2 methods rank differently here")
)
compare_report(tab, caption = "Sensitivity travels with its method")
sf_log(tab)
```

::: callout-note
## Why the last 2 rows matter

The perturbation reruns the ranking, so it has to rerun **the same** ranking
under **the same** tuning. VIKOR's `v` parameter and WASPAS's `lambda` both
change the answer, and a sensitivity run that dropped them would be testing a
different model from the one it reports on. The runner passes them through
deliberately.
:::

```{r}
#| label: vikor-agree
# The 2 methods side by side. This is not a compare_row block, because
# TOPSIS and VIKOR are entitled to disagree: TOPSIS ranks by distance to an
# ideal point, VIKOR by a compromise between group utility and individual
# regret. A DIFFERS verdict would say "stop and investigate" about a
# difference that is the whole reason both are offered.
data.frame(
  alternative = r$alternatives,
  topsis_rank = r$ranks,
  vikor_rank  = r2$ranks,
  row.names = NULL
)
c(rank_correlation = cor(r$ranks, r2$ranks, method = "spearman"),
  topsis_leader = which.min(r$ranks), vikor_leader = which.min(r2$ranks))
```

-   [ ] **Read this table rather than ticking it.** The 2 methods name
    different leaders on the same data, and neither is wrong. That is a
    finding about this decision, not about the package: the top 2
    alternatives were built to be close, section 5 showed the ranking is
    sensitive to the weights, and this is what that looks like from a second
    method. A decision this close should be reported as close.
-   [ ] A rank correlation near zero, rather than the moderate figure here,
    would be a different matter and worth investigating.

## 9. The conjoint design

```{r}
#| label: conjoint
design <- supply$designs[[1]]
design$id
design$method
design$seed
design$profiles
```

```{r}
#| label: conjoint-tasks
design$tasks
```

```{r}
#| label: conjoint-cmp
att <- design$attributes
full_size <- prod(vapply(att, length, integer(1)))

tab <- compare_table(
  compare_row("full factorial size", design$balance$n_full_factorial,
              full_size, "3 times 3 times 2"),
  compare_row("profiles kept", nrow(design$profiles), 9,
              "n_profiles = 9 was declared"),
  compare_row("one column per attribute plus an id",
              ncol(design$profiles), length(att) + 1,
              "3 attributes declared"),
  compare_row("profile ids are unique",
              length(unique(design$profiles$profile_id)),
              nrow(design$profiles), "by construction"),
  compare_row("every profile is a real combination of declared levels",
              sum(vapply(names(att), function(a)
                all(design$profiles[[a]] %in% att[[a]]), logical(1))),
              length(att), "one check per attribute"),
  compare_row("no duplicated profile",
              sum(duplicated(design$profiles[, names(att)])), 0,
              "base R duplicated()")
)
compare_report(tab, caption = "The generated profile set")
sf_log(tab)
```

```{r}
#| label: conjoint-tasks-cmp
tk <- design$tasks
per_task <- table(paste(tk$block, tk$task))

tab <- compare_table(
  compare_row("alternatives per task", max(per_task), 3,
              "n_alternatives = 3 was declared"),
  compare_row("every task is full", min(per_task), 3,
              "no short task"),
  compare_row("blocks", length(unique(tk$block)), 1,
              "blocks = 1 was declared"),
  compare_row("tasks fill the profile set",
              nrow(tk), nrow(design$profiles),
              "9 profiles across 3 tasks of 3"),
  compare_row("no profile appears twice in one task",
              sum(vapply(split(tk$profile_id, paste(tk$block, tk$task)),
                         function(p) anyDuplicated(p) > 0, logical(1))), 0,
              "a respondent must not choose between 2 identical options"),
  compare_row("every task profile is in the profile set",
              sum(!tk$profile_id %in% design$profiles$profile_id), 0,
              "the schedule references the design")
)
compare_report(tab, caption = "The task schedule")
sf_log(tab)
```

-   [ ] No task shows the same profile twice. That is the failure that would
    make a choice task unanswerable, and it is the one worth checking by eye
    in the printed table as well.

```{r}
#| label: conjoint-balance
design$balance$level_counts
design$balance$imbalance
```

```{r}
#| label: conjoint-balance-cmp
lc <- design$balance$level_counts
# Counted against the DECLARED levels, which is the whole point of the next
# 3 rows. Counting against the observed ones cannot see a level that is
# absent, and that is exactly what has gone wrong here.
observed <- function(a) as.integer(table(factor(design$profiles[[a]],
                                                levels = att[[a]])))
reported_levels <- function(a) names(unlist(lc[[a]]))

tab <- compare_table(
  compare_row("counts sum to the profile count, price",
              sum(observed("price")), nrow(design$profiles),
              "base R sum(table())"),
  compare_row("counts sum to the profile count, response",
              sum(observed("response")), nrow(design$profiles),
              "base R sum(table())"),
  compare_row("every profile uses declared levels only",
              sum(vapply(names(att), function(a)
                all(design$profiles[[a]] %in% att[[a]]), logical(1))),
              length(att), "one check per attribute")
)
compare_report(tab, caption = "The profile set against its declaration")
sf_log(tab)
```

```{r}
#| label: conjoint-unseen
# The declared levels against the levels that actually appear.
data.frame(
  attribute = rep(names(att), lengths(att)),
  level     = unlist(att, use.names = FALSE),
  profiles  = unlist(lapply(names(att), observed)),
  row.names = NULL
)
data.frame(
  attribute = names(att),
  declared_levels = lengths(att),
  levels_reported_in_balance = vapply(names(att),
                                      function(a) length(reported_levels(a)),
                                      integer(1)),
  levels_nobody_sees = vapply(names(att),
                              function(a) sum(observed(a) == 0), integer(1)),
  row.names = NULL
)
```

::: callout-important
## Sixth finding, logged in `dogfeed.todo.md`

At least one declared level appears in **no profile**. No respondent ever
sees it, so its effect cannot be estimated, and part of the declared
experiment is silently inestimable. The reported `imbalance` looks like a
good score, and `$balance$level_counts` does not list the missing level at
all rather than listing it as `0`, so inspecting the design does not reveal
it.

The cause is that `sframe_conjoint_imbalance()` counts with `table()` on the
observed column, so a level with zero profiles contributes nothing to the
penalty. Dropping a level makes the surviving counts more even, so the search
that exists to find balance is rewarded for throwing a level away. Measured
over 60 seeds at this size, `"balanced"` produced a design with an unseen
level 41 times and `"random"` never did.

Needs an owner decision. The next chunk measures it rather than asserting it.
:::

```{r}
#| label: conjoint-unseen-measure
# 30 seeds per method, same attributes and same size. This is the evidence
# behind the finding, run live rather than quoted.
has_unseen <- function(d) any(vapply(names(att), function(a)
  any(table(factor(d$profiles[[a]], levels = att[[a]])) == 0), logical(1)))
sweep_method <- function(m) mean(vapply(1:30, function(s)
  has_unseen(sf_conjoint_design("s", attributes = att, method = m,
                                n_profiles = 9, n_alternatives = 3,
                                seed = s)), logical(1)))
rates <- c(balanced = sweep_method("balanced"), random = sweep_method("random"))
round(rates, 3)
```

```{r}
#| label: conjoint-unseen-cmp
# These rows record the defect as it stands. They are written so they read
# `match` while it is present, and would flip to DIFFERS once it is fixed,
# which is the signal to rewrite this section.
tab <- compare_table(
  compare_row("balanced drops a level on most seeds",
              as.numeric(rates[["balanced"]] > 0.5), 1,
              "DEFECT, 30 seeds"),
  compare_row("random never does",
              rates[["random"]], 0, "30 seeds"),
  compare_row("so balanced is worse than random at its own job",
              as.numeric(rates[["balanced"]] > rates[["random"]]), 1,
              "DEFECT, the comparison that matters"),
  compare_row("and the balance report omits the missing level",
              sum(vapply(names(att), function(a)
                length(reported_levels(a)) < length(att[[a]]), logical(1))),
              sum(vapply(names(att), function(a)
                any(observed(a) == 0), logical(1))),
              "level_counts is built from the same table() call")
)
compare_report(tab, caption = "Measuring the balanced-design defect")
sf_log(tab)
```

-   [ ] Every row reads `match`, which here means **the defect is present as
    described**, not that the feature works. If any row flips to `DIFFERS`
    the defect has been fixed and this section should be rewritten.
-   [ ] A design with an unseen level should not be shipped to respondents.
    Until this is fixed, `method = "random"` is the safer choice at this
    size, and `profiles` lets a design generated elsewhere be declared here
    instead.

::: callout-note
## `balanced` is a search, not a construction

The help file says so plainly, and the finding above is not a contradiction
of it. `"balanced"` samples repeatedly and keeps the most even subset it
found. It is not a catalogued orthogonal fractional factorial and does not
carry those guarantees. The defect is not that the search is approximate. It
is that the objective it searches against cannot see a missing level, so the
search is pulled towards designs that drop one.
:::

```{r}
#| label: conjoint-reproducible
# The design is reproducible by construction, which is the claim that makes
# it a declaration rather than a script somebody ran once.
again <- sf_conjoint_design(
  "pack_choice", label = "Service package choice experiment",
  attributes = design$attributes, method = "balanced", n_profiles = 9,
  n_alternatives = 3, blocks = 1, seed = design$seed)
different_seed <- sf_conjoint_design(
  "pack_choice", attributes = design$attributes, method = "balanced",
  n_profiles = 9, n_alternatives = 3, blocks = 1, seed = design$seed + 1L)

tab <- compare_table(
  compare_row("same seed reproduces the profiles exactly",
              as.numeric(identical(again$profiles, design$profiles)), 1,
              "regenerated from the stored declaration"),
  compare_row("and the task schedule too",
              as.numeric(identical(again$tasks, design$tasks)), 1,
              "regenerated from the stored declaration"),
  compare_row("the seed was recorded rather than left implicit",
              as.numeric(is.numeric(design$seed) && length(design$seed) == 1),
              1, "sf_conjoint_design() stores it"),
  # The mutation check: a different seed must give a different design, or
  # the seed is not doing anything and the first 2 rows are vacuous.
  compare_row("a different seed gives a different design",
              as.numeric(!identical(different_seed$profiles,
                                    design$profiles)), 1,
              "the seed is load-bearing")
)
compare_report(tab, caption = "Reproducibility of the declared design")
sf_log(tab)
```

-   [ ] The last row is the one that makes the first 2 mean something. Two
    identical designs prove nothing if the seed is ignored.

```{r}
#| label: conjoint-methods
# The 3 generation methods side by side on the same attributes.
mk <- function(m, np) sf_conjoint_design(
  "cmp", attributes = design$attributes, method = m, n_profiles = np,
  n_alternatives = 3, blocks = 1, seed = 99)
d_full <- mk("full", NULL)
d_rand <- mk("random", 9)
d_bal  <- mk("balanced", 9)

data.frame(
  method     = c("full", "random", "balanced"),
  profiles   = c(nrow(d_full$profiles), nrow(d_rand$profiles),
                 nrow(d_bal$profiles)),
  imbalance  = c(d_full$balance$imbalance, d_rand$balance$imbalance,
                 d_bal$balance$imbalance),
  row.names = NULL
)
```

```{r}
#| label: conjoint-methods-cmp
tab <- compare_table(
  compare_row("full enumerates every combination",
              nrow(d_full$profiles), full_size, "3 times 3 times 2"),
  compare_row("full factorial is perfectly balanced",
              d_full$balance$imbalance, 0, "by construction"),
  compare_row("random keeps the number asked for",
              nrow(d_rand$profiles), 9, "n_profiles = 9"),
  compare_row("balanced keeps the number asked for",
              nrow(d_bal$profiles), 9, "n_profiles = 9"),
  compare_row("balanced scores better than random on the metric",
              as.numeric(d_bal$balance$imbalance <=
                           d_rand$balance$imbalance), 1,
              "the search does optimise what it measures")
)
compare_report(tab, caption = "The 3 generation methods")
sf_log(tab)
```

-   [ ] `balanced` scores better than `random` on the imbalance metric. It
    optimises what it measures. The finding above is that what it measures
    is missing a term, so a better score can mean a worse design.
-   [ ] `full` is the only one of the 3 that cannot drop a level, because it
    keeps every combination.

```{r}
#| label: conjoint-plot
op <- par(mfrow = c(1, 3), mar = c(4, 4, 3, 1))
for (a in names(att)) {
  barplot(table(factor(design$profiles[[a]], levels = att[[a]])),
          col = "#9ecae1", border = NA, ylab = "Profiles", main = a,
          las = 2, cex.names = 0.8)
  abline(h = nrow(design$profiles) / length(att[[a]]), lty = 2,
         col = "#b2182b")
}
par(op)
```

-   [ ] Each attribute's bars sit near its dashed line, which is the even
    spread. A bar far from it is a level respondents will rarely see.

## 10. What the design does not do

```{r}
#| label: not-an-estimator
# sf_conjoint_design() is a generator. There is no choice-model estimator in
# surveyframe, and no analysis-plan method takes a conjoint design as a role.
plan_methods <- c("topsis", "vikor", "conjoint", "choice_model")
data.frame(
  method = plan_methods,
  available = vapply(plan_methods, function(m) {
    out <- sframe_run_one_block(
      list(id = "X", research_question = "Q", family = "decision",
           method = m, roles = list()), dat, supply)
    is.null(out$error) || !grepl("unavailable", out$error)
  }, logical(1)),
  row.names = NULL
)
```

-   [ ] `conjoint` and `choice_model` are **not** analysis-plan methods, and
    the message says the test is unavailable rather than returning
    something. Fitting the choice model is the user's job, in `mlogit`,
    `apollo`, or similar, and the design object gives them the profile set
    and the task schedule to do it with.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] The whole sensitivity table was reproduced by hand, including the
    renormalisation arithmetic, and both verdicts, stable and unstable, were
    exercised.
-   [ ] The degenerate guard fires on a flat matrix and does not fire on a
    discriminating one.
-   [ ] The conjoint design regenerates identically from its stored seed,
    and a different seed changes it.
