---
title: "16. Collection routes and branching"
subtitle: "Static HTML, Shiny, Google Sheets, conditional logic, and the serialisation round trip"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Every other file in this suite runs `sf_both_routes()` and moves on. This one
stops and looks at the routes themselves.

surveyframe can collect responses 3 ways: a self-contained HTML file you host
anywhere, a Shiny app, and a Google Sheet with a generated Apps Script
endpoint. All 3 must produce a response file the package can read back, with
the same columns, or an instrument means something different depending on how
it was fielded. 0.4.0 fixed one such mismatch already, and this file checks
whether any remain.

## 1. Questionnaire, with conditional logic

```{r}
#| label: build
travel <- sf_instrument(
  title = "Commuting survey", version = "1.0.0",
  components = list(
    sf_choices("yn", c("yes", "no"), c("Yes", "No")),
    sf_choices("modes", c("car", "bus", "train", "bike", "walk"),
               c("Car", "Bus", "Train", "Bicycle", "On foot")),

    sf_item("intro", "About your journey to work", type = "section_break",
            section_intro = "Six questions, about 2 minutes."),

    sf_item("drives", "Do you drive to work?", type = "single_choice",
            choice_set = "yn", required = TRUE),
    sf_item("miles",  "Roughly how many miles do you drive each week?",
            type = "numeric"),
    sf_item("why",    "What stops you driving?", type = "text"),
    sf_item("age",    "How old are you?", type = "numeric"),
    sf_item("young",  "Do you hold a provisional licence?",
            type = "single_choice", choice_set = "yn"),
    sf_item("used",   "Which of these have you used this month?",
            type = "multiple_choice", choice_set = "modes"),

    # Four branch rules, covering 4 of the 7 operators.
    sf_branch("miles", depends_on = "drives", operator = "==",
              value = "yes", action = "show"),
    sf_branch("why",   depends_on = "drives", operator = "==",
              value = "no",  action = "show"),
    sf_branch("young", depends_on = "age", operator = "<",
              value = 25, action = "show")
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question = "How many people drive to work?",
         family = "descriptive", method = "frequency",
         roles = list(variable = "drives")),
    list(id = "RQ2", research_question =
           "Among drivers, how far do they drive?",
         family = "descriptive", method = "descriptives",
         roles = list(variables = "miles"))
  )
)
travel
```

```{r}
#| label: build-branches
data.frame(
  item       = vapply(travel$branching, function(b) b$item_id, character(1)),
  depends_on = vapply(travel$branching, function(b) b$depends_on, character(1)),
  operator   = vapply(travel$branching, function(b) b$operator, character(1)),
  value      = vapply(travel$branching, function(b) as.character(b$value),
                      character(1)),
  action     = vapply(travel$branching, function(b) b$action, character(1)),
  row.names = NULL
)
```

::: callout-note
## Branches live in `$branching`, not `$branches`

Worth knowing before you go looking. The slot is `$branching`, and
`sframe_branch_lookup()` is the accessor the collectors use.
:::

## 2. The column contract, and whether the 3 routes agree on it

```{r}
#| label: contract
sf_columns(travel)
```

```{r}
#| label: contract-static
f_static <- tempfile(fileext = ".html")
suppressMessages(export_static_survey(travel, output_path = f_static,
                                      open = FALSE, overwrite = TRUE))
html <- paste(readLines(f_static, warn = FALSE), collapse = "\n")
c(file = f_static, kb = round(file.size(f_static) / 1024, 1))
```

```{r}
#| label: contract-sheets
gdir <- tempfile(); dir.create(gdir)
suppressMessages(export_google_sheet(
  travel, sheet_url = "https://docs.google.com/spreadsheets/d/EXAMPLE/edit",
  output_dir = gdir))
gs <- readLines(file.path(gdir, "surveyframe_collector.gs"))
expected_line <- grep("^const EXPECTED_COLUMNS", gs, value = TRUE)
sheet_cols <- strsplit(gsub('.*\\[|\\].*|"', "", expected_line), ",")[[1]]
sheet_cols
```

```{r}
#| label: contract-shiny
rt <- sf_shiny_roundtrip(travel, list(
  drives = "yes", miles = 140, why = "", age = 41, young = "no",
  used = c("car", "train")), n = 1)
shiny_cols <- names(rt$collected)
shiny_cols
```

```{r}
#| label: contract-cmp
declared <- sf_columns(travel)$column
meta     <- c("respondent_id", "started_at", "submitted_at")
sheet_items <- setdiff(sheet_cols, meta)
shiny_items <- setdiff(shiny_cols, meta)

tab <- compare_table(
  compare_row("Google Sheets writes every declared column",
              sum(!declared %in% sheet_items), 0,
              "the generated Apps Script"),
  compare_row("Shiny writes every declared column",
              sum(!declared %in% shiny_items), 0,
              "sframe_response_row()"),
  compare_row("static HTML carries every declared item id",
              sum(!vapply(sf_collecting_items(travel), function(i)
                grepl(i$id, html, fixed = TRUE), logical(1))), 0,
              "the exported page"),
  compare_row("Google Sheets writes nothing extra",
              length(setdiff(sheet_items, declared)), 0,
              "the generated Apps Script")
)
compare_report(tab, caption = "Do the 3 routes cover the contract?")
sf_log(tab)
```

```{r}
#| label: contract-extra
# The other direction: does any route write a column the instrument never
# promised? This is where the routes stop agreeing.
data.frame(
  route = c("Google Sheets", "Shiny"),
  extra_columns = c(
    paste(setdiff(sheet_items, declared), collapse = ", ") %||% "",
    paste(setdiff(shiny_items, declared), collapse = ", ")),
  row.names = NULL
)
```

```{r}
#| label: contract-extra-cmp
# These rows record the fourth finding as it stands, so they read `match`
# while the defect is present and would flip to DIFFERS once it is fixed.
display_only <- vapply(
  Filter(function(i) i$type %in% c("section_break", "text_block"),
         travel$items), function(i) i$id, character(1))

tab <- compare_table(
  compare_row("Shiny writes a column for the section break",
              sum(display_only %in% shiny_items), length(display_only),
              "DEFECT, sframe_response_row()"),
  compare_row("Google Sheets does not",
              sum(display_only %in% sheet_items), 0,
              "the generated Apps Script excludes it"),
  compare_row("the static serialiser does not either",
              as.numeric(grepl(
                "item.type==='section_break'||item.type==='text_block')return",
                html, fixed = TRUE)), 1,
              "inst/static_survey/template.html"),
  compare_row("so 2 of the 3 routes agree and 1 does not",
              length(setdiff(shiny_items, sheet_items)), length(display_only),
              "DEFECT, the disagreement measured")
)
compare_report(tab, caption = "Where the 3 routes disagree")
sf_log(tab)
```

::: callout-important
## Fourth finding, logged in `dogfeed.todo.md`

A `section_break` collects nothing, and 2 of the 3 routes already treat it
that way. The Apps Script generated by `export_google_sheet()` does not list
it in `EXPECTED_COLUMNS`, so a Shiny-collected row posted to that endpoint
would have the column dropped on arrival. `quality_report()` meanwhile counts
the Shiny column as item data, so every respondent carries a missing rate
having skipped nothing. File `02`, section 11 measures that half of it.

Needs an owner decision. The rows above are written so they will flip to
`DIFFERS` when it is fixed, which is the signal to rewrite this section.
:::

## 3. Branching, on both live routes

```{r}
#| label: branch-shiny
# The same instrument, 3 respondents whose answers send them down different
# paths. Every input is supplied every time, so anything that comes back NA
# came back NA because the branch hid it.
paths <- list(
  driver     = list(drives = "yes", miles = 140, why = "no reason",
                    age = 41, young = "no", used = c("car")),
  non_driver = list(drives = "no",  miles = 999, why = "no car",
                    age = 38, young = "no", used = c("bus", "train")),
  young_one  = list(drives = "no",  miles = 999, why = "learning",
                    age = 19, young = "yes", used = c("bike"))
)
collected <- do.call(rbind, lapply(names(paths), function(nm) {
  row <- sf_shiny_roundtrip(travel, paths[[nm]], n = 1)$collected
  cbind(path = nm, row[, c("drives", "miles", "why", "age", "young")])
}))
collected
```

```{r}
#| label: branch-cmp
g <- function(path, col) collected[[col]][collected$path == path]

tab <- compare_table(
  compare_row("driver: miles kept", as.numeric(g("driver", "miles")), 140,
              "drives == yes shows miles"),
  compare_row("driver: why hidden",
              as.numeric(is.na(g("driver", "why"))), 1,
              "drives == yes hides why"),
  compare_row("non-driver: miles hidden",
              as.numeric(is.na(g("non_driver", "miles"))), 1,
              "drives == no hides miles"),
  compare_row("non-driver: why kept",
              as.numeric(!is.na(g("non_driver", "why"))), 1,
              "drives == no shows why"),
  compare_row("age 41: young hidden",
              as.numeric(is.na(g("driver", "young"))), 1,
              "age < 25 is false"),
  compare_row("age 19: young kept",
              as.numeric(!is.na(g("young_one", "young"))), 1,
              "age < 25 is true"),
  compare_row("the unconditional item is never hidden",
              sum(is.na(collected$age)), 0,
              "age has no branch rule")
)
compare_report(tab, caption = "Branching on the Shiny route")
sf_log(tab)
```

-   [ ] An answer supplied for a hidden item is **discarded**, not stored.
    That is the behaviour you want: a respondent who never saw the question
    must not have an answer to it in the file.
-   [ ] The numeric `<` operator works on a numeric answer. Branch values
    are stored as declared, so a string comparison here would silently show
    the item to everybody.

```{r}
#| label: branch-static
# The static page evaluates the same rules in the browser. The rules have to
# reach the file for that to happen.
tab <- compare_table(
  compare_row("branch rules reach the exported page",
              sum(vapply(travel$branching, function(b)
                grepl(sprintf('"item_id":"%s"', b$item_id), html,
                      fixed = TRUE), logical(1))), length(travel$branching),
              "one check per rule"),
  compare_row("the depends_on item is named in the page",
              as.numeric(grepl('"depends_on":"drives"', html, fixed = TRUE)),
              1, "the exported page"),
  # The `<` operator is JSON-escaped to < on the way out, which is
  # correct and is what stops it being read as the start of a tag when the
  # JSON sits inside a <script> block. Grepping for a literal "<" here
  # returns nothing and looks like the operator was dropped.
  compare_row("the numeric operator survives the export",
              as.numeric(grepl('"operator":"\\\\u003c"', html)), 1,
              "the exported page, JSON-escaped"),
  compare_row("and its value is still a number, not a string",
              as.numeric(grepl('"value":25', html, fixed = TRUE)), 1,
              "the exported page")
)
compare_report(tab, caption = "Branching reaches the static route")
sf_log(tab)
```

-   [ ] **Open the exported page** and check by hand. Answer "No" to the
    driving question and confirm the mileage question disappears and the
    "what stops you" question appears. Then enter an age under 25 and
    confirm the licence question appears. Reading the JSON is not the same
    as watching it work.

```{r}
#| label: branch-path
f_static
```

## 4. The Google Sheets route

```{r}
#| label: sheets-script
cat(paste(head(gs, 22), collapse = "\n"))
```

```{r}
#| label: sheets-cmp
tab <- compare_table(
  compare_row("the target sheet URL is written into the script",
              as.numeric(any(grepl("EXAMPLE", gs))), 1,
              "the sheet_url argument"),
  compare_row("the script has a doPost handler",
              as.numeric(any(grepl("function doPost", gs))), 1,
              "an Apps Script web app needs one"),
  compare_row("it caps the POST body size",
              as.numeric(any(grepl("MAX_BODY_CHARS", gs))), 1,
              "a public endpoint needs a limit"),
  compare_row("it writes a header row from the expected columns",
              as.numeric(any(grepl("appendRow\\(EXPECTED_COLUMNS\\)", gs))),
              1, "the generated script"),
  compare_row("expected columns include the meta columns",
              sum(meta %in% sheet_cols), 3,
              "respondent_id, started_at, submitted_at"),
  compare_row("multi-select expands in the sheet contract too",
              sum(grepl("^used__", sheet_cols)), 5,
              "5 travel modes offered")
)
compare_report(tab, caption = "The generated Apps Script")
sf_log(tab)
```

-   [ ] The expansion columns are in the sheet contract. A collector that
    wrote `used = "car|train"` into one cell would produce a file
    `read_responses()` cannot read, which is the defect 0.4.0 fixed on the
    Shiny side.
-   [ ] **This route cannot be tested end to end here**, because it needs a
    real Google account and a deployed web app. The generated script is
    checked; the deployment is not. If you are fielding by this route, deploy
    it once and post a test response before you send the link to anybody.

## 5. Reading a response file back

```{r}
#| label: read-strict
set.seed(16)
n <- 150
raw <- do.call(rbind, lapply(seq_len(n), function(i) {
  drives <- sample(c("yes", "no"), 1, prob = c(0.55, 0.45))
  age    <- sample(18:64, 1)
  sf_shiny_roundtrip(travel, list(
    drives = drives, miles = sample(20:300, 1), why = "prefer the train",
    age = age, young = sample(c("yes", "no"), 1),
    used = sample(c("car", "bus", "train", "bike", "walk"),
                  sample(1:3, 1))), n = 1)$collected
}))
dim(raw)
```

```{r}
#| label: read-back
dat <- read_responses(raw, travel,
                      meta_cols = c("started_at", "submitted_at"))
res <- run_analysis_plan(dat, travel)
res
```

```{r}
#| label: read-cmp
tab <- compare_table(
  compare_row("every row read back", nrow(dat), n, "150 were collected"),
  compare_row("branch-hidden mileage is missing for non-drivers",
              sum(is.na(dat$miles[dat$drives == "no"])),
              sum(dat$drives == "no"), "every non-driver"),
  compare_row("and present for every driver",
              sum(is.na(dat$miles[dat$drives == "yes"])), 0,
              "every driver saw the question"),
  compare_row("the frequency block counts the same drivers",
              res[["RQ1"]]$table$Frequency[res[["RQ1"]]$table$Value == "Yes"],
              sum(dat$drives == "yes"), "base R sum()"),
  compare_row("the descriptives block used only drivers",
              res[["RQ2"]]$table$valid_n, sum(dat$drives == "yes"),
              "branch-hidden rows are missing, not zero")
)
compare_report(tab, caption = "Reading the collected file back")
sf_log(tab)
```

::: callout-important
## The last row is the one that matters for analysis

A branch-hidden answer is `NA`, not `0` and not an empty string. Every
descriptive and every test treats it as missing, so the mileage mean is the
mean among drivers rather than a mean diluted by people who do not drive. A
collector that wrote `0` for a hidden numeric would give a plausible, wrong
answer with no error anywhere.
:::

```{r}
#| label: read-rejects
# What read_responses() refuses. Strict import is the default, and it is what
# stops a mis-shaped file being analysed as though it were fine.
attempt <- function(x, ...) {
  out <- tryCatch(read_responses(x, travel, ...), error = function(e) e)
  if (inherits(out, "error")) substr(conditionMessage(out), 1, 90)
  else sprintf("accepted, %d rows", nrow(out))
}

bad_extra <- raw; bad_extra$stray_column <- 1
# What a rewritten header looks like: the double underscore in an expansion
# column turned into a single one. read.csv() does this to any name
# make.names() dislikes, and a matrix or multi-select whose option labels
# contain spaces is the usual casualty.
bad_names <- raw
names(bad_names) <- sub("used__", "used_", names(bad_names))

data.frame(
  scenario = c("collector timestamps not declared as meta",
               "an undeclared extra column",
               "expansion columns renamed on import",
               "the same file with strict = FALSE"),
  result = c(
    attempt(raw),
    attempt(bad_extra, meta_cols = c("started_at", "submitted_at")),
    attempt(bad_names, meta_cols = c("started_at", "submitted_at")),
    attempt(bad_extra, meta_cols = c("started_at", "submitted_at"),
            strict = FALSE)),
  row.names = NULL
)
```

```{r}
#| label: read-rejects-cmp
msgs <- c(attempt(raw),
          attempt(bad_extra, meta_cols = c("started_at", "submitted_at")),
          attempt(bad_names, meta_cols = c("started_at", "submitted_at")))
ok <- attempt(bad_extra, meta_cols = c("started_at", "submitted_at"),
              strict = FALSE)

tab <- compare_table(
  compare_row("all 3 mis-shaped files are rejected",
              sum(!grepl("^accepted", msgs)), 3,
              "strict import is the default"),
  compare_row("strict = FALSE accepts the extra column",
              as.numeric(grepl("^accepted", ok)), 1,
              "the escape hatch exists"),
  # The mutation check: a correct file must still be accepted.
  compare_row("a correct file is still accepted",
              as.numeric(grepl("^accepted", attempt(
                raw, meta_cols = c("started_at", "submitted_at")))), 1,
              "the working case")
)
compare_report(tab, caption = "Strict import, both directions")
sf_log(tab)
```

-   [ ] Read the third message. `read.csv()` rewrites headers by default, so
    `used__car` survives but a column with a character `make.names()`
    dislikes does not. The fix is `check.names = FALSE`, and it is worth
    knowing before you spend an afternoon on it.
-   [ ] `strict = FALSE` is available and is not the default. That is the
    right way round.

## 6. The instrument itself round-trips

0.4.0 fixed a defect where a freshly built instrument was not a
serialisation fixed point: identical content carried 2 different hashes
depending on whether it had been through a read. If the hash is not stable,
it is not an identity, and the whole integrity claim goes with it.

::: callout-important
## Where the hash actually lives, and a trap that cost this review a round

There is **no `$integrity` element** on an `sframe` object. A freshly built
instrument does not carry a hash, and neither does one returned by
`read_sframe()`. The hash exists only inside the `.sframe` file's JSON, and
`read_sframe()` verifies it on the way in.

That matters because `instrument$integrity$hash` returns `NULL` and
`identical(NULL, NULL)` is `TRUE`. A comparison written that way reads
`match` no matter what the package does, and proves nothing at all. The
stored value is read out of the JSON here instead.
:::

```{r}
#| label: hash-where
stored_hash <- function(path)
  jsonlite::fromJSON(paste(readLines(path, warn = FALSE), collapse = "\n"),
                     simplifyVector = FALSE)$hash$value

p1 <- tempfile(fileext = ".sframe")
write_sframe(travel, p1, overwrite = TRUE)
back <- read_sframe(p1)

p2 <- tempfile(fileext = ".sframe")
write_sframe(back, p2, overwrite = TRUE)
back2 <- read_sframe(p2)

p3 <- tempfile(fileext = ".sframe")
write_sframe(back2, p3, overwrite = TRUE)

c(is.null(travel$integrity), stored = substr(stored_hash(p1), 1, 16))
```

```{r}
#| label: hash-cmp
tab <- compare_table(
  compare_row("write, read, write gives the same hash",
              as.numeric(stored_hash(p1) == stored_hash(p2)), 1,
              "the 0.4.0 fixed-point fix", tol = 0),
  compare_row("and a third pass does not move it either",
              as.numeric(stored_hash(p2) == stored_hash(p3)), 1,
              "write, read, write, read, write", tol = 0),
  compare_row("the file is byte-identical, not merely equal in hash",
              as.numeric(identical(readLines(p1, warn = FALSE),
                                   readLines(p2, warn = FALSE))), 1,
              "base R readLines()", tol = 0),
  compare_row("the items survive intact",
              length(back$items), length(travel$items), "same instrument"),
  compare_row("the branch rules survive intact",
              length(back$branching), length(travel$branching),
              "same instrument"),
  compare_row("the analysis plan survives intact",
              length(back$analysis_plan), length(travel$analysis_plan),
              "same instrument"),
  compare_row("and the conjoint and model slots are preserved",
              length(back$designs) + length(back$models),
              length(travel$designs) + length(travel$models),
              "same instrument")
)
compare_report(tab, caption = "Serialisation is a fixed point")
sf_log(tab)
```

::: callout-note
## Why this was a defect worth fixing

Before 0.4.0 a plan block built by `sf_instrument()` carried only what the
caller supplied, while the reader filled in 8 further fields. So write, read,
write changed the hash of identical content, and an instrument could carry 2
different hashes depending on whether it had been through a read. That is a
poor property for a value that is meant to be the instrument's identity.
:::

```{r}
#| label: hash-mutation
# The mutation check, in the only form that means anything here: change the
# stored file and confirm the reader refuses it. Without this the rows above
# only show that a constant equals itself.
tampered <- sub('"Do you drive to work\\?"',
                '"Do you drive to your place of work?"',
                paste(readLines(p1, warn = FALSE), collapse = "\n"))
p_bad <- tempfile(fileext = ".sframe")
writeLines(tampered, p_bad)

reads <- function(path, ...) as.numeric(!inherits(
  tryCatch(read_sframe(path, ...), error = function(e) e), "error"))

tab <- compare_table(
  compare_row("the edit did change the file",
              as.numeric(!identical(readLines(p1, warn = FALSE),
                                    readLines(p_bad, warn = FALSE))), 1,
              "one question reworded by hand", tol = 0),
  compare_row("the untouched file reads",
              reads(p1), 1, "the working case"),
  compare_row("the edited file is rejected",
              reads(p_bad), 0, "the integrity check fires"),
  compare_row("and validate = FALSE does not bypass it",
              reads(p_bad, validate = FALSE), 0,
              "integrity is checked before validation")
)
compare_report(tab, caption = "The integrity check, both directions")
sf_log(tab)
tryCatch(read_sframe(p_bad), error = function(e) substr(conditionMessage(e), 1, 160))
```

-   [ ] Rewording one question in the stored file makes it unreadable, and
    the message says the file may have been modified. That is the integrity
    claim working.
-   [ ] `validate = FALSE` does **not** turn the integrity check off. It
    relaxes schema validation only. Worth knowing before anyone reaches for
    it to open a file that will not load.

```{r}
#| label: hash-bundled
# The 3 bundled instruments. If any of these stopped verifying, every stored
# .sframe file moved with it, so this is a release gate rather than a nice
# to have.
fixtures <- c("hotel_supplier_mcdm.sframe",
              "surveyframe_input_types_demo.sframe",
              "tourism_services_demo.sframe")
check_fixture <- function(f) {
  src <- sf_fixture(f)
  a <- read_sframe(src)                       # verifies the stored hash
  q <- tempfile(fileext = ".sframe")
  write_sframe(a, q, overwrite = TRUE)
  c(reads = TRUE, stable = stored_hash(src) == stored_hash(q))
}
res_fx <- t(vapply(fixtures, check_fixture, logical(2)))
res_fx
```

```{r}
#| label: hash-bundled-cmp
tab <- compare_table(
  compare_row("all 3 bundled instruments verify on read",
              sum(res_fx[, "reads"]), 3, "read_sframe()", tol = 0),
  compare_row("and rewriting each one reproduces its stored hash",
              sum(res_fx[, "stable"]), 3,
              "the hash the file already carried", tol = 0)
)
compare_report(tab, caption = "The bundled-instrument hash gate")
sf_log(tab)
```

-   [ ] The second row is the real gate. It compares each rewritten file
    against **the hash the shipped file already carried**, so a change in the
    serialisation format would show up here rather than being hidden by
    rewriting both sides.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK`.
-   [ ] The 3 routes cover the declared contract, and the one place they
    disagree is the section-break column, logged as the fourth finding.
-   [ ] Branching hides an item on both live routes and discards the answer
    rather than storing it.
-   [ ] Serialisation is a fixed point, for the instrument built here and
    for all 3 bundled ones, and the hash still moves when the instrument
    does.
