---
title: "18. The three graphical interfaces"
subtitle: "SurveyBuilder, SurveyStudio, and the dashboard. Needs a keyboard."
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

::: callout-important
## This file cannot be finished by rendering it

Everything below section 4 is a click path. The chunks check what can be
checked from R: that the files ship, that the state functions round-trip, and
that the decision item types reach both GUIs' internals. **None of that tells
you whether the interfaces work.**

Two rules from this release apply here and are worth reading before you
start.

1. **A browser reading of a Shiny widget is unreliable.** selectize.js
   replaces the real `<select>` element, so a headless browser reads a
   working dropdown as empty. That happened twice during 0.4.0 and cost 2
   rounds of investigation. Establish a UI defect in R first, and let the
   browser corroborate.
2. **A UI claim needs a click-path run with the screenshot read back.** Not
   an inspection of the code that draws it.
:::

## 1. What ships

```{r}
#| label: files
gui_files <- c(
  builder   = system.file("builder/survey_builder.html", package = "surveyframe"),
  studio    = system.file("shiny/app.R", package = "surveyframe"),
  dashboard = system.file("shiny/dashboard/app.R", package = "surveyframe")
)
data.frame(
  gui = names(gui_files),
  found = nzchar(gui_files),
  kb = ifelse(nzchar(gui_files), round(file.size(gui_files) / 1024, 1), NA),
  row.names = NULL
)
```

```{r}
#| label: files-cmp
tab <- compare_table(
  compare_row("all 3 interface files ship",
              sum(nzchar(gui_files)), 3, "system.file()"),
  compare_row("the builder is a single self-contained HTML file",
              as.numeric(nzchar(gui_files[["builder"]]) &&
                           !grepl('src="http', paste(
                             readLines(gui_files[["builder"]], warn = FALSE),
                             collapse = ""), fixed = TRUE)), 1,
              "no external script tags"),
  compare_row("the 6 launchers are exported",
              sum(c("launch_builder", "launch_studio", "launch_dashboard",
                    "launch_builder_demo", "launch_studio_demo",
                    "launch_dashboard_demo") %in%
                    getNamespaceExports("surveyframe")), 6,
              "3 launchers and 3 demo launchers")
)
compare_report(tab, caption = "The 3 interfaces ship")
sf_log(tab)
```

::: callout-note
## Why the builder being self-contained matters

The SurveyBuilder is one HTML file with no external requests. It runs from a
file path, on a laptop with no network, inside an institution that blocks
CDNs. That is a deliberate property and the check above is what protects it.
:::

## 2. The instrument used for the click paths

```{r}
#| label: build
# 2026-08-15: expanded again, from 18 items/7 blocks to 35 items/48 blocks,
# on request, to cover every statistical family the package ships, not a
# sample of them: all 4 descriptive methods, all 4 association/categorical
# methods, all 10 group_comparison methods (including both ANOVAs,
# ANCOVA, repeated-measures ANOVA, and Friedman), all 4 correlation
# methods, all 5 prediction/regression methods, all 5 psychometrics
# methods, all 3 model-syntax generators plus mediation and moderation,
# and all 10 decision methods. Saved as a real file rather than a
# tempdir() throwaway, so it can be loaded directly:
# `read_sframe("review_040/gui_review_full.sframe")`.
#
# Verified before this went in the file, the same discipline as the
# smaller version above: builds, validates with 0 problems, round-trips
# through write_sframe()/read_sframe() with all 3 models intact, reaches
# sframe_builder_state_from_instrument() and
# sframe_builder_validate_draft() with 0 problems, and **all 48
# analysis_plan blocks ran to completion (no `error` field) against 60
# rows of simulated responses** before this was written into the file.
# See the build-plan-run chunk below for that run.
lik <- function(id, label, scale) sf_item(id, label, type = "likert",
                                          choice_set = "agree5",
                                          scale_id = scale)

demo <- sf_instrument(
  title = "GUI and statistics review instrument", version = "1.0.0",
  components = list(
    sf_choices("agree5", as.character(1:5),
               c("Strongly disagree", "Disagree", "Neutral", "Agree",
                 "Strongly agree")),
    sf_choices("site", c("north", "south"), c("Northside", "Southside")),
    sf_choices("channels", c("call", "email", "chat", "in_person"),
               c("Phone", "Email", "Live chat", "In person")),
    sf_choices("priority", c("fast", "cheap", "good"),
               c("Fast delivery", "Low price", "High quality")),
    sf_choices("time_of_day", c("morning", "afternoon", "evening"),
               c("Morning", "Afternoon", "Evening")),
    sf_choices("yn", c("yes", "no"), c("Yes", "No")),
    sf_choices("mode_cs", c("online", "in_person", "hybrid"),
               c("Online", "In person", "Hybrid")),

    sf_item("site", "Which site?", type = "single_choice",
            choice_set = "site", required = TRUE),

    # 3 three-item scales, so a real measurement model (CFA/CB-SEM/PLS-SEM)
    # and a real 3-factor EFA both have something genuine to find, not just
    # 1 scale reused everywhere.
    lik("s1", "The service met my expectations.", "sat"),
    lik("s2", "Staff were helpful.", "sat"),
    lik("s3", "I would come back.", "sat"),
    lik("v1", "This service is worth what I paid.", "val"),
    lik("v2", "I get good value here.", "val"),
    lik("v3", "The price is fair for what I get.", "val"),
    lik("l1", "I would recommend this to a friend.", "loy"),
    lik("l2", "I consider myself a loyal customer.", "loy"),
    lik("l3", "I will keep using this service.", "loy"),

    sf_item("wait", "Minutes waited", type = "numeric"),

    # The 2 item types added in 0.4.0. Getting these into both GUIs was
    # defect 3 from the independent validation round. crit_influence is
    # the same 3 criteria on the influence scale, not the Saaty scale,
    # feeding DEMATEL, so the non-interchangeable-scales guard (defect 6)
    # has a genuine second item to be tested against, not just crit_pairs
    # reused.
    sf_item("crit_pairs", "Which criterion matters more?",
            type = "pairwise_comparison",
            comparison_items = c("Cost", "Speed", "Quality"),
            comparison_scale = "saaty"),
    sf_item("crit_influence", "How strongly does one criterion drive another?",
            type = "pairwise_comparison",
            comparison_items = c("Cost", "Speed", "Quality"),
            comparison_scale = "influence"),
    sf_item("crit_points", "Split 100 points across the criteria",
            type = "criteria_weight",
            comparison_items = c("Cost", "Speed", "Quality")),
    sf_item("rate_cost", "Rate each supplier on cost", type = "matrix",
            choice_set = "agree5",
            matrix_items = c("Alpha", "Basilica", "Coral")),
    sf_item("rate_speed", "Rate each supplier on speed", type = "matrix",
            choice_set = "agree5",
            matrix_items = c("Alpha", "Basilica", "Coral")),
    sf_item("rate_quality", "Rate each supplier on quality", type = "matrix",
            choice_set = "agree5",
            matrix_items = c("Alpha", "Basilica", "Coral")),

    # The remaining 7 collecting types, so all 13 are exercised somewhere
    # in this instrument.
    sf_item("channels", "Which channels have you used?",
            type = "multiple_choice", choice_set = "channels"),
    sf_item("comment", "One word for today's visit", type = "text",
            placeholder = "e.g. Smooth"),
    sf_item("notes", "Anything else you'd like to tell us?",
            type = "textarea"),
    sf_item("visit_date", "Date of your visit", type = "date",
            date_min = "2026-01-01", date_max = "2026-12-31"),
    sf_item("effort", "How much effort did this take? (0 to 10)",
            type = "slider", slider_min = 0, slider_max = 10,
            slider_step = 1),
    sf_item("rank_priorities", "Rank these in order of importance",
            type = "ranking", choice_set = "priority"),
    sf_item("overall_rating", "Overall rating of today's visit",
            type = "rating", rating_max = 5),
    sf_item("preferred_time", "When do you usually visit?",
            type = "single_choice", choice_set = "time_of_day"),

    # Categorical and paired-binary items, for chi-square, Fisher, McNemar,
    # and Cochran's Q.
    sf_item("member", "Are you a member?", type = "single_choice",
            choice_set = "yn"),
    sf_item("used_q1", "Did you use the app in week 1?",
            type = "single_choice", choice_set = "yn"),
    sf_item("used_q2", "Did you use the app in week 2?",
            type = "single_choice", choice_set = "yn"),
    sf_item("used_q3", "Did you use the app in week 3?",
            type = "single_choice", choice_set = "yn"),

    # A 3-level grouping factor, for one-way/two-way ANOVA, Kruskal-Wallis,
    # ANCOVA, and multinomial regression.
    sf_item("mode", "How did you attend?", type = "single_choice",
            choice_set = "mode_cs"),

    # 3 repeated numeric measures, for repeated-measures ANOVA and Friedman.
    sf_item("wk4", "Confidence at week 4 (0 to 100)", type = "numeric"),
    sf_item("wk8", "Confidence at week 8 (0 to 100)", type = "numeric"),
    sf_item("wk12", "Confidence at week 12 (0 to 100)", type = "numeric"),

    # A before/after pair, for the paired t-test and Wilcoxon.
    sf_item("confidence_pre", "Confidence before training (0 to 100)",
            type = "numeric"),
    sf_item("confidence_post", "Confidence after training (0 to 100)",
            type = "numeric"),

    sf_scale("sat", "Satisfaction", items = c("s1", "s2", "s3")),
    sf_scale("val", "Value", items = c("v1", "v2", "v3")),
    sf_scale("loy", "Loyalty", items = c("l1", "l2", "l3"))
  ),
  analysis_plan = list(
    # Descriptive family, all 4 methods.
    list(id = "RQ1", research_question = "How is site distributed?",
         family = "descriptive", method = "frequency",
         roles = list(variable = "site")),
    list(id = "RQ2", research_question = "What do wait and effort look like?",
         family = "descriptive", method = "descriptives",
         roles = list(variables = c("wait", "effort"))),
    list(id = "RQ3", research_question = "How satisfied are people?",
         family = "descriptive", method = "scale_descriptives",
         roles = list(variables = "sat")),
    list(id = "RQ4", research_question = "Where are the gaps in the data?",
         family = "descriptive", method = "missing_data",
         roles = list(variables = c("wait", "effort", "s1", "s2", "s3"))),
    list(id = "RQ5", research_question = "Is the data quality acceptable?",
         family = "descriptive", method = "quality",
         roles = list()),

    # Association/categorical, all 4 methods.
    list(id = "RQ6", research_question = "Does site relate to preferred time?",
         family = "association", method = "chi_square",
         roles = list(row = "site", column = "preferred_time")),
    list(id = "RQ7", research_question = "The same with an exact test.",
         family = "association", method = "fisher_exact",
         roles = list(row = "site", column = "member")),
    list(id = "RQ8", research_question =
           "Did app use change from week 1 to week 2?",
         family = "group_comparison", method = "mcnemar",
         roles = list(before = "used_q1", after = "used_q2")),
    list(id = "RQ9", research_question =
           "Did app use change across all 3 weeks?",
         family = "group_comparison", method = "cochran_q",
         roles = list(measures = c("used_q1", "used_q2", "used_q3"))),

    # Group comparison, all 10 methods: 2 independent, 2 paired, 2 one-way,
    # two-way, ANCOVA, and 2 repeated-measures.
    list(id = "RQ10", research_question = "Does wait time differ by site?",
         family = "group_comparison", method = "t_test_ind",
         roles = list(outcome = "wait", group = "site")),
    list(id = "RQ11", research_question = "Same, without assuming normality.",
         family = "group_comparison", method = "mann_whitney",
         roles = list(outcome = "wait", group = "site")),
    list(id = "RQ12", research_question =
           "Did confidence change after training?",
         family = "group_comparison", method = "t_test_pair",
         roles = list(before = "confidence_pre", after = "confidence_post")),
    list(id = "RQ13", research_question = "Same, without assuming normality.",
         family = "group_comparison", method = "wilcoxon_pair",
         roles = list(before = "confidence_pre", after = "confidence_post")),
    list(id = "RQ14", research_question = "Does satisfaction differ by mode?",
         family = "group_comparison", method = "anova_one",
         roles = list(outcome = "sat", group = "mode")),
    list(id = "RQ15", research_question = "Same, without assuming normality.",
         family = "group_comparison", method = "kruskal_wallis",
         roles = list(outcome = "sat", group = "mode")),
    list(id = "RQ16", research_question =
           "Do mode and site interact on satisfaction?",
         family = "group_comparison", method = "anova_two",
         roles = list(factor1 = "mode", factor2 = "site", outcome = "sat")),
    list(id = "RQ17", research_question =
           "Does mode affect satisfaction, controlling for wait?",
         family = "group_comparison", method = "ancova",
         roles = list(group = "mode", covariates = "wait", outcome = "sat")),
    list(id = "RQ18", research_question =
           "Does confidence change over 3 checkpoints?",
         family = "group_comparison", method = "repeated_anova",
         roles = list(measures = c("wk4", "wk8", "wk12"))),
    list(id = "RQ19", research_question = "Same, without assuming normality.",
         family = "group_comparison", method = "friedman",
         roles = list(measures = c("wk4", "wk8", "wk12"))),

    # Correlation, all 4 methods.
    list(id = "RQ20", research_question = "Does wait relate to effort?",
         family = "association", method = "correlation_pearson",
         roles = list(x = "wait", y = "effort")),
    list(id = "RQ21", research_question = "Same, on ranks.",
         family = "association", method = "correlation_spearman",
         roles = list(x = "wait", y = "effort")),
    list(id = "RQ22", research_question = "Same, with Kendall's tau.",
         family = "association", method = "correlation_kendall",
         roles = list(x = "wait", y = "effort")),
    list(id = "RQ23", research_question =
           "Does wait relate to effort once overall rating is controlled?",
         family = "association", method = "partial_correlation",
         roles = list(x = "wait", y = "effort", controls = "overall_rating")),

    # Prediction/regression, all 5 methods.
    list(id = "RQ24", research_question = "What predicts wait time?",
         family = "prediction", method = "regression_linear",
         roles = list(outcome = "wait",
                      predictors = c("effort", "overall_rating"))),
    list(id = "RQ25", research_question = "What predicts membership?",
         family = "prediction", method = "regression_logistic_binary",
         roles = list(outcome = "member",
                      predictors = c("effort", "overall_rating"))),
    list(id = "RQ26", research_question =
           "What predicts the first satisfaction item?",
         family = "prediction", method = "regression_logistic_ordinal",
         roles = list(outcome = "s1",
                      predictors = c("effort", "overall_rating"))),
    list(id = "RQ27", research_question = "What predicts attendance mode?",
         family = "prediction", method = "regression_logistic_multinomial",
         roles = list(outcome = "mode",
                      predictors = c("effort", "overall_rating"))),
    list(id = "RQ28", research_question =
           "What predicts membership, Firth-corrected?",
         family = "prediction", method = "firth_logistic",
         roles = list(outcome = "member", predictors = "effort")),

    # Psychometrics, all 5 methods.
    list(id = "RQ29", research_question = "Are the 3 scales reliable?",
         family = "psychometrics", method = "reliability_alpha"),
    list(id = "RQ30", research_question = "Same, by omega.",
         family = "psychometrics", method = "reliability_omega"),
    list(id = "RQ31", research_question = "Does any item weaken its scale?",
         family = "psychometrics", method = "item_diagnostics"),
    list(id = "RQ32", research_question = "Is this data suitable for EFA?",
         family = "psychometrics", method = "efa_readiness"),
    list(id = "RQ33", research_question = "Do 3 factors emerge as designed?",
         family = "psychometrics", method = "efa_solution",
         roles = list(items = c("s1", "s2", "s3", "v1", "v2", "v3",
                                 "l1", "l2", "l3")),
         options = list(nfactors = 3)),

    # Model syntax (all 3 generators) plus mediation and moderation.
    list(id = "RQ34", research_question =
           "Do the 9 items measure the 3 constructs they claim to?",
         family = "model", method = "cfa_lavaan_syntax",
         roles = list(model = "m_cfa")),
    list(id = "RQ35", research_question =
           "Does value carry part of satisfaction's effect on loyalty?",
         family = "model", method = "sem_lavaan_syntax",
         roles = list(model = "m_cb")),
    list(id = "RQ36", research_question =
           "Same structural question, by partial least squares.",
         family = "model", method = "seminr_syntax",
         roles = list(model = "m_pls")),
    list(id = "RQ37", research_question =
           "Does wait reach overall rating by way of effort?",
         family = "model", method = "mediation",
         roles = list(predictor = "wait", mediator = "effort",
                      outcome = "overall_rating"),
         options = list(bootstrap = 500)),
    list(id = "RQ38", research_question =
           "Is the wait effect on rating weaker at high effort?",
         family = "model", method = "moderation",
         roles = list(predictor = "wait", moderator = "effort",
                      outcome = "overall_rating")),

    # Decision, all 10 methods: 3 weight/classification methods (AHP, ANP,
    # DEMATEL) and 7 ranking methods sharing performance_items/weights_item.
    list(id = "RQ39", research_question = "Which criterion weighs most? (AHP)",
         family = "decision", method = "ahp",
         roles = list(pairwise = "crit_pairs")),
    list(id = "RQ40", research_question =
           "Same, allowing criteria to influence each other. (ANP)",
         family = "decision", method = "anp",
         roles = list(pairwise = "crit_pairs")),
    list(id = "RQ41", research_question =
           "Which criteria drive the others? (DEMATEL)",
         family = "decision", method = "dematel",
         roles = list(pairwise = "crit_influence")),
    list(id = "RQ42", research_question =
           "Which supplier ranks best overall? (TOPSIS)",
         family = "decision", method = "topsis",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ43", research_question =
           "Same, by VIKOR's compromise ranking.",
         family = "decision", method = "vikor",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ44", research_question = "Same, by MOORA.",
         family = "decision", method = "moora",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ45", research_question = "Same, by SMART.",
         family = "decision", method = "smart",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ46", research_question = "Same, by WASPAS.",
         family = "decision", method = "waspas",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ47", research_question =
           "Same, by PROMETHEE's net flows.",
         family = "decision", method = "promethee",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points")),
    list(id = "RQ48", research_question =
           "Same, by ELECTRE's outranking.",
         family = "decision", method = "electre",
         roles = list(
           performance_items = c("rate_cost", "rate_speed", "rate_quality"),
           weights_item = "crit_points"))
  )
)

# 3 declared models, added after construction (add_model()), same pattern
# as file 10. cb_sem's paths are labelled deliberately: an unlabelled cb_sem
# model with an indirect effect is defect 5, the one dogfeed finding that is
# a hard error in the single case a structural model is most often
# declared for, and this fixture should demonstrate the working shape, not
# the broken one.
constructs_reflective <- function() list(
  sf_construct("sat", label = "Satisfaction",
               items = c("s1", "s2", "s3"), mode = "reflective"),
  sf_construct("val", label = "Value",
               items = c("v1", "v2", "v3"), mode = "reflective"),
  sf_construct("loy", label = "Loyalty",
               items = c("l1", "l2", "l3"), mode = "reflective")
)
m_cfa <- sf_model("m_cfa", "Measurement model", type = "cfa",
                  constructs = constructs_reflective())
m_cb <- sf_model(
  "m_cb", "Structural model, covariance-based", type = "cb_sem",
  constructs = constructs_reflective(),
  paths = list(
    sf_path("sat", "val", label = "a"),
    sf_path("val", "loy", label = "b"),
    sf_path("sat", "loy", label = "c")
  ),
  indirect = list(sf_indirect("sat", through = "val", to = "loy"))
)
m_pls <- sf_model(
  "m_pls", "Structural model, partial least squares", type = "pls_sem",
  constructs = list(
    sf_construct("sat", items = c("s1", "s2", "s3"), mode = "composite"),
    sf_construct("val", items = c("v1", "v2", "v3"), mode = "composite"),
    sf_construct("loy", items = c("l1", "l2", "l3"), mode = "reflective")
  ),
  paths = list(sf_path("sat", "val"), sf_path("val", "loy"))
)
demo <- add_model(demo, m_cfa)
demo <- add_model(demo, m_cb)
demo <- add_model(demo, m_pls)

# Saved as a real file, not tempdir(), so it survives this session and can
# be loaded directly: read_sframe("review_040/gui_review_full.sframe").
inst_path <- file.path(getwd(), "gui_review_full.sframe")
write_sframe(demo, inst_path, overwrite = TRUE)
inst_path
```

```{r}
#| label: build-data
set.seed(18)
sim <- sf_shiny_roundtrip(demo, function(i) {
  a  <- function() as.character(sample(1:5, 1))
  yn <- function() sample(c("yes", "no"), 1)
  w <- round(100 * prop.table(runif(3))); w[3] <- 100 - sum(w[1:2])
  list(
    site = sample(c("north", "south"), 1),
    s1 = a(), s2 = a(), s3 = a(),
    v1 = a(), v2 = a(), v3 = a(),
    l1 = a(), l2 = a(), l3 = a(),
    wait = sample(5:60, 1),
    crit_pairs__Cost__vs__Speed = sample(1:9, 1),
    crit_pairs__Cost__vs__Quality = sample(1:9, 1),
    crit_pairs__Speed__vs__Quality = sample(1:9, 1),
    crit_influence__Cost__to__Speed = sample(0:4, 1),
    crit_influence__Cost__to__Quality = sample(0:4, 1),
    crit_influence__Speed__to__Cost = sample(0:4, 1),
    crit_influence__Speed__to__Quality = sample(0:4, 1),
    crit_influence__Quality__to__Cost = sample(0:4, 1),
    crit_influence__Quality__to__Speed = sample(0:4, 1),
    crit_points__Cost = w[1], crit_points__Speed = w[2],
    crit_points__Quality = w[3],
    rate_cost__1 = a(), rate_cost__2 = a(), rate_cost__3 = a(),
    rate_speed__1 = a(), rate_speed__2 = a(), rate_speed__3 = a(),
    rate_quality__1 = a(), rate_quality__2 = a(), rate_quality__3 = a(),
    channels = sample(c("call", "email", "chat", "in_person"),
                       size = sample(1:2, 1)),
    comment = sample(c("Smooth", "Slow", "Great", "Okay"), 1),
    notes = sample(c("", "Parking was hard to find.", "Keep it up."), 1),
    visit_date = as.character(as.Date("2026-01-01") + sample(0:300, 1)),
    effort = sample(0:10, 1),
    rank_priorities = paste(sample(c("fast", "cheap", "good")),
                             collapse = "|"),
    overall_rating = sample(1:5, 1),
    preferred_time = sample(c("morning", "afternoon", "evening"), 1),
    member = yn(), used_q1 = yn(), used_q2 = yn(), used_q3 = yn(),
    mode = sample(c("online", "in_person", "hybrid"), 1),
    wk4 = sample(20:60, 1), wk8 = sample(30:75, 1), wk12 = sample(40:90, 1),
    confidence_pre = sample(20:60, 1), confidence_post = sample(30:85, 1)
  )
}, n = 60)
resp_path <- file.path(getwd(), "gui_review_full_responses.csv")
write.csv(sim$collected, resp_path, row.names = FALSE)
c(instrument = inst_path, responses = resp_path, rows = nrow(sim$collected))
```

```{r}
#| label: build-plan-run
# All 48 blocks should run to completion against the simulated data. A
# block with an `error` field here means the click-through below would hit
# the same error inside SurveyStudio's Analysis screen, so this is checked
# before, not discovered during, the manual pass.
plan_results <- if (!inherits(sim$read_back, "error")) {
  run_analysis_plan(sim$read_back, demo)
} else {
  list()
}
plan_status <- data.frame(
  block = names(plan_results),
  status = vapply(plan_results, function(b) {
    if (is.null(b$error)) "OK" else paste("ERROR:", b$error)
  }, character(1)),
  row.names = NULL
)
plan_status
```

```{r}
#| label: build-plan-run-cmp
tab <- compare_table(
  compare_row("all 48 declared blocks are present in the results",
              length(plan_results), length(demo$analysis_plan), "48 declared"),
  compare_row("every block ran with no error field",
              sum(plan_status$status == "OK"), nrow(plan_status),
              "run_analysis_plan() on the simulated data")
)
compare_report(tab, caption = "Every method family runs against this fixture")
sf_log(tab)
```

::: callout-tip
## Keep these 2 paths open

Every click path below loads one or both. Copy them somewhere you can paste
from, because you will type them into 3 different file pickers.
:::

## 3. What can be checked without a browser

```{r}
#| label: builder-state
# The builder's own state functions, round-tripped. The GUI is a thin layer
# over these, so a failure here is a failure in the GUI whatever the screen
# shows.
empty <- sframe_builder_empty_state()
state <- sframe_builder_state_from_instrument(demo)

tab <- compare_table(
  compare_row("an empty state is produced",
              as.numeric(is.list(empty)), 1,
              "sframe_builder_empty_state()"),
  compare_row("every item reaches the builder state",
              length(state$items), length(demo$items), "35 declared"),
  compare_row("every choice set reaches it",
              length(state$choices), length(demo$choices), "7 declared"),
  compare_row("every scale reaches it",
              length(state$scales), length(demo$scales), "3 declared"),
  compare_row("the analysis plan reaches it",
              length(state$analysis_plan), length(demo$analysis_plan),
              "48 blocks declared")
)
compare_report(tab, caption = "Instrument to builder state")
sf_log(tab)
```

```{r}
#| label: builder-validate
# The draft validator the builder calls on every edit.
ok <- sframe_builder_validate_draft(
  meta = demo$meta, choices = demo$choices, items = demo$items,
  scales = demo$scales, branching = demo$branching, checks = demo$checks,
  analysis_plan = demo$analysis_plan, models = demo$models,
  render = demo$render)

# A draft that should NOT validate: an item pointing at a choice set that
# was never declared.
broken_items <- demo$items
broken_items[[1]]$choice_set <- "nosuchset"
bad <- sframe_builder_validate_draft(
  meta = demo$meta, choices = demo$choices, items = broken_items,
  scales = demo$scales)

str(ok, max.level = 1)
```

```{r}
#| label: builder-validate-cmp
n_problems <- function(x) length(x$problems %||% x$errors %||% list())

tab <- compare_table(
  compare_row("a good draft validates clean", n_problems(ok), 0,
              "the instrument built above"),
  # The mutation check. A validator that accepts everything is worse than
  # none, because it tells the user their draft is fine.
  compare_row("a draft with an undeclared choice set does not",
              as.numeric(n_problems(bad) > 0), 1,
              "choice_set = 'nosuchset'")
)
compare_report(tab, caption = "The builder's draft validator")
sf_log(tab)
bad$problems %||% bad$errors
```

```{r}
#| label: studio-levels
# Defect 3 and defect 6 both landed here. studio_level_meta() had no branch
# for the 2 decision item types, so SurveyStudio showed empty MCDM
# dropdowns, and the pairwise level had to be split into Saaty and influence
# so an AHP block could not be pointed at a DEMATEL item.
studio_src <- paste(readLines(system.file("shiny/app.R",
                                          package = "surveyframe"),
                              warn = FALSE), collapse = "\n")
has <- function(x) as.numeric(grepl(x, studio_src, fixed = TRUE))

tab <- compare_table(
  compare_row("SurveyStudio knows the pairwise item type",
              has("pairwise_comparison"), 1, "inst/shiny/app.R"),
  compare_row("and the criteria-weight item type",
              has("criteria_weight"), 1, "inst/shiny/app.R"),
  compare_row("the Saaty and influence levels are separate",
              has("pairwise_saaty") * has("pairwise_influence"), 1,
              "the defect 6 guard"),
  compare_row("studio_level_meta() has a branch for them",
              as.numeric(grepl(
                "studio_level_meta[\\s\\S]{0,4000}pairwise_comparison",
                studio_src, perl = TRUE)), 1,
              "inst/shiny/app.R")
)
compare_report(tab, caption = "The decision item types reached SurveyStudio")
sf_log(tab)
```

```{r}
#| label: builder-types
builder_src <- paste(readLines(gui_files[["builder"]], warn = FALSE),
                     collapse = "\n")
bhas <- function(x) as.numeric(grepl(x, builder_src, fixed = TRUE))

tab <- compare_table(
  compare_row("the builder knows the pairwise item type",
              bhas("pairwise_comparison"), 1, "survey_builder.html"),
  compare_row("and the criteria-weight item type",
              bhas("criteria_weight"), 1, "survey_builder.html"),
  compare_row("the scale split reached the builder too",
              bhas("pairwise_saaty") * bhas("pairwise_influence"), 1,
              "the defect 6 guard"),
  compare_row("the inspector can edit comparison_items",
              bhas("comparison_items"), 1, "survey_builder.html"),
  compare_row("and comparison_scale",
              bhas("comparison_scale"), 1, "survey_builder.html"),
  # The renderer keeps an "Unsupported item type" fallback, and should: it
  # is the last branch, for a type the builder has genuinely never heard of.
  # The fix was giving the 2 decision types their own branches BEFORE it, so
  # the check is that those branches exist, not that the fallback is gone.
  compare_row("the inlined template has a pairwise renderer",
              bhas("t==='pairwise_comparison'") * bhas("renderPairwise"), 1,
              "the inline-template fix"),
  compare_row("and a criteria-weight renderer",
              bhas("t==='criteria_weight'") * bhas("renderCriteria"), 1,
              "the inline-template fix"),
  compare_row("the unsupported-type fallback is still there for real gaps",
              bhas("Unsupported item type"), 1,
              "the last branch, and it should stay"),
  compare_row("all 10 MCDM methods are offered",
              sum(vapply(c("topsis", "ahp", "anp", "dematel", "vikor",
                           "moora", "smart", "waspas", "promethee",
                           "electre"), bhas, numeric(1)) > 0), 10,
              "the JS method registry")
)
compare_report(tab, caption = "The decision item types reached the builder")
sf_log(tab)
```

-   [ ] Every row reads `match`. **These are greps, and a grep is not a
    click.** They rule out the crudest failure, a type the GUI has never
    heard of, and nothing more. Sections 4 to 6 are the actual review.

## 4. SurveyBuilder click path

```{r}
#| label: launch-builder
#| eval: false
launch_builder()
```

Run the chunk above by hand. It opens one HTML file in your browser.

**Build from nothing.**

-   [ ] Add a question. The Add-question menu opens and **the whole menu is
    visible without scrolling**, including the Decision group at the bottom.
    That group was clipped out of view before commit 266f2a1.
-   [ ] Add one of each of the **13 collecting item types**, so every one
    gets a real click rather than being covered only by section 2's
    programmatic build: Likert, single choice, multiple choice, numeric,
    text, textarea, date, slider, rating, matrix, ranking, pairwise
    comparison, and criteria weight.
-   [ ] For the pairwise item, the inspector offers **comparison items** and
    **comparison scale**. Set 3 items and choose the Saaty scale.
-   [ ] For the criteria-weight item, set the same 3 items.
-   [ ] For the ranking item, the inspector lets you attach a choice set and
    the built survey lets a respondent drag items into order.
-   [ ] For the slider and rating items, their numeric bounds (min/max/step,
    or max stars) are editable in the inspector and carry through to Preview.
-   [ ] Open the **Preview** tab. Every question draws a real widget with a
    body, not an empty box, for **all 13 types**, not only the ones added in
    0.4.0. The empty preview body was fixed in 9fdeb0e and is the obvious
    thing to regress.
-   [ ] On the criteria-weight question in Preview, enter points that do
    **not** total 100 and try to advance. It refuses, and says why.
-   [ ] Export the survey. It downloads an HTML file, and opening it shows a
    real widget for the pairwise and criteria-weight questions rather than
    the grey "Unsupported item type" line. That fallback still exists and
    should, for a type the builder has genuinely never heard of. Seeing it on
    a decision item is the regression.
-   [ ] Open the exported file and answer every question, across all 13
    types. It submits.

**Round-trip an existing instrument.**

-   [ ] Import `gui_review_full.sframe`, the path printed in section 2 (a
    real file now, not a `tempdir()` one, so you can also just open it
    directly from `review_040/` without running any R first).
-   [ ] All 35 questions arrive, with their 7 choice sets, all 3 scales,
    all 3 models, and all 48 analysis-plan blocks (RQ1 to RQ48).
-   [ ] Open the analysis-plan editor and look at **RQ39**, the AHP block.
    Its **pairwise** role dropdown offers `crit_pairs` and is **not empty**.
    That dropdown being empty was defect 3.
-   [ ] Look at **RQ41**, the DEMATEL block already in the plan. Its
    **pairwise** role dropdown offers `crit_influence`, **not** `crit_pairs`,
    because that item is on the Saaty scale. If you edit it and the dropdown
    offers `crit_pairs` too, that is defect 6 regressing: confirm the 2
    scales are not interchangeable.
-   [ ] Look at **RQ42**, the TOPSIS block. Its **performance items** role
    offers `rate_cost`, `rate_speed`, and `rate_quality` (the `matrix` items,
    task B13's fix), and its **weights item** role offers `crit_points`.
    **RQ43 to RQ48** are the same 2 roles on VIKOR, MOORA, SMART, WASPAS,
    PROMETHEE, and ELECTRE in turn, so this is 6 more chances for the same
    dropdown to misbehave, not just 1.
-   [ ] Spot-check across the other families, since 48 blocks is too many to
    click every one: **RQ14/RQ16/RQ17** (`group` populated with `mode`,
    `factor1`/`factor2` with `mode`/`site`), **RQ23** (`controls` populated
    with `overall_rating`, the 1 role none of the other correlation blocks
    have), **RQ33** (`items` offering all 9 Likert items across the 3
    scales), and **RQ34 to RQ36** (a **model** role dropdown, not an item
    role, offering `m_cfa`/`m_cb`/`m_pls` by name). A model-role dropdown
    being empty or offering the wrong model type was defect 5's shape.
-   [ ] Export back to `.sframe` and reimport. Nothing is lost, across all
    35 items, all 3 models, and all 48 blocks.

::: callout-important
## The one to look hardest at

An AHP block pointed at a DEMATEL influence item returns plausible weights
from meaningless input, with no error anywhere. Defect 6. The guard is on 4
surfaces and the builder is one of them. If the dropdown offers you the wrong
item type, stop and log it.
:::

## 5. SurveyStudio click path

```{r}
#| label: launch-studio
#| eval: false
launch_studio(instrument = inst_path, responses = resp_path,
              meta_cols = c("started_at", "submitted_at"))
```

```{r}
#| label: launch-studio-demo
#| eval: false
# Or with the bundled demo, if you would rather start from something known.
launch_studio_demo()
```

**Build screen.**

-   [ ] The instrument loads and all 35 items are listed.
-   [ ] Select the pairwise item. The inspector shows its comparison items
    and its scale, and both are editable.
-   [ ] The analysis-plan editor's role dropdowns are **populated** for
    all 48 blocks, not only the decision ones. Empty MCDM dropdowns were
    defect 3 and 6. Spot-check a few from outside the decision family too:
    RQ16's `factor1`/`factor2` dropdowns (two-way ANOVA), RQ34's `model`
    dropdown (CFA syntax), RQ37's `predictor`/`mediator`/`outcome` triple
    (mediation).

**Preview screen.**

-   [ ] Every one of the 35 items renders a working widget, including the
    pairwise grid, the constant-sum question, the ranking drag list, and the
    slider and rating controls.
-   [ ] **Check the dropdowns by clicking them, not by looking.** A
    selectize dropdown that looks empty may be populated, and one that looks
    populated may be stale. This is the single most misleading surface in
    the package.

**Data screen.**

-   [ ] The 60 responses load and the column count matches the instrument's
    contract (57 columns, per the build-data chunk in section 2).
-   [ ] The expansion columns are present, one per pair, per criterion, per
    matrix row, per multiple-choice option, and per ranking option.

**Quality screen.**

-   [ ] The quality report renders and its respondent count is 60.
-   [ ] Note the missing-data figure. Nobody skipped anything, so anything
    above 0 is worth reading against the fourth finding in
    `dogfeed.todo.md`.

**Analysis screen.**

-   [ ] All 48 plan blocks run, spanning descriptive, association,
    categorical, group comparison (including both ANOVAs, ANCOVA,
    repeated-measures ANOVA, and Friedman), correlation, regression,
    psychometrics, model syntax, mediation and moderation, and all 10
    decision methods. Confirmed to run cleanly against the simulated data
    in section 2's `build-plan-run` chunk before this file asked you to
    click through it live; that chunk's `plan_status` table names the exact
    block if any one of them fails here but not there.
-   [ ] RQ39's (AHP) consistency ratio is reported, not hidden. Random
    pairwise judgements should produce a poor one, and it should say so.
-   [ ] RQ42 to RQ48 (the 7 ranking methods) each state their
    `matrix_source` and `weights_source` (both "collected" here), so every
    result names where its 2 inputs came from, not just TOPSIS.
-   [ ] RQ34 to RQ36 return generated **syntax as text**, not a fitted
    model. That is correct: surveyframe does not fit SEM itself, only
    declares it, per file `10`.

**Dashboard screen.**

-   [ ] It renders inside Studio and the plots are legible.

## 6. The standalone dashboard

```{r}
#| label: launch-dashboard
#| eval: false
launch_dashboard(instrument = inst_path, responses = resp_path)
```

-   [ ] It launches and shows the same figures as the Studio dashboard
    screen.
-   [ ] The plots are legible and the labels are not clipped.

::: callout-note
## An open decision, not a defect

Whether to keep or deprecate the standalone dashboard is one of the open
owner decisions carried into this release. It duplicates a screen inside
SurveyStudio. Use this section to form a view: if it gives you nothing
Studio does not, that is the answer.
:::

## 7. Accessibility

0.3.4 ran 2 WCAG 2.2 AA passes and 0.4.0 added the MCDM vignette, which is
axe-core clean. The 2 new item types are new interactive surfaces and were
not part of those passes.

-   [ ] **Keyboard only.** Unplug the mouse. Tab through the exported survey
    and reach every control, including every cell of the pairwise grid and
    the constant-sum inputs. Nothing is reachable only by clicking.
-   [ ] The focus outline is visible on every control, including inside the
    matrix.
-   [ ] The constant-sum error message is announced, not only coloured. A
    red border alone fails.
-   [ ] Zoom to 200 percent. The pairwise grid reflows rather than
    overflowing horizontally.
-   [ ] On a narrow window, the matrix reflows to the mobile `<select>`
    layout and stays usable.

## 8. What to do with what you find

-   [ ] Log every observation in `dogfeed.todo.md`, one item per entry,
    status `open`. Do not fix anything while reviewing.
-   [ ] For any UI defect, establish it in R first if you can, then attach
    the screenshot. A browser observation on its own has been wrong twice in
    this release.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK` on the automated rows.
-   [ ] **Every checkbox in sections 4 to 7 has been ticked by hand.** The
    verdict line above says nothing about the interfaces. If you have not
    opened all 3, this file is not complete.
