---
title: "19. The RStudio add-in"
subtitle: "Task H2. A release blocker, and it cannot be automated."
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

::: callout-important
## This file is a release blocker

Task **H2** is the owner verifying the add-in inside a real RStudio session.
It ships in 0.4.0. Nothing in this file, and nothing in the test suite, can
stand in for that, because the add-in's whole surface is the RStudio Addins
menu and the only way to know a menu entry works is to click it.

Everything that **can** be checked from R is checked below. Then section 4 is
the click path, and it needs you at a keyboard inside RStudio.
:::

## 1. Where the add-in lives, and how to get it

::: callout-important
## This section is stale as of 2026-08-15, kept for its automated checks

`feature/rstudio-addin` merged into `main` on 2026-08-03. The add-in is on
`main` and `dev` now, in the ordinary working tree, same as everything else
in this repository. **No worktree, no branch switch, no
`git worktree add` is needed.** Just `devtools::load_all()` (or better,
`devtools::install()`, see the tip below) from this repository's normal
checkout. The callout below this one, describing a separate worktree, is
what caused real confusion in the 2026-08-15 verification session and is
left in place only because the automated `on_addin_branch` check below still
works either way.
:::

The add-in was originally built on its own branch, before it merged into
`main`.

```{r}
#| label: locate
# Under devtools::load_all() the namespace path IS the package root, while
# for an installed package it is the library subdirectory. Walking up until
# a DESCRIPTION appears covers both.
sf_root <- local({
  p <- getNamespaceInfo("surveyframe", "path")
  for (i in 1:3) {
    if (file.exists(file.path(p, "DESCRIPTION"))) break
    p <- dirname(p)
  }
  p
})
addin_files <- c(
  bindings = file.path(sf_root, "inst", "rstudio", "addins.dcf"),
  source   = file.path(sf_root, "R", "rstudio_addins.R"),
  tests    = file.path(sf_root, "tests", "testthat",
                       "test-rstudio-addins.R")
)
data.frame(
  file = names(addin_files),
  present_in_the_loaded_build = file.exists(addin_files),
  row.names = NULL
)
```

```{r}
#| label: on-branch
on_addin_branch  <- all(file.exists(addin_files))
off_addin_branch <- !on_addin_branch
on_addin_branch
```

::: callout-tip
## If the row above reads `FALSE`, that is expected

You are on `v0.5-dev`, which does not carry the add-in. To review it, work in
a checkout of `feature/rstudio-addin`:

``` bash
cd ../surveyframe-v0.5-dev
git worktree add ../surveyframe-addin feature/rstudio-addin
```

Then open `../surveyframe-addin` in RStudio and follow section 4 there.

To make the automated rows in this file run, load that build **before**
`_setup.R` does its own search, by running this in the console and then
re-rendering:

``` r
devtools::load_all("../surveyframe-addin")
```

`_setup.R` only looks for a 0.4.0 build when one is not already loaded, so
loading the add-in build first is enough. The checks below were written and
verified against `feature/rstudio-addin` at commit `0194ffb`, where all of
them pass.

The branch is 1 commit ahead of `v0.5-dev` and 1 behind, so it needs a merge
from `v0.5-dev` before it goes anywhere near a tarball. The commit it is
missing is the `read.csv()` header explanation. **Merge first, then review**,
or you will review a build that is not the one shipping.
:::

## 2. What the 4 bindings are

```{r}
#| label: bindings
declared <- data.frame(
  menu_entry = c("Open SurveyBuilder", "Open SurveyStudio",
                 "Open Dashboard", "Insert sframe skeleton"),
  binding = c("addin_launch_builder", "addin_launch_studio",
              "addin_launch_dashboard", "addin_insert_skeleton"),
  interactive = c(TRUE, TRUE, TRUE, FALSE),
  row.names = NULL
)
declared
```

```{r}
#| label: bindings-cmp
#| eval: !expr on_addin_branch
dcf <- read.dcf(addin_files[["bindings"]])
tab <- compare_table(
  compare_row("4 menu entries are declared", nrow(dcf), 4,
              "inst/rstudio/addins.dcf"),
  compare_row("every declared binding exists in the package",
              sum(dcf[, "Binding"] %in% ls(asNamespace("surveyframe"),
                                           all.names = TRUE)), 4,
              "a binding with no function is a dead menu entry"),
  compare_row("the bindings are the 4 agreed ones",
              sum(sort(dcf[, "Binding"]) == sort(declared$binding)), 4,
              "todo_rstudio_addin.md"),
  compare_row("every entry has a description",
              sum(nzchar(dcf[, "Description"])), 4,
              "the menu shows it on hover")
)
compare_report(tab, caption = "The add-in bindings")
sf_log(tab)
```

```{r}
#| label: bindings-skipped
#| eval: !expr off_addin_branch
cat("Skipped. The loaded build does not carry the add-in.\n",
    "Re-run this file from a checkout of feature/rstudio-addin.\n")
```

## 3. The 2 rules that hold this file apart

The add-in source states 2 design rules. Both are checkable and both matter
more than the menu entries do.

```{r}
#| label: rules
#| eval: !expr on_addin_branch
r_dir <- file.path(sf_root, "R")
other_files <- setdiff(list.files(r_dir, pattern = "[.]R$", full.names = TRUE),
                       addin_files[["source"]])
leaks <- vapply(other_files, function(f)
  any(grepl("rstudioapi::", readLines(f, warn = FALSE), fixed = TRUE)),
  logical(1))
addin_src <- readLines(addin_files[["source"]], warn = FALSE)

tab <- compare_table(
  compare_row("no other file in R/ calls rstudioapi",
              sum(leaks), 0,
              "surveyframe must behave identically outside RStudio"),
  compare_row("rstudioapi is guarded before every use",
              as.numeric(any(grepl("requireNamespace\\(\"rstudioapi\"",
                                   addin_src))), 1,
              "it is in Suggests, not Imports"),
  compare_row("the bindings fail soft rather than erroring",
              as.numeric(sum(grepl("invisible\\(NULL\\)", addin_src)) >= 4),
              1, "one soft return per binding"),
  compare_row("rstudioapi is in Suggests, not Imports",
              as.numeric(grepl("rstudioapi",
                               paste(readLines(file.path(sf_root,
                                                         "DESCRIPTION"),
                                               warn = FALSE),
                                     collapse = "\n"))), 1,
              "DESCRIPTION")
)
compare_report(tab, caption = "The add-in stays out of the package's way")
sf_log(tab)
```

```{r}
#| label: skeleton
#| eval: !expr on_addin_branch
# The inserted skeleton, parsed and evaluated. A skeleton that does not
# build a valid instrument is worse than no skeleton, because the user's
# first experience of the package is an error they did not write.
skel <- surveyframe:::sframe_addin_skeleton()
cat(skel)
```

```{r}
#| label: skeleton-cmp
#| eval: !expr on_addin_branch
env <- new.env(parent = globalenv())
parsed <- tryCatch({ eval(parse(text = skel), envir = env); TRUE },
                   error = function(e) { message(conditionMessage(e)); FALSE })
built <- env$instrument
prob <- tryCatch(validate_sframe(built)$problems %||% list(),
                 error = function(e) list(conditionMessage(e)))

tab <- compare_table(
  compare_row("the skeleton parses and runs", as.numeric(parsed), 1,
              "eval(parse(text = skeleton))"),
  compare_row("it builds an sframe",
              as.numeric(inherits(built, "sframe")), 1, "class check"),
  compare_row("which validates with no problems", length(prob), 0,
              "validate_sframe()"),
  compare_row("it declares an analysis plan, not just items",
              length(built$analysis_plan), 1,
              "the pre-declared plan is the package's point"),
  compare_row("and the plan block runs on data",
              as.numeric(is.null(run_analysis_plan(
                read_responses(data.frame(q1 = sample(1:5, 40, TRUE),
                                          q2 = sample(1:5, 40, TRUE)),
                               built), built)[["RQ1"]]$error)), 1,
              "a skeleton whose own plan errors is not a starting point")
)
compare_report(tab, caption = "The inserted skeleton")
sf_log(tab)
```

::: callout-note
## Why the skeleton is checked this hard

The implementation guide's original skeleton was written against an API that
no longer exists. It passed `id =` to `sf_instrument()`, named the component
list `items =`, and handed `sf_item()` an inline `choices =` argument. Every
one of those is wrong now. A skeleton is the first code a new user runs, and
it arrives with the package's authority behind it.
:::

## 4. The click path, which is task H2

Do this in **RStudio**, not in a terminal, not in VS Code, and not in this
document.

**Setup.**

-   [ ] Open the `feature/rstudio-addin` checkout as an RStudio project.
-   [ ] `devtools::load_all()`, or install the branch with
    `devtools::install()`. Installing is the truer test, because the Addins
    menu reads `inst/rstudio/addins.dcf` from the installed package.
-   [ ] Confirm `rstudioapi` is installed. If it is not, install it, then
    come back to the last check in this section and remove it again.

**The menu.**

-   [ ] Open **Addins** in the RStudio toolbar. A **surveyframe** section
    appears.
-   [ ] All 4 entries are listed, with the names in section 2.
-   [ ] The descriptions read sensibly in the menu. They are the only
    explanation a user gets.

**Each entry.**

-   [ ] **Open SurveyBuilder.** The builder opens in a browser. RStudio
    stays responsive or shows the usual Shiny busy state, and does not hang
    with no explanation.
-   [ ] **Open SurveyStudio.** Studio opens. Stop it and confirm the R
    console returns to a prompt.
-   [ ] **Open Dashboard.** The dashboard opens. Stop it and confirm the
    console returns.
-   [ ] **Insert sframe skeleton.** Put the cursor in an empty R script and
    run it. The skeleton appears **at the cursor**, correctly indented, and
    does not overwrite anything.
-   [ ] Select some existing text first, then run it again. Confirm the
    behaviour is what you want, whether that is replacing the selection or
    inserting beside it. Decide and log it either way.
-   [ ] Run the inserted skeleton. It builds an instrument and
    `validate_sframe()` passes, with no editing.

**Failing soft.**

-   [ ] Remove `rstudioapi` (`remove.packages("rstudioapi")`), restart R,
    and run one of the add-ins. You get a **message** naming `rstudioapi`
    and telling you how to install it, and **not** an error. Then reinstall
    it.
-   [ ] Run `library(surveyframe)` in a plain R console outside RStudio.
    Nothing about the add-in loads, warns, or fails.

**The one that is easy to skip.**

-   [ ] Close every Shiny app and confirm no orphaned R process is left
    running. Three launchers in one menu is 3 ways to leave a port open.

## 5. What to record

```{r}
#| label: record
data.frame(
  check = c("Addins menu shows the surveyframe section",
            "All 4 entries present and correctly described",
            "SurveyBuilder launches and stops cleanly",
            "SurveyStudio launches and stops cleanly",
            "Dashboard launches and stops cleanly",
            "Skeleton inserts at the cursor",
            "Inserted skeleton runs and validates",
            "Fails soft without rstudioapi",
            "Nothing happens outside RStudio",
            "No orphaned processes after closing"),
  result = c("confirmed, screenshot",
             "confirmed, screenshot",
             "confirmed launches (console call post-install-fix); clean stop not separately observed",
             "confirmed launches (console call post-install-fix); clean stop not separately observed",
             "confirmed: correctly errors with no instrument supplied, the designed behaviour, not a launch failure",
             "not clicked at cursor this pass; skeleton content verified by parsing/evaluating it directly",
             "confirmed after the 2-item to 3-item fix: valid, 0 problems, full test-rstudio-addins.R 26/26 passed",
             "not re-tested live; covered by the automated suite's mocked-rstudioapi test",
             "not re-tested live; covered by the automated suite's rstudioapi-grep test",
             "not observed"),
  row.names = NULL
)
```

-   [x] Owner signed off 2026-08-15. **Not every row above is a live,
    in-session observation** — see the `result` column for exactly which
    ones are a direct click/observation versus a console call or reliance
    on the automated suite. Recorded honestly rather than marked as a
    blanket pass, per this file's own house rule that an unrecorded check
    is an unverified one.
-   [x] Logged in `dogfeed.todo.md`: none of the findings from this pass
    were left as open dogfeed items, because both were fixed on the spot
    and recorded in `todo_master_0.4.md`'s H2 entry instead — the
    `load_all()`-vs-`install()` Addins-menu quirk (environmental) and the
    2-item skeleton fix (a real, verified code change).

## 6. Before this branch merges

-   [ ] `feature/rstudio-addin` has been merged **from** `v0.5-dev` so it
    carries the whole release, not a snapshot of it.
-   [ ] The full suite passes on the merged branch. It was 1486 on the
    add-in branch and 1473 on `v0.5-dev` at the last count, so expect
    roughly 1486 after the merge.
-   [ ] The diff is still new files plus one `DESCRIPTION` line. Anything
    else in it needs explaining.
-   [ ] `R CMD check --as-cran` is clean on the merged branch, which file
    `20` covers.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS, 0 CHECK` on the automated rows, or every row skipped
    because the loaded build is `v0.5-dev`.
-   [ ] **Section 4 has been done at a keyboard in RStudio.** That is H2.
    Nothing above it discharges the task.
