---
title: "20. Release safety and the open decisions"
subtitle: "The pre-flight checks, R CMD check, and everything still waiting on the owner"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

The last file. Nothing here is about whether surveyframe computes the right
numbers, which is what the 18 files before it are for. This is about whether
the tarball is fit to send.

Sections 1 to 5 run automatically and are cheap. Section 6 is the long check
and has to be run by hand. Section 7 is the list of decisions only the owner
can take, and it is the reason this file exists.

```{r}
#| label: paths
# Under devtools::load_all() the namespace path IS the package root, while
# for an installed package it is the library subdirectory. Walking up until
# a DESCRIPTION appears covers both, where dirname(dirname()) covers only
# the second and silently lands 2 levels too high on a loaded source tree.
sf_pkg_root <- function() {
  p <- getNamespaceInfo("surveyframe", "path")
  for (i in 1:3) {
    if (file.exists(file.path(p, "DESCRIPTION"))) return(p)
    p <- dirname(p)
  }
  stop("Could not locate the package root from ",
       getNamespaceInfo("surveyframe", "path"), call. = FALSE)
}
pkg_root <- sf_pkg_root()
desc <- read.dcf(file.path(pkg_root, "DESCRIPTION"))
c(root = pkg_root, version = desc[1, "Version"])
```

## 1. The version number

::: callout-important
## The local branch label and the CRAN version differ on purpose

The branch is `v0.5-dev` and the release is CRAN **0.4.0**. That is an owner
decision from 2026-07-30 and nothing should be renamed to match. There is no
0.3.5 and no 0.5.x on CRAN.

`DESCRIPTION` is bumped to `0.4.0` **at release time only**, which is task
E1. Until then it reads `0.3.4`, and that is correct rather than an
oversight.
:::

```{r}
#| label: version
version <- desc[1, "Version"]
tab <- compare_table(
  compare_row("DESCRIPTION carries a version at all",
              as.numeric(nzchar(version)), 1, "DESCRIPTION"),
  compare_row("it is either the pre-bump 0.3.4 or the release 0.4.0",
              as.numeric(version %in% c("0.3.4", "0.4.0")), 1,
              "task E1 bumps it"),
  compare_row("a 0.4.0 feature is present whatever the version says",
              as.numeric(sf_is_040()), 1,
              "sensitivity_analysis() is new in this release")
)
compare_report(tab, caption = "The version number")
sf_log(tab)
```

-   [ ] If the version reads `0.3.4`, **E1 has not been done yet** and the
    tarball must not be built. If it reads `0.4.0`, E1 is done and the rest
    of this file is the pre-flight.

## 2. The hard blocker: `inst/CITATION`

```{r}
#| label: citation
cit_path <- file.path(pkg_root, "inst", "CITATION")
cit <- if (file.exists(cit_path))
  paste(readLines(cit_path, warn = FALSE), collapse = "\n") else ""
placeholder <- grepl("TBD|TODO|XXX|PLACEHOLDER|10\\.XXXX|FIXME", cit,
                     ignore.case = TRUE)
has_doi <- grepl("doi", cit, ignore.case = TRUE)
c(exists = file.exists(cit_path), has_doi_field = has_doi,
  has_placeholder = placeholder)
```

```{r}
#| label: citation-cmp
tab <- compare_table(
  compare_row("inst/CITATION exists",
              as.numeric(file.exists(cit_path)), 1, "the file"),
  compare_row("it parses as R", as.numeric(!inherits(tryCatch(
    parse(text = cit), error = function(e) e), "error")), 1,
    "CRAN evaluates it"),
  compare_row("it carries no placeholder text",
              as.numeric(placeholder), 0, "CRAN will not accept one"),
  compare_row("the package self-citation is present",
              as.numeric(grepl("surveyframe", cit, fixed = TRUE)), 1,
              "the Manual bibentry")
)
compare_report(tab, caption = "The citation file")
sf_log(tab)
```

::: callout-important
## D6 is the hard blocker, and it is not machine-checkable

`inst/CITATION` needs the **MCDM preprint DOI**, and CRAN will not accept a
placeholder. The rows above confirm there is no placeholder string in the
file. They cannot confirm the preprint exists, is posted, or has a DOI.

Target journal is *Operations Research and Decisions*, diamond open access.
The small-sample paper moved to 0.4.1 on 2026-07-31, so this is the only
paper gating 0.4.0.
:::

-   [ ] The MCDM preprint is posted and has a DOI.
-   [ ] That DOI is in `inst/CITATION` as a second `bibentry`.
-   [ ] The DOI resolves. Paste it into a browser and check, rather than
    trusting the string.

## 3. What must not be in the tarball

```{r}
#| label: buildignore
rbi <- readLines(file.path(pkg_root, ".Rbuildignore"), warn = FALSE)
# The dev-only files, as CLAUDE.md lists them.
dev_only <- c("CLAUDE.md", "roadmap.md", "revision_todo_0.3.md",
              "cran-comments.md", "todo.md", "dogfeed.todo.md",
              "todo_master_0.4.md", "todo_0.3.5.md", "todo_0.4.md",
              "todo_text_analysis.md", "todo_0.4.1.md", "todo_rstudio_addin.md",
              "mas_review_034.md", "mas_review_040.qmd",
              "kimi_review_034.md", "qwen_review_034.md")
present <- dev_only[file.exists(file.path(pkg_root, dev_only))]
ignored <- vapply(dev_only, function(f)
  any(vapply(rbi, function(p)
    grepl(p, f, perl = TRUE), logical(1))), logical(1))

data.frame(
  file = dev_only,
  present_on_this_branch = dev_only %in% present,
  matched_by_rbuildignore = unname(ignored),
  row.names = NULL
)
```

```{r}
#| label: buildignore-cmp
# The question that actually matters is not whether .Rbuildignore is
# complete, but whether anything it misses is on this branch. A gap in the
# list is harmless if the file is not there to be included.
at_risk <- dev_only[dev_only %in% present & !ignored[dev_only]]

tab <- compare_table(
  compare_row("dev-only files present but not ignored",
              length(at_risk), 0,
              "the ones that would reach the tarball"),
  compare_row("the review suite is not on the release branch",
              as.numeric(dir.exists(file.path(pkg_root, "review_040"))), 0,
              "review_040 lives in surveyframe-dev"),
  compare_row(".Rbuildignore excludes the check artefacts",
              as.numeric(any(grepl("Rcheck", rbi))), 1,
              "surveyframe.Rcheck must not ship"),
  compare_row("and any built tarball",
              as.numeric(any(grepl("tar\\\\.gz", rbi))), 1,
              "a tarball inside a tarball")
)
compare_report(tab, caption = "Nothing dev-only reaches the build")
sf_log(tab)
```

::: callout-note
## The known `.Rbuildignore` gaps, and why they do not bite here

CLAUDE.md records 2 hygiene gaps: `main`'s `.gitignore` names only the
original 6 dev-only files, and `.Rbuildignore` still names `todo_0.4.1.md`,
which was renamed to `todo_0.5.1.md` on 2026-07-25, while omitting
`kimi_review_034.md` and `qwen_review_034.md`.

Both are real and both should be closed. Neither affects **this** tarball,
because the release branch carries no planning files at all, which is what
the first row above measures. Close them anyway, before a tarball is ever
built from a branch that does carry them.
:::

-   [ ] The first row reads 0. If it does not, name the files in
    `.Rbuildignore` before building.
-   [ ] `.Rbuildignore`'s stale `todo_0.4.1.md` entry is corrected to
    `todo_0.5.1.md`, and the 2 review files are added.

## 4. Dependencies

```{r}
#| label: deps
field <- function(f) {
  v <- if (f %in% colnames(desc)) desc[1, f] else ""
  trimws(strsplit(v, ",")[[1]])
}
deps <- list(Imports = field("Imports"), Suggests = field("Suggests"),
             Depends = field("Depends"))
str(deps)
```

```{r}
#| label: deps-cmp
# Every line of R/, read once. readLines() takes one path at a time, so the
# files are read with lapply() and flattened rather than handed the vector
# list.files() returns.
r_source <- unlist(lapply(
  list.files(file.path(pkg_root, "R"), pattern = "[.]R$", full.names = TRUE),
  readLines, warn = FALSE))

# Comments and roxygen blocks are stripped before anything is counted as a
# call. Without this, `lavaan::cfa()` inside an @examples block and
# `knitr::kable()` inside an explanatory comment both read as unguarded
# calls to a Suggests package, which is a false failure and cost this file
# a round. The rule is crude, everything from the first # on a line, and it
# is stated rather than hidden because a # inside a string would also go.
r_code <- sub("#.*$", "", r_source)

imports <- gsub("\\s*\\(.*", "", deps$Imports)
imports <- imports[nzchar(imports)]
suggests <- gsub("\\s*\\(.*", "", deps$Suggests)
suggests <- suggests[nzchar(suggests)]
expected_hard <- c("jsonlite", "rlang", "openssl")

# Which Suggests packages the source actually calls, as opposed to merely
# mentions. RMCDA is a test-time oracle and seminr is named only inside a
# generated script, so neither appears here and neither needs a guard.
called <- suggests[vapply(suggests, function(p)
  any(grepl(paste0(p, "::"), r_code, fixed = TRUE)), logical(1))]
called

tab <- compare_table(
  compare_row("the hard imports are still the agreed 3",
              sum(expected_hard %in% imports), 3,
              "jsonlite, rlang, openssl"),
  compare_row("and nothing else crept into Imports",
              length(setdiff(imports, expected_hard)), 0,
              "CLAUDE.md's dependency rule"),
  compare_row("the optional packages this release added are in Suggests",
              sum(c("RMCDA", "logistf", "naniar", "pagedown") %in% suggests),
              4, "0.4.0 and 0.3.4 additions"),
  # Guarded means: every Suggests package the source actually CALLS is
  # behind a check_installed() or requireNamespace(). A package named only
  # in a generated syntax string is not called by surveyframe, and the
  # generated script carries its own check_installed() line, so counting
  # those as unguarded gives a false failure.
  compare_row("every Suggests package the source calls is guarded",
              sum(vapply(called, function(p) any(grepl(
                sprintf("(check_installed|requireNamespace)\\(\\s*[\"']%s", p),
                r_code)), logical(1))), length(called),
              "check_installed() or requireNamespace()"),
  compare_row("and at least 5 Suggests packages are actually called",
              as.numeric(length(called) >= 5), 1,
              "a vacuous pass if the list were empty")
)
compare_report(tab, caption = "The dependency contract")
sf_log(tab)
```

-   [ ] Imports is still 3 packages. Every method added in 0.4.0 landed in
    core with an optional package guarded at the call site, which is the
    rule that keeps surveyframe installable anywhere.

## 5. Examples and vignettes

```{r}
#| label: examples
rd <- list.files(file.path(pkg_root, "man"), pattern = "[.]Rd$",
                 full.names = TRUE)
launchers <- grep("launch_", rd, value = TRUE)
has_examples <- function(f) any(grepl("\\\\examples\\{",
                                      readLines(f, warn = FALSE)))
is_guarded <- function(f) {
  x <- paste(readLines(f, warn = FALSE), collapse = "\n")
  if (!grepl("\\\\examples\\{", x)) return(TRUE)          # no example, safe
  grepl("donttest|dontrun|open = FALSE|launch.browser = FALSE", x)
}
data.frame(
  file = basename(launchers),
  has_example = vapply(launchers, has_examples, logical(1)),
  safe = vapply(launchers, is_guarded, logical(1)),
  row.names = NULL
)
```

```{r}
#| label: examples-cmp
vig <- list.files(file.path(pkg_root, "vignettes"), pattern = "[.]Rmd$")

tab <- compare_table(
  compare_row("no launcher example opens a browser on a check machine",
              sum(!vapply(launchers, is_guarded, logical(1))), 0,
              "donttest, dontrun, or a FALSE argument"),
  compare_row("the vignettes are present",
              as.numeric(length(vig) >= 3), 1,
              "surveyframe, small-sample, mcdm-analysis"),
  compare_row("the MCDM vignette shipped",
              as.numeric("mcdm-analysis.Rmd" %in% vig), 1,
              "0.4.0's headline feature needs one"),
  compare_row("the small-sample vignette shipped",
              as.numeric("small-sample.Rmd" %in% vig), 1,
              "the other half of the release")
)
compare_report(tab, caption = "Examples and vignettes")
sf_log(tab)
vig
```

::: callout-note
## Why the launcher examples matter more than they look

A CRAN check machine has no display and no patience. An example that opens a
browser, or starts a Shiny app and waits, fails the check on machines that
have nothing to do with the bug. `launch_builder()`'s example uses
`open = FALSE` and inspects the path, which is the right shape: it exercises
the function without launching anything.
:::

## 6. `R CMD check --as-cran`

This is the long one. It is not run inside this document, because it takes
minutes, needs the version bumped first, and its output belongs in
`cran-comments.md` rather than here.

```{bash}
#| label: check
#| eval: false
cd ../surveyframe-v0.5-dev
R CMD build .
R CMD check --as-cran surveyframe_0.4.0.tar.gz
```

-   [ ] `0 errors, 0 warnings`.
-   [ ] At most **1 note**, the incoming-feasibility one. Anything else is
    investigated before submission, not explained away in
    `cran-comments.md`.
-   [ ] The tarball contains no `review_040/`, no planning file, no
    `.Rcheck` directory, and no nested tarball. `tar -tzf` it and read the
    list.
-   [ ] The full test suite passes. Expect roughly **1486** after the
    add-in branch is merged, against 1473 on `v0.5-dev` alone.
-   [ ] Both vignettes build offline. The data-collection chunks are
    `eval = FALSE` and the analysis runs on simulated or bundled data, so
    there should be no network access anywhere in the build.

```{r}
#| label: check-suite
#| eval: false
# Worth running on its own first, because it is faster than a full check and
# it is what usually breaks.
devtools::test()
```

**The other machines.**

-   [ ] win-builder, both `r-devel` and `r-release`.
-   [ ] The macOS builder.
-   [ ] Both come back with the same 1 note or fewer. A Windows-only failure
    on a package with 3 hard dependencies is usually a file-path or an
    encoding problem, and it is worth finding before CRAN does.

**The branch.**

-   [ ] `feature/rstudio-addin` has been merged from `v0.5-dev` and back,
    so the tarball carries both the release and the add-in. File `19`
    section 6 has the detail.
-   [ ] `NEWS.md` states plainly that 0.3.5 and 0.5 were planned and never
    released, since the version history jumps from 0.3.4 to 0.4.0.

## 7. The open decisions

Eight findings from this suite need an owner decision. None is fixed, which
is the dogfeed protocol working as intended. All 8 are in
`dogfeed.todo.md` with the evidence.

```{r}
#| label: decisions
decisions <- data.frame(
  n = 1:8,
  finding = c(
    "item_report() ignores reverse = TRUE",
    "correlation roles: `variables` works for Kendall, fails for Pearson and Spearman",
    "assumption_report() says checks were computed when none ran",
    "display-only items get a Shiny column and count as missing data",
    "sem_lavaan_syntax() writes an indirect effect lavaan cannot parse",
    "conjoint 'balanced' is rewarded for dropping a level entirely",
    "render_results(citation_format = ) is validated and then ignored",
    "codebook_report() omits the item help text"),
  found_in = c("09", "07", "05", "02 and 16", "10", "15", "17", "17"),
  lane = c("0.4.0", "0.4.0", "0.4.1", "0.4.0", "0.4.0", "0.4.0",
           "0.4.0", "0.4.1"),
  row.names = NULL
)
decisions
```

::: callout-important
## The 3 that would change published output

**5**, the CB-SEM indirect effect, is the one to take first. The generated
syntax does not run, in the single case a structural model is most often
declared for. It is a hard error rather than a quiet wrong number, so it
surfaces immediately, but it means the mediation route through `cb_sem` is
unusable as shipped for anyone who does not label their paths.

**6**, the conjoint design, is the most expensive to discover late. A
fielded design with a level nobody ever sees cannot be repaired after
collection, and `"balanced"` is worse than `"random"` at avoiding it.

**1**, the `item_report()` keying, changes numbers the package has already
printed for users. That is why it needs a decision rather than a fix.
:::

-   [ ] Each of the 8 has a decision recorded against it: fix in 0.4.0, lane
    to 0.4.1, or document as intended behaviour.
-   [ ] Every fix taken for 0.4.0 has a `NEWS.md` entry, and the ones that
    change numbers are marked breaking.
-   [ ] **7 has a cheap honest option.** Reducing `citation_format` to
    `"apa"` and saying in NEWS that the other 2 were never implemented costs
    almost nothing and removes a false claim from the manual. Leaving it
    documented and inert is the option that should not survive review.

## 8. The carried items, unrelated to a single release

-   [ ] The `.bib` reference carry-in decision. Leaning defer.
-   [ ] The Codecov badge.
-   [ ] Keep or deprecate the standalone dashboard. File `18` section 6 is
    where to form a view.
-   [ ] An interactive pass over `R/survey_module.R`.
-   [ ] **README's Roadmap section still promises small-sample at 0.4 and
    MCDM at 0.5.** Both ship in 0.4.0. That is on a public page and it is a
    5-minute fix.

```{r}
#| label: readme
readme <- paste(readLines(file.path(pkg_root, "README.md"), warn = FALSE),
                collapse = "\n")
tab <- compare_table(
  compare_row("README no longer promises MCDM at 0.5",
              as.numeric(grepl("0\\.5[^0-9]{0,40}MCDM|MCDM[^.]{0,40}0\\.5",
                               readme)), 0,
              "both ship in 0.4.0"),
  compare_row("README mentions the current release",
              as.numeric(grepl("0\\.4\\.0", readme, fixed = TRUE)), 1,
              "the version being shipped")
)
compare_report(tab, caption = "The public README")
sf_log(tab)
```

-   [ ] Both rows read `match` before `main` is pushed. The README is the
    first page anyone sees, and pkgdown rebuilds it on every push.

## 9. Sign-off

```{r}
#| label: signoff
data.frame(
  gate = c("All 19 review files run with 0 DIFFERS and 0 CHECK",
           "Every manual checkbox in 18, 19, and 20 is ticked",
           "H2: the add-in verified in a real RStudio session",
           "D6: the MCDM preprint DOI is in inst/CITATION",
           "All 8 open findings have a recorded decision",
           "E1: DESCRIPTION reads 0.4.0",
           "R CMD check --as-cran: 0 errors, 0 warnings, at most 1 note",
           "win-builder and macOS builder agree",
           "feature/rstudio-addin merged both ways",
           "NEWS.md complete, breaking changes marked",
           "README roadmap corrected"),
  done = "",
  row.names = NULL
)
```

-   [ ] Fill this in. An unrecorded gate is an unmet gate.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] `0 DIFFERS` on the automated rows, **except** the version row while
    E1 is outstanding and the README rows until that 5-minute fix is made.
    Both are expected to differ before release and to match after.
-   [ ] Section 6 has been run on a real machine and its output pasted into
    `cran-comments.md`.
-   [ ] Sections 7 and 8 have been read by the owner, not skimmed.
