---
title: "0.5 text-analysis review suite: start here"
subtitle: "Read this file first, then work through 01 and 02"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: false
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

## What this suite is for

surveyframe 0.5 adds one file, `R/text_analysis.R`, and 9 analysis-plan
method ids for open-ended text: `term_freq`, `ngram_freq`, `term_context`,
`co_occurrence`, `co_occurrence_network`, `tidy_sentiment`, `quanteda_dfm`,
`topic_model_lda`, `stm_topics`, plus the standalone `extract_quotes()`. No
new item types, no serialisation changes: a text/textarea item that already
existed just gained analysis coverage.

The scope is narrower than 0.4.0's, so this suite is proportionally smaller:
**2 files instead of 20.** The goal is the same: confirm every number
surveyframe reports for a text method against an independent computation,
built from first principles rather than by calling the package's own
internal helpers a second time.

`review_040/HANDOVER.md` documents the file template, the shared helpers,
and the token-saving rules this suite reuses unchanged. Read that once if
you have not; it is not repeated here.

## Order to work through

| File | Covers | Comparisons |
|---|---|---|
| `01_base_text_methods` | `term_freq` (incl. `group` role, word cloud), `ngram_freq`, `term_context`, `co_occurrence` — the 4 ids needing zero optional packages | 20 |
| `02_guarded_text_methods` | `tidy_sentiment`, `quanteda_dfm`, `co_occurrence_network` (seed determinism), `topic_model_lda` (skips, see below), `stm_topics`, `extract_quotes()` | 17 |

37 checked numbers total. There is no file `03`: this scope does not need
one, and padding the file count to look more like `review_040` would not add
review value.

## Reading a comparison

Same 3-word verdict column and log-scale plot as `review_040`. `match` means
the difference is at or below tolerance (default `1e-6`); `DIFFERS` means
stop and investigate; `CHECK` means one side returned `NA`, usually a
reference-call problem rather than a package one. See `review_040/HANDOVER.md`
section 4 (the reference-call traps table) before treating any `DIFFERS` in
this suite as a package defect — it may be the comparison's own reference
call that is wrong, the same way it was 9 times out of 17 in the 0.4.0 suite.

## Environment check

```{r}
#| label: env
data.frame(
  item = c("R", "platform", "surveyframe loaded from",
           "0.5 text-analysis feature present"),
  value = c(R.version.string, R.version$platform,
            local({
              p <- getNamespaceInfo("surveyframe", "path")
              for (i in 1:3) {
                if (file.exists(file.path(p, "DESCRIPTION"))) break
                p <- dirname(p)
              }
              p
            }),
            as.character(sf_is_050())),
  stringsAsFactors = FALSE
)
```

```{r}
#| label: refs
pk <- c("tidytext", "quanteda", "stm", "topicmodels", "igraph", "ggplot2")
data.frame(
  package = pk,
  version = vapply(pk, function(p)
    if (requireNamespace(p, quietly = TRUE))
      as.character(utils::packageVersion(p)) else "MISSING", character(1)),
  row.names = NULL
)
```

-   [ ] `0.5 text-analysis feature present` reads `TRUE`. If it reads
    `FALSE`, `_setup.R` fell back to an installed build with none of this
    work in it, and nothing below means anything.
-   [ ] `tidytext`, `quanteda`, `stm`, `igraph`, and `ggplot2` all report a
    version. **`topicmodels` is expected to read `MISSING` in this
    environment** — it is not installed here, confirmed via
    `Rscript -e 'requireNamespace("topicmodels", quietly=TRUE)'` returning
    `FALSE`. File `02`'s `topic_model_lda` section is guarded with the same
    `eval = requireNamespace(...)` pattern the vignette uses and skips
    cleanly rather than erroring; that skip is expected, not a finding.

::: callout-note
## Why there is no HTML/Shiny collection-route section here

`review_040`'s file template opens every file with both collection routes
(`sf_both_routes()`), because 0.4.0 changed how the Shiny collector wrote
matrix/ranking/multi-select columns. 0.5 adds no item type and no collector
change: a `text`/`textarea` item collects exactly as it always has, on both
routes, and that is already covered by `review_040`'s own field-type file.
Files `01` and `02` still build a real instrument and a real analysis plan
(the package's central claim), they just do not re-prove routes this release
did not touch.
:::

## Where to record what you find

Log findings in `dogfeed.todo.md`, one entry each, status `open`. Do not fix
anything while reviewing.
