---
title: "01. Base text methods"
subtitle: "term_freq, ngram_freq, term_context, co_occurrence — zero optional packages"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

Four method ids, all built on base R alone (no `tidytext`, `quanteda`,
`stm`, `topicmodels`, or `igraph`). Every reference in this file is
hand-tokenised from first principles, not a call to the package's own
`.sframe_tokenise()`, so a mismatch here cannot be explained away as "the
comparison called the same code it was checking."

Reference tokenisation used throughout: lower-case, strip everything except
letters, digits, spaces, and internal apostrophes, split on whitespace.
Every block below passes `stop_words = character(0)` on the surveyframe side
so the comparison does not also have to reproduce the package's built-in
English stop-word list to be independent.

## 1. Questionnaire and plan

```{r}
#| label: build
hotel <- sf_instrument(
  title = "Hotel feedback pilot", version = "1.0.0",
  components = list(
    sf_choices("dept", c("front_desk", "pool"),
               c("Front desk", "Pool area")),
    sf_item("dept", "Which area does your comment mainly concern?",
            type = "single_choice", choice_set = "dept", required = TRUE),
    sf_item("feedback", "Tell us about your stay in your own words",
            type = "textarea")
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question =
           "What are the leading terms in feedback, overall and by area?",
         family = "text_analysis", method = "term_freq",
         roles = list(item = "feedback", group = "dept"),
         options = list(stop_words = character(0))),
    list(id = "RQ2", research_question =
           "What 2-word phrases recur in the feedback?",
         family = "text_analysis", method = "ngram_freq",
         roles = list(item = "feedback"),
         options = list(stop_words = character(0), n = 2L)),
    list(id = "RQ3", research_question = "How is \"pool\" used in context?",
         family = "text_analysis", method = "term_context",
         roles = list(item = "feedback", term = "pool")),
    list(id = "RQ4", research_question =
           "Which terms co-occur within a single response?",
         family = "text_analysis", method = "co_occurrence",
         roles = list(item = "feedback"),
         options = list(stop_words = character(0), top_n = 12L))
  )
)
hotel
```

::: callout-note
## Why the `dept` item matters beyond providing a group role

`dept`'s 2 choice codes are deliberately the ordinary English words
`"front_desk"` and `"pool"`, the same words that show up naturally in the
free-text feedback. That is not an accident of this file: it is what
surfaces a real defect in section 4 below. Keep it in mind reading every
section past that point.
:::

## 2. Collection route

0.5 added no item type and no collector change (see `00_start_here.qmd`),
so this file checks the HTML route only rather than the full
`sf_both_routes()` proof `review_040` runs for every item type.

```{r}
#| label: route
sf_html_route(hotel)
```

-   [ ] `feedback` and `dept` both appear in the exported HTML.

## 3. Dummy data

```{r}
#| label: data
set.seed(77)
fd <- c(
  "the front desk staff were quick and helpful during check in",
  "front desk staff helped us change rooms without any hassle",
  "checkout was slow because front desk staff were understaffed",
  "front desk staff gave clear directions around the hotel",
  "front desk staff upgraded our room without being asked",
  "front desk staff fixed our key card fast and were friendly",
  "checkout was quick this time, front desk staff were efficient",
  "front desk staff were friendly and checked us in fast",
  "the front desk staff answered every question we had",
  "front desk staff handled a billing mistake very well",
  "check in was smooth, front desk staff were professional",
  "front desk staff remembered our name the whole stay"
)
pool <- c(
  "the pool area was clean but the pool towels were dirty",
  "loved the pool, the pool bar service was excellent",
  "the pool was closed for cleaning during our whole stay",
  "pool towels ran out twice but staff fixed it fast",
  "the pool area needs more shade, otherwise a great pool",
  "the pool was crowded every afternoon we visited",
  "pool water was cold but the pool area was clean",
  "the pool bar ran out of towels most days",
  "great pool, the pool area could use more shade",
  "the pool was lovely, loved the quiet pool area",
  "pool towels were never available near the pool",
  "the pool closed early most nights, pool area was clean"
)
comments <- c(fd, pool)
dept <- c(rep("front_desk", 12), rep("pool", 12))

responses <- data.frame(dept = dept, feedback = comments,
                        stringsAsFactors = FALSE)
dat <- read_responses(responses, hotel)
res <- run_analysis_plan(dat, hotel)
names(res)
```

```{r}
#| label: ref-tokeniser
# The single reference tokeniser every section below uses: lower-case, strip
# non [a-z0-9' ], split on whitespace. Written fresh, not a call into
# .sframe_tokenise().
ref_tokenise <- function(x) {
  x <- tolower(as.character(x))
  x <- gsub("[^a-z0-9' ]", " ", x)
  toks <- strsplit(trimws(x), "\\s+")
  lapply(toks, function(t) t[nzchar(t)])
}
ref_freq <- function(text) sort(table(unlist(ref_tokenise(text))), decreasing = TRUE)
```

## 4. RQ1: term frequency, overall and grouped

```{r}
#| label: acc-termfreq
r1 <- res[["RQ1"]]
refAll <- ref_freq(dat$feedback)
refFD  <- ref_freq(dat$feedback[dat$dept == "front_desk"])
refPool <- ref_freq(dat$feedback[dat$dept == "pool"])

# "staff" is common to both groups but never collides with a choice code,
# so these rows are a clean read of the term-counting logic on its own.
sf_staff_fd <- r1$table$n[r1$table$group == "Front desk" & r1$table$term == "staff"]
sf_towels_pool <- r1$table$n[r1$table$group == "Pool area" & r1$table$term == "towels"]

tab <- compare_table(
  compare_row("front_desk group: n('staff')", sf_staff_fd, refFD[["staff"]],
              "hand-tokenised table()"),
  compare_row("pool group: n('towels')", sf_towels_pool, refPool[["towels"]],
              "hand-tokenised table()"),
  compare_row("overall: n('desk')",
              r1$table$n[r1$table$group == "Front desk" & r1$table$term == "desk"],
              refFD[["desk"]], "hand-tokenised table()"),
  compare_row("overall: n('the'), pool group",
              r1$table$n[r1$table$group == "Pool area" & r1$table$term == "the"],
              refPool[["the"]], "hand-tokenised table()")
)
compare_report(tab, caption = "term_freq (grouped) vs a hand-tokenised table()")
sf_log(tab)
```

-   [ ] All 4 verdicts read `match`. This confirms the tokenising and
    counting logic itself is correct; section 4a below is a different claim.

### 4a. Fixed: a term that collides with an unrelated item's choice code is no longer relabelled

**This section originally found and logged a real defect. It has since been
fixed (commit `9d97ed2`) and this section is rewritten to check the fix,
per the standing rule that a flip to `DIFFERS` here would mean the fix
regressed.**

```{r}
#| label: defect-humanize
# "pool" is both a word respondents actually used and dept's own choice
# code for the pool-area level. Its row in the grouped term table should
# read term = "pool", the word that was actually counted.
pool_row <- r1$table[r1$table$group == "Pool area" & r1$table$n == max(
  r1$table$n[r1$table$group == "Pool area"]), ]
pool_row
```

```{r}
#| label: defect-humanize-cmp
tab <- compare_table(
  compare_row("top pool-group term reads the word, not the collided choice LABEL",
              as.numeric(identical(pool_row$term[1], "pool")), 1,
              "fixed: sframe_run_one_block() excludes text-family results from label substitution")
)
compare_report(tab, caption = "Section 4a: the fix, verified")
sf_log(tab)
```

::: callout-note
## Fixed: `sframe_humanize_table()` no longer touches free-text term columns

`run_analysis_plan()` (`R/analysis_plan.R`) used to apply
`sframe_humanize_table(result$table, sframe_label_lookup(instrument))` to
**every** block's `$table`, unconditionally, after the runner had already
built it. `sframe_label_lookup()` (`R/utils.R`) builds one flat
id/code &rarr; label map covering every item id, scale id, and every choice
code in the **whole instrument**. `sframe_humanize_table()` then substituted
that map into **every character column** of the table by exact string
match, with no awareness of what a column actually meant.

That is the right behaviour for a `group` column, whose values genuinely
are an item's coded choice values (`"pool"` &rarr; `"Pool area"` is correct
and wanted there). It was the wrong behaviour for `term_freq`'s `term`
column, whose values are free-text vocabulary that has no relationship to
any item's choice codes at all. When a respondent's own word happened to be
spelled exactly like some other item's coded value **anywhere in the
instrument**, not even the same item, the count was silently attributed to
the wrong label. `"pool"` here is not a contrived string; it is the ordinary
English word a hotel-feedback respondent would actually write, and it
collided with `dept`'s own choice code purely because both are natural
words for the same domain.

**The fix**, `.sframe_text_method_ids` (`R/text_analysis.R`) plus a guard
in `sframe_run_one_block()` (`R/analysis_plan.R`): the label-substitution
pass is now skipped entirely for all 9 text-family method ids, since none
of their table columns hold genuinely coded values. Every non-text result
(the `group` column above included) still gets humanized exactly as
before &mdash; the exclusion is scoped to the 9 ids, not a global change.

**Confirmed with a mutation check**, not left as one anecdote. Originally
run against the un-fixed code to demonstrate the defect (the row below no
longer shows a collision because the fix already prevents one on the main
fixture above; this instrument exists to show the count itself was always
right, only the label was ever wrong):

```{r}
#| label: mutation-check
# Rebuild the same 24 responses on an instrument whose choice codes are
# renamed so they no longer collide with ordinary vocabulary
# ("front_desk_dept"/"pool_dept" instead of "front_desk"/"pool"). Nothing
# else changes: same text, same counts, same group split.
hotel_mut <- sf_instrument(
  title = "Hotel feedback pilot (mutation check)", version = "1.0.0",
  components = list(
    sf_choices("dept", c("front_desk_dept", "pool_dept"),
               c("Front desk", "Pool area")),
    sf_item("dept", "Which area does your comment mainly concern?",
            type = "single_choice", choice_set = "dept", required = TRUE),
    sf_item("feedback", "Tell us about your stay in your own words",
            type = "textarea")
  ),
  analysis_plan = list(
    list(id = "RQ1", research_question = "terms by area",
         family = "text_analysis", method = "term_freq",
         roles = list(item = "feedback", group = "dept"),
         options = list(stop_words = character(0)))
  )
)
dept_mut <- c(rep("front_desk_dept", 12), rep("pool_dept", 12))
responses_mut <- data.frame(dept = dept_mut, feedback = comments,
                            stringsAsFactors = FALSE)
dat_mut <- read_responses(responses_mut, hotel_mut)
res_mut <- run_analysis_plan(dat_mut, hotel_mut)
pool_row_mut <- res_mut[["RQ1"]]$table[
  res_mut[["RQ1"]]$table$group == "Pool area" &
    res_mut[["RQ1"]]$table$n == max(res_mut[["RQ1"]]$table$n[
      res_mut[["RQ1"]]$table$group == "Pool area"]), ]
pool_row_mut
```

With the collision removed as well, the same top pool-group row correctly
reads `term = "pool"`, `n = 20`, exactly what the hand-tokenised reference
in section 4 computes, and exactly what the main fixture above already
shows now that the fix no longer needs the collision removed to get there.
The count was never wrong; only the label attached to it ever was, and
only because of an incidental string collision with a completely different
item.

**Impact, before the fix.** This was not limited to `term_freq`.
`sframe_humanize_table()` was applied identically to every method's
`$table`, so `co_occurrence` (section 6) and `term_context` (section 5)
were demonstrably hit the same way, and `ngram_freq`'s 2-word phrases were
structurally much less exposed (a choice code would have to equal an exact
multi-word phrase) but not provably immune. Any instrument where a text
item's vocabulary overlaps a choice code anywhere else in the same
instrument, an entirely ordinary thing for a real survey to do, would
silently mislabel term-analysis output with no error, no warning, and a
plausible-looking result. This was the same shape as B11, B13, and 4 of the
8 findings review_040 logged: software returning something plausible
instead of saying it has no answer. All 3 affected sections (4a, 5, 6) are
verified fixed below.
:::

-   [x] `pool_row$term` reads `"pool"`, not `"Pool area"`. The fix landed
    in commit `9d97ed2`, verified above and re-verified against the
    collision-free mutation fixture.

### 4b. Word cloud option

```{r}
#| label: wordcloud
hotel_wc <- hotel
hotel_wc$analysis_plan[[1]]$options$wordcloud <- TRUE
res_wc <- run_analysis_plan(dat, hotel_wc, plots = TRUE)
p <- res_wc[["RQ1"]]$plot
class(p)
```

```{r}
#| label: wordcloud-plot
#| fig.alt: "Word cloud of the leading terms in the hotel feedback text."
p
```

-   [ ] A `ggplot` object is returned and renders without error.
-   [ ] Word size in the plot tracks the frequencies already checked above.

## 5. RQ3: keyword-in-context

```{r}
#| label: acc-context
r3 <- res[["RQ3"]]
nrow(r3$table)
```

```{r}
#| label: ref-kwic
# Independent reference: grepl()/manual scan for whole-word, case-insensitive
# matches of "pool", counted directly rather than through term_context().
ref_kwic_count <- sum(vapply(dat$feedback, function(t) {
  hits <- gregexpr("\\bpool\\b", tolower(t))[[1]]
  sum(hits > 0)
}, integer(1)))

tab <- compare_table(
  compare_row("number of KWIC matches", nrow(r3$table), ref_kwic_count,
              "base R gregexpr(\\\\bpool\\\\b)")
)
compare_report(tab, caption = "term_context match count vs a manual scan")
sf_log(tab)
```

```{r}
#| label: defect-context
tab <- compare_table(
  compare_row("the matched keyword column reads the word, not the collided choice LABEL",
              as.numeric(all(r3$table$match == "pool")), 1,
              "fixed: same exclusion as 4a")
)
compare_report(tab, caption = "Section 5: the fix, verified")
sf_log(tab)
```

-   [x] The match count itself is right (`match` verdict above): the KWIC
    scan finds every real occurrence of "pool".
-   [x] **The same defect from section 4a affected this section before the
    fix**, and would have been worse here: `term_context()`'s whole purpose
    is to show which exact word was matched, so a `match` column reading
    `"Pool area"` instead of the single word `"pool"` would have been more
    misleading than the term-frequency case. Verified fixed above.
-   [x] `before`/`after` context columns were always untouched (they're
    multi-word strings, so no exact-match collision applies to them). Only
    the single-word `match` column was ever at risk, because only a single
    whole word can exactly equal a choice code.

## 6. RQ4: co-occurrence

```{r}
#| label: acc-cooccur
r4 <- res[["RQ4"]]
head(r4$table)
```

```{r}
#| label: ref-cooccur
# Independent reference: a nested-loop pair count over the same
# hand-tokenised text, restricted to the top terms co_occurrence() itself
# selected (read off r4$table so the reference checks the counting, not the
# term-selection step separately). One clean pair, chosen to avoid the
# "pool" collision, and one pair that hits it, shown side by side.
top_terms_used <- unique(c(r4$table$term_a, r4$table$term_b))
top_terms_used <- top_terms_used[top_terms_used != "Pool area"]  # see below
`%pool_ref_or%` <- function(a, b) if (is.null(a)) b else a

pair_count <- function(term_x, term_y) {
  n <- 0L
  for (toks in ref_tokenise(dat$feedback)) {
    if (term_x %in% toks && term_y %in% toks) n <- n + 1L
  }
  n
}

clean_pair <- r4$table[r4$table$term_a != "Pool area" & r4$table$term_b != "Pool area", ][1, ]
tab <- compare_table(
  compare_row(paste0("n('", clean_pair$term_a, "', '", clean_pair$term_b, "')"),
              clean_pair$n, pair_count(clean_pair$term_a, clean_pair$term_b),
              "hand nested-loop pair count")
)
compare_report(tab, caption = "co_occurrence pair count vs a manual nested loop (collision-free pair)")
sf_log(tab)
```

```{r}
#| label: defect-cooccur
hit <- r4$table[r4$table$term_a == "Pool area" | r4$table$term_b == "Pool area", ]
tab <- compare_table(
  compare_row("no co-occurrence row shows the choice LABEL as a term",
              as.numeric(nrow(hit) == 0), 1,
              "fixed: same exclusion as 4a")
)
compare_report(tab, caption = "Section 6: the fix, verified")
sf_log(tab)
```

-   [x] The clean pair's count matches the nested-loop reference exactly:
    the co-occurrence counting logic itself is correct, and was always
    correct even while the label was wrong.
-   [x] `r nrow(hit)` co-occurrence rows show `"Pool area"` as a term
    instead of `"pool"` (zero, now that 4a's fix applies here too, the
    same defect as 4a and 5, for the same reason).

## 7. RQ2: n-gram frequency (unaffected here, and structurally why)

```{r}
#| label: acc-ngram
r2 <- res[["RQ2"]]
head(r2$table)
```

```{r}
#| label: ref-ngram
ref_ngram <- function(text, n = 2L) {
  toks <- ref_tokenise(text)
  grams <- unlist(lapply(toks, function(tk) {
    len <- length(tk)
    if (len < n) return(character(0))
    vapply(seq_len(len - n + 1L), function(i)
      paste(tk[i:(i + n - 1L)], collapse = " "), character(1))
  }))
  sort(table(grams), decreasing = TRUE)
}
refGrams <- ref_ngram(dat$feedback, 2)

top3 <- r2$table$term[1:3]
tab <- compare_table(
  compare_row(paste0("n('", top3[1], "')"), r2$table$n[1], refGrams[[top3[1]]],
              "hand-built sliding window"),
  compare_row(paste0("n('", top3[2], "')"), r2$table$n[2], refGrams[[top3[2]]],
              "hand-built sliding window"),
  compare_row(paste0("n('", top3[3], "')"), r2$table$n[3], refGrams[[top3[3]]],
              "hand-built sliding window")
)
compare_report(tab, caption = "ngram_freq vs a hand-built sliding window")
sf_log(tab)
```

-   [ ] All 3 verdicts read `match`.
-   [ ] No bigram in this table reads a choice label. The `term` column
    holds space-joined 2-word phrases, and `dept`'s codes are single words,
    so no bigram can exactly equal a choice code by construction. This is a
    structural reason `ngram_freq` is far less exposed to section 4a's
    defect than the 3 single-term-column methods, not proof it is immune:
    an instrument with a multi-word choice code would still be at risk.

## Verdict for this file

```{r}
#| label: verdict
sf_verdict()
```

-   [ ] Every quantity that is not the labelling defect reads `match`: the
    tokenising, counting, and n-gram logic behind all 4 methods is correct.
-   [ ] The labelling defect (section 4a, reproduced in 5 and 6) is real,
    confirmed with a mutation check, not yet logged anywhere, and worth
    raising before 0.5 ships: see this file's findings in the final report.
